### 2012-12-17 article.tex
Article template

\documentclass[a4paper,12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{fullpage,url}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{listings}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{color}

\title{How to make a delicious pizza}
\author{Yuto Takei \\ The University of Tokyo}

\begin{document}

\maketitle

\begin{abstract}
  This is the abstract.
\end{abstract}

\section{Introduction}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis nec felis eros. Phasellus tristique ultrices mi eget molestie. Nulla turpis augue, adipiscing et dapibus nec, porttitor sed ante. Donec ut arcu augue. Nulla facilisi. Duis tempor nisl enim. Nulla facilisi. Vivamus adipiscing consequat turpis sed dignissim.

\section{Conclusion}

\bibliographystyle{plain}
\bibliography{./biblio}

\end{document}

### 2012-12-17 article.tex

@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis nec felis eros.
-Phasellus tristique ultrices mi eget molestie.
-Nulla turpis augue, adipiscing et dapibus nec, porttitor sed ante.
-Donec ut arcu augue.
-Nulla facilisi.
-Duis tempor nisl enim.
-Nulla facilisi.
-Vivamus adipiscing consequat turpis sed dignissim.
+* Formal verification
+--> Dependent typing of SHP
+--> Horn clauses
 
+* Some
+Background for Horn clause solving
+* Dependent type determination
+
+Definition of Horn clause solving (Informal)
+
+Advantages
+* Solution space consideration
+
+* Why do we limit to integer space?
+
+\section{Algorithm Overview}
+
+\subsection{Preliminaries}
+
+$ <num> ::= ... | -2 | -1 | 0 | 1 | 2 | ... $ .
+$ <var> ::= x | y | ... $
+$ <term> ::= <num> \cdot <var> $
+$ <expr> ::=  <term> + <term> + \cdots + <num> >= 0 $
+$ <la> ::= <expr>
+         | <la> \wedge <la> $
+$ <pvar> ::= A | B | ... $
+$ <pred> ::= <pvar> ( <var>, <var>, ... ) $
+$ <hornTerm> ::= <pvar> | <pred> $
+$ <horn> ::= <hornTerm> \wedge <hornTerm> \wedge \cdots \longrightarrow
+         <hornTerm> $
+
+* We consider inequalities.
+** NOTE: Even we have =, <>, >, <=, we can normalize to >=
+         over integer space
+
+\subsection{Generating a tree}
+
+By receiving Horn clauses as an input for the algorithm, we build a
+implication tree.
+
+Def <Implication Tree>:
+
+Implication tree is such a tree that every node is labeled by logical
+atoms. The label of every node is implied by the conjunction of its
+children's label.
+
+* Renaming?
+
+\subsection{Application of Farkas' Lemma}
+
+Correct statement of Farkas' Lemma
+Relationship to Craig's Interpolation.
+
+\subsection{Solution space representation}
+
+\section{Extensions}
+
+\subsection{Extension to DAG structure}
+
+* Building graph
+
+* Splitting a graph into tree
+
+* Merging with the same-name predicate conjunctively
+
+\subsection{Linear arithmetics with disjunction}
+
+In the previous section, we assumed that the input inequalities are
+conjunctively connected.
+
+* Why do we need this extension? *
+
+We extend our inequalities to include disjunctions in the linear
+equations.
+
+$ <la> ::= <expr>
+         | <la> \wedge <la>
+         | <la> \vee <la> $
+
+* Predicates on the direct path from the leaf with disjunction to the
+root are connected by disjunction
+* Other predicates are connected by conjunction
+
+* Why do we process LA-disjunction before DAG?
+** DAG to LA-Disjunction are invalid???
+
+\subsection{Merging of pred-pred}
+
+* Simplification between clauses
+
+\subsection{General Horn clauses with Loops}
+
+No limitation Horn clause solving
+
+* Merging pred-pred between looped same name
+
+\section{Experiment}
+\section{Future Work}
 \section{Conclusion}

### 2012-12-19 article.tex

@@
-\documentclass[a4paper,12pt]{article}
+\documentclass{llncs}
 
-\usepackage[utf8]{inputenc}
 \usepackage{fullpage,url}
@@
 \usepackage{listings}
-\usepackage{amssymb,amsmath,amsthm}
-\usepackage{color}
+\usepackage{syntax}
 
 \title{How to make a delicious pizza}
-\author{Yuto Takei \\ The University of Tokyo}
+\author{Yuto Takei}
+\institute{The University of Tokyo}
 
@@
 
-* Why do we limit to integer space?
+Discussion on Real space
 
@@
 
-$ <num> ::= ... | -2 | -1 | 0 | 1 | 2 | ... $ .
-$ <var> ::= x | y | ... $
-$ <term> ::= <num> \cdot <var> $
-$ <expr> ::=  <term> + <term> + \cdots + <num> >= 0 $
-$ <la> ::= <expr>
-         | <la> \wedge <la> $
-$ <pvar> ::= A | B | ... $
-$ <pred> ::= <pvar> ( <var>, <var>, ... ) $
-$ <hornTerm> ::= <pvar> | <pred> $
-$ <horn> ::= <hornTerm> \wedge <hornTerm> \wedge \cdots \longrightarrow
-         <hornTerm> $
+\setlength{\grammarparsep}{-2pt}
+\setlength{\grammarindent}{6em}
+\begin{grammar}
+
+<num> ::= All integers
+
+<var> ::= Arithmetic variables
+
+<term> ::= $<num> \cdot <var>$
+
+<expr> ::=  $<term> + <term> + \cdots + <num> \leq 0$
+
+<la> ::= $<expr>$ | $<la> \wedge <la>$
+
+<pvar> ::= Predicate variables
+
+<pred> ::= $<pvar> ( <var>, <var>, ... )$
+
+<hTerm> ::= <la> | <pred>
+
+<horn> ::= $<hTerm> \wedge <hTerm> \wedge \cdots \longrightarrow <hTerm>$
+
+\end{grammar}
 
@@
 
-Def <Implication Tree>:
+Def <Resolution Tree>:
 
@@
 
-$ <la> ::= <expr>
-         | <la> \wedge <la>
-         | <la> \vee <la> $
+\setlength{\grammarindent}{2em}
+\begin{grammar}
+<la> ::= <expr> | $<la> \wedge <la>$ | $<la> \vee <la>$
+\end{grammar}
 
@@
 
+\section{Completeness and Complexity}
+
 \section{Experiment}
+
+We limited to integer space
+
 \section{Future Work}

### 2012-12-19 article.tex

@@
 
-\usepackage{fullpage,url}
 \usepackage{algorithm}
@@
 
-\title{How to make a delicious pizza}
+\title{Solving Horn clauses over LI}
 \author{Yuto Takei}
@@
 \begin{document}
-
 \maketitle
-
 \begin{abstract}
@@
 
-Definition of Horn clause solving (Informal)
+Horn clause solving is a problem of constraint solving that is consisted of multiple Horn clauses with unknown predicate symbols.
+The solution is an assignment for such symbols so that all given Horn clauses stay valid.
+There may be multiple solutions for one problem, and the algorithm is expected to show at least one.
 
-Advantages
-* Solution space consideration
+In this paper, we present a novel algorithm to solve Horn clauses.
+As the unique feature of the algorithm, we preserve the solution space throughout the computation.
+We are able to provide multiple assignments for predicate symbols.
+Additionally, this technique enables us to handle general Horn clauses rather than the recursion-free ones, which was the major restriction of the previous work.
 
-Discussion on Real space
+The rest of our paper is structured as follows: Section 2 shows the basic algorithm to solve recursion-free Horn clauses with solution space consideration.
+Section 3 extends this algorithm so that it works on general settings.
+Section 4 discusses the complexity and the completeness of the extended algorithm.
+Section 5 illustrates the experimental result.
+Finally Section 6 brings the conclusion.
 
-\section{Algorithm Overview}
+\section{Algorithm}
 
@@
 
-\setlength{\grammarparsep}{-2pt}
 \setlength{\grammarindent}{6em}
@@
 
-<num> ::= All integers
+<t> ::= $n$ | $nv$ | $t + t$
 
-<var> ::= Arithmetic variables
+<e> ::= $t \leq 0$ | $e \wedge e$
 
-<term> ::= $<num> \cdot <var>$
+<p> ::= $P (v, v, ... )$
 
-<expr> ::=  $<term> + <term> + \cdots + <num> \leq 0$
+<x> ::= $e$ | $p$
 
-<la> ::= $<expr>$ | $<la> \wedge <la>$
-
-<pvar> ::= Predicate variables
-
-<pred> ::= $<pvar> ( <var>, <var>, ... )$
-
-<hTerm> ::= <la> | <pred>
-
-<horn> ::= $<hTerm> \wedge <hTerm> \wedge \cdots \longrightarrow <hTerm>$
+<h> ::= $x \wedge x \wedge \cdots \longrightarrow x$
 
@@
 * We consider inequalities.
-** NOTE: Even we have =, <>, >, <=, we can normalize to >=
+** NOTE: Even we have =, $\neq$, >, <=, we can normalize to >=
          over integer space
@@
 
-Def <Resolution Tree>:
+Def <Implication Tree>:
 
@@
 
+?
+
 \section{Experiment}

### 2012-12-20 article.tex

@@
 
+The input for our algorithm is a set of Horn clauses.
+They contain linear inequalities on real numbers and unknown predicate symbols.
+With having real numbers $n$ and variables $v$, the syntax is defined as follows.
+Predicate symbols $P$ are represented by capital alphabet through this paper.
+
 \setlength{\grammarindent}{6em}
@@
 
-<x> ::= $e$ | $p$
+<x> ::= $e$ | $p$  % Horn term
 
@@
 
-* We consider inequalities.
-** NOTE: Even we have =, $\neq$, >, <=, we can normalize to >=
-         over integer space
+Although we limit the input only to linear inequalities with $\leq$, we will later extend to any linear arithmetic expressions by introducing disjunctions in $e$.
 
-\subsection{Generating a tree}
+\subsection{Generating a implication tree}
 
@@
 
+---
 Def <Implication Tree>:
+Implication tree is such a tree that every node is labeled by a Horn term. As the semantics of the tree, the Horn term of every node is implied by the conjunction of its children's Horn terms.
+---
 
-Implication tree is such a tree that every node is labeled by logical
-atoms. The label of every node is implied by the conjunction of its
-children's label.
+In the implication tree, every relationship from a certain node to its children corresponds to one Horn clause of the original input.
+All leaves and the root node have linear expressions $e$ while all other have predicate symbols $p$.
+Therefore
+
+We show the pseudo code for generating a tree.
+
+\begin{algorithm}
+\caption{$\mbox{\sc BuildTree}$}
+\begin{algorithmic}
+\item pseudo-code
+\end{algorithmic}
+\end{algorithm}
+
+While building a tree, renaming should be accordingly to make the construct.
 

### 2012-12-24 article.tex

--- 2012-12-19 article.tex
@@
 
+In this section, we present a Horn clause solving algorithm with the conventional restriction, i.e., the input Horn clauses are recursion-less.
+However unlike the existing methods, our algorithm returns the set of solutions against a problem.
+
+<The problem example>
+
+For example, the above program input has at least two solutions as follows (but not limited to these).
+
+<Solution 1>     <Soltuion 2>
+
+Our method yields a set of solutions which includes two above. It is consisted of following major steps.
+
+\begin{enumerate}
+\item Creation of a implication tree out of input Horn clauses
+\item Generation of linear constraints under consideration of all input linear inequalities
+\item Output of a solution space with an assignment for all predicate symbols
+\end{enumerate}
+
+We give the definition of our problem and input/output of our algorithm with its syntax and semantics.
+
 \subsection{Preliminaries}
+
+In our problem setting, we are given a set of Horn clauses over linear arithmetic.
+They are consisted of linear inequalities on real space and unknown predicate symbols with parameters.
+Given real numbers $n$ and the binary relationship $leq$, the syntax is defined as follows.
+Let $v$ denote the real variables and $P$ denote the unknown predicate variables.
 
@@
 
-* We consider inequalities.
-** NOTE: Even we have =, $\neq$, >, <=, we can normalize to >=
-         over integer space
+For the above syntax, we give the following interpretations.
 
-\subsection{Generating a tree}
+<semantics definition>
+[0=0] Top
+[0=1] Bot
 
-By receiving Horn clauses as an input for the algorithm, we build a
-implication tree.
+We further impose following constraints on our input.
 
-Def <Implication Tree>:
+\begin{enumerate}
+\item The input does not have a Horn clause with only linear inequalities on both left-hand and right-hand of the clause.
+Even if there exists one, it does not affect the satisfiability of other clauses regardless of a predicate assignment if it is valid, and the whole problem does not have a solution otherwise.
+Thus, this constraint does not affect the generality of the problem.
 
-Implication tree is such a tree that every node is labeled by logical
-atoms. The label of every node is implied by the conjunction of its
-children's label.
+\item (Relaxed in section 3.1) Each predicate symbols appears exactly once in both left-hand and right-hand on different Horn clauses (recursion-less Horn clauses).
 
-* Renaming?
+\item For the convenience of the algorithm explanation, exactly one Horn clause has linear inequality on its right-hand.
+As the background of this constraint, we can split a problem with more than one such inequality into two smaller problems without a loss of generality by the previous constraint.
 
-\subsection{Application of Farkas' Lemma}
+\item (Relaxed in section 3.4) When considering the relationship among Horn clauses by implications based on predicate symbols, it must not form a loop.
 
-Correct statement of Farkas' Lemma
-Relationship to Craig's Interpolation.
+(TODO: We want to customize the bullet head to something like (A-1)...(A-4) for future reference.)
 
-\subsection{Solution space representation}
+\end{enumerate}
 
-\section{Extensions}
+Now we define the solution format for the problem.
+A solution to Horn clause solving problem is expressed by a set of assignment to unknown predicate symbols.
+Each assignment substitutes a predicate symbol with a linear equation.
+It must only includes variables which are bound by the parameter list of the predicate symbol.
+Given the set of assignments as a solution, all input Horn clauses must be valid after substitution.
 
-\subsection{Extension to DAG structure}
+<definition of solution>
 
-* Building graph
+Our algorithm receives a problem as defined above and, unlike existing algorithms, yields a pair of linear constraints on parameter variables and a set of assignments of parameterized linear expressions.
 
-* Splitting a graph into tree
+<output specification>
+$$x := linear constraint on parameters
+\theta := { [parameterized expressions on P / P] | P }
+out := < x, \theta >
+[ \theta P ] = [ parameterized expressions ]$$
 
-* Merging with the same-name predicate conjunctively
+In order to generate a solution to a problem out of our algorithm output, it is required to somehow find a set of values that satisfy linear constraints on parameters, and instantiate the parameterized linear expressions in the assignments.
 
-\subsection{Linear arithmetics with disjunction}
+\subsection{Generation of Implication Tree}
 
-In the previous section, we assumed that the input inequalities are
-conjunctively connected.
+Our algorithm first generates an implication tree from the input Horn clauses.
 
-* Why do we need this extension? *
+\paragraph{Definition<Implication tree>} An implication tree is such a tree that all nodes are labeled by a Horn term.
+As the semantics of the tree, a Horn term on each node is implied by the conjunction of Horn terms on its children nodes.
 
-We extend our inequalities to include disjunctions in the linear
-equations.
+In this paper, we may denote by a term "Horn term" a node itself in a implication tree unless it becomes ambiguous.
+
+From a given Horn clause, we create a implication tree by letting the right-hand Horn term  be a parent node and each left-hand terms be children nodes.
+Because all predicate symbols appears exactly once on both left-hand and right-hand side of different clauses by the given restriction, it is possible to graft two different implication trees T1 and T2 at the predicate symbol P, given T1 which is generated from the Horn clause with P on its left-hand, and T2 equally that P on right-hand.
+Renaming on variables should accordingly take place when grafting trees.
+We show the pseudo-code for tree generating algorithm.
+
+\begin{algorithm}
+\caption{$\mbox{\sc BuildTree}$}
+\begin{algorithmic}
+\item pseudo-code
+\end{algorithmic}
+\end{algorithm}
+
+The following statement is true about a generated implication tree.
+
+\paragraph{Characteristics} Let $T$ denote an implication tree generated from a problem expressed by the set of Horn clauses $HC$.
+Every Horn term $x$ in $T$ that has children $Y$ corresponds to exactly one Horn clause Y |- x in HC.
+In the other way, a Horn clause (a, b, c |- z) in HC corresponds to exactly one sub-tree in T that consisted of node z with its children a, b, c.
+
+All leaves and the root in the implication tree have a linear expression $e$, while others have a predicate variable $P$.
+Let $e_star$ denote the root.
+The following statement must be satisfied for a predicate symbol $P$, given $e_1,cdots,e_m$ as $P$'s descendants and $e_(m+1),cdots,e_n$ as other leaves.
+
+$$E1,\cdots,em |- P. P \wedge (e(m+1), \cdots en) |- e_\star$$    	(1)
+
+(TODO: How do we add expression numberings?)
+
+For the later convenience, we represent $\neg e_\star$ as $e_(n+1)$.
+
+\subsection{Craig's Interpolation}
+
+Given two inconsistent prepositions A and B, we call a new preposition I a Craig's interpolant of A and B such that $I$ is implied by $A$, $I$ and $B$ are inconsistent, and $I$ includes only free variables that are used in both $A$ and $B$.
+The preposition P in \ref{expr:1} is a Craig's interpolant of $e_1,\cdots,e_m$ and $e_(m+1),\cdots,e_n,e_(n+1)$ by using the fact $e_1,\cdots,e_n,e_(n+1) |- bot$.
+We compute such interpolant by Farkas' Lemma.
+
+\paragraph{Farkas' Lemma on linear inequalities} Let a linear inequality $e_i$ be represented as $a_i1 x_1 + \cdots + a_im x_m <= a_i0$.
+Assuming that $e_1,\cdots,e_n$ implies $e_0$, there exists $\lambda_1,\cdots,\lambda_n$ that satisfy $a_0j =\sum_(i=1)^n \lambda_i * a_ij (j=0...m)$.
+
+(TODO: what should I do with the operator for $e_(n+1)$ ? It is a negated expression.)
+
+Some variables in some inequalities may not exists in other inequalities.
+In such a case, we consider that the variables has a coefficient of $0$.
+According to the lemmma above, $\lambda_i$ exists with satisfying the constraint below when considering linear inequalities $e_1,\cdots,e_m$ and $e_(m+1),\cdots,e_n,e_(n+1)$.
+
+<expr: Coeff's constraint
+Const's constraint>
+
+The interpolant I is then written as $c_1 x_1 + \cdots + c_m x_m <= c_0$ with parameterized coefficients $c_j$ satisfying $c_j =\sum_(i=1)^m \lambda _i * a_ij (j=0...m)$.
+For intuitive understanding, it is the linear combination of $e_1,...,e_m$ by the weight $\lambda _1,...\lambda _m$.
+Note that, with any assignments for $\lambda _i$ under the constraint, I is a correct interpolant.
+
+\subsection{Solution Space Generation}
+
+Let the predicate symbol $P$ have linear inequalities $e_1,...,e_m$ as its decendants in the implication tree.
+We denote all other leaves by $e_(m+1),...,e_n$.
+Our algorithm assigns to $P$ an interpolant between $e_1,...,e_m$ and $e_(m+1),...,e_n,e_(n+1)$.
+We hereby note that the constraint for $\lambda_i$ (see \ref{expr:2}) to obtain such interpolant.
+That constraint is common among computing the assignments to all predicate symbols.
+Thus we return the parameterized assignments and the constraint as the algorithm's output.
+
+As we described in \ref{subsection:Preliminaries}, the solution for the Horn clause solving problem is the instanciation of all parameterized substitutions by solving the linear constraints of parameters.
+The correctness of the output as the problem solution can be verified by the induction on implication trees from characteristics \ref{}.
+We let $<a_1,...,a_(n+1)>$ represent one solution of $\lambda_i$ after solving the constraint.
+Indices in the statements below can be changed without loosing generality.
+
+\begin{enumerate}
+\item We consider the predicate symbol node P with only linear inequality nodes as children $e_1,...,e_m$.
+The algorithm assigns an interpolant between $e_1,...,e_m$ and all other inequalities $e_(m+1),...,e_(n+1)$ to $P$.
+Therefore $\theta P$ is a linear combination among $e_1,...,e_m$ with the weights $a_1,...,a_m$, and thus implied by the conjunction of $e_1,...,e_m$.
+The Horn clause implying P is valid.
+
+\item We consider other predicate symbold P, having predicate symbols $Q_1,...,Q_k$ and linear expressions $e_1,...,e_m$ (if any) as its children.
+Assume that Horn clauses implying $Q_1,...,Q_k$ respectively are all valid.
+When we denote $Q_i$'s linear inequality decendants as $e_(m+t_1+\cdots+t_(i-1)+1),...,e_(m+t_1+\cdots+t_i)$, $P$'s assignment by the algorithm is an interpolant between $e_1,...,e_m+a_1+\cdots+a_k$ and $e_m+a_1+\cdots+a_k,\cdots,e_(n+1)$.
+While this interpolant is a linear combination among $e_1,...,e_m+a_1+\cdots+a_k$ by weights $a_1,...,a_m+a_1+\cdots+a_k$, when we note $\theta Q_i$ is a linear combination of $e_(m+t_1+\cdots+t_(i-1)+1),...,e_(m+t_1+\cdots+t_i)$ by weight $a_(m+t_1+\cdots+t_(i-1)+1),...,a_(m+t_1+\cdots+t_i)$, then $\theta P$ is clearly a linear combination among $e_1,...,e_m$ by weights $a_1,...,a_m$ and $\theta Q_1,...,\theta Q_k$ by weight all $1$.
+Therefore, the Horn clause implying P is valid.
+
+\item We consider the root node $e_star$, having predicate symbols $Q_1,...,Q_k$ (if any) and linear expressions $e_1,...,e_m$ (if any) as its children.
+Let $e_(m+1),...,e_n$ be the all decendants of $Q_1,...,Q_k$.
+We can obtain the following expression from the $a_i$'s linear constraint.
+$$ \theta Q_1 \wedge \cdots \wedge \theta Q_k \wedge e_1 \wedge \cdots e_m \wedge \neg e_\star |- \bot$$
+By moving the $e_\star$ to the right-hand yields the Horn clause that we want.
+\end{enumerate}
+
+Therefore, we can say that the our alogrithm returns a valid solution space.
+As for the example shown in the beginning of this section, our algorithm returns the following linear constraint and parameterized assignments.
+
+<algo output>
+
+By solving a linear constraint, we may obtain following solutions for example.
+
+(a1,....) = (~~~)
+P: foo
+Q: bar
+
+\section{Applications}
+
+\subsection{Supporting DAG-structured Horn clauses}
+
+In the previous section, we imposed a restriction that each predicate symbol appears exactly once on both sides of different Horn clauses without making a looped dependency (A-2).
+Now we relax this by allowing the multiple appearance of predicate symbols without a looped dependency.
+Accompanied by this change, we relax (A-3) with consideration of multiple Horn clauses that the same predicate symbols implies different linear inequalities.
+
+With the above relaxations, we perform the following changes to our algorithms.
+
+\begin{itemize}
+\item After receiving the input, we build an implication graph rather than an implication tree.
+\item From the inequality node without outgoing edges, we traverse the implication graph to make implication trees.
+If we encounter the same predicate symbols multiple times during the traversal, we duplicate treat the symbol as a different one with renaming.
+\item Finally we unify the duplicated symbols into one.
+\end{itemize}
+
+Firstly, we define an implication graph as follows.
+
+\paragraph{Definition<Implication graph>} An implication graph is a directed multigraph that all nodes are labeled by a Horn term and all edges are labeled by a set of variable renaming.
+Semantically, a Horn term on each node is implied by Horn terms on its predecessor nodes with all variables on each node renamed by the edge's renaming rules.
+
+The modified algorithm in this section generates an implication graph from input Horn clauses.
+
+<pseudo-code>
+
+The following statement is true about a generated graph.
+
+\paragraph{Characteristics} Let $G$ denote an implication graph generated from a problem expressed by the set of Horn clauses $HC$.
+Every Horn term $x$ in $G$ that has children $Y$ corresponds to exactly one Horn clause Y |- x in HC.
+In the other way, a Horn clause (a, b, c |- z) in HC corresponds to exactly one sub-graph in $G$ that consisted of node z with its children a, b, c.
+(TODO: Do we need this description?
+Do we use this characteristics later?)
+
+\paragraph{Characteristics} All the linear inequality nodes have degree 1.
+That is, each node has exactly one outgoing edge to a predicate symbol node, or otherwise exactly one incoming edge from a predicate symbol node.
+
+For ease of algorithm definition, we introduce a new constraint (A-3').
+
+(A-3') When generating an implication graph from input Horn clauses, the implication graph must be connected.
+If we have a non-connected graph for the implication graph, we are able to split the problem into two smalle problems which do not share unknown predicate symbols.
+
+The constraint (A-4) remains valid in this subsection. This is equivalent to that the implication graph is acyclic.
+
+Next, we present the traversing algorithm on the implication graph.
+It starts traversing from the linear inequality node without outgoing edges, and generates a set of implication trees, that are processed by the same algorithm as the previous section.
+Although we may visit the same predicate symbol multiple times, we handle it by renaming the predicate symbols differently every time and unify all those copies in the end.
+Shown below is the pseudo-code for the traversal.
+
+<pseudo-code>
+Input: Implication graph
+Output: Mapping from pred symbol to a set of its copies, and a set of impl trees
+
+Implication trees generated from the algorithm is processed by the same method in the previous section.
+We obtain a linear constraints of parameters and an parameterized assignment for each implication tree.
+
+Finally, we explain about the unification.
+We duplicated the predicate symbols for every appearance during the traversal, and we notice that all of them are implied by the same hypothesis.
+Therefore, we create a conjunction of all parameterized assignments and let it be the new assignment for the original predicate symbol.
+By this method, we can merge several several solution spaces out of different implication trees into one solution space of the original problem.
+
+Merging two solution spaces $(c_1,\theta_1)$ and $(c_2,\theta_2)$ into $(c, \theta)$ is done as follows.
+
+$$P_1, P_2, ... \in Copies of P
+c := c_1 /\ c_2
+\theta := { [ x1 /\ x2 / P ] | [x1/P1], [x2/P2] \in \theta_1 \cup \theta_2 }$$
+
+\section{Linear arithmetic with disjunction}
+
+We assumed that all the linear inequalities were connected conjunctively. However, when we try to infer dependent types for the following SHP, it gives raise to the occurrence of disjunction.
+
+<example of SHP>
+<Horn clauses with disjunctions>
+
+For this reason, we must extend our algorithm to support disjunctions in linear inequalities. The extended grammar syntax is shown below.
 
@@
 
-* Predicates on the direct path from the leaf with disjunction to the
-root are connected by disjunction
-* Other predicates are connected by conjunction
+By this change, our algorithm is affected for following points.
 
-* Why do we process LA-disjunction before DAG?
-** DAG to LA-Disjunction are invalid???
+(TODO)
+\begin{itemize}
+\item Linear inequality node becomes a disjunction of linear inequalities.
+\item For pred symbols on the path from such node with disjunction to the root, the assignment is connected by $\vee$
+\item Other preds are connected by $\wedge$
+\item First: Graph traversal to trees. Then: Choose one from disjunction.
+	(TODO: Why the other way is NG?)
+\end{itemize}
 
-\subsection{Merging of pred-pred}
+\section{Unification of different predicate symbols}
+Our suggested algorithm only handles LA and not UIF.
+For this reason, we cannot treat Boolean values in predicate symbol parameters.
+For an intuitive example, the following Horn clause solving problem is not valid.
 
-* Simplification between clauses
+<Horn clauses with bool>
 
-\subsection{General Horn clauses with Loops}
+Therefore, we tried to work around this problem by extending Horn clauses with Boolean parameters as follows.
 
-No limitation Horn clause solving
+<expanded Horn clauses>
 
-* Merging pred-pred between looped same name
+In this case, the original predicate symbol $P$ is passed to our algorithm as $P_T$ and $P_F$, and we want to obtain the same preposition for them if possible.
+
+(TODO)
+
+\subsection{Restriction-less Horn clauses}
+
+TODO:
+Unification of predicate symbols over expanded loop.
+$P_1$, $P_2$, ...
+Failure to unify means the absence of solution (complete)
+
+citation from PPDP paper
 

### 2013-04-18 article/algo.tex
Separated articles * May create 'src' directory in future

--- 2012-12-19 article.tex
@@
-\documentclass{llncs}
+\documentclass[a4paper,12pt]{article}
 
@@
 
-\title{Solving Horn clauses over LI}
-\author{Yuto Takei}
-\institute{The University of Tokyo}
+\title{Algorithm explanation}
+\author{Yuto Takei \\ The University of Tokyo}
 
@@
 \maketitle
-\begin{abstract}
-  This is the abstract.
-\end{abstract}
-
-\section{Introduction}
-
-* Formal verification
---> Dependent typing of SHP
---> Horn clauses
-
-* Some
-Background for Horn clause solving
-* Dependent type determination
-
-Horn clause solving is a problem of constraint solving that is consisted of multiple Horn clauses with unknown predicate symbols.
-The solution is an assignment for such symbols so that all given Horn clauses stay valid.
-There may be multiple solutions for one problem, and the algorithm is expected to show at least one.
-
-In this paper, we present a novel algorithm to solve Horn clauses.
-As the unique feature of the algorithm, we preserve the solution space throughout the computation.
-We are able to provide multiple assignments for predicate symbols.
-Additionally, this technique enables us to handle general Horn clauses rather than the recursion-free ones, which was the major restriction of the previous work.
-
-The rest of our paper is structured as follows: Section 2 shows the basic algorithm to solve recursion-free Horn clauses with solution space consideration.
-Section 3 extends this algorithm so that it works on general settings.
-Section 4 discusses the complexity and the completeness of the extended algorithm.
-Section 5 illustrates the experimental result.
-Finally Section 6 brings the conclusion.
 
@@
 
-\subsection{Preliminaries}
-
-\setlength{\grammarindent}{6em}
-\begin{grammar}
-
-<t> ::= $n$ | $nv$ | $t + t$
-
-<e> ::= $t \leq 0$ | $e \wedge e$
-
-<p> ::= $P (v, v, ... )$
-
-<x> ::= $e$ | $p$
-
-<h> ::= $x \wedge x \wedge \cdots \longrightarrow x$
-
-\end{grammar}
-
-* We consider inequalities.
-** NOTE: Even we have =, $\neq$, >, <=, we can normalize to >=
-         over integer space
-
-\subsection{Generating a tree}
-
-By receiving Horn clauses as an input for the algorithm, we build a
-implication tree.
-
-Def <Implication Tree>:
-
-Implication tree is such a tree that every node is labeled by logical
-atoms. The label of every node is implied by the conjunction of its
-children's label.
-
-* Renaming?
-
-\subsection{Application of Farkas' Lemma}
-
-Correct statement of Farkas' Lemma
-Relationship to Craig's Interpolation.
-
-\subsection{Solution space representation}
-
-\section{Extensions}
-
-\subsection{Extension to DAG structure}
-
-* Building graph
-
-* Splitting a graph into tree
-
-* Merging with the same-name predicate conjunctively
-
-\subsection{Linear arithmetics with disjunction}
-
-In the previous section, we assumed that the input inequalities are
-conjunctively connected.
-
-* Why do we need this extension? *
-
-We extend our inequalities to include disjunctions in the linear
-equations.
-
-\setlength{\grammarindent}{2em}
-\begin{grammar}
-<la> ::= <expr> | $<la> \wedge <la>$ | $<la> \vee <la>$
-\end{grammar}
-
-* Predicates on the direct path from the leaf with disjunction to the
-root are connected by disjunction
-* Other predicates are connected by conjunction
-
-* Why do we process LA-disjunction before DAG?
-** DAG to LA-Disjunction are invalid???
-
-\subsection{Merging of pred-pred}
-
-* Simplification between clauses
-
-\subsection{General Horn clauses with Loops}
-
-No limitation Horn clause solving
-
-* Merging pred-pred between looped same name
-
-\section{Completeness and Complexity}
-
-?
-
-\section{Experiment}
-
-We limited to integer space
-
-\section{Future Work}
-\section{Conclusion}
-
-\bibliographystyle{plain}
-\bibliography{./biblio}
-
 \end{document}

### 2013-04-18 article/algo.tex
Algorithm explanaion added * Mostly quoted from the email by Prof. Kobayashi

@@
 
+\usepackage{amsmath}
 \usepackage{algorithm}
@@
 \usepackage{listings}
+\usepackage{stmaryrd}
 \usepackage{syntax}
+
+\newcommand{\sembrack}[1]{[\![#1]\!}
 
@@
 
+The Horn clause solving algorithm receives an input problem in Horn
+graph. A Horn graph is a labelled directed graph $G=(V,E,\varphi)$
+where V is a set of vertices and E is a set of edges. $\varphi: V
+\rightarrow L$ is a labelling function for nodes, where L is a set of
+elements of a predicate variable $P(x_1, ..., x_n)$ or a linear
+inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.
+
+The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
+Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V$.
+
+\begin{align*}
+\mathcal{HC}_v &= \left( \bigwedge_{u \in pred(v)} \varphi(u) \right)
+\Longrightarrow \varphi(v) \\ \llbracket G \rrbracket &= \bigwedge_{v
+  \in V} \mathcal{HC}_v
+\end{align*}
+
+The solution of $G$ is a map $\rho$ from predicate variables to linear
+predicates of the form $\lambda x_1, \cdots ,x_n. \psi $ (where $\psi$
+is a formula constructed from
+\begin{align*}
+\psi ::= &a_1 x_1 + \cdots + a_n x_n \leq b \vert \\ &\psi \wedge \psi
+\vert \psi \vee \psi \vert true \vert false
+\end{align*}
+), where $\rho[G]$ is a tautology. A solution is \textit{simple} if
+$\rho[P]$ is of the form $a_1 x_1 + \cdots + a_n x_n \leq b$ for every
+$P$ in $dom(\rho)$.
+
+The problem of solving Horn clauses (represented by $G$) is to find a
+solution of $G$ (and a simple solution if possible).
+
+We consider below a Horn graph that is acyclic.
+
+Given an acyclic Horn graph (or called a Horn DAG), we define the
+relation $v <^+ v'$ by
+\[ v <^+ v' \Longleftrightarrow (v,v') \in E \]
+By the assumption that $G$ is acyclic, $<^+$ is a well-founded
+relation.  We assume without loss of generality that $V$ contains the
+least element $v_\bot$ with respect to $<^+$; otherwise we can always
+ignore $v_\bot$ with $\varphi(v_\bot) = true$.
+
+For each $v \in V$, we define $\Delta(v) = (a_v x \leq b_v, C_v)$,
+where
+\begin{itemize}
+\item $a_v x \leq b_v$ is a linear inequality that may contain
+  coefficient variables
+\item $C_v$ is a set of linear inequalities on coefficient variables.
+\end{itemize}
+Intuitively, the pair $(\mathbf{a} \mathbf{x} \leq b, C)$ denotes the
+set of inequalities:
+
+\[ \left\lbrace \mathbf{a} \mathbf{x} \leq b | \exists \neg a,\neg
+b. C \right\rbrace \]
+
+(where $\exists \neg a, \neg b.$ is the existential quantification
+over coefficient variables other than $a$ and $b$.
+
+$\Delta(v)$ is defined by well-founded induction on $v$ with respect
+to the relation $<^+$.
+
+Then,
+
+\begin{align*}
+\Delta(v) =
+\begin{cases}
+\begin{split}
+\Bigg(
+\sum_i { a_i a_{0,i} x_i } \leq b b_0,
+\bigwedge_i \left( - a_i a_{0,i} + \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} a_{u,i} = 0 \right) \wedge \\
+\left( \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} b_u - b (b_0+1) < 0 \right) \wedge \\
+\bigwedge_{(u,v) \in E} C_u
+\Bigg)
+& \mbox{if } \varphi(v) = \mathbf{a}_v \mathbf{x} \leq b_v
+\end{split}
+ \\
+\begin{split}
+\Bigg(
+\mathbf{a} \mathbf{x} \leq b,
+\bigwedge_i \left( - a_i + \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} a_{u,i} = 0 \right) \wedge \\
+\left( \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} b_u - (b+1) < 0 \right) \wedge \\
+\bigwedge_{(u,v) \in E} C_u
+\Bigg)
+& \mbox{if } \varphi(v) = P(\mathbf{x}).
+\end{split}
+\end{cases}
+\end{align*}
+
+where $\mathbf{a}$ and $b$ are fresh coefficient variables.
+
+If $C_{v_\bot}$ is satisfiable, then any solution $\sigma$ for it
+gives a simple solution for $G$, i.e., we should have:
+
+If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
+\[ \rho(v)= \sigma(a_v x <= b_v) \]
+is a solution of $G$.
+
 \end{document}

### 2013-04-18 article/algo.tex
Revised explanation for clarity

@@
 least element $v_\bot$ with respect to $<^+$; otherwise we can always
-ignore $v_\bot$ with $\varphi(v_\bot) = true$.
+ignore $v_\bot$ with $\varphi(v_\bot) = 0 \leq 0$.
 
@@
   coefficient variables
-\item $C_v$ is a set of linear inequalities on coefficient variables.
+\item $C_v$ is a set of linear expressions on coefficient variables.
 \end{itemize}
@@
 
-\[ \left\lbrace \mathbf{a} \mathbf{x} \leq b | \exists \neg a,\neg
-b. C \right\rbrace \]
-
-(where $\exists \neg a, \neg b.$ is the existential quantification
-over coefficient variables other than $a$ and $b$.
+\begin{align*}
+\left\lbrace
+ \mathbf{a} \mathbf{x} \leq b |
+ \exists \left( FV(C)
+  \setminus (\mathbf{a} \cup \left\lbrace b \right\rbrace
+ \right). C
+\right\rbrace
+\end{align*}
 
 $\Delta(v)$ is defined by well-founded induction on $v$ with respect
-to the relation $<^+$.
-
-Then,
+to the relation $<^+$. Then,
+$\Delta(v) = \left( \mathbf{a}_v \mathbf{x} \leq b_v, C_v \right)$,
+where
 
 \begin{align*}
-\Delta(v) =
+C_v =
 \begin{cases}
-\begin{split}
-\Bigg(
-\sum_i { a_i a_{0,i} x_i } \leq b b_0,
-\bigwedge_i \left( - a_i a_{0,i} + \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} a_{u,i} = 0 \right) \wedge \\
-\left( \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} b_u - b (b_0+1) < 0 \right) \wedge \\
-\bigwedge_{(u,v) \in E} C_u
-\Bigg)
-& \mbox{if } \varphi(v) = \mathbf{a}_v \mathbf{x} \leq b_v
-\end{split}
- \\
-\begin{split}
-\Bigg(
-\mathbf{a} \mathbf{x} \leq b,
-\bigwedge_i \left( - a_i + \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} a_{u,i} = 0 \right) \wedge \\
-\left( \sum_{\substack{(u,v) \in E \\ \Delta(u)=\mathbf{a}_u \mathbf{x} \leq b_u}} b_u - (b+1) < 0 \right) \wedge \\
-\bigwedge_{(u,v) \in E} C_u
-\Bigg)
-& \mbox{if } \varphi(v) = P(\mathbf{x}).
-\end{split}
+\left\lbrace \mathbf{a}_v = \sum_{(u,v) \in E} \mathbf{a}_u \right\rbrace \cup
+\left\lbrace b_v >= \sum_{(u,v) \in E} b_u \right\rbrace \cup
+\bigcup_{(u,v) \in E} C_u
+& \mbox{if } \varphi(v) = P(\mathbf{x}) \vee v = v_\bot \\
+\left\lbrace \mathbf{a}_v = \lambda_v \mathbf{a}_0 \right\rbrace \cup
+\left\lbrace b_v = \lambda_v b_0 \right\rbrace
+& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x} \leq b_0
 \end{cases}
@@
 
-where $\mathbf{a}$ and $b$ are fresh coefficient variables.
+Given $\varphi(v_\bot) = \mathbf{a}_0 \mathbf{x} \leq b_0$, any
+solution $\sigma$ to the constraint $C_\star = C_{v_\bot} \cup {
+  \mathbf{a}_0 = \mathbf{a}_{v_\bot} } \cup { b_0 = b_{v_\bot} }$
+gives a simple solution for $G$ if $C_\star$ is satisfiable, i.e., we
+should have:
 
-If $C_{v_\bot}$ is satisfiable, then any solution $\sigma$ for it
-gives a simple solution for $G$, i.e., we should have:
-
-If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
-\[ \rho(v)= \sigma(a_v x <= b_v) \]
+\quote
+If $\models \sigma(C_\star)$, then $\rho$ defined by
+\begin{align*}
+ \rho = \left\lbrace
+  \left( P, \sigma(\mathbf{a}_v \mathbf{x} \leq b_v) \right) |
+  \forall v \in V. \varphi(v) = P(\mathbf{x})
+ \right\rbrace
+\end{align*}
 is a solution of $G$.
+\endquote
 

### 2013-04-18 article/algo.tex
Explanation of DAG-splitting * Equivalent to copying constraints

@@
 \begin{align*}
-\psi ::= &a_1 x_1 + \cdots + a_n x_n \leq b \vert \\ &\psi \wedge \psi
-\vert \psi \vee \psi \vert true \vert false
+\psi ::= & a_1 x_1 + \cdots + a_n x_n \leq b \mid \\
+& \psi \wedge \psi \mid \psi \vee \psi \mid true \mid false
 \end{align*}
@@
 \left\lbrace
- \mathbf{a} \mathbf{x} \leq b |
+ \mathbf{a} \mathbf{x} \leq b \middle|
  \exists \left( FV(C)
@@
 
-\quote
+\begin{quote}
 If $\models \sigma(C_\star)$, then $\rho$ defined by
@@
  \rho = \left\lbrace
-  \left( P, \sigma(\mathbf{a}_v \mathbf{x} \leq b_v) \right) |
+  \left( P, \sigma(\mathbf{a}_v \mathbf{x} \leq b_v) \right) \middle|
   \forall v \in V. \varphi(v) = P(\mathbf{x})
@@
 is a solution of $G$.
-\endquote
+\end{quote}
+
+If $C_\star$ is unsatisfiable, there are two possibility for them.
+
+\begin{itemize}
+\item No solution exists for $G$.
+\item No simple solution exists for $G$.
+\end{itemize}
+
+If $G$ has node(s) which has multiple successors, the second case is
+possible. We now consider to give a solution to the such case.
+
+Let $\hat u$ be the (least) vertex on $<^+$ with multiple successors
+denoted by $succ(u)$. For each $v_i \in succ(u)$, we transform the graph
+$G$ with the following operation:
+
+\begin{align*}
+\frac{
+(u, v_i) \in E \wedge \varphi(u) = P(\mathbf{x}) \wedge
+\Delta(u) = \left( \mathbf{a} \mathbf{x} \leq b, C \right)
+}{
+\begin{aligned}
+E & := { (u, v_i'), (v_i', v_i) } \cup ( E \setminus (u, v_i) ) \\
+\varphi(v_i') & := P_i(\mathbf{x}) \\
+\Delta(v_i') & := \left( \mathbf{a}' \mathbf{x} \leq b', \textsc{Copy}(C) \right)
+\end{aligned}
+}
+\end{align*}
+
+The $\textsc{Copy}$ duplicates a given linear constraint by renaming
+all variables to fresh ones. $\mathbf{a'}$ and $b'$ are renamed
+variables from $\mathbf{a}$ and $b$ respectively.
+
+Obtaining $G'$ after the transformation above, if $C_{v_\bot}$ becomes
+satisfiable, a solution $\rho$ to $G$ is given by using solution
+$\rho'$ to $G'$:
+
+\begin{align*}
+ \rho = \rho' \cup \left\lbrace P = \bigwedge_i \rho'[P_i] \middle|
+ \varphi(\hat u) = P(\mathbf{x}) \right\rbrace
+\end{align*}
+
+Note that predicate variables on greater nodes than $\hat u$ may
+also have duplicated assignments, which should be combined in the
+same manner as $P$ above.
+
+If $C_{v_\bot}$ stays unsatisfiable in $G'$, we can repeat the
+transformation until no vertices have multiple successors.
 

### 2013-04-19 article/algo.tex
Slight fix of definition * Rename mapping on edges

@@
 \newcommand{\sembrack}[1]{[\![#1]\!}
+\newcommand{\edge}[2]{#1\rightarrow#2}
+\newcommand{\path}[2]{#1\rightarrow^\star#2}
 
@@
 The Horn clause solving algorithm receives an input problem in Horn
-graph. A Horn graph is a labelled directed graph $G=(V,E,\varphi)$
-where V is a set of vertices and E is a set of edges. $\varphi: V
-\rightarrow L$ is a labelling function for nodes, where L is a set of
-elements of a predicate variable $P(x_1, ..., x_n)$ or a linear
-inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.
+graph. A Horn graph is a labelled directed graph $G=(V,E,\varphi)$,
+where V is a set of vertices and E is a set of labelled
+edges. $\varphi: V \rightarrow L$ is a labelling function for
+vertices, where L is a set of elements of a predicate variable $P(x_1,
+..., x_n)$ or a linear inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.
+An edge from a vertex $u$ to $v$ is expressed as a 3-tuple
+$(u,v,\theta)$ where $\theta$ is a finite map from variables to
+variables.
 
@@
 \begin{align*}
-\mathcal{HC}_v &= \left( \bigwedge_{u \in pred(v)} \varphi(u) \right)
-\Longrightarrow \varphi(v) \\ \llbracket G \rrbracket &= \bigwedge_{v
-  \in V} \mathcal{HC}_v
+\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(v) \\
+\llbracket G \rrbracket & = \bigwedge_{v \in V} \mathcal{HC}_v
 \end{align*}
@@
 Given an acyclic Horn graph (or called a Horn DAG), we define the
-relation $v <^+ v'$ by
-\[ v <^+ v' \Longleftrightarrow (v,v') \in E \]
+relation $u <^+ v$ as follows.
+\[ u <^+ v \Longleftrightarrow \exists \theta. (u,v,\theta) \in E \]
 By the assumption that $G$ is acyclic, $<^+$ is a well-founded
@@
 least element $v_\bot$ with respect to $<^+$; otherwise we can always
-ignore $v_\bot$ with $\varphi(v_\bot) = 0 \leq 0$.
+ignore $v_\bot$ with $\varphi(v_\bot) = 0 \leq -1$.
 
@@
 C_v =
-\begin{cases}
-\left\lbrace \mathbf{a}_v = \sum_{(u,v) \in E} \mathbf{a}_u \right\rbrace \cup
-\left\lbrace b_v >= \sum_{(u,v) \in E} b_u \right\rbrace \cup
-\bigcup_{(u,v) \in E} C_u
-& \mbox{if } \varphi(v) = P(\mathbf{x}) \vee v = v_\bot \\
-\left\lbrace \mathbf{a}_v = \lambda_v \mathbf{a}_0 \right\rbrace \cup
-\left\lbrace b_v = \lambda_v b_0 \right\rbrace
+\bigcup_{(u,v) \in E} C_u & \cup
+\left\lbrace
+ \mathbf{a}_v = \sum_{(u,v) \in E} \mathbf{a}_u,
+ b_v \geq \sum_{(u,v) \in E} b_u
+\right\rbrace \cup
+\\
+& \begin{cases}
+\emptyset
+& \mbox{if } \varphi(v) = P(\mathbf{x}) \\
+\left\lbrace
+ \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
+ b_v = \lambda_v b_0
+\right\rbrace
 & \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x} \leq b_0
@@
 
-If $G$ has node(s) which has multiple successors, the second case is
+If $G$ has a vertex which has multiple successors, the second case is
 possible. We now consider to give a solution to the such case.
 
-Let $\hat u$ be the (least) vertex on $<^+$ with multiple successors
-denoted by $succ(u)$. For each $v_i \in succ(u)$, we transform the graph
-$G$ with the following operation:
-
-\begin{align*}
-\frac{
-(u, v_i) \in E \wedge \varphi(u) = P(\mathbf{x}) \wedge
-\Delta(u) = \left( \mathbf{a} \mathbf{x} \leq b, C \right)
-}{
-\begin{aligned}
-E & := { (u, v_i'), (v_i', v_i) } \cup ( E \setminus (u, v_i) ) \\
-\varphi(v_i') & := P_i(\mathbf{x}) \\
-\Delta(v_i') & := \left( \mathbf{a}' \mathbf{x} \leq b', \textsc{Copy}(C) \right)
-\end{aligned}
-}
-\end{align*}
+Let $\hat u$ be a vertex with multiple successors denoted by
+$succ(u)$. We now duplicate to relax the linear constraints.
 
@@
 
-Note that predicate variables on greater nodes than $\hat u$ may
+$\forall u \in V. \path{u}{v} \wedge \varphi(u) = P(\mathbf{x}).
+$
+
+Note that predicate variables on  than $\hat u$ may
 also have duplicated assignments, which should be combined in the

### 2013-04-20 article/algo.tex
Fix for consistency on symbols

@@
 least element $v_\bot$ with respect to $<^+$; otherwise we can always
-ignore $v_\bot$ with $\varphi(v_\bot) = 0 \leq -1$.
+implicitly assume $\varphi(v_\bot) = 0 \leq -1$.
 
-For each $v \in V$, we define $\Delta(v) = (a_v x \leq b_v, C_v)$,
-where
+For each $v \in V$, we define
+$\Delta(v) = (\mathbf{a}_v \mathbf{x} \leq b_v, C_v)$
+, where
 \begin{itemize}
-\item $a_v x \leq b_v$ is a linear inequality that may contain
-  coefficient variables
+\item $\mathbf{a}_v \mathbf{x} \leq b_v$ is a linear inequality that
+  may contain coefficient variables
 \item $C_v$ is a set of linear expressions on coefficient variables.
@@
 
-Given $\varphi(v_\bot) = \mathbf{a}_0 \mathbf{x} \leq b_0$, any
-solution $\sigma$ to the constraint $C_\star = C_{v_\bot} \cup {
-  \mathbf{a}_0 = \mathbf{a}_{v_\bot} } \cup { b_0 = b_{v_\bot} }$
-gives a simple solution for $G$ if $C_\star$ is satisfiable, i.e., we
+Given a set of maximal nodes
+$V_\top = \left\lbrace v \in V \middle| \neg(\exists v'. v <^+ v') \right\rbrace$,
+any model $\sigma$ to the constraint
+$C_{V_\top} = \bigwedge_{v \in V_{\top}} C_v$ gives a
+simple solution for $G$ if $C_{V_\top}$ is satisfiable, i.e., we
 should have:
@@
 \begin{quote}
-If $\models \sigma(C_\star)$, then $\rho$ defined by
+If $\models \sigma(C_{V_\top})$, then $\rho$ defined by
 \begin{align*}
@@
 
-If $C_\star$ is unsatisfiable, there are two possibility for them.
+If $C_{V_\top}$ is unsatisfiable, there are two possibilities for them.
 

### 2013-04-21 article/algo.tex
Modifying the least element to the greatest element

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{algorithm}
@@
 relation. We assume without loss of generality that $V$ contains the
-least element $v_\bot$ with respect to $<^+$; otherwise we can always
-implicitly assume $\varphi(v_\bot) = 0 \leq -1$.
+greatest element $v_\bot$ with respect to $<^+$; otherwise we can
+always implicitly assume $\varphi(v_\bot) = 0 \leq 0$.
 
@@
 \begin{align*}
-C_v =
-\bigcup_{(u,v) \in E} C_u & \cup
+C_v = & \bigcup_{(u,v) \in E} C_u
+\\
+& \cup \begin{cases}
+\emptyset
+& \mbox{if } \nexists u. u <^+ v \\
 \left\lbrace
- \mathbf{a}_v = \sum_{(u,v) \in E} \mathbf{a}_u,
- b_v \geq \sum_{(u,v) \in E} b_u
-\right\rbrace \cup
+ \mathbf{a}_v = \sum_{u <^+ v} \mathbf{a}_u,
+ b_v \geq \sum_{u <^+ v} b_u
+\right\rbrace
+& \mbox{otherwise}
+\end{cases}
 \\
-& \begin{cases}
+& \cup \begin{cases}
 \emptyset
@@
 
-Given a set of maximal nodes
-$V_\top = \left\lbrace v \in V \middle| \neg(\exists v'. v <^+ v') \right\rbrace$,
-any model $\sigma$ to the constraint
-$C_{V_\top} = \bigwedge_{v \in V_{\top}} C_v$ gives a
-simple solution for $G$ if $C_{V_\top}$ is satisfiable, i.e., we
-should have:
+Any model $\sigma$ to the constraint $C_{v_\bot}$ gives a simple
+solution for $G$ if $C_{v_\bot}$ is satisfiable, i.e., we should have:
 
 \begin{quote}
-If $\models \sigma(C_{V_\top})$, then $\rho$ defined by
+If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
 \begin{align*}
@@
 
-If $C_{V_\top}$ is unsatisfiable, there are two possibilities for them.
+If $C_{v_\bot}$ is unsatisfiable, there are two possibilities for them.
 
@@
 
-Obtaining $G'$ after the transformation above, if $C_{v_\bot}$ becomes
+Obtaining $G'$ after the transformation above, if $C_{v_\top}$ becomes
 satisfiable, a solution $\rho$ to $G$ is given by using solution

### 2013-04-21 article/algo.tex
Explanation on constraint duplication

@@
 \begin{align*}
-C_v = & \bigcup_{(u,v) \in E} C_u
+C_v = & \bigcup_{u <^+ v} C_u
 \\
@@
 If $G$ has a vertex which has multiple successors, the second case is
-possible. We now consider to give a solution to the such case.
+possible. We now try to give a solution by relaxing the constraint
+$C_{v_\bot}$.
 
-Let $\hat u$ be a vertex with multiple successors denoted by
-$succ(u)$. We now duplicate to relax the linear constraints.
-
-The $\textsc{Copy}$ duplicates a given linear constraint by renaming
-all variables to fresh ones. $\mathbf{a'}$ and $b'$ are renamed
-variables from $\mathbf{a}$ and $b$ respectively.
-
-Obtaining $G'$ after the transformation above, if $C_{v_\top}$ becomes
-satisfiable, a solution $\rho$ to $G$ is given by using solution
-$\rho'$ to $G'$:
+We slightly modify the definition $C_v$. Let $\hat V$ a subset of
+vertices with multiple successors.
 
 \begin{align*}
- \rho = \rho' \cup \left\lbrace P = \bigwedge_i \rho'[P_i] \middle|
- \varphi(\hat u) = P(\mathbf{x}) \right\rbrace
+C_v = & \bigcup_{\substack{u <^+ v \\ u \not\in \hat V}} C_u \cup
+\bigcup_{\substack{u <^+ v \\ u \in \hat V}} \textsc{Copy}_v (C_u)
+\\
+& \cup \begin{cases}
+\emptyset
+& \mbox{if } \nexists u. u <^+ v \\
+\left\lbrace
+ \mathbf{a}_{v,\emptyset} =
+  \sum_{\substack{u <^+ v \\ u \not\in \hat V}} \mathbf{a}_{u,U} +
+  \sum_{\substack{u <^+ v \\ u \in \hat V}} \mathbf{a}_{u,U \cup \left\lbrace v \right\rbrace},
+ b_{v,\emptyset} \geq
+  \sum_{\substack{u <^+ v \\ u \not\in \hat V}} b_{u,U} +
+  \sum_{\substack{u <^+ v \\ u \in \hat V}} b_{u,U \cup \left\lbrace v \right\rbrace}
+\right\rbrace
+& \mbox{otherwise}
+\end{cases}
+\\
+& \cup \begin{cases}
+\emptyset
+& \mbox{if } \varphi(v) = P(\mathbf{x}) \\
+\left\lbrace
+ \lambda_v \geq 0, \mathbf{a}_{v,\emptyset} = \lambda_v \mathbf{a}_0,
+ b_{v,\emptyset} = \lambda_v b_0
+\right\rbrace
+& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x} \leq b_0
+\end{cases}
 \end{align*}
 
-$\forall u \in V. \path{u}{v} \wedge \varphi(u) = P(\mathbf{x}).
-$
+The procedure $\textsc{Copy}_v$ renames all free variables to avoid
+name collision. Its definition is given as follows.
 
-Note that predicate variables on  than $\hat u$ may
-also have duplicated assignments, which should be combined in the
-same manner as $P$ above.
+\[\textsc{Copy}_{\hat v}(\mathbf{a}_{u,V}) = \mathbf{a}_{u,(V \cup
+  \left\lbrace \hat v \right\rbrace)} \]
 
-If $C_{v_\bot}$ stays unsatisfiable in $G'$, we can repeat the
-transformation until no vertices have multiple successors.
+All variables are then distinguished by the label of copy origin
+nodes. This intuitively means that constraints generated for nodes in
+$\hat V$ are duplicated at their successors. Note that the variables
+without the label should be interpreted as that it has $\emptyset$ for
+origin nodes.
+
+When $C_{v_\bot}$ becomes satisfiable by specifing an appropriate
+$\hat V$, a solution $\rho$ to $G$ is given as follows.
+
+\begin{align*}
+ \rho = \left\lbrace
+  \left( P, \bigwedge_S \sigma(\mathbf{a}_{v,S} \mathbf{x} \leq b_{v,S}) \right) \middle|
+  \forall v \in V. \varphi(v) = P(\mathbf{x})
+ \right\rbrace
+\end{align*}
+
+If $C_{v_\bot}$ stays unsatisfiable, we can add a vertex with multiple
+successors to $\hat V$ until all such nodes are in $\hat V$.
 

### 2013-04-24 article/algo.tex
Small changes

@@
 vertices, where L is a set of elements of a predicate variable $P(x_1,
-..., x_n)$ or a linear inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.
-An edge from a vertex $u$ to $v$ is expressed as a 3-tuple
+..., x_{n_P})$ or a linear inequality $a_1 x_1 + \cdots + a_n x_n \leq
+b$.  An edge from a vertex $u$ to $v$ is expressed as a 3-tuple
 $(u,v,\theta)$ where $\theta$ is a finite map from variables to
@@
 
-Any model $\sigma$ to the constraint $C_{v_\bot}$ gives a simple
+Any model $\sigma$ of the constraint $C_{v_\bot}$ gives a simple
 solution for $G$ if $C_{v_\bot}$ is satisfiable, i.e., we should have:

### 2013-04-27 article/algo.tex
Changes of Horn graph structure * For supporting disjunctive Horn clauses

--- 2013-04-21 article/algo.tex
@@
 The Horn clause solving algorithm receives an input problem in Horn
-graph. A Horn graph is a labelled directed graph $G=(V,E,\varphi)$,
-where V is a set of vertices and E is a set of labelled
-edges. $\varphi: V \rightarrow L$ is a labelling function for
-vertices, where L is a set of elements of a predicate variable $P(x_1,
-..., x_n)$ or a linear inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.
-An edge from a vertex $u$ to $v$ is expressed as a 3-tuple
+graph. A Horn graph is a labelled directed graph $G=(V,E,\varphi)$.
+\begin{itemize}
+\item $V$ is a union of two sets of vertices; Horn term vertices $V_T$
+  and arrow vertices $V_\Rightarrow$. Each vertex $u \in
+  V_\Rightarrow$ have eactly one outgoing edge to some $v \in V_T$,
+  denoted as $succ(u)$.
+\item $E$ is a set of labelled edges, which are expressed in the form
   $(u,v,\theta)$ where $\theta$ is a finite map from variables to
   variables.
+\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
+  term vertices, where L is a set of elements of a predicate variable
+  in the form $P(x_1, ..., x_{arity(P)})$ or a linear inequality $a_1
+  x_1 + \cdots + a_n x_n \leq b$.
+\end{itemize}
+
+There are two kinds of edges in a Horn graph;
+\begin{itemize}
+\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
+  $(u,v,\theta)$, or,
+\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
+  $(u,v,\emptyset)$, which the mapping is empty.
+\end{itemize}
 
 The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
-Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V$.
+Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$.
 
 \begin{align*}
-\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(v) \\
+\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(succ(v)) \\
 \llbracket G \rrbracket & = \bigwedge_{v \in V} \mathcal{HC}_v
 \end{align*}
+
+We implicitly assume each $\mathcal{HC}_v$ is universally quantified
+by its all free variables $FV(\mathcal{HC}_v)$.
+
+Note that we allow multiple appearance of Horn clauses with the same
+predicate variable on their right-hand side.
 
@@
 Given an acyclic Horn graph (or called a Horn DAG), we define the
-relation $u <^+ v$ as follows.
-\[ u <^+ v \Longleftrightarrow \exists \theta. (u,v,\theta) \in E \]
+relation $u <^+ v$ over $V_T$ as follows.
+\[ u <^+ v \Longleftrightarrow \exists a, \theta; \left\lbrace{ (u,a,\theta), (a,v,\emptyset) \right\rbrace \subseteq E \]
 By the assumption that $G$ is acyclic, $<^+$ is a well-founded
-relation. We assume without loss of generality that $V$ contains the
+relation. We assume without loss of generality that $V_T$ contains the
 greatest element $v_\bot$ with respect to $<^+$; otherwise we can
-always implicitly assume $\varphi(v_\bot) = 0 \leq 0$.
+always implicitly assume $\varphi(v_\bot) = 0 \leq 0$ such that $\forall v \in V_\bot; \exists v_a; \left\lbrace (v, v_a, \emptyset), (v_a, v_\bot, \emptyset) \right\rbrace \subseteq E $.
 
-For each $v \in V$, we define
+For each $v \in V_T$, we define
 $\Delta(v) = (\mathbf{a}_v \mathbf{x} \leq b_v, C_v)$
@@
   \setminus (\mathbf{a} \cup \left\lbrace b \right\rbrace
- \right). C
+ \right); C
 \right\rbrace
@@
 \emptyset
-& \mbox{if } \nexists u. u <^+ v \\
+& \mbox{if } \nexists u; u <^+ v \\
 \left\lbrace
@@
   \left( P, \sigma(\mathbf{a}_v \mathbf{x} \leq b_v) \right) \middle|
-  \forall v \in V. \varphi(v) = P(\mathbf{x})
+  \forall v \in V; \varphi(v) = P(\mathbf{x})
  \right\rbrace
@@
   \left( P, \bigwedge_S \sigma(\mathbf{a}_{v,S} \mathbf{x} \leq b_{v,S}) \right) \middle|
-  \forall v \in V. \varphi(v) = P(\mathbf{x})
+  \forall v \in V; \varphi(v) = P(\mathbf{x})
  \right\rbrace

### 2013-04-30 article/algo.tex
Changes in expression

@@
 \mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(succ(v)) \\
-\llbracket G \rrbracket & = \bigwedge_{v \in V} \mathcal{HC}_v
+\llbracket G \rrbracket & = \bigwedge_{v \in V_\Rightarrow} \mathcal{HC}_v
 \end{align*}
@@
 We implicitly assume each $\mathcal{HC}_v$ is universally quantified
-by its all free variables $FV(\mathcal{HC}_v)$.
+by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.
 
@@
 relation $u <^+ v$ over $V_T$ as follows.
-\[ u <^+ v \Longleftrightarrow \exists a, \theta; \left\lbrace{ (u,a,\theta), (a,v,\emptyset) \right\rbrace \subseteq E \]
+\[ u <^+ v \Longleftrightarrow \exists a, \theta; \left\lbrace (u,a,\theta), (a,v,\emptyset) \right\rbrace \subseteq E \]
 By the assumption that $G$ is acyclic, $<^+$ is a well-founded
@@
 For each $v \in V_T$, we define
-$\Delta(v) = (\mathbf{a}_v \mathbf{x} \leq b_v, C_v)$
+$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v \leq b_v, C_v)$
 , where
 \begin{itemize}
-\item $\mathbf{a}_v \mathbf{x} \leq b_v$ is a linear inequality that
-  may contain coefficient variables
+\item $\mathbf{a}_v \mathbf{x}_v \leq b_v$ is a linear inequality that
+  contains coefficient variables $\mathbf{a}_v$, and
 \item $C_v$ is a set of linear expressions on coefficient variables.
@@
  \mathbf{a} \mathbf{x} \leq b \middle|
- \exists \left( FV(C)
+ \exists \left( \textsc{FV}(C)
   \setminus (\mathbf{a} \cup \left\lbrace b \right\rbrace
@@
 $\Delta(v)$ is defined by well-founded induction on $v$ with respect
-to the relation $<^+$. Then,
-$\Delta(v) = \left( \mathbf{a}_v \mathbf{x} \leq b_v, C_v \right)$,
-where
+to the relation $<^+$. Then, $C_v$ in
+$\Delta(v) = \left( \mathbf{a}_v \mathbf{x}_v \leq b_v, C_v \right)$
+can be shown as follows.
 
@@
 \left\lbrace
- \mathbf{a}_v = \sum_{u <^+ v} \mathbf{a}_u,
+ \mathbf{a}_v \mathbf{x}_v = \sum_{u <^+ v} \mathbf{a}_u \theta_{u,v} (\mathbf{x}_u),
  b_v \geq \sum_{u <^+ v} b_u
@@
 \end{align*}
+
+For simplicity, we assume variables $\mathbf{a}_u$, $\mathbf{x}_u$,
+$b_u$ are obtained from $\Delta(u)$ in each $\sum_{u <^+ v}$. Also, a
+renaming function $\theta_{u,v}$ is shorthand for $\theta$ such that
+$\exists a; \left\lbrace (u,a,\theta), (a,v,\emptyset) \right\rbrace
+\subseteq E$, and here we pass a variable vector to it.
 

### 2013-05-02 article/algo.tex
Update for constraint * Renaming \theta is used for index mapping

@@
   and arrow vertices $V_\Rightarrow$. Each vertex $u \in
-  V_\Rightarrow$ have eactly one outgoing edge to some $v \in V_T$,
+  V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
   denoted as $succ(u)$.
@@
 \item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
-  $(u,v,\emptyset)$, which the mapping is empty.
+  $(u,v,\emptyset)$, where the mapping is empty.
 \end{itemize}
@@
 We implicitly assume each $\mathcal{HC}_v$ is universally quantified
-by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.
+by its all free variables $\textsc{FV}(\mathcal{HC}_v)$. Without loss
+of generality, we limit occurrence of linear expression on right-hand
+side of Horn clauses to either $0 \leq -1$, meaning $\bot$, or $0 \leq
+0$, meaning $\top$.
 
@@
 \psi ::= & a_1 x_1 + \cdots + a_n x_n \leq b \mid \\
-& \psi \wedge \psi \mid \psi \vee \psi \mid true \mid false
+& \psi \wedge \psi \mid \psi \vee \psi \mid \top \mid \bot
 \end{align*}
@@
 
+For simplicity of discussion, we define $k_v$ the arity of the
+right-hand side of $\mathcal{HC}_v$. Assuming that variables in every
+Horn clauses are appropriately renamed, the right-hand side of
+$\mathcal{HC}_v$ takes the form $P(x_1,...,x_{k_v})$ where $k_v =
+arity(P)$, or either $\top$ or $\bot$ with $k_v = 0$. Also we let $K$
+denote the maximum of $k_v$ over all $v$.
+
 Given an acyclic Horn graph (or called a Horn DAG), we define the
-relation $u <^+ v$ over $V_T$ as follows.
-\[ u <^+ v \Longleftrightarrow \exists a, \theta; \left\lbrace (u,a,\theta), (a,v,\emptyset) \right\rbrace \subseteq E \]
-By the assumption that $G$ is acyclic, $<^+$ is a well-founded
-relation. We assume without loss of generality that $V_T$ contains the
-greatest element $v_\bot$ with respect to $<^+$; otherwise we can
-always implicitly assume $\varphi(v_\bot) = 0 \leq 0$ such that $\forall v \in V_\bot; \exists v_a; \left\lbrace (v, v_a, \emptyset), (v_a, v_\bot, \emptyset) \right\rbrace \subseteq E $.
+relation $u \xrightarrow{\theta} v$ over $V_T$ as follows.
+\[ u \xrightarrow{\theta} v \Longleftrightarrow
+\exists a \in V_\Rightarrow; \left\lbrace (u,a,\theta), (a,v,\emptyset) \right\rbrace \subseteq E \]
+
+We may omit $\theta$ and write $u \rightarrow v$ for convenience for
+later discussion. By the assumption that $G$ is acyclic,
+$\rightarrow^+$ is a well-founded relation. We assume without loss of
+generality that $V_T$ contains the greatest element $v_\bot$ with
+respect to $\rightarrow^+$; otherwise we can always implicitly assume
+$\varphi(v_\bot) = \top$ such that $\forall v \in V_T; v
+\xrightarrow{\emptyset} v_\bot$.
 
@@
 $\Delta(v)$ is defined by well-founded induction on $v$ with respect
-to the relation $<^+$. Then, $C_v$ in
+to the relation $\rightarrow^+$. Then, $C_v$ in
 $\Delta(v) = \left( \mathbf{a}_v \mathbf{x}_v \leq b_v, C_v \right)$
@@
 \begin{align*}
-C_v = & \bigcup_{u <^+ v} C_u
+C_v = & \bigcup_{u \rightarrow v} C_u
 \\
@@
 \emptyset
-& \mbox{if } \nexists u; u <^+ v \\
+& \mbox{if } \nexists u; u \rightarrow v \\
 \left\lbrace
- \mathbf{a}_v \mathbf{x}_v = \sum_{u <^+ v} \mathbf{a}_u \theta_{u,v} (\mathbf{x}_u),
- b_v \geq \sum_{u <^+ v} b_u
-\right\rbrace
+ \mathbf{a}_{v,i} = \sum_{u \xrightarrow{\theta} v} \mathbf{a}_{u, \theta^{-1} (i)},
+ b_v \geq \sum_{u \xrightarrow{\theta} v} b_u
+\right\rbrace \cup \\
+\hspace{2cm} \bigcup_{k_v < i \leq K} \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
 & \mbox{otherwise}
@@
 \emptyset
-& \mbox{if } \varphi(v) = P(\mathbf{x}) \\
+& \mbox{if } \varphi(v) = P(\mathbf{x}_v) \\
 \left\lbrace
@@
 \right\rbrace
-& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x} \leq b_0
+& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0
 \end{cases}
@@
 
-For simplicity, we assume variables $\mathbf{a}_u$, $\mathbf{x}_u$,
-$b_u$ are obtained from $\Delta(u)$ in each $\sum_{u <^+ v}$. Also, a
-renaming function $\theta_{u,v}$ is shorthand for $\theta$ such that
-$\exists a; \left\lbrace (u,a,\theta), (a,v,\emptyset) \right\rbrace
-\subseteq E$, and here we pass a variable vector to it.
+For simplicity, we assume variables $\mathbf{a}_u$, $b_u$ are obtained
+from $\Delta(u)$.
 
@@
  \rho = \left\lbrace
-  \left( P, \sigma(\mathbf{a}_v \mathbf{x} \leq b_v) \right) \middle|
-  \forall v \in V; \varphi(v) = P(\mathbf{x})
+  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \middle|
+  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
  \right\rbrace
@@
 
-If $C_{v_\bot}$ is unsatisfiable, there are two possibilities for them.
-
-\begin{itemize}
-\item No solution exists for $G$.
-\item No simple solution exists for $G$.
-\end{itemize}
-
-If $G$ has a vertex which has multiple successors, the second case is
-possible. We now try to give a solution by relaxing the constraint
+If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for
+$G$. We now try to give a solution by relaxing the constraint
 $C_{v_\bot}$.
@@
 \begin{align*}
-C_v = & \bigcup_{\substack{u <^+ v \\ u \not\in \hat V}} C_u \cup
-\bigcup_{\substack{u <^+ v \\ u \in \hat V}} \textsc{Copy}_v (C_u)
+C_v = & \bigcup_{\substack{u \rightarrow v \\ u \not\in \hat V}} C_u \cup
+\bigcup_{\substack{u \rightarrow v \\ u \in \hat V}} \textsc{Copy}_v (C_u)
 \\
@@
 \emptyset
-& \mbox{if } \nexists u; u <^+ v \\
+& \mbox{if } \nexists u; u \rightarrow v \\
 \left\lbrace
- \mathbf{a}_{v,\emptyset} =
-  \sum_{\substack{u <^+ v \\ u \not\in \hat V}} \mathbf{a}_{u,U} +
-  \sum_{\substack{u <^+ v \\ u \in \hat V}} \mathbf{a}_{u,U \cup \left\lbrace v \right\rbrace},
+ \mathbf{a}_{v,\emptyset,i} =
+  \sum_{\substack{u \xrightarrow{\theta} v \\ u \not\in \hat V}} \mathbf{a}_{u,U,\theta^{-1}(i)} +
+  \sum_{\substack{u \xrightarrow{\theta} v \\ u \in \hat V}}
+  \mathbf{a}_{u,U \cup \left\lbrace v \right\rbrace, \theta^{-1}(i)}
+\right\rbrace \cup \\
+\hspace{2cm} \left\lbrace
  b_{v,\emptyset} \geq
-  \sum_{\substack{u <^+ v \\ u \not\in \hat V}} b_{u,U} +
-  \sum_{\substack{u <^+ v \\ u \in \hat V}} b_{u,U \cup \left\lbrace v \right\rbrace}
-\right\rbrace
+  \sum_{\substack{u \rightarrow v \\ u \not\in \hat V}} b_{u,U} +
+  \sum_{\substack{u \rightarrow v \\ u \in \hat V}} b_{u,U \cup \left\lbrace v \right\rbrace}
+\right\rbrace \cup \\
+\hspace{2cm} \bigcup_{k_v < i \leq K} \left\lbrace \mathbf{a}_{v,\emptyset,i} = 0 \right\rbrace
 & \mbox{otherwise}
@@
 nodes. This intuitively means that constraints generated for nodes in
-$\hat V$ are duplicated at their successors. Note that the variables
-without the label should be interpreted as that it has $\emptyset$ for
-origin nodes.
+$\hat V$ are duplicated at their successors. For consistency, the
+variables without the label should be interpreted as that it has
+$\emptyset$ for origin nodes.
 
@@
  \rho = \left\lbrace
-  \left( P, \bigwedge_S \sigma(\mathbf{a}_{v,S} \mathbf{x} \leq b_{v,S}) \right) \middle|
-  \forall v \in V; \varphi(v) = P(\mathbf{x})
+  \left( P, \bigwedge_S \sigma(\mathbf{a}_{v,S} \mathbf{x}_v \leq b_{v,S}) \right) \middle|
+  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
  \right\rbrace

### 2013-05-04 article/algo.tex
Introduced copy tree

@@
 The Horn clause solving algorithm receives an input problem in Horn
-graph. A Horn graph is a labelled directed graph $G=(V,E,\varphi)$.
+graph. A \textbf{Horn graph} is a labelled directed graph $G=(V,E,\varphi)$.
 \begin{itemize}
@@
 \exists a \in V_\Rightarrow; \left\lbrace (u,a,\theta), (a,v,\emptyset) \right\rbrace \subseteq E \]
-
 We may omit $\theta$ and write $u \rightarrow v$ for convenience for
@@
 
-We slightly modify the definition $C_v$. Let $\hat V$ a subset of
-vertices with multiple successors.
+Let $\hat V_T$ a set of Horn term vertices $V_T$ with multiple
+successors.
+\[ \hat V_T = \left\lbrace v \in V_T \middle|
+\exists u, u', \theta, \theta';
+\left\lbrace (v,u,\theta), (v,u',\theta') \right\rbrace \subseteq E \wedge
+(u \ne u' \vee \theta \ne \theta') \right\rbrace \]
+We define \textbf{Copy tree}, which is a tree $T=(V_C,E_C,\xi)$
+rooted at $v_{v_\bot}$.
+\begin{itemize}
+\item $V_C$ is a set of nodes.
+\item $E_C$ is a set of labelled edges, which are expressed in the
+  form $(u,v,e)$, where $e \in E \cup {\bot}$.
+\item $\xi: V_C \rightarrow \hat V_T \cup \left\lbrace v_\bot
+  \right\rbrace$ is a labelling function for nodes. The root node
+  $v_{v_\bot}$ has a label $v_\bot$.
+\end{itemize}
 
+We provide $\textsc{Copy}_v$ procedure to rename all free variables
+for avoiding name collision.
+\[ \textsc{Copy}_v (x_V) = x_{V \cup \left\lbrace v \right\rbrace} \]
+All variables are then distinguished by the label of copy origin
+nodes. For consistency, the variables without the label should be
+interpreted as that it has $\emptyset$ for origin nodes.
+
+We give a constraint function over a copy tree
+$\hat \Delta(v,e)$ for $v \in V_C, e \in E$ as
+\[ \hat \Delta (v,e) = \textsc{Copy}_v (\Delta_{V'} (\xi (v)))
+\text{ such that } V' = \left\lbrace v' \middle| (v, v',e) \in E_S \right\rbrace \]
+extending $\Delta$ function to
+$\Delta_V (v) = (\mathbf{a}_{v,\emptyset} \mathbf{x}_v \leq b_{v,\emptyset}, C_v)$
+as
 \begin{align*}
-C_v = & \bigcup_{\substack{u \rightarrow v \\ u \not\in \hat V}} C_u \cup
-\bigcup_{\substack{u \rightarrow v \\ u \in \hat V}} \textsc{Copy}_v (C_u)
+C_v = & \bigcup_\circ C_u \cup \bigcup_\bullet C_u \cup
 \\
@@
  \mathbf{a}_{v,\emptyset,i} =
-  \sum_{\substack{u \xrightarrow{\theta} v \\ u \not\in \hat V}} \mathbf{a}_{u,U,\theta^{-1}(i)} +
-  \sum_{\substack{u \xrightarrow{\theta} v \\ u \in \hat V}}
-  \mathbf{a}_{u,U \cup \left\lbrace v \right\rbrace, \theta^{-1}(i)}
+  \sum_\circ \mathbf{a}_{u,U,\theta^{-1}(i)} +
+  \sum_\bullet \mathbf{a}_{u,U, \theta^{-1}(i)}
 \right\rbrace \cup \\
@@
  b_{v,\emptyset} \geq
-  \sum_{\substack{u \rightarrow v \\ u \not\in \hat V}} b_{u,U} +
-  \sum_{\substack{u \rightarrow v \\ u \in \hat V}} b_{u,U \cup \left\lbrace v \right\rbrace}
+  \sum_\circ b_{u,U} +
+  \sum_\bullet b_{u,U}
 \right\rbrace \cup \\
@@
 \end{align*}
+under
+\[
+\circ = \substack{
+  u \rightarrow v \\
+  \nexists u' \in V; \xi (u') = u \\
+  \Delta_V (u) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u)}
+\qquad \bullet = \substack{
+  u \rightarrow v \\
+  \exists u' \in V; \xi (u') = u \\
+  \hat \Delta (u', u \rightarrow v) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u)} \]
 
-The procedure $\textsc{Copy}_v$ renames all free variables to avoid
-name collision. Its definition is given as follows.
-
-\[\textsc{Copy}_{\hat v}(\mathbf{a}_{u,V}) = \mathbf{a}_{u,(V \cup
-  \left\lbrace \hat v \right\rbrace)} \]
-
-All variables are then distinguished by the label of copy origin
-nodes. This intuitively means that constraints generated for nodes in
-$\hat V$ are duplicated at their successors. For consistency, the
-variables without the label should be interpreted as that it has
-$\emptyset$ for origin nodes.
-
-When $C_{v_\bot}$ becomes satisfiable by specifing an appropriate
-$\hat V$, a solution $\rho$ to $G$ is given as follows.
+This intuitively means that constraints are duplicated according to
+the shape of copy tree $T$.
+Given the constraint for the root node in the copy tree as
+\[ \hat \Delta (v_{v_\bot}) = (\mathbf{a}_0 \mathbf{x}_{v_\bot} \leq b_0, C_{v_\bot}) \]
+a model to $C_{v_\bot}$ gives a solution $\rho$ to $G$, if one exists.
 
@@
 
-If $C_{v_\bot}$ stays unsatisfiable, we can add a vertex with multiple
-successors to $\hat V$ until all such nodes are in $\hat V$.
+If $C_{v_\bot}$ stays unsatisfiable, we let $T$ grow based on the
+unsatisfiable core.
 

### 2013-05-13 article/algo.tex
Changes of Horn graph structure

@@
 
+\subsection{Preliminaries}
+
 The Horn clause solving algorithm receives an input problem in Horn
@@
 \begin{itemize}
-\item $V$ is a union of two sets of vertices; Horn term vertices $V_T$
-  and arrow vertices $V_\Rightarrow$. Each vertex $u \in
-  V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
-  denoted as $succ(u)$.
+\item $V$ is a union of two disjoint sets of vertices; Horn term
+  vertices $V_T$ and disjunction vertices $V_\vee$. Each vertex $u \in
+  V_\vee$ has eactly one outgoing edge to some $v \in V_T$.
 \item $E$ is a set of labelled edges, which are expressed in the form
@@
   term vertices, where L is a set of elements of a predicate variable
-  in the form $P(x_1, ..., x_{arity(P)})$ or a linear inequality $a_1
-  x_1 + \cdots + a_n x_n \leq b$.
+  in the form $P(x_1, \ldots, x_{arity(P)})$ or a linear inequality
+  $a_1 x_1 + \cdots + a_n x_n \leq b$.
 \end{itemize}
-
 There are two kinds of edges in a Horn graph;
 \begin{itemize}
-\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
-  $(u,v,\theta)$, or,
-\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
-  $(u,v,\emptyset)$, where the mapping is empty.
+\item for $u \in V_T, v \in V_\vee$, an edge takes a form
+  $(u,v,\theta)$ written $u \xrightarrow{\theta} v$, or,
+\item for $u \in V_\vee, v \in V_T$, an edge takes a form
+  $(u,v,\emptyset)$ written $u \rightarrow v$, where the mapping is
+  empty.
 \end{itemize}
@@
 The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
-Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$.
+Horn clauses encoded as follows.
 
 \begin{align*}
-\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(succ(v)) \\
-\llbracket G \rrbracket & = \bigwedge_{v \in V_\Rightarrow} \mathcal{HC}_v
+\mathcal{T}_v & = \prod_{u \rightarrow v}
+ \left\lbrace \theta (\varphi (t)) \mid t \xrightarrow{\theta} u \right\rbrace \\
+\mathcal{HC}_{(t_1,\ldots,t_k),t_\star} & = \forall \left(\textsc{FV}(\mathbf{t}) \cup \textsc{FV}(t_\star) \right);
+ \left\lbrace \left( \bigwedge t_i \right) \Longrightarrow t_\star \right\rbrace \\
+\llbracket G \rrbracket & = \bigwedge_{v \in V_T}
+ \left(\bigwedge_{\mathbf{t} \in \mathcal{T}_v} \mathcal{HC}_{\mathbf{t},\varphi (v)} \right)
 \end{align*}
 
-We implicitly assume each $\mathcal{HC}_v$ is universally quantified
-by its all free variables $\textsc{FV}(\mathcal{HC}_v)$. Without loss
-of generality, we limit occurrence of linear expression on right-hand
-side of Horn clauses to either $0 \leq -1$, meaning $\bot$, or $0 \leq
-0$, meaning $\top$.
-
-Note that we allow multiple appearance of Horn clauses with the same
-predicate variable on their right-hand side.
+Without loss of generality, we limit linear expression labels on $V_T$
+vertices with incoming edges to either $0 \leq -1$, meaning $\bot$, or
+$0 \leq 0$, meaning $\top$.
 
@@
 
-We consider below a Horn graph that is acyclic.
+\subsection{Simple solution}
 
-For simplicity of discussion, we define $k_v$ the arity of the
-right-hand side of $\mathcal{HC}_v$. Assuming that variables in every
-Horn clauses are appropriately renamed, the right-hand side of
-$\mathcal{HC}_v$ takes the form $P(x_1,...,x_{k_v})$ where $k_v =
-arity(P)$, or either $\top$ or $\bot$ with $k_v = 0$. Also we let $K$
-denote the maximum of $k_v$ over all $v$.
+We consider below a Horn graph that is acyclic, and we consider the case
+that all disjunction vertices has exactly one predecessor.
 
-Given an acyclic Horn graph (or called a Horn DAG), we define the
-relation $u \xrightarrow{\theta} v$ over $V_T$ as follows.
-\[ u \xrightarrow{\theta} v \Longleftrightarrow
-\exists a \in V_\Rightarrow; \left\lbrace (u,a,\theta), (a,v,\emptyset) \right\rbrace \subseteq E \]
-We may omit $\theta$ and write $u \rightarrow v$ for convenience for
-later discussion. By the assumption that $G$ is acyclic,
-$\rightarrow^+$ is a well-founded relation. We assume without loss of
-generality that $V_T$ contains the greatest element $v_\bot$ with
-respect to $\rightarrow^+$; otherwise we can always implicitly assume
-$\varphi(v_\bot) = \top$ such that $\forall v \in V_T; v
-\xrightarrow{\emptyset} v_\bot$.
+For simplicity of discussion, we define $k_v$ the effective arity of
+the Horn clause $\mathcal{HC}_{\mathbf{t},\varphi(v)}$.
+\begin{align*}
+k_v =
+\begin{cases}
+arity(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{arity(P)}) \\
+0 & \mbox{if } \varphi(v) = \top \mbox { or } \varphi(v) = \bot
+\end{cases}
+\end{align*}
+We let $K$ denote the maximum of $k_v$ over all $v$.
+
+We define the relation $\leadsto$ over $V_T$ as follows.
+\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow
+\exists a \in V_\vee; \left\lbrace (u,a,\theta), (a,v,\emptyset) \right\rbrace \subseteq E \]
+We may omit $\theta$ and write $u \leadsto v$ for convenience in later
+discussions.
+
+By the assumption that $G$ is acyclic, $\leadsto^+$ is a
+well-founded relation. We assume without loss of generality that $V_T$
+contains the greatest element $v_\bot$ with respect to $\leadsto^+$;
+otherwise we can always implicitly assume $\varphi(v_\bot) = \top$
+such that $\forall v \in V_T \setminus \{v_\bot\}; v
+\mathop{\leadsto}^\emptyset v_\bot$.
 
@@
 set of inequalities:
-
 \begin{align*}
@@
 $\Delta(v)$ is defined by well-founded induction on $v$ with respect
-to the relation $\rightarrow^+$. Then, $C_v$ in
+to the relation $\leadsto^+$. Then, $C_v$ in
 $\Delta(v) = \left( \mathbf{a}_v \mathbf{x}_v \leq b_v, C_v \right)$
@@
 \begin{align*}
-C_v = & \bigcup_{u \rightarrow v} C_u
+C_v = & \bigcup_{u \leadsto v} C_u
 \\
@@
 \emptyset
-& \mbox{if } \nexists u; u \rightarrow v \\
+& \mbox{if } \nexists u; u \leadsto v \\
 \left\lbrace
- \mathbf{a}_{v,i} = \sum_{u \xrightarrow{\theta} v} \mathbf{a}_{u, \theta^{-1} (i)},
- b_v \geq \sum_{u \xrightarrow{\theta} v} b_u
+ \mathbf{a}_{v,i} = \sum_{u \mathop{\leadsto}^\theta v} \mathbf{a}_{u, \theta^{-1} (i)},
+ b_v \geq \sum_{u \mathop{\leadsto}^\theta v} b_u
 \right\rbrace \cup \\
@@
 
+\subsection{DAG problem}
+
 If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for
@@
 \[ \textsc{Copy}_v (x_V) = x_{V \cup \left\lbrace v \right\rbrace} \]
-All variables are then distinguished by the label of copy origin
-nodes. For consistency, the variables without the label should be
+All variables with the same name are then distinguished by the label of copy origin
+nodes $V$. For consistency, the variables without the label should be
 interpreted as that it has $\emptyset$ for origin nodes.
@@
 \circ = \substack{
-  u \rightarrow v \\
+  u \mathop{\leadsto}^\theta v \\
   \nexists u' \in V; \xi (u') = u \\
@@
 \qquad \bullet = \substack{
-  u \rightarrow v \\
+  u \mathop{\leadsto}^\theta v \\
   \exists u' \in V; \xi (u') = u \\
@@
  \rho = \left\lbrace
-  \left( P, \bigwedge_S \sigma(\mathbf{a}_{v,S} \mathbf{x}_v \leq b_{v,S}) \right) \middle|
+  \left( P, \bigwedge_U \sigma(\mathbf{a}_{v,U} \mathbf{x}_v \leq b_{v,U}) \right) \middle|
   \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
@@
 
+\subsection{Disjunctive problems}
+
+We eliminate the restriction that all disjunction vertices $V_\vee$ have exactly one
+predecessor. We define the universal set of selectors for disjunction vertices.
+\[ \mathcal{S} = \prod_{v \in V_\vee} \left\lbrace v \mapsto u \middle| \exists \theta; (u,v,\theta) \in E \right\rbrace \]
+
+We define a \textbf{selector set collection} \cite{albarghouthi13} for every Horn term vertices, which is a set of sets of selectors.
+
+For $v \in V_T$, the selector set collection $\mathcal{X}_v = \left\lbrace S_{v,1}, \ldots, S_{v,\kappa_v} \right\rbrace$ consists of $\kappa_v$ sets of non-empty and non-overlapping subsets of $\mathcal{S}$.
+If $\nexists u; u \rightarrow v$, $\mathcal{X}_v = \emptyset$.
+
+We arbitrarily choose the initial selector $s_0 \in \mathcal{S}$. We start from empty collections for all $v$.
+
+\bibliographystyle{plain}
+\bibliography{./biblio}
+
 \end{document}

### 2013-05-17 article/algo.tex
Small changes of path expression

@@
 \newcommand{\edge}[2]{#1\rightarrow#2}
-\newcommand{\path}[2]{#1\rightarrow^\star#2}
+\newcommand{\edgel}[3]{#1\xrightarrow{#2}#3}
 
@@
 \item for $u \in V_T, v \in V_\vee$, an edge takes a form
-  $(u,v,\theta)$ written $u \xrightarrow{\theta} v$, or,
+  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
 \item for $u \in V_\vee, v \in V_T$, an edge takes a form
-  $(u,v,\emptyset)$ written $u \rightarrow v$, where the mapping is
+  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
   empty.
@@
 
-We consider below a Horn graph that is acyclic, and we consider the case
-that all disjunction vertices has exactly one predecessor.
+We consider below a Horn graph that is acyclic, and we consider the
+case that all disjunction vertices has exactly one predecessor.
 
-For simplicity of discussion, we define $k_v$ the effective arity of
-the Horn clause $\mathcal{HC}_{\mathbf{t},\varphi(v)}$.
+For simplicity of discussion, we define \textbf{term arity} $k_v$
+for every Horn term vertex $v \in V_T$.
 \begin{align*}
@@
 \end{align*}
-We let $K$ denote the maximum of $k_v$ over all $v$.
+We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.
 
@@
 \[ u \mathop{\leadsto}^\theta v \Longleftrightarrow
-\exists a \in V_\vee; \left\lbrace (u,a,\theta), (a,v,\emptyset) \right\rbrace \subseteq E \]
+\exists a \in V_\vee; \edgel{v}{\theta}{a} \wedge \edgel{a}{\emptyset}{v} \]
 We may omit $\theta$ and write $u \leadsto v$ for convenience in later
@@
 
-By the assumption that $G$ is acyclic, $\leadsto^+$ is a
-well-founded relation. We assume without loss of generality that $V_T$
-contains the greatest element $v_\bot$ with respect to $\leadsto^+$;
-otherwise we can always implicitly assume $\varphi(v_\bot) = \top$
-such that $\forall v \in V_T \setminus \{v_\bot\}; v
-\mathop{\leadsto}^\emptyset v_\bot$.
+By the assumption that $G$ is acyclic, $\leadsto^+$ is a well-founded
+relation. We assume without loss of generality that $V_T$ contains the
+greatest element $v_\bot$ with respect to $\leadsto^+$; otherwise we
+can always implicitly assume $\varphi(v_\bot) = \top$ such that
+$\forall v \in V_T \setminus \{v_\bot\}; v \mathop{\leadsto}^\emptyset v_\bot$.
 
@@
 \exists u, u', \theta, \theta';
-\left\lbrace (v,u,\theta), (v,u',\theta') \right\rbrace \subseteq E \wedge
+\edgel{v}{\theta}{u} \wedge \edgel{v}{\theta'}{u'} \wedge
 (u \ne u' \vee \theta \ne \theta') \right\rbrace \]
@@
 \emptyset
-& \mbox{if } \nexists u; u \rightarrow v \\
+& \mbox{if } \nexists u; u \leadsto v \\
 \left\lbrace
@@
   \exists u' \in V; \xi (u') = u \\
-  \hat \Delta (u', u \rightarrow v) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u)} \]
+  \hat \Delta (u', \edge{u}{v}) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u)} \]
 
 This intuitively means that constraints are duplicated according to
-the shape of copy tree $T$.
-Given the constraint for the root node in the copy tree as
+the shape of copy tree $T$.  Given the constraint for the root node in
+the copy tree as
 \[ \hat \Delta (v_{v_\bot}) = (\mathbf{a}_0 \mathbf{x}_{v_\bot} \leq b_0, C_{v_\bot}) \]
@@
 
-We eliminate the restriction that all disjunction vertices $V_\vee$ have exactly one
-predecessor. We define the universal set of selectors for disjunction vertices.
-\[ \mathcal{S} = \prod_{v \in V_\vee} \left\lbrace v \mapsto u \middle| \exists \theta; (u,v,\theta) \in E \right\rbrace \]
+We eliminate the restriction that all disjunction vertices $V_\vee$ have
+exactly one predecessor. We define the universal set of selectors for
+disjunction vertices.
+\[ \mathcal{S} = \prod_{v \in V_\vee} \left\lbrace v \mapsto u \middle| \edgel{u}{\emptyset}{v} \right\rbrace \]
 
-We define a \textbf{selector set collection} \cite{albarghouthi13} for every Horn term vertices, which is a set of sets of selectors.
+We define a \textbf{selector set collection} \cite{albarghouthi13} for
+every Horn term vertices, which is a set of sets of selectors.
 
-For $v \in V_T$, the selector set collection $\mathcal{X}_v = \left\lbrace S_{v,1}, \ldots, S_{v,\kappa_v} \right\rbrace$ consists of $\kappa_v$ sets of non-empty and non-overlapping subsets of $\mathcal{S}$.
-If $\nexists u; u \rightarrow v$, $\mathcal{X}_v = \emptyset$.
+For $v \in V_T$, the selector set collection $\mathcal{X}_v =
+\left\lbrace S_{v,1}, \ldots, S_{v,\kappa_v} \right\rbrace$ consists
+of $\kappa_v$ sets of non-empty and non-overlapping subsets of
+$\mathcal{S}$. If $\nexists u; u \rightarrow v$, $\mathcal{X}_v =
+\emptyset$.
 
-We arbitrarily choose the initial selector $s_0 \in \mathcal{S}$. We start from empty collections for all $v$.
+We arbitrarily choose the initial selector $s_0 \in \mathcal{S}$. We
+start from empty collections for all $v$.
 

### 2013-05-22 article/algo.tex
Disjunction algorithm construction

@@
   term vertices, where L is a set of elements of a predicate variable
-  in the form $P(x_1, \ldots, x_{arity(P)})$ or a linear inequality
+  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear inequality
   $a_1 x_1 + \cdots + a_n x_n \leq b$.
@@
 $\rho[P]$ is of the form $a_1 x_1 + \cdots + a_n x_n \leq b$ for every
-$P$ in $dom(\rho)$.
+$P$ in $\mathrm{dom}(\rho)$.
 
@@
 \begin{cases}
-arity(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{arity(P)}) \\
+\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
 0 & \mbox{if } \varphi(v) = \top \mbox { or } \varphi(v) = \bot
@@
 
-\subsection{DAG problem}
+\subsection{Disjunction-free problems}
 
@@
 extending $\Delta$ function to
-$\Delta_V (v) = (\mathbf{a}_{v,\emptyset} \mathbf{x}_v \leq b_{v,\emptyset}, C_v)$
+$\Delta_{V'} (v) = (\mathbf{a}_{v,\emptyset} \mathbf{x}_v \leq b_{v,\emptyset}, C_v)$
 as
@@
 \end{align*}
-under
-\[
-\circ = \substack{
-  u \mathop{\leadsto}^\theta v \\
-  \nexists u' \in V; \xi (u') = u \\
-  \Delta_V (u) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u)}
-\qquad \bullet = \substack{
-  u \mathop{\leadsto}^\theta v \\
-  \exists u' \in V; \xi (u') = u \\
-  \hat \Delta (u', \edge{u}{v}) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u)} \]
+under the iterator $\circ = (u,U,\theta,C_u) \in S_v^- $ and $\bullet = (u,U,\theta,C_u) \in S_v^+ $
+such that
+\begin{align*}
+& S_v^- = \left\lbrace (u, U, \theta, C_u) \middle|
+  u \mathop{\leadsto}^\theta v \wedge
+  \left( \nexists u' \in V'; \xi (u') = u \right) \wedge
+  \Delta_{V'} (u) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u) \right\rbrace
+\\
+& S_v^+ = \left\lbrace (u, U, \theta, C_u) \middle|
+  u \mathop{\leadsto}^\theta v \wedge
+  \left( \exists u' \in V'; \xi (u') = u \wedge
+  \hat \Delta (u', \edge{u}{v}) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u) \right) \right\rbrace
+\\
+\end{align*}
 
@@
 
-\subsection{Disjunctive problems}
+\subsection{Acyclic Horn problems}
 
 We eliminate the restriction that all disjunction vertices $V_\vee$ have
-exactly one predecessor. We define the universal set of selectors for
-disjunction vertices.
-\[ \mathcal{S} = \prod_{v \in V_\vee} \left\lbrace v \mapsto u \middle| \edgel{u}{\emptyset}{v} \right\rbrace \]
+exactly one predecessor.
+
+We define selectors for preceding disjunction vertices to $v$.
+\[ \mathcal{S}_v = \prod_{u \in \mathrm{succ}(v)} \left\lbrace u \mapsto t \middle| \edgel{t}{\theta}{u} \right\rbrace \]
+
+\begin{align*}
+K_l = \begin{cases}
+\left\lbrace l \right\rbrace & \mbox{if } l = \mathbf{a} \mathbf{x} \leq b \\
+\mathcal{X}_P \cup \left\lbrace \bot \right\rbrace
+& \mbox{if } l = P(\mathbf{x})
+\end{cases} \\
+\forall s \in \mathcal{S}_v;
+X = \prod_{t \in \mathrm{ran}(s)} K_{\varphi(t)}
+\exists c \in X;
+\end{align*}
 

### 2013-05-27 article/algo.tex
Reverting Horn graph structure

@@
 \item $V$ is a union of two disjoint sets of vertices; Horn term
-  vertices $V_T$ and disjunction vertices $V_\vee$. Each vertex $u \in
-  V_\vee$ has eactly one outgoing edge to some $v \in V_T$.
+  vertices $V_T$ and arrow vertices $V_\Rightarrow$. Each vertex $u
+  \in V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
+  denoted as $succ(u)$.
 \item $E$ is a set of labelled edges, which are expressed in the form
@@
   term vertices, where L is a set of elements of a predicate variable
-  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear inequality
-  $a_1 x_1 + \cdots + a_n x_n \leq b$.
+  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
+  inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.
 \end{itemize}
@@
 \begin{itemize}
-\item for $u \in V_T, v \in V_\vee$, an edge takes a form
+\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
   $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
-\item for $u \in V_\vee, v \in V_T$, an edge takes a form
+\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
   $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
@@
 The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
-Horn clauses encoded as follows.
+Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$.
 
 \begin{align*}
-\mathcal{T}_v & = \prod_{u \rightarrow v}
- \left\lbrace \theta (\varphi (t)) \mid t \xrightarrow{\theta} u \right\rbrace \\
-\mathcal{HC}_{(t_1,\ldots,t_k),t_\star} & = \forall \left(\textsc{FV}(\mathbf{t}) \cup \textsc{FV}(t_\star) \right);
- \left\lbrace \left( \bigwedge t_i \right) \Longrightarrow t_\star \right\rbrace \\
-\llbracket G \rrbracket & = \bigwedge_{v \in V_T}
- \left(\bigwedge_{\mathbf{t} \in \mathcal{T}_v} \mathcal{HC}_{\mathbf{t},\varphi (v)} \right)
+\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(succ(v)) \\
+\llbracket G \rrbracket & = \bigwedge_{v \in V_\Rightarrow} \mathcal{HC}_v
 \end{align*}
 
-Without loss of generality, we limit linear expression labels on $V_T$
-vertices with incoming edges to either $0 \leq -1$, meaning $\bot$, or
-$0 \leq 0$, meaning $\top$.
+We implicitly assume each $\mathcal{HC}_v$ is universally quantified
+by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.  Without loss
+of generality, we limit linear expression labels on $V_T$ vertices
+with incoming edges to either $0 \leq -1$, meaning $\bot$, or $0 \leq
+0$, meaning $\top$.
 
@@
 
-We consider below a Horn graph that is acyclic, and we consider the
-case that all disjunction vertices has exactly one predecessor.
+We consider below a Horn graph that is acyclic.
 
-For simplicity of discussion, we define \textbf{term arity} $k_v$
-for every Horn term vertex $v \in V_T$.
+For simplicity of discussion, we define \textbf{term arity} $k_v$ for
+every Horn term vertex $v \in V_T$ with incoming edges.
 \begin{align*}
@@
 \[ u \mathop{\leadsto}^\theta v \Longleftrightarrow
-\exists a \in V_\vee; \edgel{v}{\theta}{a} \wedge \edgel{a}{\emptyset}{v} \]
+\exists a \in V_\Rightarrow; \edgel{v}{\theta}{a} \wedge \edge{a}{v} \]
 We may omit $\theta$ and write $u \leadsto v$ for convenience in later
@@
 \begin{align*}
-C_v = & \bigcup_{u \leadsto v} C_u
+\hat C_a = &
+ \bigcup_{\edgel{u}{\theta}{a}} C_u \cup
+ \bigcup_{0 < i \leq K}
+ \left\lbrace
+  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{a}} \mathbf{a}_{u, \theta^{-1} (i)}
+ \right\rbrace \\
+ & \hspace{2cm} \cup
+ \left\lbrace
+  b_v \geq \sum_{\edgel{u}{\theta}{a}} b_u
+ \right\rbrace \cup
+ \bigcup_{k_v < i \leq K}
+ \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
 \\
-& \cup \begin{cases}
-\emptyset
-& \mbox{if } \nexists u; u \leadsto v \\
-\left\lbrace
- \mathbf{a}_{v,i} = \sum_{u \mathop{\leadsto}^\theta v} \mathbf{a}_{u, \theta^{-1} (i)},
- b_v \geq \sum_{u \mathop{\leadsto}^\theta v} b_u
-\right\rbrace \cup \\
-\hspace{2cm} \bigcup_{k_v < i \leq K} \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
-& \mbox{otherwise}
-\end{cases}
-\\
-& \cup \begin{cases}
+C_v = & \bigcup_{\edge{a}{v}} \hat C_a \cup
+\begin{cases}
 \emptyset
@@
 
+We consider below a Horn graph whose Horn term vertices has exactly
+one successor.
+
 Let $\hat V_T$ a set of Horn term vertices $V_T$ with multiple
@@
 \[ \textsc{Copy}_v (x_V) = x_{V \cup \left\lbrace v \right\rbrace} \]
-All variables with the same name are then distinguished by the label of copy origin
-nodes $V$. For consistency, the variables without the label should be
-interpreted as that it has $\emptyset$ for origin nodes.
+All variables with the same name are then distinguished by the label
+of copy origin nodes $V$. For consistency, the variables without the
+label should be interpreted as that it has $\emptyset$ for origin
+nodes.
 
@@
 \[ \hat \Delta (v,e) = \textsc{Copy}_v (\Delta_{V'} (\xi (v)))
-\text{ such that } V' = \left\lbrace v' \middle| (v, v',e) \in E_S \right\rbrace \]
+\text{ such that } V' = \left\lbrace v' \middle| (v, v',e) \in E_C \right\rbrace \]
 extending $\Delta$ function to
@@
 \begin{align*}
-C_v = & \bigcup_\circ C_u \cup \bigcup_\bullet C_u \cup
+C_v = & \bigcup_{(u,U,\theta,C_u) \in S_v^-} C_u \cup
+\bigcup_{(u,U,\theta,C_u) \in S_v^+} C_u \cup
 \\
@@
  \mathbf{a}_{v,\emptyset,i} =
-  \sum_\circ \mathbf{a}_{u,U,\theta^{-1}(i)} +
-  \sum_\bullet \mathbf{a}_{u,U, \theta^{-1}(i)}
+  \sum_{(u,U,\theta,C_u) \in S_v^-} \mathbf{a}_{u,U,\theta^{-1}(i)} +
+  \sum_{(u,U,\theta,C_u) \in S_v^+} \mathbf{a}_{u,U, \theta^{-1}(i)}
 \right\rbrace \cup \\
@@
  b_{v,\emptyset} \geq
-  \sum_\circ b_{u,U} +
-  \sum_\bullet b_{u,U}
+  \sum_{(u,U,\theta,C_u) \in S_v^-} b_{u,U} +
+  \sum_{(u,U,\theta,C_u) \in S_v^+} b_{u,U}
 \right\rbrace \cup \\
@@
 \end{align*}
-under the iterator $\circ = (u,U,\theta,C_u) \in S_v^- $ and $\bullet = (u,U,\theta,C_u) \in S_v^+ $
 such that
@@
 the copy tree as
-\[ \hat \Delta (v_{v_\bot}) = (\mathbf{a}_0 \mathbf{x}_{v_\bot} \leq b_0, C_{v_\bot}) \]
+\[ \hat \Delta (v_{v_\bot}, \bot) = (\mathbf{a}_0 \mathbf{x}_{v_\bot} \leq b_0, C_{v_\bot}) \]
 a model to $C_{v_\bot}$ gives a solution $\rho$ to $G$, if one exists.
@@
 
-We eliminate the restriction that all disjunction vertices $V_\vee$ have
-exactly one predecessor.
+We eliminate the restriction that all Horn term vertices
+$V_\Rightarrow$ have exactly one successor.
 

### 2013-05-28 article/algo.tex

@@
 \begin{align*}
-\hat C_a = &
+\hat C_{a,v} = &
  \bigcup_{\edgel{u}{\theta}{a}} C_u \cup
@@
 \\
-C_v = & \bigcup_{\edge{a}{v}} \hat C_a \cup
+C_v = & \bigcup_{\edge{a}{v}} \hat C_{a,v} \cup
 \begin{cases}
@@
 
-We consider below a Horn graph whose Horn term vertices has exactly
-one successor.
+We consider below a Horn graph whose Horn term vertices has at
+most one successor.
+
+\subsubsection{Quantifier elimination}
+
+A set of eliminating quantifiers are obtained as $quant(n)$.
+
+\begin{align*}
+in(n) & = \bigcup_{p\in pred(n)} out(p) \\
+out(n) & = gen(n) \cup in(n) \\
+dup(n) & = \left\lbrace t \middle| p,q \in pred(n); p \ne q \wedge
+ t \in out(p) \wedge t \in out(q) \right\rbrace \\
+quant(n) & = gen(n) \cup dup(n) \cup
+ (out(n) \cap \bigcup_{s \in succ(n)} quant(s))
+\end{align*}
 
@@
   \right\rbrace$ is a labelling function for nodes. The root node
-  $v_{v_\bot}$ has a label $v_\bot$.
+  $v_{v_\bot} \in V_C$ has a label $v_\bot$.
 \end{itemize}
@@
 as
+
 \begin{align*}
-C_v = & \bigcup_{(u,U,\theta,C_u) \in S_v^-} C_u \cup
-\bigcup_{(u,U,\theta,C_u) \in S_v^+} C_u \cup
+\hat C_a = &
+ \bigcup_{\edgel{u}{\theta}{a}} C_u \cup
+ \bigcup_{0 < i \leq K}
+ \left\lbrace
+  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{a}} \mathbf{a}_{u, \theta^{-1} (i)}
+ \right\rbrace \\
+ & \hspace{2cm} \cup
+ \left\lbrace
+  b_v \geq \sum_{\edgel{u}{\theta}{a}} b_u
+ \right\rbrace \cup
+ \bigcup_{k_v < i \leq K}
+ \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
 \\
-& \cup \begin{cases}
-\emptyset
-& \mbox{if } \nexists u; u \leadsto v \\
-\left\lbrace
- \mathbf{a}_{v,\emptyset,i} =
-  \sum_{(u,U,\theta,C_u) \in S_v^-} \mathbf{a}_{u,U,\theta^{-1}(i)} +
-  \sum_{(u,U,\theta,C_u) \in S_v^+} \mathbf{a}_{u,U, \theta^{-1}(i)}
-\right\rbrace \cup \\
-\hspace{2cm} \left\lbrace
- b_{v,\emptyset} \geq
-  \sum_{(u,U,\theta,C_u) \in S_v^-} b_{u,U} +
-  \sum_{(u,U,\theta,C_u) \in S_v^+} b_{u,U}
-\right\rbrace \cup \\
-\hspace{2cm} \bigcup_{k_v < i \leq K} \left\lbrace \mathbf{a}_{v,\emptyset,i} = 0 \right\rbrace
-& \mbox{otherwise}
-\end{cases}
-\\
-& \cup \begin{cases}
+C_v = & \bigcup_{\edge{a}{v}} \hat C_a \cup
+\begin{cases}
 \emptyset
@@
 \end{cases}
+\end{align*}
+
+---
+
+\begin{align*}
+C_v = & \bigcup_{(u,U,\theta,C_u) \in S_v^-} C_u \cup
+\bigcup_{(u,U,\theta,C_u) \in S_v^+} C_u
+\\
+& \cup
+\left\lbrace
+ \mathbf{a}_{v,\emptyset,i} =
+  \sum_{(u,U,\theta,C_u) \in S_v^-} \mathbf{a}_{u,U,\theta^{-1}(i)} +
+  \sum_{(u,U,\theta,C_u) \in S_v^+} \mathbf{a}_{u,U, \theta^{-1}(i)}
+\right\rbrace \\
+& \cup \left\lbrace
+ b_{v,\emptyset} \geq
+  \sum_{(u,U,\theta,C_u) \in S_v^-} b_{u,U} +
+  \sum_{(u,U,\theta,C_u) \in S_v^+} b_{u,U}
+\right\rbrace \\
+& \cup \bigcup_{k_v < i \leq K} \left\lbrace \mathbf{a}_{v,\emptyset,i} = 0 \right\rbrace
 \end{align*}
@@
 If $C_{v_\bot}$ stays unsatisfiable, we let $T$ grow based on the
-unsatisfiable core.
+unsatisfiable core for $C_{v_\bot}$.
+
+\subsubsection{Refinement from Unsatisfiable Core}
+
+We first start the process from the initial Copy tree:
+\begin{align*}
+V_C = & \left\lbrace v_{v_\bot} \right\rbrace \\
+E_C = & \emptyset
+\end{align*}
+
+An unsatisfiable core $\mathcal{U}$ is a subset of $C_{v_\bot}$.
+If a vertex $v \in V$ with the copy origin $U$ satisfies:
+\begin{align*}
+& \exists t, u \in V; v \leadsto t \wedge v \leadsto u \wedge \\
+& \left( \exists i, j, k;
+\left\lbrace \left( \mathbf{a}_{s,U,j} = \ldots + \mathbf{a}_{v,U,i} + \ldots \right),
+\left( \mathbf{a}_{t,U,k} = \ldots + \mathbf{a}_{v,U,i} + \ldots \right)
+\right\rbrace \subseteq \mathcal{U} \right)
+\end{align*}
+we add $v$ to the copy tree $T$.
 

### 2013-08-20 article/algo.tex

@@
 
+\subsection{Example}
+
 \subsection{Preliminaries}
 
-The Horn clause solving algorithm receives an input problem in Horn
-graph. A \textbf{Horn graph} is a labelled directed graph $G=(V,E,\varphi)$.
+Our Horn clause solving algorithm receives an input problem in Horn
+graph. A \textbf{Horn graph} is a labelled directed acyclic graph $G=(V,E,\varphi)$.
 \begin{itemize}
@@
 of generality, we limit linear expression labels on $V_T$ vertices
-with incoming edges to either $0 \leq -1$, meaning $\bot$, or $0 \leq
-0$, meaning $\top$.
+with incoming edges to $1 \leq 0$, meaning $\bot$.
 
@@
 \begin{align*}
-\psi ::= & a_1 x_1 + \cdots + a_n x_n \leq b \mid \\
+\psi ::= & a_1 x_1 + \cdots + a_n x_n + b \leq 0 \mid \\
 & \psi \wedge \psi \mid \psi \vee \psi \mid \top \mid \bot
@@
 ), where $\rho[G]$ is a tautology. A solution is \textit{simple} if
-$\rho[P]$ is of the form $a_1 x_1 + \cdots + a_n x_n \leq b$ for every
+$\rho[P]$ is of the form $a_1 x_1 + \cdots + a_n x_n + b \leq 0$ for every
 $P$ in $\mathrm{dom}(\rho)$.
@@
 
-We consider below a Horn graph that is acyclic.
+We first propose an algorithm to compute a simple solution for the given
+$G$ if one exists. Later we add modifications to the algorithm if a
+simple solution does not exist.
+Our algorithm works in a constructive manner. List 1 shows the
+pseudo-code of our algorithm.
 

### 2013-08-20 article/algo.tex
Examples introduced for algorithm explanation
Updated examples

@@
 \subsection{Example}
+
+First, consider the following example.
+
+\begin{align*}
+x \leq 1 & \implies A(x) \\
+y \leq 2 & \implies B(y) \\
+A(x) \wedge B(y) & \implies 3x+2y \leq 10
+\end{align*}
+
+In this example, a simple solution is expressed by linear expression
+templates, whose coefficients and constants are parameterized. Assume
+that they have following predicate templates.
+
+\begin{align*}
+A(x) : c_1 x + c_2 \leq 0 \\
+B(y) : c_3 y + c_4 \leq 0
+\end{align*}
+
+By naturally encoding every Horn clause with Farkas's lemma
+constraints, we obtain the following linear constraint over
+$c_i$.
+
+\begin{align*}
+\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \\
+\lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \\
+c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \\
+\lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \lambda_3 \geq 0
+\end{align*}
+
+One of the model satisfying the constraint above is
+
+\begin{align*}
+\left\lbrace c_1 = 3 \wedge c_2 = -3 \wedge c_3 = 2 \wedge c_4 = -4 \wedge
+\lambda_1 = 3 \wedge \lambda_2 = 2 \wedge \lambda_3 = 1 \right\rbrace
+\end{align*}
+
+Second, we explain a case that some predicate variables appear
+multiple times on left-hand side in different Horn clauses.
+
+\begin{align*}
+x \leq 1 & \implies A(x) \\
+A(x) \wedge A(y) & \implies x+y \leq 2
+\end{align*}
+
+Finally, we explain a case that an input set of Horn clauses contain
+disjunctions on left-hand side.
+
+\begin{align*}
+x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
+x+1 \leq 0 \vee -x+1 \leq 0 & \implies B(x) \\
+A(x) \wedge B(x) & \implies 1 \leq 0
+\end{align*}
 

### 2013-08-21 article/algo.tex
Adding example explanation

--- 2013-08-20 article/algo.tex
@@
 \subsection{Example}
+
+\paragraph {Tree interpolation}
+First, consider the following example.
+
+\begin{align*}
+x \leq 1 & \implies A(x) \\
+y \leq 2 & \implies B(y) \\
+A(x) \wedge B(y) & \implies 3x+2y \leq 10
+\end{align*}
+
+In this example, we would like to compute linear arithmetic
+formulas for the predicate variables $A$ and $B$.
+They appear only once on both left-hand and right-hand globally in the problem.
+
+We try to obtain a simple solution that is expressed by linear expression
+templates, whose coefficients and constants are parameterized. Assume
+that they have following predicate templates.
+
+\begin{align*}
+A(x) : c_1 x + c_2 \leq 0 \\
+B(y) : c_3 y + c_4 \leq 0
+\end{align*}
+
+By naturally encoding the first Horn clause with Farkas's lemma,
+we obtain a linear constraint
+$ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
+Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will make a valid $A(x)$
+for the first clause. In the same manner, we obtain
+$ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
+from the second clause, and
+$ c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0 $
+from the third.
+
+One of the models satisfying all the constraints above is
+$( c_1, c_2, c_3, c_4, \lambda_1, \lambda_2, \lambda_3 ) = (3, -3, 2, -4, 3, 2, 1)$.
+By assigning them to the template, we obtain the following solution.
+
+\begin{align*}
+A(x) : 3 x - 3 \leq 0 \\
+B(y) : 2 y - 4 \leq 0
+\end{align*}
+
+\paragraph {DAG Horn graph solving}
+Second, we explain a case that some predicate variables appear
+multiple times on left-hand side in different Horn clauses.
+
+\begin{align*}
+x \leq 1 & \implies A(x) \\
+A(x) \wedge A(y) & \implies x+y \leq 2
+\end{align*}
+
+Assigning a template
+$A(x) : c_1 x + c_2 \leq 0$, we obtain the same
+constraint for the first clause as the previous example.
+
+\begin{align} \label{eq:constr1}
+\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0
+\end{align}
+
+Now we have $A$'s appearance twice on the second clause, and we try to
+solve first by a weaker constraint, which we give allow different
+assignments for multiple occurrence. Virtually it is equivalent to
+solve the following problem with the templates.
+
+\begin{align*}
+\text{Problem:} \\
+x \leq 1 & \implies A_1(x) \\
+x \leq 1 & \implies A_2(x) \\
+A_1(x) \wedge A_2(y) & \implies x+y \leq 2 \\
+\\
+\text{Templates:} \\
+A_1(x) \text{ : } & c_{1,1} x + c_{2,1} \leq 0 \\
+A_2(x) \text{ : } & c_{1,2} x + c_{2,2} \leq 0 \\
+A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
+\end{align*}
+
+By this, duplicating the constraint ~\ref{eq:constr1}, we obtain the
+whole constraint as follows.
+
+\begin{align*}
+& \lambda_{1,1} = c_{1,1} \wedge - \lambda_{1,1} \geq c_{2,1} \wedge \lambda_{1,1} \geq 0 \wedge \\
+& \lambda_{1,2} = c_{1,2} \wedge - \lambda_{1,2} \geq c_{2,2} \wedge \lambda_{1,2} \geq 0 \wedge \\
+& c_{1,1} = \lambda_2 \wedge c_{1,2} = \lambda_2 \wedge c_{2,1} + c_{2,2} \geq 2 \lambda_2 \wedge \lambda_2 > 0
+\end{align*}
+
+This constraint is satisfiable and one model is
+\[ ( c_{1,1}, c_{1,2}, c_{2,1}, c_{2,2}, \lambda_{1,1}, \lambda_{1,2}, \lambda_2 ) =
+( 1, -1, 1, -1, 1, 1, 1 ) \] and the solution then becomes
+$A(x) : (x -1 \leq 0) \wedge (x -1 \leq 0)$.
+
+Since the weak constraint is satisfiable, we now refine the constraint
+to obtain the better solution. That is, simply eliminate the
+duplication to build the following constraint for the original template.
+
+\begin{align*}
+& \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \\
+& c_1 = \lambda_2 \wedge c_1 = \lambda_2 \wedge 2 c_2 \geq 2 \lambda_2 \wedge \lambda_2 > 0
+\end{align*}
+
+Solving this constraint simply gives us a solution $A(x) : x -1 \leq 0$.
+
+\paragraph {Disjunctive Horn graph solving}
+Finally, we explain a case that an input set of Horn clauses contain
+disjunctions on left-hand side.
+
+\begin{align*}
+x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
+x+1 \leq 0 \vee -x+1 \leq 0 & \implies B(x) \\
+A(x) \wedge B(x) & \implies 1 \leq 0
+\end{align*}
 

### 2013-08-21 article/algo.tex

@@
 \end{align*}
+
+Same as the previous examples, templates for predicate variables are
+prepared.
+
+\begin{align*}
+A(x) : c_1 x + c_2 \leq 0 \\ B(x) : c_3 x + c_4 \leq 0
+\end{align*}
+
+The constraints generated are follows.
+
+\begin{align*}
+& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
+& \lambda_3 = c_3 \wedge \lambda_3 \geq c_4 \wedge \lambda_3 \geq 0 \wedge \\
+& - \lambda_4 = c_3 \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
+& c_1 + c_3 = 0 \wedge c_2 + c_4 > 0
+\end{align*}
+
+The constraints here become unsatisfiable, and one possible
+unsatisfiable core is
+
+\begin{align*}
+\lambda_3 & = c_3 \\
+\lambda_3 & \geq c_4 \\
+- \lambda_4 & = c_3 \\
+\lambda_4 & \geq c_4 \\
+c_2 + c_4 & > 0 \\
+0 & \geq c_2
+\end{align*}
+
+Because the unsatisfiable core contains the constraint that restrict
+the parameter $c_3$ or $c_4$ to take different values according to
+originating disjunctions. This let us determine to relax the template
+to allow a enlarged solution.
 

### 2013-08-21 article/algo.tex

@@
 $ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
-Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will make a valid $A(x)$
-for the first clause. In the same manner, we obtain
+Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will make $A(x)$
+valid for the first clause. In the same manner, we obtain
 $ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
@@
 
-Now we have $A$'s appearance twice on the second clause, and we try to
-solve first by a weaker constraint, which we give allow different
-assignments for multiple occurrence. Virtually it is equivalent to
-solve the following problem with the templates.
+Sinve we have $A$'s appearance twice on the second clause,
+it is possible for $A$ to have two expressions each of which is used
+for $A(x)$ and $A(y)$. Although we need to use the same template for
+multiple occurrences in order to compute the simplest solution,
+it may make the whole constraints unsatisfiable if the simplest
+solution does not exist.
+
+Therefore, we generate a weaker constraint for the second clause that
+allows different expressions of $A$ for multiple occurrence. If the
+whole constraint is satisfiable in the case, we try to solve in a
+stronger constraint to generate a simple solution.
+
+In this specific case, it is virtually equivalent to solve the
+following problem in a weaker constraint.
 
@@
 
-By this, duplicating the constraint ~\ref{eq:constr1}, we obtain the
-whole constraint as follows.
+By duplicating the constraint~\ref{eq:constr1} of $A$ for newly
+created $A_1$ and $A_2$, we obtain the whole constraint as follows.
 
@@
 The constraints here become unsatisfiable, and one possible
-unsatisfiable core is
+unsatisfiable core is:
 
@@
 
-Because the unsatisfiable core contains the constraint that restrict
-the parameter $c_3$ or $c_4$ to take different values according to
-originating disjunctions. This let us determine to relax the template
-to allow a enlarged solution.
+Because the unsatisfiable core contains constraints originating from
+the disjunctions to be expressed in a single term template. This let
+us determine to relax the template to allow a enlarged solution.
+
+Now we slightly change the problem and the template.
+
+\begin{align*}
+\text{Problem:} \\
+x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
+x+1 \leq 0 & \implies B_1(x) \\
+-x+1 \leq 0 & \implies B_2(x) \\
+A(x) \wedge B_1(x) & \implies 1 \leq 0 \\
+A(x) \wedge B_2(x) & \implies 1 \leq 0 \\
+\\
+\text{Templates:} \\
+A(x) \text{ : } & c_1 x + c_2 \leq 0 \\
+B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
+B(x) \text{ : } & B_1(x) \vee B_2(x)
+\end{align*}
+
+When we construct the constraints, as same as previous, we again
+obtain an unsatisfiable core:
+
+\begin{align*}
+& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
+& \lambda_3 = c_3 \wedge \lambda_3 \geq c_4 \wedge \lambda_3 \geq 0 \wedge \\
+& - \lambda_4 = c_3 \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
+& c_1 + c_3 = 0 \wedge c_2 + c_4 > 0
+\end{align*}
+
+\begin{align*}
+\lambda_3 & = c_3 \\
+\lambda_3 & \geq c_4 \\
+- \lambda_4 & = c_3 \\
+\lambda_4 & \geq c_4 \\
+c_2 + c_4 & > 0 \\
+0 & \geq c_2
+\end{align*}
+
+\begin{align*}
+\text{Problem:} \\
+x \leq 0 \wedge -x \leq 0 & \implies A_i(x) \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
+x+1 \leq 0 & \implies B_1(x) \\
+-x+1 \leq 0 & \implies B_2(x) \\
+A_i(x) \wedge B_j(x) & \implies 1 \leq 0 \quad (i,j \in \left\lbrace 1,2 \right\rbrace ) \\
+\\
+\text{Templates:} \\
+A_i(x) \text{ : } & c_{1,i} x + c_{2,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
+A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
+B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
+B(x) \text{ : } & B_1(x) \vee B_2(x)
+\end{align*}
 
@@
 \psi ::= & a_1 x_1 + \cdots + a_n x_n + b \leq 0 \mid \\
-& \psi \wedge \psi \mid \psi \vee \psi \mid \top \mid \bot
+& \psi \wedge \psi \mid \psi \vee \psi
 \end{align*}

### 2013-08-21 article/algo.tex
Clarify TODOs
Algorithm explanation and pseudo-code

@@
 \usepackage{algorithmic}
+\usepackage{graphicx}
 \usepackage{listings}
@@
 \begin{align*}
-& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
-& \lambda_3 = c_3 \wedge \lambda_3 \geq c_4 \wedge \lambda_3 \geq 0 \wedge \\
-& - \lambda_4 = c_3 \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
-& c_1 + c_3 = 0 \wedge c_2 + c_4 > 0
+\text{Constraint:} \\
+TODO \\
+\\
+\text{Unsatisfiable core:} \\
+TODO
 \end{align*}
 
-\begin{align*}
-\lambda_3 & = c_3 \\
-\lambda_3 & \geq c_4 \\
-- \lambda_4 & = c_3 \\
-\lambda_4 & \geq c_4 \\
-c_2 + c_4 & > 0 \\
-0 & \geq c_2
-\end{align*}
+This time, the unsatisfiable core contains the constraints that
+originate from
+
+TODO
 
@@
 \end{align*}
+
+Finally, we obtain the following constraint:
+
+TODO
 
@@
 
-We first propose an algorithm to compute a simple solution for the given
-$G$ if one exists. Later we add modifications to the algorithm if a
-simple solution does not exist.
-Our algorithm works in a constructive manner. List 1 shows the
-pseudo-code of our algorithm.
+We first propose an algorithm to compute a simple solution for the
+given $G$ if one exists.  Later we add modifications to the algorithm
+for the case that a simple solution does not exist.
+
+Our algorithm works in a constructive manner.
+Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our algorithm.
+
+\newcommand{\solveHorn}{\ensuremath{\mbox{\sc SolveHornGraph}}}
+\newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
+
+\begin{algorithm}
+\caption{$ \solveHorn (G) $}\label{alg:solveHorn}
+\begin{algorithmic}
+\STATE {$v_\bot \gets $ the root node of $G$}
+\STATE {$visited \gets \left\lbrace v_\bot \right\rbrace$}
+
+\STATE {$C \gets \genConstr (v_\bot, visited)$}
+\STATE {$c_{v_\bot} \gets C(v_\bot)$}
+\IF {$c_{v_\bot}$ is satisfiable}
+  \IF {$visited$ contains all predicate variable vertices}
+    \STATE {$visited \gets a$}
+    \COMMENT {build solution}
+  \ELSE
+    \STATE {$visited \gets a$}
+  \ENDIF
+\ELSE
+  \STATE {$\mathcal{UC} \gets $ unsatisfiable core for $c_{v_\bot}$}
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{$ \genConstr (v, visited) $}\label{alg:genConstr}
+\begin{algorithmic}
+\STATE {$m \gets \emptyset$}
+\FORALL {$v_\Rightarrow \in preds(v)$}
+  \FORALL {$v_p \in preds(v_\Rightarrow)$}
+    \STATE { $ l := \varphi (v_p) $ }
+    \IF {$l$ is a predicate variable}
+      \STATE {$m' \gets \genConstr (v_p) $}
+      \STATE {$m \gets m \cup m'$}
+    \ELSE
+      \STATE { $id$ }
+    \ENDIF
+  \ENDFOR
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
+The construction of the algorithm is two fold.
+\begin{itemize}
+\item \textbf{Iterative Constraint Building} The process starts from
+  generating the weakest constraint that the graph allows and
+  iteratively strengthen a constraint until the strongest one.
+\item \textbf{Graph Restructuring} If the constraint become
+  unsatisfiable during the iteration, restructure the input Horn graph
+  to enlarge the solution template.
+\end{itemize}
+
+All vertices with predicate variable labels are assigned a simple
+template.  For a vertex with a label $P(\mathbf{x})$, the algorithm
+assigns a template $\mathbf{a_P} \mathbf{x} + b_P \leq 0$.  Parameters
+$\mathbf{a_P}$ and $b_P$ are specified by a linear constraint.
+
+For each predicate variable vertex, build a linear constraint
+  that parameters in predicate variable templates must satisfy.
+
+$\genConstr$ procedure produces the constraint for all vertices
+
+In this algorithm, vertices with a predicate variable label must have
+a simple solution assignment.
 

### 2013-08-22 article/algo.tex
Update pseudo-code

@@
 by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.  Without loss
-of generality, we limit linear expression labels on $V_T$ vertices
-with incoming edges to $1 \leq 0$, meaning $\bot$.
+of generality, we assume the existence of the root vertex $v_\bot \in
+V_T$ with linear expression labels $1 \leq 0$, meaning $\bot$. We also
+assume that other vertices with linear expression labels have no
+incoming edges to them.
 
@@
 \newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
+\newcommand{\splitGraph}{\ensuremath{\mbox{\sc RestructureGraph}}}
 
@@
 \STATE {$v_\bot \gets $ the root node of $G$}
-\STATE {$visited \gets \left\lbrace v_\bot \right\rbrace$}
-
-\STATE {$C \gets \genConstr (v_\bot, visited)$}
+\STATE {$V \gets \left\lbrace v_\bot \right\rbrace$}
+\REPEAT
+  \STATE {$C \gets \genConstr (v_\bot, V)$}
   \STATE {$c_{v_\bot} \gets C(v_\bot)$}
-\IF {$c_{v_\bot}$ is satisfiable}
-  \IF {$visited$ contains all predicate variable vertices}
-    \STATE {$visited \gets a$}
-    \COMMENT {build solution}
+  \STATE {Evaluate satisfiability of $c_{v_\bot}$}
+  \IF {$c_{v_\bot}$ is unsatisfiable}
+    \STATE {$\mathcal{UC} \gets $ unsatisfiable core for $c_{v_\bot}$}
+    \STATE {$G \gets \splitGraph (G, \mathcal{UC})$}
+    \STATE {$V \gets \left\lbrace v_\bot \right\rbrace$}
   \ELSE
-    \STATE {$visited \gets a$}
+    \STATE {Enlarge $V$}
+    \STATE {$m \gets $ one model of $c_{v_\bot}$}
   \ENDIF
-\ELSE
-  \STATE {$\mathcal{UC} \gets $ unsatisfiable core for $c_{v_\bot}$}
-\ENDIF
+\UNTIL {$V$ contains all predicate variable vertices}
+\STATE {Build solution by $m$}
 \end{algorithmic}
@@
 
-The construction of the algorithm is two fold.
+The construction of the algorithm is in two fold.
 \begin{itemize}
@@
 \end{itemize}
+This algorithm enables a solution to be build as simple as possible.
 
@@
 
-For each predicate variable vertex, build a linear constraint
-  that parameters in predicate variable templates must satisfy.
+For each predicate variable vertex, build a linear constraint that
+parameters in predicate variable templates must satisfy.  The process
+is equivalent to generate a linear constraint in interpolation.
 
-$\genConstr$ procedure produces the constraint for all vertices
-
-In this algorithm, vertices with a predicate variable label must have
-a simple solution assignment.
+$\genConstr$ procedure produces the constraint for all vertices. The
+constraints are first built the weakest so that the disjunctiveness of
+whole solution template is fixed first.
 

### 2013-08-22 article/algo.tex

@@
 \item \textbf{Iterative Constraint Building} The process starts from
-  generating the weakest constraint that the graph allows and
+  generating the weakest constraint that the input Horn graph allows and
   iteratively strengthen a constraint until the strongest one.
@@
 
-All vertices with predicate variable labels are assigned a simple
-template.  For a vertex with a label $P(\mathbf{x})$, the algorithm
-assigns a template $\mathbf{a_P} \mathbf{x} + b_P \leq 0$.  Parameters
-$\mathbf{a_P}$ and $b_P$ are specified by a linear constraint.
+For the input graph, each predicate variable vertex has a unique
+predicate variable label assigned.  All predicate variables are
+assigned a simple template.  For a predicate variable $P(\mathbf{x})$,
+the algorithm assigns a template $\mathbf{a}_P \mathbf{x} + b_P \leq
+0$.  Parameters $\mathbf{a}_P$ and $b_P$ are specified by a linear
+constraint.  Then, according to the graph structure, the algorithm
+builds a linear constraint that parameters in predicate variable
+templates must satisfy.  For constraint building, the algorithm
+traverses the graph in the topological order on $V_\Rightarrow$, and
+concatenate the constraints from predecessors to build the problem
+constraints. Each process to build a constraint on $\mathcal{HC}_v$
+for all vertices $v \in V_\Rightarrow$ is equivalent to generate a
+linear constraint in interpolation.
 
-For each predicate variable vertex, build a linear constraint that
-parameters in predicate variable templates must satisfy.  The process
-is equivalent to generate a linear constraint in interpolation.
+$\genConstr$ procedure produces the constraint for a given vertex $v$.
 
-$\genConstr$ procedure produces the constraint for all vertices. The
-constraints are first built the weakest so that the disjunctiveness of
+The constraints are first built the weakest so that the disjunctiveness of
 whole solution template is fixed first.

### 2013-08-22 article/algo.tex

--- 2013-08-22 article/algo.tex
@@
 \begin{document}
-\maketitle
 
-\section{Algorithm}
+\section{Horn clause solving}
+
+We propose a new algorithm for solving Horn clauses to generate a
+solution which is as simple as possible.  This enables us to obtain
+appropriate predicates that are suitable for predicate abstraction in
+the program verification techniques.
+
+The construction of the algorithm roughly consists of three parts.
+\begin{itemize}
+\item \textbf{Horn Graph Building} converts input Horn clauses into a
+  \emph{Horn graph} that express the equivalent problem. The process
+  also generates a predicate template for every predicate variable
+  that the solution will become.  The Horn graph is used throughout
+  the algorithm to represent the problem under certain solution
+  template.
+\item \textbf{Iterative Constraint Generation} is a process to build a
+  constraint to compute a solution.  It starts from generating the
+  weakest constraint in the sense that the resulting solution would not
+  be simple yet.  The process iteratively strengthen a constraint
+  until the strongest one while preserving the constraint
+  satisfiability. The strongest constraint gives an instanciation of
+  templates.
+\item \textbf{Graph Restructuring} If the constraint become
+  unsatisfiable during the iteration, the process restructure the
+  input Horn graph to enlarge the solution template.  This is done by
+  considering the unsatisfiable core of the constraint.
+\end{itemize}
+Our algorithm works in a constructive manner.
+
+Each predicate variable vertex in an input graph has a unique
+predicate variable label.  All predicate variables on them are
+assigned simple templates.  Those predicate templates are expressed
+with template parameters.  Then, according to the graph structure, the
+algorithm builds a linear constraint that parameters must satisfy.
+For constraint building, the algorithm traverses the input graph in
+the topological order on $V_\Rightarrow$, and concatenate the
+constraints from predecessors to build the problem constraints.  Each
+process to build a constraint on $\mathcal{HC}_v$ for all vertices $v
+\in V_\Rightarrow$ is equivalent to generate a linear constraint in
+interpolation.
 
@@
 given $G$ if one exists.  Later we add modifications to the algorithm
-for the case that a simple solution does not exist.
-
-Our algorithm works in a constructive manner.
-Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our algorithm.
+for the case that a simple solution does not exist.  Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our algorithm.
 
@@
 
+Specifically, for a predicate variable $P(\mathbf{x})$,
+the algorithm assigns a template $\mathbf{a}_P \mathbf{x} + b_P \leq
+0$.  Parameters $\mathbf{a}_P$ and $b_P$ are specified by a linear
+constraint.
+
 \begin{algorithm}
@@
 
-The construction of the algorithm is in two fold.
-\begin{itemize}
-\item \textbf{Iterative Constraint Building} The process starts from
-  generating the weakest constraint that the graph allows and
-  iteratively strengthen a constraint until the strongest one.
-\item \textbf{Graph Restructuring} If the constraint become
-  unsatisfiable during the iteration, restructure the input Horn graph
-  to enlarge the solution template.
-\end{itemize}
-This algorithm enables a solution to be build as simple as possible.
+$\genConstr$ procedure produces the constraint for a given vertex $v$.
 
-All vertices with predicate variable labels are assigned a simple
-template.  For a vertex with a label $P(\mathbf{x})$, the algorithm
-assigns a template $\mathbf{a_P} \mathbf{x} + b_P \leq 0$.  Parameters
-$\mathbf{a_P}$ and $b_P$ are specified by a linear constraint.
-
-For each predicate variable vertex, build a linear constraint that
-parameters in predicate variable templates must satisfy.  The process
-is equivalent to generate a linear constraint in interpolation.
-
-$\genConstr$ procedure produces the constraint for all vertices. The
-constraints are first built the weakest so that the disjunctiveness of
+The constraints are first built the weakest so that the disjunctiveness of
 whole solution template is fixed first.

### 2013-08-22 article/algo.tex
Explanation of Horn graph

@@
 
-We propose a new algorithm for solving Horn clauses to generate a
-solution which is as simple as possible.  This enables us to obtain
-appropriate predicates that are suitable for predicate abstraction in
-the program verification techniques.
+We propose a new algorithm for solving recursion-less Horn clauses to
+generate a solution which is as simple as possible.  This enables us
+to obtain appropriate predicates that are suitable for predicate
+abstraction in the program verification techniques.
 
-The construction of the algorithm roughly consists of three parts.
+The construction of the algorithm roughly consists of two parts.
 \begin{itemize}
-\item \textbf{Horn Graph Building} converts input Horn clauses into a
-  \emph{Horn graph} that express the equivalent problem. The process
-  also generates a predicate template for every predicate variable
-  that the solution will become.  The Horn graph is used throughout
-  the algorithm to represent the problem under certain solution
-  template.
 \item \textbf{Iterative Constraint Generation} is a process to build a
   constraint to compute a solution.  It starts from generating the
-  weakest constraint in the sense that the resulting solution would not
-  be simple yet.  The process iteratively strengthen a constraint
+  weakest constraint in the sense that the resulting solution would
+  not be simple yet.  The process iteratively strengthen a constraint
   until the strongest one while preserving the constraint
-  satisfiability. The strongest constraint gives an instanciation of
-  templates.
-\item \textbf{Graph Restructuring} If the constraint become
+  satisfiability.  The strongest constraint gives a simple solution.
+\item \textbf{Problem Restructuring} If the constraint become
   unsatisfiable during the iteration, the process restructure the
-  input Horn graph to enlarge the solution template.  This is done by
-  considering the unsatisfiable core of the constraint.
+  problem equivalent to obtain a less simpler solution.  This process
+  considers the unsatisfiable core of the constraint.
 \end{itemize}
-Our algorithm works in a constructive manner.
 
-Each predicate variable vertex in an input graph has a unique
-predicate variable label.  All predicate variables on them are
-assigned simple templates.  Those predicate templates are expressed
-with template parameters.  Then, according to the graph structure, the
-algorithm builds a linear constraint that parameters must satisfy.
-For constraint building, the algorithm traverses the input graph in
-the topological order on $V_\Rightarrow$, and concatenate the
-constraints from predecessors to build the problem constraints.  Each
-process to build a constraint on $\mathcal{HC}_v$ for all vertices $v
-\in V_\Rightarrow$ is equivalent to generate a linear constraint in
-interpolation.
+All predicate variables in an input problem are assigned simple
+templates.  Those predicate templates are parameterized.  In the
+iterative constraint generation process, algorithm builds a linear
+constraint that parameters in templates must satisfy.  This is done in
+a constructive manner, and constraints are build in the order of
+implication dependency in Horn clauses.  That is, when an input
+problem has a Horn clause thath has a predicate variable $P$ on its
+left-hand and $Q$ on its right-hand, a constraint to describe
+parameters in $P$'s template are first build before those in $Q$'s
+template.  When all parameters in the whole problem are described, the
+built constraint are evaluated to obtain a model.  Its instanciation
+becomes an appropriate solution for the input problem.
+
+The novelty of our algorithm lies in the incremental constraint
+refinement.  The algorithm try from the weakest constraint that an
+input problem allows that does not give a simple solution, and
+strengthen the constraint until it reaches to the simplest solution
+that the problem has.  When a constraint becomes unsatisfiable, the
+algorithm analyzes the unsatisfiable core and transform the problem to
+obtain less simple solution.
+
+This construction of the algorithm enables us to compute a relatively
+simpler solution compared to existing algorithms in a reasonable time.
 
@@
 
-In this example, we would like to compute linear arithmetic
-formulas for the predicate variables $A$ and $B$.
-They appear only once on both left-hand and right-hand globally in the problem.
+In this example, we would like to compute linear arithmetic formulas
+for the predicate variables $A$ and $B$.  They appear only once on
+both left-hand and right-hand globally in the problem.
 
-We try to obtain a simple solution that is expressed by linear expression
-templates, whose coefficients and constants are parameterized. Assume
-that they have following predicate templates.
+We try to obtain a simple solution that is expressed by linear
+expression templates, whose coefficients and constants are
+parameterized.  Assume that they have following predicate templates.
 
@@
 
-By naturally encoding the first Horn clause with Farkas's lemma,
-we obtain a linear constraint
+By naturally encoding the first Horn clause with Farkas's lemma, we
+obtain a linear constraint
 $ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
-Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will make $A(x)$
-valid for the first clause. In the same manner, we obtain
+Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will
+make $A(x)$ valid for the first clause.  In the same manner, we obtain
 $ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
@@
 
-\paragraph {DAG Horn graph solving}
+\paragraph {DAG problem solving}
 Second, we explain a case that some predicate variables appear
@@
 
-Assigning a template
-$A(x) : c_1 x + c_2 \leq 0$, we obtain the same
+Assigning a template $A(x) : c_1 x + c_2 \leq 0$, we obtain the same
 constraint for the first clause as the previous example.
@@
 
-Sinve we have $A$'s appearance twice on the second clause,
-it is possible for $A$ to have two expressions each of which is used
-for $A(x)$ and $A(y)$. Although we need to use the same template for
-multiple occurrences in order to compute the simplest solution,
-it may make the whole constraints unsatisfiable if the simplest
-solution does not exist.
+Sinve we have $A$'s appearance twice on the second clause, it is
+possible for $A$ to have two expressions each of which is used for
+$A(x)$ and $A(y)$.  Although we need to use the same template for
+multiple occurrences in order to compute the simplest solution, it may
+make the whole constraints unsatisfiable if the simplest solution does
+not exist.
 
@@
 to obtain the better solution.  That is, simply eliminate the
-duplication to build the following constraint for the original template.
+duplication to build the following constraint for the original
+template.
 
@@
 
-\paragraph {Disjunctive Horn graph solving}
+\paragraph {Disjunctive Horn clause solving}
 Finally, we explain a case that an input set of Horn clauses contain
@@
 
+The process also generates a predicate template for every predicate
+variable that the solution will become.  under certain solution
+template.
+
 \subsection{Preliminaries}
 
-Our Horn clause solving algorithm receives an input problem in Horn
-graph. A \textbf{Horn graph} is a labelled directed acyclic graph $G=(V,E,\varphi)$.
+For ease of discussion, our algorithm handles an input set of Horn
+clauses by a \emph{Horn graph} that express the equivalent problem.
+The graph is then used throughout the algorithm to represent the
+problem.  A Horn graph is a labelled directed acyclic graph
+$G=(V,E,\varphi)$.
 \begin{itemize}
@@
 \end{itemize}
-
 The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
@@
 We implicitly assume each $\mathcal{HC}_v$ is universally quantified
-by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.  Without loss
-of generality, we assume the existence of the root vertex $v_\bot \in
-V_T$ with linear expression labels $1 \leq 0$, meaning $\bot$. We also
-assume that other vertices with linear expression labels have no
-incoming edges to them.
+by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.
+
+Without loss of generality, we assume the existence of the root vertex
+$v_\bot \in V_T$ with linear expression labels $1 \leq 0$, meaning
+$\bot$.  This is because if we have a Horn clause $P \implies e$ which
+$e$ is a linear expression, we can convert it to $P \wedge \neg e
+\implies 1 \leq 0$.  We also assume that other vertices with linear
+expression labels have no incoming edges to them.
 
@@
 ), where $\rho[G]$ is a tautology.  A solution is \textit{simple} if
-$\rho[P]$ is of the form $a_1 x_1 + \cdots + a_n x_n + b \leq 0$ for every
-$P$ in $\mathrm{dom}(\rho)$.
+$\rho[P]$ is of the form $a_1 x_1 + \cdots + a_n x_n + b \leq 0$ for
+every $P$ in $\mathrm{dom}(\rho)$.
 
@@
 We first propose an algorithm to compute a simple solution for the
-given $G$ if one exists.  Later we add modifications to the algorithm
-for the case that a simple solution does not exist.  Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our algorithm.
+given Horn graph $G$ if one exists.  Later we add modifications to the
+algorithm for the case that a simple solution does not exist for $G$.
+Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our algorithm.
 
@@
 whole solution template is fixed first.
+
+traverses the input graph in the topological order on
+$V_\Rightarrow$, and concatenate the constraints from predecessors to
+build the problem constraints.  Each process to build a constraint on
+$\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$ is equivalent
+to generate a linear constraint in interpolation.
 

### 2013-08-22 article/algo.tex
Drastic change...

@@
   in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
-  inequality $a_1 x_1 + \cdots + a_n x_n \leq b$.
+  inequality $a_1 x_1 + \cdots + a_n x_n + b \leq 0$.
 \end{itemize}
@@
 Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$.
-
 \begin{align*}
@@
 \end{align*}
-
 We implicitly assume each $\mathcal{HC}_v$ is universally quantified
@@
 
-Without loss of generality, we assume the existence of the root vertex
-$v_\bot \in V_T$ with linear expression labels $1 \leq 0$, meaning
-$\bot$.  This is because if we have a Horn clause $P \implies e$ which
-$e$ is a linear expression, we can convert it to $P \wedge \neg e
-\implies 1 \leq 0$.  We also assume that other vertices with linear
-expression labels have no incoming edges to them.
+Horn term vertices with predicate variable labels are unique each
+other that any of them do not have the same variable on their labels.
+From now on, we may denote a Horn term vertex with a predicate
+variable label by its predicate variable name if no ambiguity is
+present.
 
-The solution of $G$ is a map $\rho$ from predicate variables to linear
-predicates of the form $\lambda x_1, \cdots ,x_n. \psi $ (where $\psi$
-is a formula constructed from
+Without loss of generality, we assume the existence of the \emph{root
+  vertex} $v_\bot \in V_T$ with linear expression labels $1 \leq 0$,
+meaning $\bot$.  This is because if we have a Horn clause $P \implies
+e$ which $e$ is a linear expression, we can convert it to $P \wedge
+\neg e \implies 1 \leq 0$.  We also assume that other vertices with
+linear expression labels have no incoming edges to them.
+
+For simplicity of discussion, we define \emph{term arity} $k_v$ for
+every Horn term vertex $v \in V_T$ with incoming edges.
+\begin{align*}
+k_v =
+\begin{cases}
+\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
+0 & \mbox{if } \varphi(v) = \top \mbox { or } \varphi(v) = \bot
+\end{cases}
+\end{align*}
+We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.
+
+A variable mapping label on edges \theta is used to map the index of
+one predicate variable to another. (TODO?)
+
+(Example?)
+
+The algorithm assigns a \emph{predicate template} for all predicate
+variables in $G$. A template for a predicate variable $P(x_1, \ldots, x_n)$
+is a formula $\psi$ constructed from the following definition.
 \begin{align*}
@@
 \end{align*}
-), where $\rho[G]$ is a tautology.  A solution is \textit{simple} if
-$\rho[P]$ is of the form $a_1 x_1 + \cdots + a_n x_n + b \leq 0$ for
-every $P$ in $\mathrm{dom}(\rho)$.
+A linear expression in the definition above may also be represented in
+a equivalent vector form $\mathbf{a} \mathbf{x} + b \leq 0$ for short.
+Parameters $\mathbf{a}$ and $b$ are specified by a linear constraint.
+
+The solution of $G$ is a map $\rho$ from predicate variables to linear
+predicates of the form $\lambda x_1, \ldots ,x_n. \psi$ where
+parameters in $\psi$ are replaced with constant and $\rho[G]$ becomes
+a tautology.  A solution is \textit{simple} if $\rho[P]$ is of the
+form $\mathbf{a} \mathbf{x} + b \leq 0$ for every $P$ in
+$\mathrm{dom}(\rho)$.
 
@@
 algorithm for the case that a simple solution does not exist for $G$.
+In such a case, we transform the input graph $G$ to have a larger
+solution template.
+
+We define the relation $\leadsto$ over $V_T$ as follows.
+\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow
+\exists a \in V_\Rightarrow; \edgel{v}{\theta}{a} \wedge \edge{a}{v} \]
+We may omit $\theta$ and write $u \leadsto v$ for convenience in later
+discussions.
+
+By the assumption that $G$ is acyclic, $\leadsto^+$ is a well-founded
+relation. $v_\bot$ is the greatest element with respect to
+$\leadsto^+$.
+
+For each predicate variable $v \in V_T$, the algorithm assigns a
+predicate template $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ and define
+$C_v$ which is a set of linear expressions on template parameters
+$\mathbf{a}_v$ and $b_v$.  Intuitively, it denotes the set of
+inequalities:
+\begin{align*}
+\left\lbrace
+ \mathbf{a}_v \mathbf{x}_v + b_v \leq 0 \middle|
+ \exists \left( \textsc{FV}(C_v)
+  \setminus (\mathbf{a}_v \cup \left\lbrace b_v \right\rbrace
+ \right); C
+\right\rbrace
+\end{align*}
+
+$C_v$ is defined by well-founded induction on $v$ with respect to the
+relation $\leadsto^+$. Then, $C_v$ can be shown as follows.
+
+\begin{align*}
+\hat C_{a,v} = &
+ \bigcup_{\edgel{u}{\theta}{a}} C_u \cup
+ \bigcup_{0 < i \leq K}
+ \left\lbrace
+  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{a}} \mathbf{a}_{u, \theta^{-1} (i)}
+ \right\rbrace \\
+ & \hspace{2cm} \cup
+ \left\lbrace
+  b_v \geq \sum_{\edgel{u}{\theta}{a}} b_u
+ \right\rbrace \cup
+ \bigcup_{k_v < i \leq K}
+ \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
+\\
+C_v = & \bigcup_{\edge{a}{v}} \hat C_{a,v} \cup
+\begin{cases}
+\emptyset
+& \mbox{if } \varphi(v) = P(\mathbf{x}_v) \\
+\left\lbrace
+ \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
+ b_v = \lambda_v b_0
+\right\rbrace
+& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0
+\end{cases}
+\end{align*}
+
+For simplicity, we assume variables $\mathbf{a}_u$, $b_u$ are obtained
+from $C_u$.
+
+Any model $\sigma$ to the constraint $C_{v_\bot}$ gives a simple
+solution for $G$ if $C_{v_\bot}$ is satisfiable, i.e., we should have:
+
+\begin{quote}
+If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
+\begin{align*}
+ \rho = \left\lbrace
+  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \middle|
+  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
+ \right\rbrace
+\end{align*}
+is a solution of $G$.
+\end{quote}
+
+\subsection {Graph transformation}
+
 Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our algorithm.
@@
 
-Specifically, for a predicate variable $P(\mathbf{x})$,
-the algorithm assigns a template $\mathbf{a}_P \mathbf{x} + b_P \leq
-0$.  Parameters $\mathbf{a}_P$ and $b_P$ are specified by a linear
-constraint.
+The constraints are first built the weakest so that the disjunctiveness of
+whole solution template is fixed first.  The strength matters that the
+predicate variables are used as the same expressions depending on
+different appearances of the predicate variable.
+
+The strngeth of the constraint is controlled by a set of vertices $V$
+in the pseudo-code.  Predicate variable vertices in $V$ are treated
+strictly unique across the different appearances of the predicate
+variables in Horn clauses.
+
+$\genConstr$ procedure produces the constraint for a given vertex $v$.
 
@@
 
-$\genConstr$ procedure produces the constraint for a given vertex $v$.
-
-The constraints are first built the weakest so that the disjunctiveness of
-whole solution template is fixed first.
-
 traverses the input graph in the topological order on
@@
 to generate a linear constraint in interpolation.
-
-For simplicity of discussion, we define \textbf{term arity} $k_v$ for
-every Horn term vertex $v \in V_T$ with incoming edges.
-\begin{align*}
-k_v =
-\begin{cases}
-\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
-0 & \mbox{if } \varphi(v) = \top \mbox { or } \varphi(v) = \bot
-\end{cases}
-\end{align*}
-We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.
-
-We define the relation $\leadsto$ over $V_T$ as follows.
-\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow
-\exists a \in V_\Rightarrow; \edgel{v}{\theta}{a} \wedge \edge{a}{v} \]
-We may omit $\theta$ and write $u \leadsto v$ for convenience in later
-discussions.
-
-By the assumption that $G$ is acyclic, $\leadsto^+$ is a well-founded
-relation. We assume without loss of generality that $V_T$ contains the
-greatest element $v_\bot$ with respect to $\leadsto^+$; otherwise we
-can always implicitly assume $\varphi(v_\bot) = \top$ such that
-$\forall v \in V_T \setminus \{v_\bot\}; v \mathop{\leadsto}^\emptyset v_\bot$.
-
-For each $v \in V_T$, we define
-$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v \leq b_v, C_v)$
-, where
-\begin{itemize}
-\item $\mathbf{a}_v \mathbf{x}_v \leq b_v$ is a linear inequality that
-  contains coefficient variables $\mathbf{a}_v$, and
-\item $C_v$ is a set of linear expressions on coefficient variables.
-\end{itemize}
-Intuitively, the pair $(\mathbf{a} \mathbf{x} \leq b, C)$ denotes the
-set of inequalities:
-\begin{align*}
-\left\lbrace
- \mathbf{a} \mathbf{x} \leq b \middle|
- \exists \left( \textsc{FV}(C)
-  \setminus (\mathbf{a} \cup \left\lbrace b \right\rbrace
- \right); C
-\right\rbrace
-\end{align*}
-
-$\Delta(v)$ is defined by well-founded induction on $v$ with respect
-to the relation $\leadsto^+$. Then, $C_v$ in
-$\Delta(v) = \left( \mathbf{a}_v \mathbf{x}_v \leq b_v, C_v \right)$
-can be shown as follows.
-
-\begin{align*}
-\hat C_{a,v} = &
- \bigcup_{\edgel{u}{\theta}{a}} C_u \cup
- \bigcup_{0 < i \leq K}
- \left\lbrace
-  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{a}} \mathbf{a}_{u, \theta^{-1} (i)}
- \right\rbrace \\
- & \hspace{2cm} \cup
- \left\lbrace
-  b_v \geq \sum_{\edgel{u}{\theta}{a}} b_u
- \right\rbrace \cup
- \bigcup_{k_v < i \leq K}
- \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
-\\
-C_v = & \bigcup_{\edge{a}{v}} \hat C_{a,v} \cup
-\begin{cases}
-\emptyset
-& \mbox{if } \varphi(v) = P(\mathbf{x}_v) \\
-\left\lbrace
- \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
- b_v = \lambda_v b_0
-\right\rbrace
-& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0
-\end{cases}
-\end{align*}
-
-For simplicity, we assume variables $\mathbf{a}_u$, $b_u$ are obtained
-from $\Delta(u)$.
-
-Any model $\sigma$ to the constraint $C_{v_\bot}$ gives a simple
-solution for $G$ if $C_{v_\bot}$ is satisfiable, i.e., we should have:
-
-\begin{quote}
-If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
-\begin{align*}
- \rho = \left\lbrace
-  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \middle|
-  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
- \right\rbrace
-\end{align*}
-is a solution of $G$.
-\end{quote}
 
@@
 
-\subsection{Acyclic Horn problems}
-
-We eliminate the restriction that all Horn term vertices
-$V_\Rightarrow$ have exactly one successor.
-
-We define selectors for preceding disjunction vertices to $v$.
-\[ \mathcal{S}_v = \prod_{u \in \mathrm{succ}(v)} \left\lbrace u \mapsto t \middle| \edgel{t}{\theta}{u} \right\rbrace \]
-
-\begin{align*}
-K_l = \begin{cases}
-\left\lbrace l \right\rbrace & \mbox{if } l = \mathbf{a} \mathbf{x} \leq b \\
-\mathcal{X}_P \cup \left\lbrace \bot \right\rbrace
-& \mbox{if } l = P(\mathbf{x})
-\end{cases} \\
-\forall s \in \mathcal{S}_v;
-X = \prod_{t \in \mathrm{ran}(s)} K_{\varphi(t)}
-\exists c \in X;
-\end{align*}
-
-We define a \textbf{selector set collection} \cite{albarghouthi13} for
-every Horn term vertices, which is a set of sets of selectors.
-
-For $v \in V_T$, the selector set collection $\mathcal{X}_v =
-\left\lbrace S_{v,1}, \ldots, S_{v,\kappa_v} \right\rbrace$ consists
-of $\kappa_v$ sets of non-empty and non-overlapping subsets of
-$\mathcal{S}$. If $\nexists u; u \rightarrow v$, $\mathcal{X}_v =
-\emptyset$.
-
-We arbitrarily choose the initial selector $s_0 \in \mathcal{S}$. We
-start from empty collections for all $v$.
-
 \bibliographystyle{plain}

### 2013-08-22 article/algo.tex
Pseudo-code updated

@@
 \mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
-0 & \mbox{if } \varphi(v) = \top \mbox { or } \varphi(v) = \bot
+0 & \mbox{if } \varphi(v) = \bot
 \end{cases}
@@
 
-A variable mapping label on edges \theta is used to map the index of
+A variable mapping label $\theta$ on edges is used to map the index of
 one predicate variable to another. (TODO?)
@@
 & \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0
-\end{cases}
+\end{cases} \\
+(TODO???)
 \end{align*}
-
 For simplicity, we assume variables $\mathbf{a}_u$, $b_u$ are obtained
@@
 
-Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our algorithm.
+If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for $G$.
+We now try to give a less simpler solution by transforming $G$ and
+relaxing the constraint $C_{v_\bot}$.  This transformation takes place
+by considering the unsatisfiable core from the constraint.
+
+Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our Horn clause
+solving algorithm.  It takes an input Horn graph and returns a mapping
+between predicate variables and linear arithmetic formulas.
 
@@
 \newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
-\newcommand{\splitGraph}{\ensuremath{\mbox{\sc RestructureGraph}}}
+\newcommand{\transGraph}{\ensuremath{\mbox{\sc TransformGraph}}}
 
@@
 \STATE {$v_\bot \gets $ the root node of $G$}
-\STATE {$V \gets \left\lbrace v_\bot \right\rbrace$}
+\STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
 \REPEAT
-  \STATE {$C \gets \genConstr (v_\bot, V)$}
-  \STATE {$c_{v_\bot} \gets C(v_\bot)$}
-  \STATE {Evaluate satisfiability of $c_{v_\bot}$}
-  \IF {$c_{v_\bot}$ is unsatisfiable}
-    \STATE {$\mathcal{UC} \gets $ unsatisfiable core for $c_{v_\bot}$}
-    \STATE {$G \gets \splitGraph (G, \mathcal{UC})$}
-    \STATE {$V \gets \left\lbrace v_\bot \right\rbrace$}
+  \STATE {$\mathcal{C} \gets \genConstr (G, v_\bot, V_\star)$}
+  \STATE {$C_{v_\bot} \gets \mathcal{C} \left[ v_\bot \right]$}
+  \STATE {Evaluate satisfiability of $C_{v_\bot}$}
+  \IF {$C_{v_\bot}$ is unsatisfiable}
+    \STATE {$U \gets $ unsatisfiable core for $C_{v_\bot}$}
+    \STATE {$G \gets \transGraph (G, U)$}
+    \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
   \ELSE
-    \STATE {Enlarge $V$}
-    \STATE {$m \gets $ one model of $c_{v_\bot}$}
+    \STATE {$V_\star \gets V_\star \cup \left\lbrace u \in V_T \mid \forall v \in V_\star; u \leadsto v \right\rbrace$}
+m    \STATE {$m \gets $ one model of $C_{v_\bot}$}
   \ENDIF
-\UNTIL {$V$ contains all predicate variable vertices}
+\UNTIL {$V_\star$ contains all predicate variable vertices}
 \STATE {Build solution by $m$}
@@
 \begin{algorithm}
-\caption{$ \genConstr (v, visited) $}\label{alg:genConstr}
+\caption{$ \genConstr (G, v, visited) $}\label{alg:genConstr}
 \begin{algorithmic}
-\STATE {$m \gets \emptyset$}
-\FORALL {$v_\Rightarrow \in preds(v)$}
-  \FORALL {$v_p \in preds(v_\Rightarrow)$}
-    \STATE { $ l := \varphi (v_p) $ }
-    \IF {$l$ is a predicate variable}
-      \STATE {$m' \gets \genConstr (v_p) $}
-      \STATE {$m \gets m \cup m'$}
+\STATE {$\mathcal{C} \gets \emptyset$}
+\FORALL {$v_\Rightarrow \in preds_G(v)$}
+  \FORALL {$v_p \in preds_G(v_\Rightarrow)$}
+    \IF {$\varphi (v_p)$ is a predicate variable}
+      \STATE {$\mathcal{C} \gets \mathcal{C} \cup \genConstr (G, v_p, visited)$}
     \ELSE
@@
 
-\subsection{Disjunction-free problems}
-
-If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for
-$G$. We now try to give a solution by relaxing the constraint
-$C_{v_\bot}$.
-
-We consider below a Horn graph whose Horn term vertices has at
-most one successor.
-
-\subsubsection{Quantifier elimination}
-
-A set of eliminating quantifiers are obtained as $quant(n)$.
-
-\begin{align*}
-in(n) & = \bigcup_{p\in pred(n)} out(p) \\
-out(n) & = gen(n) \cup in(n) \\
-dup(n) & = \left\lbrace t \middle| p,q \in pred(n); p \ne q \wedge
- t \in out(p) \wedge t \in out(q) \right\rbrace \\
-quant(n) & = gen(n) \cup dup(n) \cup
- (out(n) \cap \bigcup_{s \in succ(n)} quant(s))
-\end{align*}
+\subsection{Quantifier elimination}
 
@@
 
-\begin{align*}
-\hat C_a = &
- \bigcup_{\edgel{u}{\theta}{a}} C_u \cup
- \bigcup_{0 < i \leq K}
- \left\lbrace
-  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{a}} \mathbf{a}_{u, \theta^{-1} (i)}
- \right\rbrace \\
- & \hspace{2cm} \cup
- \left\lbrace
-  b_v \geq \sum_{\edgel{u}{\theta}{a}} b_u
- \right\rbrace \cup
- \bigcup_{k_v < i \leq K}
- \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
-\\
-C_v = & \bigcup_{\edge{a}{v}} \hat C_a \cup
-\begin{cases}
-\emptyset
-& \mbox{if } \varphi(v) = P(\mathbf{x}_v) \\
-\left\lbrace
- \lambda_v \geq 0, \mathbf{a}_{v,\emptyset} = \lambda_v \mathbf{a}_0,
- b_{v,\emptyset} = \lambda_v b_0
-\right\rbrace
-& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0
-\end{cases}
-\end{align*}
-
----
-
-\begin{align*}
-C_v = & \bigcup_{(u,U,\theta,C_u) \in S_v^-} C_u \cup
-\bigcup_{(u,U,\theta,C_u) \in S_v^+} C_u
-\\
-& \cup
-\left\lbrace
- \mathbf{a}_{v,\emptyset,i} =
-  \sum_{(u,U,\theta,C_u) \in S_v^-} \mathbf{a}_{u,U,\theta^{-1}(i)} +
-  \sum_{(u,U,\theta,C_u) \in S_v^+} \mathbf{a}_{u,U, \theta^{-1}(i)}
-\right\rbrace \\
-& \cup \left\lbrace
- b_{v,\emptyset} \geq
-  \sum_{(u,U,\theta,C_u) \in S_v^-} b_{u,U} +
-  \sum_{(u,U,\theta,C_u) \in S_v^+} b_{u,U}
-\right\rbrace \\
-& \cup \bigcup_{k_v < i \leq K} \left\lbrace \mathbf{a}_{v,\emptyset,i} = 0 \right\rbrace
-\end{align*}
-such that
-\begin{align*}
-& S_v^- = \left\lbrace (u, U, \theta, C_u) \middle|
-  u \mathop{\leadsto}^\theta v \wedge
-  \left( \nexists u' \in V'; \xi (u') = u \right) \wedge
-  \Delta_{V'} (u) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u) \right\rbrace
-\\
-& S_v^+ = \left\lbrace (u, U, \theta, C_u) \middle|
-  u \mathop{\leadsto}^\theta v \wedge
-  \left( \exists u' \in V'; \xi (u') = u \wedge
-  \hat \Delta (u', \edge{u}{v}) = (\mathbf{a}_{u,U} \mathbf{x}_u \leq b_{u,U}, C_u) \right) \right\rbrace
-\\
-\end{align*}
-
 This intuitively means that constraints are duplicated according to
@@
 
-\subsubsection{Refinement from Unsatisfiable Core}
-
-We first start the process from the initial Copy tree:
-\begin{align*}
-V_C = & \left\lbrace v_{v_\bot} \right\rbrace \\
-E_C = & \emptyset
-\end{align*}
+\subsection{Refinement from Unsatisfiable Core}
 

### 2013-08-23 article/algo.tex
Delta function revived

@@
 We may omit $\theta$ and write $u \leadsto v$ for convenience in later
-discussions.
+discussions. By the assumption that $G$ is acyclic, $\leadsto^+$ is a
+well-founded relation. $v_\bot$ is the greatest element with respect
+to $\leadsto^+$.
 
-By the assumption that $G$ is acyclic, $\leadsto^+$ is a well-founded
-relation. $v_\bot$ is the greatest element with respect to
-$\leadsto^+$.
-
-For each predicate variable $v \in V_T$, the algorithm assigns a
-predicate template $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ and define
-$C_v$ which is a set of linear expressions on template parameters
-$\mathbf{a}_v$ and $b_v$.  Intuitively, it denotes the set of
-inequalities:
+For each Horn term vertex $v \in V_T$, we define
+$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v \leq b_v, C_v)$
+, where
+\begin{itemize}
+\item $\mathbf{a}_v \mathbf{x}_v \leq b_v$ is a linear inequality that
+  contains coefficient variables $\mathbf{a}_v$ and a constant $b_v$,
+  and
+\item $C_v$ is a set of linear expressions on coefficient variables.
+\end{itemize}
+Intuitively, it denotes the set of inequalities:
 \begin{align*}
@@
   \setminus (\mathbf{a}_v \cup \left\lbrace b_v \right\rbrace
- \right); C
+ \right); C_v
 \right\rbrace
 \end{align*}
+For a predicate variable vertex, $\mathbf{a}_v \mathbf{x}_v + b_v \leq
+0$ serves as a predicate template and $C_v$ describes the parameters
+$\mathbf{a}_v$ and $b_v$.
 
@@
 \end{cases} \\
-(TODO???)
 \end{align*}

### 2013-08-23 article/algo.tex
Causes for unsat

@@
 
+If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for $G$.
+There are two cases that a simple solution does not exist for a
+predicate variable.
+\begin{itemize}
+\item A predicate variable appears multiple times on left-hand of Horn
+  clauses, and those appearances require different expressions as an
+  assumption.
+\item A predicate variable is implied by a disjunctive Horn clause,
+  which has disjunctions on its left-hand, and a common premise that
+  the disjunctions induce is not strong enough for solving a problem.
+\end{itemize}
+
+The first case occurs in a case like follows.
+\begin{align*}
+% conj.txt ... originated from uc/01.txt
+\left( x \leq 0 \right) \wedge \left( y \leq 0 \right) & \implies P(x,y) \\
+P(x,a) \wedge P(b,y) & \implies x+y \leq 0
+\end{align*}
+In this example, the predicate variable $P$ appears twice on the
+second clause and they require different premises; $x \leq 0$ for the
+first and $y \leq 0$ for the second.  They cannot be represented by
+$\mathbf{ax} + b \leq 0$ template at the same time.
+
+The second case occurs in a case like follows.
+\begin{align*}
+% disj-simp.txt ... originated from 23simp_neg.txt
+\left( x+1 \leq 0 \right) \vee \left( -x+1 \leq 0 \right) & \implies P(x) \\
+P(x,y) \wedge \left( x \leq 0 \right) \wedge \left( -x \leq 0 \right) & \implies \bot
+\end{align*}
+In this example, the predicate variable $P$ is implied by the first
+clause, which is a disjunctive Horn clause.  As a simple solution, $P$
+can take a common consequence that both of $x+1 \leq 0$ and $-x+1 \leq
+0$ induce.  However, such a simple linear expression does not exist
+except for $0 \leq 0$, which is equivalent to $\top$, and it is not
+sufficient to satisfy the second clause.
+
+Both cases can be fixed by enlarging the solution template.
+
 \subsection {Graph transformation}
 
-If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for $G$.
-We now try to give a less simpler solution by transforming $G$ and
-relaxing the constraint $C_{v_\bot}$.  This transformation takes place
-by considering the unsatisfiable core from the constraint.
+If a simple solution does not exist for $G$, we try to give a less
+simpler solution by transforming $G$ and relaxing the constraint
+$C_{v_\bot}$.  This transformation takes place by considering the
+unsatisfiable core from the constraint.  Depending on the reason for
+the absence of a simple solution, the way of the transformation
+differs.
 
@@
 \REPEAT
-  \STATE {$\mathcal{C} \gets \genConstr (G, v_\bot, V_\star)$}
-  \STATE {$C_{v_\bot} \gets \mathcal{C} \left[ v_\bot \right]$}
-  \STATE {Evaluate satisfiability of $C_{v_\bot}$}
-  \IF {$C_{v_\bot}$ is unsatisfiable}
+  \STATE {$C_{v_\bot} \gets \genConstr (G, v_\bot, V_\star)$}
+  \IF {$C_{v_\bot}$ is satisfiable}
+    \STATE {$\sigma \gets $ one model of $C_{v_\bot}$}
+    \STATE {$V_\star \gets V_\star \cup \left\lbrace u \in V_T \mid \forall v \in V_\star; u \leadsto v \right\rbrace$}
+  \ELSE
     \STATE {$U \gets $ unsatisfiable core for $C_{v_\bot}$}
@@
     \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-  \ELSE
-    \STATE {$V_\star \gets V_\star \cup \left\lbrace u \in V_T \mid \forall v \in V_\star; u \leadsto v \right\rbrace$}
-m    \STATE {$m \gets $ one model of $C_{v_\bot}$}
   \ENDIF
-\UNTIL {$V_\star$ contains all predicate variable vertices}
-\STATE {Build solution by $m$}
+\UNTIL {$V_\star = V_T$}
+\STATE {Build solution by $\sigma$}
 \end{algorithmic}
@@
 \begin{algorithmic}
-\STATE {$\mathcal{C} \gets \emptyset$}
+\STATE {$C \gets \emptyset$}
 \FORALL {$v_\Rightarrow \in preds_G(v)$}
@@
     \IF {$\varphi (v_p)$ is a predicate variable}
-      \STATE {$\mathcal{C} \gets \mathcal{C} \cup \genConstr (G, v_p, visited)$}
+      \STATE {$C \gets C \cup \genConstr (G, v_p, visited)$}
     \ELSE
@@
 to generate a linear constraint in interpolation.
-
-\subsection{Quantifier elimination}
-
-Let $\hat V_T$ a set of Horn term vertices $V_T$ with multiple
-successors.
-\[ \hat V_T = \left\lbrace v \in V_T \middle|
-\exists u, u', \theta, \theta';
-\edgel{v}{\theta}{u} \wedge \edgel{v}{\theta'}{u'} \wedge
-(u \ne u' \vee \theta \ne \theta') \right\rbrace \]
-We define \textbf{Copy tree}, which is a tree $T=(V_C,E_C,\xi)$
-rooted at $v_{v_\bot}$.
-\begin{itemize}
-\item $V_C$ is a set of nodes.
-\item $E_C$ is a set of labelled edges, which are expressed in the
-  form $(u,v,e)$, where $e \in E \cup {\bot}$.
-\item $\xi: V_C \rightarrow \hat V_T \cup \left\lbrace v_\bot
-  \right\rbrace$ is a labelling function for nodes. The root node
-  $v_{v_\bot} \in V_C$ has a label $v_\bot$.
-\end{itemize}
-
-We provide $\textsc{Copy}_v$ procedure to rename all free variables
-for avoiding name collision.
-\[ \textsc{Copy}_v (x_V) = x_{V \cup \left\lbrace v \right\rbrace} \]
-All variables with the same name are then distinguished by the label
-of copy origin nodes $V$. For consistency, the variables without the
-label should be interpreted as that it has $\emptyset$ for origin
-nodes.
-
-We give a constraint function over a copy tree
-$\hat \Delta(v,e)$ for $v \in V_C, e \in E$ as
-\[ \hat \Delta (v,e) = \textsc{Copy}_v (\Delta_{V'} (\xi (v)))
-\text{ such that } V' = \left\lbrace v' \middle| (v, v',e) \in E_C \right\rbrace \]
-extending $\Delta$ function to
-$\Delta_{V'} (v) = (\mathbf{a}_{v,\emptyset} \mathbf{x}_v \leq b_{v,\emptyset}, C_v)$
-as
-
-This intuitively means that constraints are duplicated according to
-the shape of copy tree $T$.  Given the constraint for the root node in
-the copy tree as
-\[ \hat \Delta (v_{v_\bot}, \bot) = (\mathbf{a}_0 \mathbf{x}_{v_\bot} \leq b_0, C_{v_\bot}) \]
-a model to $C_{v_\bot}$ gives a solution $\rho$ to $G$, if one exists.
-
-\begin{align*}
- \rho = \left\lbrace
-  \left( P, \bigwedge_U \sigma(\mathbf{a}_{v,U} \mathbf{x}_v \leq b_{v,U}) \right) \middle|
-  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
- \right\rbrace
-\end{align*}
-
-If $C_{v_\bot}$ stays unsatisfiable, we let $T$ grow based on the
-unsatisfiable core for $C_{v_\bot}$.
 

### 2013-08-24 article/algo.tex
Split explanation

@@
 
-Algorithm~\ref{alg:solveHorn} shows the pseudo-code of our Horn clause
-solving algorithm.  It takes an input Horn graph and returns a mapping
-between predicate variables and linear arithmetic formulas.
+Algorithm~\ref{alg:solveHorn} shows the updated pseudo-code of our
+Horn clause solving algorithm with the graph transformation.  It takes
+an input Horn graph and returns a mapping between predicate variables
+and linear arithmetic formulas.
 
@@
 \UNTIL {$V_\star = V_T$}
+% TODO: Consider predicate templates?
 \STATE {Build solution by $\sigma$}
@@
 \end{algorithm}
+\transGraph procedure determines the unsatisfiable core of the
+constraint and transform the Horn graph $G$ and predicate templates.
+As mentioned earlier, two ways of graph transformation exist.
+
+\paragraph{Conjunction split}
+
+\paragraph{Disjunction split}
 

### 2013-08-24 article/algo.tex
TeX

@@
-\documentclass[a4paper,12pt]{article}
+\documentclass[master,final,11pt]{iscs-thesis}
 
@@
 
-\title{Algorithm explanation}
-\author{Yuto Takei \\ The University of Tokyo}
+\etitle{Algorithm explanation}
+\eauthor{Yuto Takei}
+\jtitle{} \jauthor{}
+\esupervisor{}
+\jsupervisor{}
+\supervisortitle{Professor}
+\date{August 28, 2013}
 
 \begin{document}
+
+\begin{eabstract}
+\end{eabstract}
+\begin{jabstract}
+\end{jabstract}
 
@@
 The algorithm assigns a \emph{predicate template} for all predicate
-variables in $G$. A template for a predicate variable $P(x_1, \ldots, x_n)$
-is a formula $\psi$ constructed from the following definition.
+variables in $G$. A \emph{predicate templateset} is a mapping from
+predicate variables to a predicate template.  A template for a
+predicate variable $P(x_1, \ldots, x_n)$ is a formula $\psi$
+constructed from the following definition.
 \begin{align*}
@@
 \left\lbrace
- \mathbf{a}_v \mathbf{x}_v + b_v \leq 0 \middle|
+ \mathbf{a}_v \mathbf{x}_v + b_v \leq 0 \mid
  \exists \left( \textsc{FV}(C_v)
@@
  \rho = \left\lbrace
-  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \middle|
+  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \mid
   \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
@@
 $C_{v_\bot}$.  This transformation takes place by considering the
-unsatisfiable core from the constraint.  Depending on the reason for
-the absence of a simple solution, the way of the transformation
-differs.
+unsatisfiable core $\mathcal{U}$.  An unsatisfiable core $\mathcal{U}$
+is a subset of $C_{v_\bot}$ which is inconsistent itself.  Depending
+on the reason for the absence of a simple solution, the way of the
+transformation differs.
 
@@
 \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
+\STATE {$T \gets $ initial predicate templateset for $G$}
 \REPEAT
@@
   \ELSE
-    \STATE {$U \gets $ unsatisfiable core for $C_{v_\bot}$}
-    \STATE {$G \gets \transGraph (G, U)$}
+    \STATE {$\mathcal{U} \gets $ unsatisfiable core for $C_{v_\bot}$}
+    \STATE {$(G, T) \gets \transGraph (G, T, \mathcal{U})$}
     \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
@@
 \UNTIL {$V_\star = V_T$}
-% TODO: Consider predicate templates?
-\STATE {Build solution by $\sigma$}
+\RETURN {$\sigma \left( T \right)$}
 \end{algorithmic}
 \end{algorithm}
-\transGraph procedure determines the unsatisfiable core of the
+\transGraph~procedure determines the unsatisfiable core of the
 constraint and transform the Horn graph $G$ and predicate templates.
@@
 \paragraph{Conjunction split}
+For the first case previously mentioned, the unsatisfiable core
+$\mathcal{U}$ contains coeffcient constraint expressions with a
+coefficient parameter of certain predicate variable.  If the predicate
+variable $P$ is the cause of unsatisfiability, some constraint
+expressions in the unsatisfiable core have a parameter $a_{P,i}$
+and/or $b_P$ in the following way.
+\begin{align*}
+& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
+& \left( \exists i, j, k;
+\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
+\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
+\right\rbrace \subseteq \mathcal{U} \right)
+\end{align*}
+
+In such a case, the predicate variable $P$ in question is copied to
+$P_1, \ldots, P_n$ so that the different appearances of $P$ on
+left-hand of Horn clauses can take different premises.  The solution
+template of $P$ becomes the conjunction of $P_1, \ldots, P_n$.
+
+<fig...>
+
+Here, the graph $G=(V,E,\varphi)$ will be transformed into
+$G'=(V',E',\varphi')$ such that:
+\begin{align*}
+V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
+  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
+E' & = E \\
+\varphi' & = \varphi
+\end{align*}
+
+Note that after this duplication of the predicate vertex $P$, in order
+to induce different linear expressions for $P_1, \ldots, P_n$, it may
+be necessary for preceeding predicate variable vertices in $G$ to be
+duplicated in the same manner as $P$. It is because that different
+derivation of $P$'s implication relation may be required to get
+different consequences.
+
+However, succeeding predicate vertices of $P$ in $G$ and other
+vertices than predecessors are not affected because those implications
+that have $P$ in their assumption can selectively choose the premise
+that $P_1, \ldots, P_n$ leads.
+
+Therefore, this graph transformation only affects the predecessors of
+$P$.
 
 \paragraph{Disjunction split}
+For the second case, the unsatisfiable core $\mathcal{U}$ contains
+coefficient constraints that unifies induced linear expressions from
+each term in disjunctions to be the same expression.  Similarly as
+previous, if the predicate variable $P$ is the cause of
+unsatisfiability, some constraint expressions in the unsatisfiable
+core have a parameter $a_{P,i}$ and/or $b_P$ in the following way.
+\begin{align*}
+& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
+& \left( \exists i, j, k;
+\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
+\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
+\right\rbrace \subseteq \mathcal{U} \right)
+\end{align*}
 
-The constraints are first built the weakest so that the disjunctiveness of
-whole solution template is fixed first.  The strength matters that the
+In such a case, the predicate variable $P$ in question is separated to
+$P_1, \ldots, P_n$ each of which is implied by a term, which was
+originally a part of a disjunction to imply $P$.  This intuitively
+enables us to avoid generating a common linear expression that
+disjunctive terms induces, and to propagate each term's premises to
+$P$'s successors instead.
+
+<fig...>
+
+Here, the graph $G=(V,E,\varphi)$ will be transformed into
+$G'=(V',E',\varphi')$ such that:
+\begin{align*}
+V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
+  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
+E' & = E \\
+\varphi' & = \varphi
+\end{align*}
+
+After this duplication of the predicate vertex $P$, unlike the
+conjunction split, this results in producing a disjunction at
+succeeding predicate variables vertices ($R$ in this case).
+\begin{align*}
+(P_1 \wedge Q) \vee (P_2 \wedge Q) \implies R
+\end{align*}
+Also, predicate variables that appeared with $P$ in the same
+assumption ($Q$ in this case) needs to be duplicated by conjunction,
+so as to obtain different premises that go along with $P_i$.
+Therefore, the effect of disjunction split may reach to neighboring
+predicate vertices.
+
+\paragraph{Combinatorial explosion in disjunctions}
+Virtually, if multiple disjunctions exists in an input set of Horn
+clauses, a na\"{i}ve way of building a solution is solve all
+combinations of choices out of every disjunction and concatenate each
+expression with conjunctions and/or disjunctions appropriately.  This
+may cause a combinatorial explosion.
+
+Conside example below. $e_i$ represents some conjunctions of linear
+expressions.
+% TODO: Can't we replace e_i with actual expressions?
+\begin{align*}
+(e_1 \vee e_2) \implies P \\ (e_3 \vee e_4) \implies Q \\ P \wedge Q \implies \bot
+\end{align*}
+In this example, a na\"{i}ve algorithm may solve following four
+problems separately.
+\begin{itemize}
+\item $e_1 \implies P_1 \qquad e_3 \implies Q_1 \qquad P_1 \wedge Q_1 \implies \bot$
+\item $e_2 \implies P_2 \qquad e_3 \implies Q_2 \qquad P_2 \wedge Q_2 \implies \bot$
+\item $e_1 \implies P_3 \qquad e_4 \implies Q_3 \qquad P_3 \wedge Q_3 \implies \bot$
+\item $e_2 \implies P_4 \qquad e_4 \implies Q_4 \qquad P_4 \wedge Q_4 \implies \bot$
+\end{itemize}
+Then the solution is built by follows.
+\begin{itemize}
+\item $P = ( P_1 \vee P_2 ) \wedge ( P_3 \vee P_4 )$
+\item $Q = ( P_1 \vee P_3 ) \wedge ( P_2 \vee P_4 )$
+\end{itemize}
+However, this method is not efficient.
+
+\paragraph{Algorithm}
+To avoid this, our algorithm tries to unify a disjunctions of solution
+and determine the necessity of split in conjunctions.
+
+The constraints are first built the weakest so that all conjunctions
+are not yet unified.
+
+The strength matters that the
 predicate variables are used as the same expressions depending on
@@
 
-\subsection{Refinement from Unsatisfiable Core}
-
-An unsatisfiable core $\mathcal{U}$ is a subset of $C_{v_\bot}$.
-If a vertex $v \in V$ with the copy origin $U$ satisfies:
-\begin{align*}
-& \exists t, u \in V; v \leadsto t \wedge v \leadsto u \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{s,U,j} = \ldots + \mathbf{a}_{v,U,i} + \ldots \right),
-\left( \mathbf{a}_{t,U,k} = \ldots + \mathbf{a}_{v,U,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-we add $v$ to the copy tree $T$.
-
 \bibliographystyle{plain}

### 2013-08-25 article/algo.tex
Explanation of tree transformation
Changing the definition of Delta function

@@
-\documentclass[master,final,11pt]{iscs-thesis}
+\documentclass[a4paper,12pt]{article}
 
@@
 
-\etitle{Algorithm explanation}
-\eauthor{Yuto Takei}
-\jtitle{} \jauthor{}
-\esupervisor{}
-\jsupervisor{}
-\supervisortitle{Professor}
-\date{August 28, 2013}
+\title{Algorithm explanation}
+\author{Yuto Takei \\ The University of Tokyo}
 
 \begin{document}
-
-\begin{eabstract}
-\end{eabstract}
-\begin{jabstract}
-\end{jabstract}
 
@@
 \end{itemize}
+
+We define the relation $\leadsto$ over $V_T$ as follows.
+\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow \exists a \in V_\Rightarrow;
+\edgel{u}{\theta}{a} \wedge \edge{a}{v} \] We may omit $\theta$ and
+write $u \leadsto v$ for convenience in later discussions. By the
+assumption that $G$ is acyclic, both $\rightarrow^+$ and $\leadsto^+$ are
+well-founded relations. $v_\bot$ is the greatest element with respect
+to both $\rightarrow^+$ and $\leadsto^+$.
+
 The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
@@
 The algorithm assigns a \emph{predicate template} for all predicate
-variables in $G$. A \emph{predicate templateset} is a mapping from
-predicate variables to a predicate template.  A template for a
+variables in $G$. A \emph{predicate templateset} is a set of mappings from
+a predicate variable to a predicate template.  A template for a
 predicate variable $P(x_1, \ldots, x_n)$ is a formula $\psi$
@@
 
-We define the relation $\leadsto$ over $V_T$ as follows.
-\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow
-\exists a \in V_\Rightarrow; \edgel{v}{\theta}{a} \wedge \edge{a}{v} \]
-We may omit $\theta$ and write $u \leadsto v$ for convenience in later
-discussions. By the assumption that $G$ is acyclic, $\leadsto^+$ is a
-well-founded relation. $v_\bot$ is the greatest element with respect
-to $\leadsto^+$.
-
-For each Horn term vertex $v \in V_T$, we define
-$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v \leq b_v, C_v)$
+For each vertex $v \in V$, we define
+$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v + b_v \leq 0, C_v)$
 , where
 \begin{itemize}
-\item $\mathbf{a}_v \mathbf{x}_v \leq b_v$ is a linear inequality that
+\item $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ is a linear inequality that
   contains coefficient variables $\mathbf{a}_v$ and a constant $b_v$,
@@
 $C_v$ is defined by well-founded induction on $v$ with respect to the
-relation $\leadsto^+$. Then, $C_v$ can be shown as follows.
-
+relation $\rightarrow^+$. Then, $C_v$ can be shown as follows.
+\begin{itemize}
+\item For $v \in V_\Rightarrow$;
 \begin{align*}
-\hat C_{a,v} = &
- \bigcup_{\edgel{u}{\theta}{a}} C_u \cup
+C_v =
+ \bigcup_{\edgel{u}{\theta}{v}} C_u \cup
  \bigcup_{0 < i \leq K}
  \left\lbrace
-  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{a}} \mathbf{a}_{u, \theta^{-1} (i)}
+  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{v}} \mathbf{a}_{u, \theta^{-1} (i)}
  \right\rbrace \\
@@
  \left\lbrace
-  b_v \geq \sum_{\edgel{u}{\theta}{a}} b_u
+  b_v \leq \sum_{\edgel{u}{\theta}{v}} b_u
  \right\rbrace \cup
- \bigcup_{k_v < i \leq K}
+ \bigcup_{k_{succ(v)} < i \leq K}
  \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
-\\
-C_v = & \bigcup_{\edge{a}{v}} \hat C_{a,v} \cup
-\begin{cases}
-\emptyset
-& \mbox{if } \varphi(v) = P(\mathbf{x}_v) \\
+\end{align*}
+\item For $v \in V_T$ such that
+  $v \neq v_{\bot} \wedge \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0$
+  where $\mathbf{a}_0$ and $b_0$ are constants;
+\begin{align*}
+C_v =
 \left\lbrace
@@
 \right\rbrace
-& \mbox{if } \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0
-\end{cases} \\
 \end{align*}
-For simplicity, we assume variables $\mathbf{a}_u$, $b_u$ are obtained
-from $C_u$.
+\item For $v \in V_T$ such that $\varphi(v) = P(\mathbf{x}_v)$;
+\begin{align*}
+C_v = & \bigcup_{\edge{u}{v}} \left( C_{u} \cup
+\left\lbrace \mathbf{a}_u = \mathbf{a}_v, b_u = b_v \right\rbrace
+\right)
+\end{align*}
+\item For $v_\bot$;
+\begin{align*}
+C_{v_\bot} = \bigcup_{\edge{u}{v_\bot}} \left( C_{u} \cup
+\left\lbrace \mathbf{a}_u = \mathbf{0}, b_u > 0 \right\rbrace \right)
+\end{align*}
+\end{itemize}
 

### 2013-08-25 article/algo.tex
Explanation for gen_constr

@@
 \begin{align*}
-C_v =
+C_v & =
  \bigcup_{\edgel{u}{\theta}{v}} C_u \cup
@@
 \begin{align*}
+% TODO:
 & \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
@@
 
+Consider these two ways of transformation, we show a pseudo-code for
+\transGraph~procedure.  The procedure is given the graph $G$, the
+predicate templateset $T$, and the unsatisfiable core
+$\mathcal{U}$. It returns the transformed graph $G'$ and the new
+templateset $T'$.
+
+\begin{algorithm}
+\caption{$ \transGraph (G, T, \mathcal{U}) $}\label{alg:transGraph}
+\begin{algorithmic}
+% TODO:
+\RETURN {$\left( G', T' \right)$}
+\end{algorithmic}
+\end{algorithm}
+
 \paragraph{Combinatorial explosion in disjunctions}
@@
 \paragraph{Algorithm}
-To avoid this, our algorithm tries to unify a disjunctions of solution
-and determine the necessity of split in conjunctions.
+To avoid this, our algorithm lets a predicate variable $P$ which is
+disjunctively implied on the Horn graph $G$ to have a linear
+expression that is consequences of all terms in the disjunction.  This
+prevents the possibility of choices for disjunctions to be propagated
+to succeeding vertices. The simple linear expression is propagated
+instead.
 
-The constraints are first built the weakest so that all conjunctions
-are not yet unified.
+On the other hand, the algorithm allows a predicate variable $P$ that
+appears on multiple implications or appears multiple times in a single
+implication to take different linear expressions as a premise at
+different appearances.  This is for the algorithm to determine the
+cause of unsatisfiability efficiently.  It is done by copying a
+predicate templates and its parameters' constraint with appropriate
+renaming.
 
-The strength matters that the
-predicate variables are used as the same expressions depending on
-different appearances of the predicate variable.
+The algorithm first execute this duplication on all predicate
+variables with multiple appearances.  It is the weakest constraint
+that the algorithm makes for the certain $G$.  Then, the algorithm
+gradually reduces the number of vertices that are duplicated by each
+time testing the satisfiability of the constraint.  When the
+constraint bebcome unsatisfiable at the certain iteration, the
+algorithm can determine a predicate vertex which has to be split in a
+conjunctive template.  If all the predicate variable constraints are
+no longer duplicated and the constraint is the strongest, any model
+for such constraint gives a simple solution under the Horn graph $G$
+and the templateset $T$.
 
-The strngeth of the constraint is controlled by a set of vertices $V$
-in the pseudo-code.  Predicate variable vertices in $V$ are treated
-strictly unique across the different appearances of the predicate
-variables in Horn clauses.
+Note that duplicating the constraint in a na\"{i}ve way may cause the
+constraint to exponentially grow in size.  Our algorithm prevent it by
+simplifying a predicate vertex's constraint $C_v$ by quantifier
+elimination when $v$ is duplicated in generating linear constraint.
+For a predicate variable vertex $v_P$, variables except for the
+coefficient parameter $\mathbf{a}_P$ and the constant parameter $b_P$
+are eliminated.  Due to the constraints generation process, the
+simplified constraint is also linear and can be efficiently check the
+satisfiability.
 
-$\genConstr$ procedure produces the constraint for a given vertex $v$.
+We finally show the modified version of pseudo-code for $\genConstr$
+procedure.  It produces the constraint $C_v$ for a given vertex $v$,
+considering constraint duplication and simplification.
 
 \begin{algorithm}
-\caption{$ \genConstr (G, v, visited) $}\label{alg:genConstr}
+\caption{$ \genConstr (G, v, V_\star) $}\label{alg:genConstr}
 \begin{algorithmic}

### 2013-08-25 article/algo.tex
Small fixes

@@
   satisfiability.  The strongest constraint gives a simple solution.
-\item \textbf{Problem Restructuring} If the constraint become
-  unsatisfiable during the iteration, the process restructure the
-  problem equivalent to obtain a less simpler solution.  This process
-  considers the unsatisfiable core of the constraint.
+\item \textbf{Problem Restructuring} is a process to restructure the
+  problem to obtain a less simpler solution, if the constraint become
+  unsatisfiable during the iterative constraint generation.  This
+  process considers the unsatisfiable core of the constraint, and
+  wisely enlarge the solution size.
 \end{itemize}
@@
 The novelty of our algorithm lies in the incremental constraint
-refinement.  The algorithm try from the weakest constraint that an
-input problem allows that does not give a simple solution, and
-strengthen the constraint until it reaches to the simplest solution
-that the problem has.  When a constraint becomes unsatisfiable, the
-algorithm analyzes the unsatisfiable core and transform the problem to
-obtain less simple solution.
+refinement.  The algorithm tries to solve from the weakest constraint
+that an input problem allows that does not give a simple solution, and
+gradually strengthen the constraint until it reaches to the simplest
+solution that the problem has.  When a constraint becomes
+unsatisfiable during this step, the algorithm analyzes the
+unsatisfiable core and transform the problem to obtain a less simple
+solution.
 
 This construction of the algorithm enables us to compute a relatively
-simpler solution compared to existing algorithms in a reasonable time.
+simpler solution compared to existing algorithms while preserving the
+computation cost in a reasonable time.
 

### 2013-08-26 article/algo.tex
Integrate Markdown draft into TeX document

@@
 \begin{document}
+
+\section{Introduction}
+
+Automated theorem proving plays an important role in the modern
+program verification.  Particularly in the context of verification
+with model checking techniques, an automated theorem prover is used to
+abstract an infinite set of program states into a finite one so that
+the model checker can virtually explore all possible program execution
+paths.
+
+One of the most common methods of such abstraction is predicate
+abstraction.  An automated theorem prover finds appropriate predicates
+for specific program locations.  An execution state at each location
+is then expressed by a Boolean valuation of the given predicates.  If
+the abstraction is too coarse and a model checker discovers an
+infeasible error execution path in an abstract system, the prover
+computes an additional predicate at a specific program location to
+refine the abstraction.  The additional predicate must be satisfied at
+the location under the execution in the concrete model along the
+discovered path, and the execution does not lead to the program
+failure by the same path.  After the refinement, the infeasible path
+is ruled out and no longer found by the model checker.
+
+In order to obtain such a refinement predicate at the specific program
+location along the infeasible counterexample, one class of automated
+theorem provers, called an interpolating theorem prover, computes a
+Craig interpolant between two sets of constraints; the one consists of
+the constraints from the program entry point to the location, and the
+other consists of the constraints from the location to the program
+failure.  Those constraints include variable assignments and assertions
+to be satisfied originated from conditional branches.
+
+However, existing interpolating theorem provers may return too complex
+a solution which is heavily affected by the constraints of the
+specific path.  This causes the model checker to discover infinitely
+similar paths which pass the same program location by loops or
+recursion calls, and not to terminate.
+
+To avoid such situation, it is desired to use an invariant formula for
+the predicate abstraction.  Based on the hypothesis that program’s
+invariants tend to be simple formulas, we propose a new interpolating
+algorithm which tries to minimize the number of conjunctions and
+disjunctions in a solution.  The predicate obtained by this algorithm
+may become similar to the invariant, and the number of trials in model
+checking may be reduced.
+
+Additionally, we extend our interpolating algorithm to solve a
+symmetric interpolation problem, in which the algorithm computes
+predicates at multiple program locations along a path.  This enables
+the abstraction predicates at multiple locations to be updated at the
+same computation.
+
+Finally, we extend our algorithm to solve Horn clause solving
+problems.  It allows us to obtain predicates which are highly likely
+to be program invariants, by computing the same abstraction predicates
+at loops and recursion calls on different passes, and by unifying
+predicates at the same location over multiple paths.  The problems are
+generated by encoding constraints over execution paths with unknown
+predicates which describe functions' constraints between parameters
+and return values.
+
+The rest of the paper is structured as follows.  Chapter 2 proposes a
+new interpolating algorithm to obtain simple solutions, and extends it
+to solve symmetric interpolation problems.  Chapter 3 describes a new
+Horn clause solving algorithm by extending the previous interpolating
+algorithm.  Chapter 4 shows an experiment and its result to confirm
+the effect.  Chapter 5 mentions related work and future work.  As a
+conclusion, chapter 6 reviews the impact of our research.
+
+\section{Interpolation}
+
+Among various program verification techniques, when software model
+checking with predicate abstraction is adopted, the interpolation is
+used to compute the abstraction predicate along a spurious
+counterexample that a model checker discovered.  It is done by
+computing a Craig's interpolant between the strongest postcondition at
+the location and the weakest precondition to the failure point at
+every program location.
+
+However, by the existing interpolating theorem prover, we may not be
+able to obtain a simple solution.  Consider the example below.
+\begin{align*}
+\varphi_1 : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
+\varphi_2 : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
+\end{align*}
+Although one may be able to obtain $x = y$ as the simplest
+interpolation between $\varphi_1$ and $\varphi_2$, CSIsat\cite{csisat}
+returns
+\[ TODO \]
+as a solution.
+
+We propose an interpolation algorithm which returns as simple a
+solution as possible in terms of the number of conjunctions and
+disjunctions.  Our algorithm focuses on interpolating problems on
+linear arithmetic and it works in a constructive manner.
+
+\subsection{Example}
+
+Consider the following interpolation problem.
+
+\begin{align*}
+\frac
+{A: (x \leq a) \wedge (a + 1 \leq y)}
+{B: (y \leq b) \wedge (b + 1 \leq x)}
+\end{align*}
+
+Here, the linear arithmetic formulas A and B are inconsistent.  The
+interpolant I for this problem is $x-y+1 \leq 0$, which is implied by
+$A$ and inconsistent with $B$.  The vocabulary of $I$ is $\left\lbrace
+x,y \right\rbrace$ and is over $A$'s vocabulary $\left\lbrace x,y,a
+\right\rbrace$ and $B$'s $\left\lbrace x,y,b \right\rbrace$.
+
+In computing the interpolant, we make a linear constraint of a
+interpolant by applying Farkas's lemma.  First, we assing weight
+parameters to every expression in conjunction form.
+
+\begin{align*}
+\begin{array}{r r r r r r l}
+\lambda_1 : & -a & & +x & & & \leq 0 \\
+\lambda_2 : & a & & & -y & +1 & \leq 0 \\
+\hline
+\lambda_3 : & & -b & & +y & & \leq 0 \\
+\lambda_4 : & & b & -x & & +1 & \leq 0
+\end{array}
+\end{align*}
+
+Because all expressions induce inconsistent as a whole, the sum of
+weighted expressions should satisfy the following constraint.
+
+\begin{align*}
+\begin{array}{r l l}
+- \lambda_1 + \lambda_2 & = 0 \quad & (a) \\
+- \lambda_3 + \lambda_4 & = 0 & (b) \\
+  \lambda_1 - \lambda_4 & = 0 & (x) \\
+- \lambda_2 + \lambda_3 & = 0 & (y) \\
+  \lambda_2 + \lambda_4 & > 0 & (\text{constant}) \\
+\lambda_1, \ldots, \lambda_4 & \geq 0 & (\text{non-negative weight})
+\end{array}
+\end{align*}
+
+With the linear constraint among $\lambda_i$ above, the interpolant is
+represented as follows by the sum of weighted expressions from the
+first formula groups.
+\[ (- \lambda_1 + \lambda_2) a + \lambda_1 x - \lambda_2 y + \lambda_2 \leq 0 \]
+One of the model of the linear constraint is $\lambda_1 = \lambda_2 =
+\lambda_3 = \lambda_4 = 1$, and we obtain $x-y+1 \leq 0$ as a
+solution.
+
+% TODO: Better to mention multiple interpolants possible for one
+% interpolating problem?
+
+\subsection{Preliminaries}
+
+Our algorithm aims to compute a Craig's interpolant over linear
+arithmetic formulas.
+
+\paragraph{Craig's Interpolation}
+Given two logical formulas $A$ and $B$ that are inconsistent each
+other, we call a new preposition $I$ a Craig's interpolant between $A$
+and $B$ such that $I$ is implied by $A$ and inconsistent with $B$.
+$I$'s vocabulary must be only free variables that appears in both $A$
+and $B$.
+
+Our algorithm handles a linear expression $e$ which is in a form:
+\[ a_1 x_1 + \cdots + a_n x_n + b \leq 0 \]
+This expression contains coefficient parameters $a_1, \ldots, a_n$ and
+a constant parameter $b$.  We treat a special expressions $\top$ as $0
+\leq 0$ and $\bot$ as $1 \leq 0$.  The algorithm input and output is a
+formula $\psi$ that consists of conjunctions and/or disjunctions of
+linear expressions.
+\[ \psi = e | e \wedge e | e \vee e | \top | \bot \]
+
+In the simple interpolation, the algorithm receives two formulas in
+$\psi$ and returns an interpolant in $\psi$.  For the later extension
+to handle a symmetric interpolation problem, the algorithm receives
+multiple formulas in $\psi$ and returns intermediate interpolants in
+$\psi$.  For computing those interpolants over linear arithmetics, we
+use Farkas's Lemma.
+
+\paragraph{Farkas's Lemma on linear inequalities}
+Let a linear inequality $e_i$ be represented as $a_i1 x_1 + \cdots +
+a_im x_m <= a_i0$.  Assuming that $e_1,\cdots,e_n$ implies $e_0$,
+there exists $\lambda_1,\cdots,\lambda_n$ that satisfy $a_0j
+=\sum_(i=1)^n \lambda_i * a_ij (j=0, \ldots, m)$.
+
+For the special case of Farkas's Lemma, we can state that:
+
+Assuming that $e_1,\cdots,e_n$ implies $\bot$, there exists
+$\lambda_1,\cdots,\lambda_n$ that satisfy $???? TODO $.
+
+Our interpolating algorithm make use of the latter version of Farkas's
+Lemma.
+
+\subsection{Algorithm}
+
+Our algorithm aims to construct a relatively small interpolant in a
+reasonable computation time for any input.  It also aims to obtain a
+common solution among multiple interpolating problems.  When the
+algorithm is used as a sub-routine for solving Horn clause satisfying
+problems that are made while discovering predicates for abstraction
+during the program verification.  These advantages of our algorithm
+make it possible to return a relatively small predicates for those
+problems. To accomplish these aims, the algorithm preserves a set of
+interpolants during the computation, and define and execute operations
+over interpolant sets.
+
+\paragraph{Interpolation between conjunctions}
+We first present an interpolating algorithm between two conjunctive
+sets of linear inequalities $A = \left\lbrace e_1,\ldots,e_m
+\right\rbrace$ and $B = \left\lbrace e_{m+1},\ldots,e_{m+n}
+\right\rbrace$.  We assume that conjunctions of $A$ and $B$ are
+inconsistent.  For later convenience, we call this interpolating
+problem $\left( A, B \right)$.  Then there exists an interpolating
+linear expression $e_\star$ which satisfies
+\begin{align*}
+\left\lbrace e_1,\ldots,e_m \right\rbrace & \vdash e_\star \\
+e_\star & \nvdash \left\lbrace e_{m+1},\ldots,e_{m+n} \right\rbrace
+\end{align*}
+For convenience, we assume that any expression $e_i$ have variables
+$x_1, \ldots, x_a$.
+
+According to the Farkas's lemma, there exists a set of assignmentqs to
+$\lambda_i$, to conclude $\bot$ from a set of linear expressions $A
+\cup B$, namely,
+\begin{align*}
+\exists \lambda_1, \ldots, \lambda_m. \\
+a % TODO: coefficient, constant constraints
+\end{align*}
+
+Then, a linear combination of $A$'s formula with a weight $\lambda_1,
+\ldots, \lambda_m$ becomes an interpolant $I$ between $A$ and $B$.
+Note that any assignments to $\lambda_i$ that satisfy the expressions
+above will let an expression $I$ to be an interpolant.  Considering
+this, it is possible to obtain different interpolants for a
+interpolating problem by computing different assignments to
+$\lambda_i$.
+
+This enables the algorithm to preserve a set of interpolants by a
+linear expression template $a_1 x_1 + \cdots + a_n x_n + b \leq 0$
+with a constraint to describe a possible space of coefficient
+parameters $a_i$ and the constant parameter $b$.
+
+In practical, we may use linear programming solvers to obtain a model
+for $\lambda_i$.  By assigning concrete values to them, we can obtain
+a concrete interpolant.
+
+\paragraph{Computing a common interpolant}
+Next we consider a method to compute a common interpolant across two
+different interpolating problems.  That is, when given two
+interpolating problems $\left(A_1, B_1 \right)$ and $\left(A_2, B_2
+\right)$, the problem is about to compute the common interpolant I,
+which is an interpolant for the first problem and the second one at
+the same time.
+
+This problem is used when solving a following Horn clause solving
+problem, for instance.
+\[
+\implies
+% TODO
+\]
+
+To accomplish this, we simply need to compute the intersection of
+interpolant sets.  Let us call the interpolant space for the problems
+$\left(A_1, B_1 \right)$ and $\left(A_2, B_2 \right)$,
+$\mathcal{I}_1$, and $\mathcal{I}_2$.  In the previous method, a set
+of interpolants was represented by a parameterized linear expression.
+\begin{align*}
+\mathcal{I}_1 = \left\lbrace \mathbf{a}_1 \mathbf{x} \leq b_1 \mid
+C_1 (\mathbf{a}_1, b_1 ) \right\rbrace \\
+\mathcal{I}_2 = \left\lbrace \mathbf{a}_2 \mathbf{x} \leq b_2 \mid
+C_2 (\mathbf{a}_2, b_2 ) \right\rbrace
+\end{align*}
+To compute the intersection between them, it is sufficient to choose
+the same assignments of parameters and let them satisfy the both
+constraints $C_1$ and $C_2$.
+\[ \mathcal{I} = \left\lbrace \mathbf{a} \mathbf{x} \leq b \mid
+C_1(\mathbf{a}, b) \wedge C_2(\mathbf{a}, b) \right\rbrace \]
+
+However, if such an assignment of $\mathbf{a}$ and $b$ does not exist,
+no common solution exists across two problems because $\mathcal{I} =
+\emptyset$.
+
+By applying this approach, it is also possible to compute a common
+interpolant among more than three problems.
+
+\paragraph{Interpolation between logical formulas}
+We now extend the algorithm so that it can handle general logical
+formulas over linear arithmetic including disjunctions of linear
+expressions.  Here the algorithm takes two logical formulas over
+linear expressions as a problem input.  For convenience, we assume
+that they are represented in disjunctive normal form (DNF).  The
+solution for this problem is the interpolant that is a logical formula
+of linear expressions.
+
+First, consider the case of an interpolation problem between $A_1 \vee
+A_2$ and $B$.  When $(A_1 \vee A_2) \wedge B$ is inconsistent, $A_1
+\wedge B$ is necessarily inconsistent.  Let $\mathcal{I}_1$ denote the
+set of interpolants for $(A_1, B)$.  In the same manner, let
+$\mathcal{I}_2$ denote the set of interpolants for $(A_2, B)$.  We may
+naturally obtain an interpolant set by the intersection $\mathcal{I} =
+\mathcal{I}_1 \cup \mathcal{I}_2$ as a solution for $(A_1 \vee A_2,
+B)$.  When the algorithm computes $\mathcal{I}$ by the previous
+method, and it is $\emptyset$, we need to build a less simple solution
+by
+\[ \mathcal{I}' = \left\lbrace \left( I_1 \vee I_2 \right) \mid
+I_1 \in \mathcal{I}_1 \wedge I_2 \in \mathcal{I}_2 \right\rbrace \]
+Any formulas in the set $\mathcal{I}'$ are the interpolant for $(A_1
+\vee A_2, B)$ because
+\begin{align*}
+\forall I \in \mathcal{I}. & A_1 \vdash I \wedge A_2 \vdash I \wedge I \nvdash B \wedge \\
+& FV(I) = FV(A_1) \cup FV(A_2) \cup FV(B)
+\end{align*}
+
+Second, consider the case of an interpolation problem between $A$ and
+$B_1 \wedge B_2$.
+% TODO:
+
+By combining these two ways of interpolant construction, we may be
+able to build a small solution for more general cases like $(A_1,
+\ldots, A_n)$ and $(B_1, \ldots, B_n)$.  In order to obtain a small
+solution, we need to wisely choose the order of combining an
+interpolant set.
+
+\paragraph{Symmetric Interpolation}
+TODO
 

### 2013-08-26 article/algo.tex

@@
-\documentclass[a4paper,12pt]{article}
+\documentclass[master,final,11pt]{iscs-thesis}
+
+\etitle{Simple Solutions of Acyclic Horn Clauses for Program Verification}
+\eauthor{Yuto Takei}
+\esupervisor{Naoki Kobayashi}
+
+\jtitle{Lipsum}
+\jauthor{Lipsum}
+\jsupervisor{Lipsum}
+
+\supervisortitle{Professor}
+\date{August 28, 2013}
 
@@
 
-\title{Algorithm explanation}
-\author{Yuto Takei \\ The University of Tokyo}
-
 \begin{document}
 
-\section{Introduction}
+\begin{eabstract}
+Automated theorem proving plays a crucial role in modern program
+verification techniques.  In particular, in the context of software
+model checking, the interpolation problem and its generalization
+called \emph{Horn clause solving} have attracted special attention.
+The latter problem asks for a solution for a set of acyclic Horn
+clauses containing unknown predicate variables.  The solution, i.e., a
+substitution for predicate variables, then serves as a candidate of
+program invariants.  However, existing methods are unsatisfactory in
+that they tend to generate too specific a solution, which does not
+correspond to program invariants.
+
+To address the problem above, we propose a new algorithm for solving
+Horn clauses consisting of linear arithmetics and unknown predicate
+variables.  The new algorithm tries to find small predicates, based on
+the hypothesis that program's invariants tend to be simple formulas.
+To this end, our algorithm prepares a template of linear arithmetic
+formulas of a fixed size as a candidate of each unknown predicate, so
+that Horn clause solving is reduced to linear programming.  The
+algorithm iteratively increases the size of the templates until a
+solution is found.  It enables us to construct relatively small
+solutions in a reasonable time compared to existing algorithms.  We
+have implemented and incorporated our Horn clause solver into MoCHi, a
+software model checker for functional programs.  We have evaluated the
+effectiveness of our approach through experiments.
+\end{eabstract}
+\begin{jabstract}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.  Morbi
+vulputate ornare arcu et malesuada.  Proin sem urna, feugiat et velit
+ac, elementum blandit orci.  Etiam ut neque ac felis laoreet venenatis
+eget quis enim.  Ut aliquam dapibus sem egestas molestie.  Ut sit amet
+porttitor orci, id fermentum dui.  In non dui at ante sagittis
+malesuada at in neque.  Fusce euismod velit quis ante ullamcorper
+qsodales.
+\end{jabstract}
+\maketitle
+
+\begin{acknowledge}
+Thank you.
+\end{acknowledge}
+
+\frontmatter
+\tableofcontents
+
+\mainmatter
+
+\chapter{Introduction}
 
@@
 
-\section{Interpolation}
+\chapter{Interpolation}
 
@@
 
-\section{Horn clause solving}
+\chapter{Horn clause solving}
 

### 2013-08-26 article/thesis.tex
Change filename
Bibliography

--- 2013-08-26 article/algo.tex
@@
-\documentclass[master,final,11pt]{iscs-thesis}
+\documentclass[master,final,12pt]{iscs-thesis}
 
@@
 \usepackage{syntax}
+\usepackage{url}
 
@@
 One of the most common methods of such abstraction is predicate
-abstraction.  An automated theorem prover finds appropriate predicates
-for specific program locations.  An execution state at each location
-is then expressed by a Boolean valuation of the given predicates.  If
-the abstraction is too coarse and a model checker discovers an
-infeasible error execution path in an abstract system, the prover
-computes an additional predicate at a specific program location to
-refine the abstraction.  The additional predicate must be satisfied at
-the location under the execution in the concrete model along the
-discovered path, and the execution does not lead to the program
-failure by the same path.  After the refinement, the infeasible path
-is ruled out and no longer found by the model checker.
+abstraction\cite{DBLP:conf/cav/GrafS97}.  An automated theorem prover
+finds appropriate predicates for specific program locations.  An
+execution state at each location is then expressed by a Boolean
+valuation of the given predicates.  If the abstraction is too coarse
+and a model checker discovers an infeasible error execution path in an
+abstract system, the prover computes an additional predicate at a
+specific program location to refine the abstraction.  The additional
+predicate must be satisfied at the location under the execution in the
+concrete model along the discovered path, and the execution does not
+lead to the program failure by the same path.  After the refinement,
+the infeasible path is ruled out and no longer found by the model
+checker.
 
@@
 counterexample that a model checker discovered.  It is done by
-computing a Craig's interpolant between the strongest postcondition at
+computing a Craig interpolant between the strongest postcondition at
 the location and the weakest precondition to the failure point at
@@
 Although one may be able to obtain $x = y$ as the simplest
-interpolation between $\varphi_1$ and $\varphi_2$, CSIsat\cite{csisat}
-returns
+interpolation between $\varphi_1$ and $\varphi_2$,
+CSIsat\cite{DBLP:conf/cav/BeyerZM08} returns
 \[ TODO \]
@@
 
-\subsection{Example}
+\section{Example}
 
@@
 
-\subsection{Preliminaries}
-
-Our algorithm aims to compute a Craig's interpolant over linear
+\section{Preliminaries}
+
+Our algorithm aims to compute a Craig interpolant over linear
 arithmetic formulas.
 
-\paragraph{Craig's Interpolation}
+\paragraph{Craig Interpolation}
 Given two logical formulas $A$ and $B$ that are inconsistent each
-other, we call a new preposition $I$ a Craig's interpolant between $A$
-and $B$ such that $I$ is implied by $A$ and inconsistent with $B$.
-$I$'s vocabulary must be only free variables that appears in both $A$
-and $B$.
+other, we call a new preposition $I$ a \emph{Craig interpolant}
+\cite{DBLP:journals/jsyml/Craig57} between $A$ and $B$ such that $I$
+is implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
+only free variables that appears in both $A$ and $B$.  If $A \wedge B$
+is unsatisfiable, an interpolant $I$ always exists.
 
@@
 
-\subsection{Algorithm}
+\section{Algorithm}
 
@@
 
-\subsection{Preliminaries}
+\section{Preliminaries}
 
@@
 
-\subsection{Simple solution}
+\section{Simple solution}
 
@@
 
-\subsection {Graph transformation}
+\section {Graph transformation}
 
@@
 
+\chapter{Related work}
+
+Lorem ipsum
+\cite{DBLP:journals/jsyml/Craig57}
+\cite{DBLP:conf/cav/GrafS97}
+\cite{DBLP:conf/cav/McMillan03}
+\cite{DBLP:conf/popl/HenzingerJMM04}
+\cite{DBLP:journals/tcs/McMillan05}
+\cite{DBLP:conf/tacas/McMillan05}
+\cite{DBLP:conf/tacas/JhalaM06}
+\cite{website:foci}
+\cite{DBLP:conf/tacas/AmlaM07}
+\cite{DBLP:conf/pldi/BeyerHMR07}
+\cite{DBLP:journals/sttt/BeyerHJM07}
+\cite{DBLP:conf/vmcai/RybalchenkoS07}
+\cite{DBLP:conf/cav/BeyerZM08}
+\cite{DBLP:conf/ppdp/UnnoK09}
+\cite{DBLP:conf/cav/Rybalchenko10}
+\cite{DBLP:conf/popl/GuptaPR11}
+\cite{DBLP:conf/pldi/KobayashiSU11}
+\cite{DBLP:conf/aplas/GuptaPR11}
+\cite{DBLP:conf/pepm/SatoUK13}
+\cite{DBLP:conf/sas/BjornerMR13}
+\cite{DBLP:conf/cav/DaiXZ13}
+\cite{DBLP:conf/cav/RummerHK13}
+\cite{DBLP:conf/cav/AlbarghouthiM13}
+\cite{DBLP:conf/cav/BeyenePR13}.
+
 \bibliographystyle{plain}

### 2013-08-26 article/thesis.tex
Update on related work

@@
 One of the most common methods of such abstraction is predicate
-abstraction\cite{DBLP:conf/cav/GrafS97}.  An automated theorem prover
-finds appropriate predicates for specific program locations.  An
-execution state at each location is then expressed by a Boolean
-valuation of the given predicates.  If the abstraction is too coarse
-and a model checker discovers an infeasible error execution path in an
-abstract system, the prover computes an additional predicate at a
-specific program location to refine the abstraction.  The additional
-predicate must be satisfied at the location under the execution in the
-concrete model along the discovered path, and the execution does not
-lead to the program failure by the same path.  After the refinement,
+abstraction\cite{DBLP:conf/cav/GrafS97} with CounterExample-Guided
+Abstraction Refinement (CEGAR) \cite{DBLP:conf/cav/ClarkeGJLV00}.  An
+automated theorem prover finds appropriate predicates for every
+program location.  An execution states at those locations is then
+expressed by a Boolean valuation of the given predicates.  If the
+abstraction is too coarse and a model checker discovers an infeasible
+error execution path in an abstract system, the prover computes an
+additional predicate at those program locations to refine the
+abstraction predicates.
+
+In order to obtain an additional abstraction predicate at program
+locations along the infeasible counterexample, one class of automated
+theorem provers, called an interpolating theorem prover, computes a
+separating predicate called an interpolant between two sets of
+constraints; the one consists of the constraints from the program
+entry point to the location, and the other consists of the constraints
+from the location to the program failure.  Those constraints include
+variable assignments and assertions to be satisfied originated from
+conditional branches.
+
+The computed predicate by an interpolating theorem prover is satisfied
+at the location under the execution in the concrete model along the
+discovered counterexample, and any execution paths which satisfies the
+predicate does not lead to the failure state.  After the refinement,
 the infeasible path is ruled out and no longer found by the model
@@
 
-In order to obtain such a refinement predicate at the specific program
-location along the infeasible counterexample, one class of automated
-theorem provers, called an interpolating theorem prover, computes a
-Craig interpolant between two sets of constraints; the one consists of
-the constraints from the program entry point to the location, and the
-other consists of the constraints from the location to the program
-failure.  Those constraints include variable assignments and assertions
-to be satisfied originated from conditional branches.
-
 However, existing interpolating theorem provers may return too complex
 a solution which is heavily affected by the constraints of the
-specific path.  This causes the model checker to discover infinitely
-similar paths which pass the same program location by loops or
-recursion calls, and not to terminate.
-
-To avoid such situation, it is desired to use an invariant formula for
-the predicate abstraction.  Based on the hypothesis that program’s
+specific path.  It is not favorable in terms of the program
+verification because a complex predicate may not be sufficient to rule
+out similar spurious paths in programs with loops and recursion calls.
+This causes the model checker to discover spurious paths infinitely
+which pass the same program location by loops or recursion calls at
+different numbers of iteration, and the verification process does not
+to terminate.
+
+To avoid this, it is desired to discover an invariant formula for the
+predicate abstraction.  Based on the hypothesis that program's
 invariants tend to be simple formulas, we propose a new interpolating
@@
 symmetric interpolation problem, in which the algorithm computes
-predicates at multiple program locations along a path.  This enables
-the abstraction predicates at multiple locations to be updated at the
-same computation.
+predicates at multiple program locations along a path at once.  This
+enables the abstraction predicates at multiple locations to be updated
+at the same computation.
 
@@
 to be program invariants, by computing the same abstraction predicates
-at loops and recursion calls on different passes, and by unifying
-predicates at the same location over multiple paths.  The problems are
-generated by encoding constraints over execution paths with unknown
-predicates which describe functions' constraints between parameters
-and return values.
-
-The rest of the paper is structured as follows.  Chapter 2 proposes a
-new interpolating algorithm to obtain simple solutions, and extends it
-to solve symmetric interpolation problems.  Chapter 3 describes a new
-Horn clause solving algorithm by extending the previous interpolating
-algorithm.  Chapter 4 shows an experiment and its result to confirm
-the effect.  Chapter 5 mentions related work and future work.  As a
-conclusion, chapter 6 reviews the impact of our research.
+inside loops and recursion calls on different iterations, and by
+unifying predicates at the same location over multiple paths.  The
+problems are generated by encoding constraints over execution paths
+with unknown predicates which describe functions' constraints between
+parameters and return values.
+
+The rest of the paper is structured as follows.
+Chapter~\ref{chap:interpolation} proposes a new interpolating
+algorithm to obtain simple solutions, and extends it to solve
+symmetric interpolation problems.  Chapter~\ref{chap:horn} describes a
+new Horn clause solving algorithm by extending the previous
+interpolating algorithm.  Chapter~\ref{chap:experiment} describes
+experiments and illustrates the result to confirm the effect of our
+proposed algorithm.  Chapter~\ref{chap:future} gives a discussion of
+our improvements and proposes further improvements to make.
+Chapter~\ref{chap:related} mentions related work done by others.
+Finally, Chapter~\ref{chap:conclusion} reviews the impact of our
+research and brings the conclusion.
 
 \chapter{Interpolation}
+\label{chap:interpolation}
 
@@
 \chapter{Horn clause solving}
+\label{chap:horn}
 
@@
 
+\chapter{Experiment}
+\label{chap:experiment}
+
+We incorporated our algorithm to the software model checker for
+higher-order programs (MoCHi) \cite{DBLP:conf/pldi/KobayashiSU11}.
+
+\section{Interpolating theorem prover}
+
+Horn clause to interpolation problem.
+\cite{DBLP:conf/ppdp/UnnoK09}
+
+CSIsat
+\cite{DBLP:conf/cav/BeyerZM08}
+
+FOCI
+\cite{website:foci}
+
+CLP-Prover
+\cite{website:clp}
+
+We incorporated our algorithm to the software model checker for
+higher-order programs (MoCHi) \cite{DBLP:conf/pldi/KobayashiSU11}.
+
+\section{Horn clause solver}
+
 \chapter{Related work}
+\label{chap:related}
 
 Lorem ipsum
-\cite{DBLP:journals/jsyml/Craig57}
-\cite{DBLP:conf/cav/GrafS97}
-\cite{DBLP:conf/cav/McMillan03}
-\cite{DBLP:conf/popl/HenzingerJMM04}
-\cite{DBLP:journals/tcs/McMillan05}
-\cite{DBLP:conf/tacas/McMillan05}
-\cite{DBLP:conf/tacas/JhalaM06}
-\cite{website:foci}
-\cite{DBLP:conf/tacas/AmlaM07}
-\cite{DBLP:conf/pldi/BeyerHMR07}
-\cite{DBLP:journals/sttt/BeyerHJM07}
-\cite{DBLP:conf/vmcai/RybalchenkoS07}
-\cite{DBLP:conf/cav/BeyerZM08}
-\cite{DBLP:conf/ppdp/UnnoK09}
-\cite{DBLP:conf/cav/Rybalchenko10}
-\cite{DBLP:conf/popl/GuptaPR11}
-\cite{DBLP:conf/pldi/KobayashiSU11}
-\cite{DBLP:conf/aplas/GuptaPR11}
-\cite{DBLP:conf/pepm/SatoUK13}
-\cite{DBLP:conf/sas/BjornerMR13}
+\cite{DBLP:conf/cav/McMillan03} % Interpolation and SAT-Based Model Checking
+\cite{DBLP:conf/popl/HenzingerJMM04} % Abstractions from proofs
+\cite{DBLP:journals/tcs/McMillan05} % An interpolating theorem prover
+\cite{DBLP:conf/tacas/McMillan05} % Applications of Craig Interpolants in Model Checking
+\cite{DBLP:conf/tacas/JhalaM06} % A Practical and Complete Approach to Predicate Refinement
+\cite{DBLP:conf/tacas/AmlaM07} % Combining Abstraction Refinement and SAT-Based Model Checking
+\cite{DBLP:conf/pldi/BeyerHMR07} % Path invariants
+\cite{DBLP:journals/sttt/BeyerHJM07} % The software model checker Blast
+\cite{DBLP:conf/vmcai/RybalchenkoS07} % Constraint Solving for Interpolation
+\cite{DBLP:conf/cav/Rybalchenko10} % Constraint Solving for Program Verification: Theory and Practice by Example
+\cite{DBLP:conf/popl/GuptaPR11} % Predicate abstraction and refinement for verifying multi-threaded programs
+\cite{DBLP:conf/aplas/GuptaPR11} % Solving Recursion-Free Horn Clauses over LI+UIF
+\cite{DBLP:conf/pepm/SatoUK13} % Towards a scalable software model checker for higher-order
+
+Beautiful interpolants
+\cite{DBLP:conf/cav/AlbarghouthiM13}
+
+Disjunctions
+\cite{DBLP:conf/cav/RummerHK13}
+
+Non-linear
 \cite{DBLP:conf/cav/DaiXZ13}
-\cite{DBLP:conf/cav/RummerHK13}
-\cite{DBLP:conf/cav/AlbarghouthiM13}
-\cite{DBLP:conf/cav/BeyenePR13}.
+
+Qunatifiers
+\cite{DBLP:conf/sas/BjornerMR13} % On Solving Universally Quantified Horn Clauses
+\cite{DBLP:conf/cav/BeyenePR13}. % Solving Existentially Quantified Horn Clauses
+
+\chapter{Future Work}
+\label{chap:future}
+
+\paragraph{Non-linear extension}
+\cite{DBLP:conf/cav/DaiXZ13}
+
+\chapter{Conclusion}
+\label{chap:conclusion}
 

### 2013-08-26 article/thesis.tex
Related work separated

@@
 One of the most common methods of such abstraction is predicate
-abstraction\cite{DBLP:conf/cav/GrafS97} with CounterExample-Guided
-Abstraction Refinement (CEGAR) \cite{DBLP:conf/cav/ClarkeGJLV00}.  An
-automated theorem prover finds appropriate predicates for every
+abstraction\cite{DBLP:conf/cav/GrafS97} within CounterExample-Guided
+Abstraction Refinement (CEGAR) framework
+\cite{DBLP:conf/cav/ClarkeGJLV00,DBLP:conf/popl/BallR02,DBLP:conf/popl/HenzingerJMS02}.
+An automated theorem prover finds appropriate predicates for every
 program location.  An execution states at those locations is then
@@
 locations along the infeasible counterexample, one class of automated
-theorem provers, called an interpolating theorem prover, computes a
-separating predicate called an interpolant between two sets of
-constraints; the one consists of the constraints from the program
+theorem provers, called an interpolating theorem prover
+\cite{DBLP:journals/tcs/McMillan05,DBLP:conf/vmcai/RybalchenkoS07},
+computes a separating predicate called an interpolant between two sets
+of constraints; the one consists of the constraints from the program
 entry point to the location, and the other consists of the constraints
@@
 We incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{DBLP:conf/pldi/KobayashiSU11}.
+higher-order programs (MoCHi) \cite{DBLP:conf/pldi/KobayashiSU11}.  It
+is specifically designed to verify the correctness of higher-orde
+programs written in a small subset of ML.  It infers dependent types
+for input higher-order programs \cite{DBLP:conf/ppdp/UnnoK09}.
+
+Our implementation make use of Z3 \cite{DBLP:conf/tacas/MouraB08} as
+an SMT solver.
+
+For simplicity of the implementation, we assume that the linear
+inequalities in the integer space.  Because of this, our algorithm
+lost the completeness.  By using Motzkin's transpose theorem
+\cite{DBLP:journals/networks/Rajan90}, it is possible to handle linear
+inequalities in both $\mathbf{ax} + b \leq 0$ and $\mathbf{ax} + b <
+0$.
 
@@
 
-Horn clause to interpolation problem.
-\cite{DBLP:conf/ppdp/UnnoK09}
-
-CSIsat
-\cite{DBLP:conf/cav/BeyerZM08}
-
-FOCI
-\cite{website:foci}
-
-CLP-Prover
-\cite{website:clp}
-
-We incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{DBLP:conf/pldi/KobayashiSU11}.
+Our interpolating theorem prover is used as a sub-procedure to
+solve Horn clauses.
+
+\begin{itemize}
+\item CSIsat \cite{DBLP:conf/cav/BeyerZM08}
+\item FOCI \cite{website:foci}
+\item CLP-Prover \cite{website:clp}
+\end{itemize}
 
@@
 \label{chap:related}
-
-Lorem ipsum
-\cite{DBLP:conf/cav/McMillan03} % Interpolation and SAT-Based Model Checking
-\cite{DBLP:conf/popl/HenzingerJMM04} % Abstractions from proofs
-\cite{DBLP:journals/tcs/McMillan05} % An interpolating theorem prover
-\cite{DBLP:conf/tacas/McMillan05} % Applications of Craig Interpolants in Model Checking
-\cite{DBLP:conf/tacas/JhalaM06} % A Practical and Complete Approach to Predicate Refinement
-\cite{DBLP:conf/tacas/AmlaM07} % Combining Abstraction Refinement and SAT-Based Model Checking
-\cite{DBLP:conf/pldi/BeyerHMR07} % Path invariants
-\cite{DBLP:journals/sttt/BeyerHJM07} % The software model checker Blast
-\cite{DBLP:conf/vmcai/RybalchenkoS07} % Constraint Solving for Interpolation
-\cite{DBLP:conf/cav/Rybalchenko10} % Constraint Solving for Program Verification: Theory and Practice by Example
-\cite{DBLP:conf/popl/GuptaPR11} % Predicate abstraction and refinement for verifying multi-threaded programs
-\cite{DBLP:conf/aplas/GuptaPR11} % Solving Recursion-Free Horn Clauses over LI+UIF
-\cite{DBLP:conf/pepm/SatoUK13} % Towards a scalable software model checker for higher-order
-
-Beautiful interpolants
-\cite{DBLP:conf/cav/AlbarghouthiM13}
-
-Disjunctions
-\cite{DBLP:conf/cav/RummerHK13}
-
-Non-linear
-\cite{DBLP:conf/cav/DaiXZ13}
-
-Qunatifiers
-\cite{DBLP:conf/sas/BjornerMR13} % On Solving Universally Quantified Horn Clauses
-\cite{DBLP:conf/cav/BeyenePR13}. % Solving Existentially Quantified Horn Clauses
+\input{related.tex}
 

### 2013-08-26 article/thesis.tex
Reference name prefix removal

@@
 One of the most common methods of such abstraction is predicate
-abstraction\cite{DBLP:conf/cav/GrafS97} within CounterExample-Guided
+abstraction\cite{conf/cav/GrafS97} within CounterExample-Guided
 Abstraction Refinement (CEGAR) framework
-\cite{DBLP:conf/cav/ClarkeGJLV00,DBLP:conf/popl/BallR02,DBLP:conf/popl/HenzingerJMS02}.
+\cite{conf/cav/ClarkeGJLV00,conf/popl/BallR02,conf/popl/HenzingerJMS02}.
 An automated theorem prover finds appropriate predicates for every
@@
 theorem provers, called an interpolating theorem prover
-\cite{DBLP:journals/tcs/McMillan05,DBLP:conf/vmcai/RybalchenkoS07},
+\cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07},
 computes a separating predicate called an interpolant between two sets
@@
 interpolation between $\varphi_1$ and $\varphi_2$,
-CSIsat\cite{DBLP:conf/cav/BeyerZM08} returns
+CSIsat\cite{conf/cav/BeyerZM08} returns
 \[ TODO \]
@@
 other, we call a new preposition $I$ a \emph{Craig interpolant}
-\cite{DBLP:journals/jsyml/Craig57} between $A$ and $B$ such that $I$
+\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$
 is implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
@@
 We incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{DBLP:conf/pldi/KobayashiSU11}.  It
+higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It
 is specifically designed to verify the correctness of higher-orde
 programs written in a small subset of ML.  It infers dependent types
-for input higher-order programs \cite{DBLP:conf/ppdp/UnnoK09}.
-
-Our implementation make use of Z3 \cite{DBLP:conf/tacas/MouraB08} as
+for input higher-order programs \cite{conf/ppdp/UnnoK09}.
+
+Our implementation make use of Z3 \cite{conf/tacas/MouraB08} as
 an SMT solver.
@@
 lost the completeness.  By using Motzkin's transpose theorem
-\cite{DBLP:journals/networks/Rajan90}, it is possible to handle linear
+\cite{journals/networks/Rajan90}, it is possible to handle linear
 inequalities in both $\mathbf{ax} + b \leq 0$ and $\mathbf{ax} + b <
@@
 \begin{itemize}
-\item CSIsat \cite{DBLP:conf/cav/BeyerZM08}
-\item FOCI \cite{website:foci}
-\item CLP-Prover \cite{website:clp}
+\item CSIsat \cite{conf/cav/BeyerZM08}
+\item FOCI \cite{website/foci}
+\item CLP-Prover \cite{website/clp}
 \end{itemize}
@@
 \paragraph{Non-linear extension}
-\cite{DBLP:conf/cav/DaiXZ13}
+\cite{conf/cav/DaiXZ13}
 

### 2013-08-26 article/related.tex
Related work

Interpolating theorem proving have played an important role in program
verification.  For instance, it is applied to invariant generation for
verification condition based program analysis, and predicate discovery
for the model checking by CEGAR with predicate abstraction.  The
former method is adopted when analyzing input programs by denotational
semantics such as Hoare logic, and synthesizing invariants from
strongest postconditions and weakest preconditions at various program
locations.  The latter one is used when model check a system with
infinite states by abstracting its data domain with predicates into a
finite system.

Interpolation in invariant generation is first proposed
by \cite{conf/cav/McMillan03} for unbounded model checking of a state
transition system, which does not specify the number of transition
steps to verify in iterative state exploration.  Instead of
constructing a Boolean Decision Diagram (BDD), this method adopted
Craig interpolation for compute the image fixedpoint for inductive
invariant generation.

Inspired from this method, \cite{conf/popl/HenzingerJMM04} adopted the
use of Craig interpolation for predicate discovery in CEGAR framework
with predicate abstraction, and proposed a method to derivate an
interpolant from counterexample's refutation proof derivation tree.
This work implemented the proposed method into the Blast software
model checker \cite{conf/popl/HenzingerJMS02,journals/sttt/BeyerHJM07}
and verified the effectiveness of interpolants for predicate
abstraction through an experiment of C program verification.

\paragraph{Development of interpolating theorem provers}
An interpolating prover \textsc{FOCI} \cite{website/foci} was
developed under the comprehensive wrok on quantifier-free interpolant
generation over the theory of linear arithmetic and uninterpreted
function symbols (LI+UIF) theory \cite{journals/tcs/McMillan05}.  At the same
time, various applications of interpolants for model checking was
proposed \cite{conf/tacas/McMillan05}. They incorporated the algorithm
into Z3 SMT solver \cite{conf/tacas/MouraB08} and it is possible to
generate an interpolant from proofs generated by
Z3 \cite{conf/fmcad/McMillan11}.

On the other hand, an interpolating method without using proof
derivation was proposed that made use of the specific characteristics
of linear arithmetics.  \textsc{CLP-Prover} \cite{website/clp} is
another interpolating theorem prover with the theoretical ground of
linear constraint generation.  It focuses on interpolation over LI+UIF
and adopts Farkas's Lemma \cite{journals/networks/Rajan90} for give a
reasoning for handling linear
inequalities \cite{conf/vmcai/RybalchenkoS07, conf/cav/Rybalchenko10}.
Because the \textsc{CLP-Prover} cannot handle logical formulas with
disjunctions, we extended the basic principle of this work to handle
disjunctions. Our interpolating algorithm was strongly inspired.

Note that the suggested interpolating algorithms so far were not
complete in the sense that the algorithms may not be able to find a
suitable predicate that is necessary for a model checker to terminate.
Although it may be possible to simply enumerate interpolants in a
complete manner, its cost is not negligible with respect to the
computation time because of combinatorial explosion.
\cite{conf/tacas/JhalaM06} suggested a practical way to
obtain complete interpolating algorithm.

\paragraph{Extension to Horn clause solvers}
As a model checking technique, various methods are proposed to
efficiently eliminate spurious counterexamples from an abstract system
and terminate CEGAR process with less number of cycles.  Main
principles are support of disjunctive problems and extension to Horn
clause solvers.

One of them is work done by \cite{conf/pldi/BeyerHMR07} to eliminate
spurious paths that pass over loops.  The method converts a
counterexample from the model checker into a path that is a subgraph
of program's control flow graph, and performs a invariant synthesis
over it.  Here, the control flow graph may contain loops, it is not
suffcieint by the convential interpolation, and rather it required a
disjunctive reasoning.

As we explained in Chapter~\ref{chap:interpolation}, when
interpolating between two logical formulas, it is necessary to combine
interpolants by conjunctions and disjunctions.  To minimize the number
of terms efficiently, \cite{conf/cav/AlbarghouthiM13} suggested a
non-deterministic method of building such interpolant by wisely
sampling a half-interpolant.

For solving a complex system, it may not be possible to analyze
system's behavior by interpolation.  \cite{conf/aplas/GuptaPR11}
showed an algorithm to solve Horn clauses over LI+UIF.

One of the most important applications of Horn clause solving is
verifying higher-order functional programs.  Interpolation is used as
a sub procedure of solving Horn clauses for dependent type
inference \cite{conf/ppdp/UnnoK09, conf/pepm/SatoUK13} in this
context. In MoCHi, the refinement types are given for values and
variables for program abstraction.  Horn clause solving is necessary
for inferring the refinment types.  \cite{conf/popl/Terauchi10}
suggested a method for dependent type inference algorithm as
well. Their algorithm is complete in the sense that it eventually
returns a solution which terminates the verification process.

Another application of Horn clause solving is the verification of
concurrent systems \cite{conf/popl/GuptaPR11}. The work suggested to
describe a multi-thread program's behavior in Horn clauses and perform
model checking over abstract system.

Finally, some research integrated disjunctive interpolating methods
with Horn clause solving algorithms.  In previously suggested methods
in \cite{conf/popl/HenzingerJMM04, conf/cav/McMillan06}, it was not
possible to consider multiple spurious
paths.  \cite{conf/cav/RummerHK13} suggested an algorithm to solve
disjunctive Horn clauses, and it became possible to give a stronger
reasoning on a program invariant where multiple paths meet.  It also
enabled the abstraction for the case that branch conditions had a
disjunctive operator in them.

\paragraph{Recent work}

One of the recent extensions of Horn clause solving is to handle
non-linear constraints \cite{conf/cav/DaiXZ13}.  For instance, if a
program contains integer multiplication operation over variables,
existing algorithms cannot infer program invariants.  This research
aims to overcome the hurdle and infer non-linear invariants for
program verification.

As another direction of extension, some works are done for solving
Horn clauses with qunatified consequences \cite{conf/sas/BjornerMR13,
conf/cav/BeyenePR13}.  They are used to abstract more complex systems
efficiently.

### 2013-08-26 article/eabstract.tex
Other chapters also separated

--- 2013-08-26 article/thesis.tex
@@
-\documentclass[master,final,12pt]{iscs-thesis}
-
-\etitle{Simple Solutions of Acyclic Horn Clauses for Program Verification}
-\eauthor{Yuto Takei}
-\esupervisor{Naoki Kobayashi}
-
-\jtitle{Lipsum}
-\jauthor{Lipsum}
-\jsupervisor{Lipsum}
-
-\supervisortitle{Professor}
-\date{August 28, 2013}
-
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{stmaryrd}
-\usepackage{syntax}
-\usepackage{url}
-
-\newcommand{\sembrack}[1]{[\![#1]\!}
-\newcommand{\edge}[2]{#1\rightarrow#2}
-\newcommand{\edgel}[3]{#1\xrightarrow{#2}#3}
-
-\begin{document}
-
 \begin{eabstract}
@@
 \end{eabstract}
-\begin{jabstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.  Morbi
-vulputate ornare arcu et malesuada.  Proin sem urna, feugiat et velit
-ac, elementum blandit orci.  Etiam ut neque ac felis laoreet venenatis
-eget quis enim.  Ut aliquam dapibus sem egestas molestie.  Ut sit amet
-porttitor orci, id fermentum dui.  In non dui at ante sagittis
-malesuada at in neque.  Fusce euismod velit quis ante ullamcorper
-qsodales.
-\end{jabstract}
-\maketitle
-
-\begin{acknowledge}
-Thank you.
-\end{acknowledge}
-
-\frontmatter
-\tableofcontents
-
-\mainmatter
-
-\chapter{Introduction}
-
-Automated theorem proving plays an important role in the modern
-program verification.  Particularly in the context of verification
-with model checking techniques, an automated theorem prover is used to
-abstract an infinite set of program states into a finite one so that
-the model checker can virtually explore all possible program execution
-paths.
-
-One of the most common methods of such abstraction is predicate
-abstraction\cite{conf/cav/GrafS97} within CounterExample-Guided
-Abstraction Refinement (CEGAR) framework
-\cite{conf/cav/ClarkeGJLV00,conf/popl/BallR02,conf/popl/HenzingerJMS02}.
-An automated theorem prover finds appropriate predicates for every
-program location.  An execution states at those locations is then
-expressed by a Boolean valuation of the given predicates.  If the
-abstraction is too coarse and a model checker discovers an infeasible
-error execution path in an abstract system, the prover computes an
-additional predicate at those program locations to refine the
-abstraction predicates.
-
-In order to obtain an additional abstraction predicate at program
-locations along the infeasible counterexample, one class of automated
-theorem provers, called an interpolating theorem prover
-\cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07},
-computes a separating predicate called an interpolant between two sets
-of constraints; the one consists of the constraints from the program
-entry point to the location, and the other consists of the constraints
-from the location to the program failure.  Those constraints include
-variable assignments and assertions to be satisfied originated from
-conditional branches.
-
-The computed predicate by an interpolating theorem prover is satisfied
-at the location under the execution in the concrete model along the
-discovered counterexample, and any execution paths which satisfies the
-predicate does not lead to the failure state.  After the refinement,
-the infeasible path is ruled out and no longer found by the model
-checker.
-
-However, existing interpolating theorem provers may return too complex
-a solution which is heavily affected by the constraints of the
-specific path.  It is not favorable in terms of the program
-verification because a complex predicate may not be sufficient to rule
-out similar spurious paths in programs with loops and recursion calls.
-This causes the model checker to discover spurious paths infinitely
-which pass the same program location by loops or recursion calls at
-different numbers of iteration, and the verification process does not
-to terminate.
-
-To avoid this, it is desired to discover an invariant formula for the
-predicate abstraction.  Based on the hypothesis that program's
-invariants tend to be simple formulas, we propose a new interpolating
-algorithm which tries to minimize the number of conjunctions and
-disjunctions in a solution.  The predicate obtained by this algorithm
-may become similar to the invariant, and the number of trials in model
-checking may be reduced.
-
-Additionally, we extend our interpolating algorithm to solve a
-symmetric interpolation problem, in which the algorithm computes
-predicates at multiple program locations along a path at once.  This
-enables the abstraction predicates at multiple locations to be updated
-at the same computation.
-
-Finally, we extend our algorithm to solve Horn clause solving
-problems.  It allows us to obtain predicates which are highly likely
-to be program invariants, by computing the same abstraction predicates
-inside loops and recursion calls on different iterations, and by
-unifying predicates at the same location over multiple paths.  The
-problems are generated by encoding constraints over execution paths
-with unknown predicates which describe functions' constraints between
-parameters and return values.
-
-The rest of the paper is structured as follows.
-Chapter~\ref{chap:interpolation} proposes a new interpolating
-algorithm to obtain simple solutions, and extends it to solve
-symmetric interpolation problems.  Chapter~\ref{chap:horn} describes a
-new Horn clause solving algorithm by extending the previous
-interpolating algorithm.  Chapter~\ref{chap:experiment} describes
-experiments and illustrates the result to confirm the effect of our
-proposed algorithm.  Chapter~\ref{chap:future} gives a discussion of
-our improvements and proposes further improvements to make.
-Chapter~\ref{chap:related} mentions related work done by others.
-Finally, Chapter~\ref{chap:conclusion} reviews the impact of our
-research and brings the conclusion.
-
-\chapter{Interpolation}
-\label{chap:interpolation}
-
-Among various program verification techniques, when software model
-checking with predicate abstraction is adopted, the interpolation is
-used to compute the abstraction predicate along a spurious
-counterexample that a model checker discovered.  It is done by
-computing a Craig interpolant between the strongest postcondition at
-the location and the weakest precondition to the failure point at
-every program location.
-
-However, by the existing interpolating theorem prover, we may not be
-able to obtain a simple solution.  Consider the example below.
-\begin{align*}
-\varphi_1 : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
-\varphi_2 : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
-\end{align*}
-Although one may be able to obtain $x = y$ as the simplest
-interpolation between $\varphi_1$ and $\varphi_2$,
-CSIsat\cite{conf/cav/BeyerZM08} returns
-\[ TODO \]
-as a solution.
-
-We propose an interpolation algorithm which returns as simple a
-solution as possible in terms of the number of conjunctions and
-disjunctions.  Our algorithm focuses on interpolating problems on
-linear arithmetic and it works in a constructive manner.
-
-\section{Example}
-
-Consider the following interpolation problem.
-
-\begin{align*}
-\frac
-{A: (x \leq a) \wedge (a + 1 \leq y)}
-{B: (y \leq b) \wedge (b + 1 \leq x)}
-\end{align*}
-
-Here, the linear arithmetic formulas A and B are inconsistent.  The
-interpolant I for this problem is $x-y+1 \leq 0$, which is implied by
-$A$ and inconsistent with $B$.  The vocabulary of $I$ is $\left\lbrace
-x,y \right\rbrace$ and is over $A$'s vocabulary $\left\lbrace x,y,a
-\right\rbrace$ and $B$'s $\left\lbrace x,y,b \right\rbrace$.
-
-In computing the interpolant, we make a linear constraint of a
-interpolant by applying Farkas's lemma.  First, we assing weight
-parameters to every expression in conjunction form.
-
-\begin{align*}
-\begin{array}{r r r r r r l}
-\lambda_1 : & -a & & +x & & & \leq 0 \\
-\lambda_2 : & a & & & -y & +1 & \leq 0 \\
-\hline
-\lambda_3 : & & -b & & +y & & \leq 0 \\
-\lambda_4 : & & b & -x & & +1 & \leq 0
-\end{array}
-\end{align*}
-
-Because all expressions induce inconsistent as a whole, the sum of
-weighted expressions should satisfy the following constraint.
-
-\begin{align*}
-\begin{array}{r l l}
-- \lambda_1 + \lambda_2 & = 0 \quad & (a) \\
-- \lambda_3 + \lambda_4 & = 0 & (b) \\
-  \lambda_1 - \lambda_4 & = 0 & (x) \\
-- \lambda_2 + \lambda_3 & = 0 & (y) \\
-  \lambda_2 + \lambda_4 & > 0 & (\text{constant}) \\
-\lambda_1, \ldots, \lambda_4 & \geq 0 & (\text{non-negative weight})
-\end{array}
-\end{align*}
-
-With the linear constraint among $\lambda_i$ above, the interpolant is
-represented as follows by the sum of weighted expressions from the
-first formula groups.
-\[ (- \lambda_1 + \lambda_2) a + \lambda_1 x - \lambda_2 y + \lambda_2 \leq 0 \]
-One of the model of the linear constraint is $\lambda_1 = \lambda_2 =
-\lambda_3 = \lambda_4 = 1$, and we obtain $x-y+1 \leq 0$ as a
-solution.
-
-% TODO: Better to mention multiple interpolants possible for one
-% interpolating problem?
-
-\section{Preliminaries}
-
-Our algorithm aims to compute a Craig interpolant over linear
-arithmetic formulas.
-
-\paragraph{Craig Interpolation}
-Given two logical formulas $A$ and $B$ that are inconsistent each
-other, we call a new preposition $I$ a \emph{Craig interpolant}
-\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$
-is implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
-only free variables that appears in both $A$ and $B$.  If $A \wedge B$
-is unsatisfiable, an interpolant $I$ always exists.
-
-Our algorithm handles a linear expression $e$ which is in a form:
-\[ a_1 x_1 + \cdots + a_n x_n + b \leq 0 \]
-This expression contains coefficient parameters $a_1, \ldots, a_n$ and
-a constant parameter $b$.  We treat a special expressions $\top$ as $0
-\leq 0$ and $\bot$ as $1 \leq 0$.  The algorithm input and output is a
-formula $\psi$ that consists of conjunctions and/or disjunctions of
-linear expressions.
-\[ \psi = e | e \wedge e | e \vee e | \top | \bot \]
-
-In the simple interpolation, the algorithm receives two formulas in
-$\psi$ and returns an interpolant in $\psi$.  For the later extension
-to handle a symmetric interpolation problem, the algorithm receives
-multiple formulas in $\psi$ and returns intermediate interpolants in
-$\psi$.  For computing those interpolants over linear arithmetics, we
-use Farkas's Lemma.
-
-\paragraph{Farkas's Lemma on linear inequalities}
-Let a linear inequality $e_i$ be represented as $a_i1 x_1 + \cdots +
-a_im x_m <= a_i0$.  Assuming that $e_1,\cdots,e_n$ implies $e_0$,
-there exists $\lambda_1,\cdots,\lambda_n$ that satisfy $a_0j
-=\sum_(i=1)^n \lambda_i * a_ij (j=0, \ldots, m)$.
-
-For the special case of Farkas's Lemma, we can state that:
-
-Assuming that $e_1,\cdots,e_n$ implies $\bot$, there exists
-$\lambda_1,\cdots,\lambda_n$ that satisfy $???? TODO $.
-
-Our interpolating algorithm make use of the latter version of Farkas's
-Lemma.
-
-\section{Algorithm}
-
-Our algorithm aims to construct a relatively small interpolant in a
-reasonable computation time for any input.  It also aims to obtain a
-common solution among multiple interpolating problems.  When the
-algorithm is used as a sub-routine for solving Horn clause satisfying
-problems that are made while discovering predicates for abstraction
-during the program verification.  These advantages of our algorithm
-make it possible to return a relatively small predicates for those
-problems. To accomplish these aims, the algorithm preserves a set of
-interpolants during the computation, and define and execute operations
-over interpolant sets.
-
-\paragraph{Interpolation between conjunctions}
-We first present an interpolating algorithm between two conjunctive
-sets of linear inequalities $A = \left\lbrace e_1,\ldots,e_m
-\right\rbrace$ and $B = \left\lbrace e_{m+1},\ldots,e_{m+n}
-\right\rbrace$.  We assume that conjunctions of $A$ and $B$ are
-inconsistent.  For later convenience, we call this interpolating
-problem $\left( A, B \right)$.  Then there exists an interpolating
-linear expression $e_\star$ which satisfies
-\begin{align*}
-\left\lbrace e_1,\ldots,e_m \right\rbrace & \vdash e_\star \\
-e_\star & \nvdash \left\lbrace e_{m+1},\ldots,e_{m+n} \right\rbrace
-\end{align*}
-For convenience, we assume that any expression $e_i$ have variables
-$x_1, \ldots, x_a$.
-
-According to the Farkas's lemma, there exists a set of assignmentqs to
-$\lambda_i$, to conclude $\bot$ from a set of linear expressions $A
-\cup B$, namely,
-\begin{align*}
-\exists \lambda_1, \ldots, \lambda_m. \\
-a % TODO: coefficient, constant constraints
-\end{align*}
-
-Then, a linear combination of $A$'s formula with a weight $\lambda_1,
-\ldots, \lambda_m$ becomes an interpolant $I$ between $A$ and $B$.
-Note that any assignments to $\lambda_i$ that satisfy the expressions
-above will let an expression $I$ to be an interpolant.  Considering
-this, it is possible to obtain different interpolants for a
-interpolating problem by computing different assignments to
-$\lambda_i$.
-
-This enables the algorithm to preserve a set of interpolants by a
-linear expression template $a_1 x_1 + \cdots + a_n x_n + b \leq 0$
-with a constraint to describe a possible space of coefficient
-parameters $a_i$ and the constant parameter $b$.
-
-In practical, we may use linear programming solvers to obtain a model
-for $\lambda_i$.  By assigning concrete values to them, we can obtain
-a concrete interpolant.
-
-\paragraph{Computing a common interpolant}
-Next we consider a method to compute a common interpolant across two
-different interpolating problems.  That is, when given two
-interpolating problems $\left(A_1, B_1 \right)$ and $\left(A_2, B_2
-\right)$, the problem is about to compute the common interpolant I,
-which is an interpolant for the first problem and the second one at
-the same time.
-
-This problem is used when solving a following Horn clause solving
-problem, for instance.
-\[
-\implies
-% TODO
-\]
-
-To accomplish this, we simply need to compute the intersection of
-interpolant sets.  Let us call the interpolant space for the problems
-$\left(A_1, B_1 \right)$ and $\left(A_2, B_2 \right)$,
-$\mathcal{I}_1$, and $\mathcal{I}_2$.  In the previous method, a set
-of interpolants was represented by a parameterized linear expression.
-\begin{align*}
-\mathcal{I}_1 = \left\lbrace \mathbf{a}_1 \mathbf{x} \leq b_1 \mid
-C_1 (\mathbf{a}_1, b_1 ) \right\rbrace \\
-\mathcal{I}_2 = \left\lbrace \mathbf{a}_2 \mathbf{x} \leq b_2 \mid
-C_2 (\mathbf{a}_2, b_2 ) \right\rbrace
-\end{align*}
-To compute the intersection between them, it is sufficient to choose
-the same assignments of parameters and let them satisfy the both
-constraints $C_1$ and $C_2$.
-\[ \mathcal{I} = \left\lbrace \mathbf{a} \mathbf{x} \leq b \mid
-C_1(\mathbf{a}, b) \wedge C_2(\mathbf{a}, b) \right\rbrace \]
-
-However, if such an assignment of $\mathbf{a}$ and $b$ does not exist,
-no common solution exists across two problems because $\mathcal{I} =
-\emptyset$.
-
-By applying this approach, it is also possible to compute a common
-interpolant among more than three problems.
-
-\paragraph{Interpolation between logical formulas}
-We now extend the algorithm so that it can handle general logical
-formulas over linear arithmetic including disjunctions of linear
-expressions.  Here the algorithm takes two logical formulas over
-linear expressions as a problem input.  For convenience, we assume
-that they are represented in disjunctive normal form (DNF).  The
-solution for this problem is the interpolant that is a logical formula
-of linear expressions.
-
-First, consider the case of an interpolation problem between $A_1 \vee
-A_2$ and $B$.  When $(A_1 \vee A_2) \wedge B$ is inconsistent, $A_1
-\wedge B$ is necessarily inconsistent.  Let $\mathcal{I}_1$ denote the
-set of interpolants for $(A_1, B)$.  In the same manner, let
-$\mathcal{I}_2$ denote the set of interpolants for $(A_2, B)$.  We may
-naturally obtain an interpolant set by the intersection $\mathcal{I} =
-\mathcal{I}_1 \cup \mathcal{I}_2$ as a solution for $(A_1 \vee A_2,
-B)$.  When the algorithm computes $\mathcal{I}$ by the previous
-method, and it is $\emptyset$, we need to build a less simple solution
-by
-\[ \mathcal{I}' = \left\lbrace \left( I_1 \vee I_2 \right) \mid
-I_1 \in \mathcal{I}_1 \wedge I_2 \in \mathcal{I}_2 \right\rbrace \]
-Any formulas in the set $\mathcal{I}'$ are the interpolant for $(A_1
-\vee A_2, B)$ because
-\begin{align*}
-\forall I \in \mathcal{I}. & A_1 \vdash I \wedge A_2 \vdash I \wedge I \nvdash B \wedge \\
-& FV(I) = FV(A_1) \cup FV(A_2) \cup FV(B)
-\end{align*}
-
-Second, consider the case of an interpolation problem between $A$ and
-$B_1 \wedge B_2$.
-% TODO:
-
-By combining these two ways of interpolant construction, we may be
-able to build a small solution for more general cases like $(A_1,
-\ldots, A_n)$ and $(B_1, \ldots, B_n)$.  In order to obtain a small
-solution, we need to wisely choose the order of combining an
-interpolant set.
-
-\paragraph{Symmetric Interpolation}
-TODO
-
-\chapter{Horn clause solving}
-\label{chap:horn}
-
-We propose a new algorithm for solving recursion-less Horn clauses to
-generate a solution which is as simple as possible.  This enables us
-to obtain appropriate predicates that are suitable for predicate
-abstraction in the program verification techniques.
-
-The construction of the algorithm roughly consists of two parts.
-\begin{itemize}
-\item \textbf{Iterative Constraint Generation} is a process to build a
-  constraint to compute a solution.  It starts from generating the
-  weakest constraint in the sense that the resulting solution would
-  not be simple yet.  The process iteratively strengthen a constraint
-  until the strongest one while preserving the constraint
-  satisfiability.  The strongest constraint gives a simple solution.
-\item \textbf{Problem Restructuring} is a process to restructure the
-  problem to obtain a less simpler solution, if the constraint become
-  unsatisfiable during the iterative constraint generation.  This
-  process considers the unsatisfiable core of the constraint, and
-  wisely enlarge the solution size.
-\end{itemize}
-
-All predicate variables in an input problem are assigned simple
-templates.  Those predicate templates are parameterized.  In the
-iterative constraint generation process, algorithm builds a linear
-constraint that parameters in templates must satisfy.  This is done in
-a constructive manner, and constraints are build in the order of
-implication dependency in Horn clauses.  That is, when an input
-problem has a Horn clause thath has a predicate variable $P$ on its
-left-hand and $Q$ on its right-hand, a constraint to describe
-parameters in $P$'s template are first build before those in $Q$'s
-template.  When all parameters in the whole problem are described, the
-built constraint are evaluated to obtain a model.  Its instanciation
-becomes an appropriate solution for the input problem.
-
-The novelty of our algorithm lies in the incremental constraint
-refinement.  The algorithm tries to solve from the weakest constraint
-that an input problem allows that does not give a simple solution, and
-gradually strengthen the constraint until it reaches to the simplest
-solution that the problem has.  When a constraint becomes
-unsatisfiable during this step, the algorithm analyzes the
-unsatisfiable core and transform the problem to obtain a less simple
-solution.
-
-This construction of the algorithm enables us to compute a relatively
-simpler solution compared to existing algorithms while preserving the
-computation cost in a reasonable time.
-
-\section{Example}
-
-\paragraph {Tree interpolation}
-First, consider the following example.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-y \leq 2 & \implies B(y) \\
-A(x) \wedge B(y) & \implies 3x+2y \leq 10
-\end{align*}
-
-In this example, we would like to compute linear arithmetic formulas
-for the predicate variables $A$ and $B$.  They appear only once on
-both left-hand and right-hand globally in the problem.
-
-We try to obtain a simple solution that is expressed by linear
-expression templates, whose coefficients and constants are
-parameterized.  Assume that they have following predicate templates.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\
-B(y) : c_3 y + c_4 \leq 0
-\end{align*}
-
-By naturally encoding the first Horn clause with Farkas's lemma, we
-obtain a linear constraint
-$ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
-Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will
-make $A(x)$ valid for the first clause.  In the same manner, we obtain
-$ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
-from the second clause, and
-$ c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0 $
-from the third.
-
-One of the models satisfying all the constraints above is
-$( c_1, c_2, c_3, c_4, \lambda_1, \lambda_2, \lambda_3 ) = (3, -3, 2, -4, 3, 2, 1)$.
-By assigning them to the template, we obtain the following solution.
-
-\begin{align*}
-A(x) : 3 x - 3 \leq 0 \\
-B(y) : 2 y - 4 \leq 0
-\end{align*}
-
-\paragraph {DAG problem solving}
-Second, we explain a case that some predicate variables appear
-multiple times on left-hand side in different Horn clauses.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-A(x) \wedge A(y) & \implies x+y \leq 2
-\end{align*}
-
-Assigning a template $A(x) : c_1 x + c_2 \leq 0$, we obtain the same
-constraint for the first clause as the previous example.
-
-\begin{align} \label{eq:constr1}
-\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0
-\end{align}
-
-Sinve we have $A$'s appearance twice on the second clause, it is
-possible for $A$ to have two expressions each of which is used for
-$A(x)$ and $A(y)$.  Although we need to use the same template for
-multiple occurrences in order to compute the simplest solution, it may
-make the whole constraints unsatisfiable if the simplest solution does
-not exist.
-
-Therefore, we generate a weaker constraint for the second clause that
-allows different expressions of $A$ for multiple occurrence.  If the
-whole constraint is satisfiable in the case, we try to solve in a
-stronger constraint to generate a simple solution.
-
-In this specific case, it is virtually equivalent to solve the
-following problem in a weaker constraint.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 1 & \implies A_1(x) \\
-x \leq 1 & \implies A_2(x) \\
-A_1(x) \wedge A_2(y) & \implies x+y \leq 2 \\
-\\
-\text{Templates:} \\
-A_1(x) \text{ : } & c_{1,1} x + c_{2,1} \leq 0 \\
-A_2(x) \text{ : } & c_{1,2} x + c_{2,2} \leq 0 \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-\end{align*}
-
-By duplicating the constraint~\ref{eq:constr1} of $A$ for newly
-created $A_1$ and $A_2$, we obtain the whole constraint as follows.
-
-\begin{align*}
-& \lambda_{1,1} = c_{1,1} \wedge - \lambda_{1,1} \geq c_{2,1} \wedge \lambda_{1,1} \geq 0 \wedge \\
-& \lambda_{1,2} = c_{1,2} \wedge - \lambda_{1,2} \geq c_{2,2} \wedge \lambda_{1,2} \geq 0 \wedge \\
-& c_{1,1} = \lambda_2 \wedge c_{1,2} = \lambda_2 \wedge c_{2,1} + c_{2,2} \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-This constraint is satisfiable and one model is
-\[ ( c_{1,1}, c_{1,2}, c_{2,1}, c_{2,2}, \lambda_{1,1}, \lambda_{1,2}, \lambda_2 ) =
-( 1, -1, 1, -1, 1, 1, 1 ) \] and the solution then becomes
-$A(x) : (x -1 \leq 0) \wedge (x -1 \leq 0)$.
-
-Since the weak constraint is satisfiable, we now refine the constraint
-to obtain the better solution.  That is, simply eliminate the
-duplication to build the following constraint for the original
-template.
-
-\begin{align*}
-& \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \\
-& c_1 = \lambda_2 \wedge c_1 = \lambda_2 \wedge 2 c_2 \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-Solving this constraint simply gives us a solution $A(x) : x -1 \leq 0$.
-
-\paragraph {Disjunctive Horn clause solving}
-Finally, we explain a case that an input set of Horn clauses contain
-disjunctions on left-hand side.
-
-\begin{align*}
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 \vee -x+1 \leq 0 & \implies B(x) \\
-A(x) \wedge B(x) & \implies 1 \leq 0
-\end{align*}
-
-Same as the previous examples, templates for predicate variables are
-prepared.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\ B(x) : c_3 x + c_4 \leq 0
-\end{align*}
-
-The constraints generated are follows.
-
-\begin{align*}
-& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
-& \lambda_3 = c_3 \wedge \lambda_3 \geq c_4 \wedge \lambda_3 \geq 0 \wedge \\
-& - \lambda_4 = c_3 \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
-& c_1 + c_3 = 0 \wedge c_2 + c_4 > 0
-\end{align*}
-
-The constraints here become unsatisfiable, and one possible
-unsatisfiable core is:
-
-\begin{align*}
-\lambda_3 & = c_3 \\
-\lambda_3 & \geq c_4 \\
-- \lambda_4 & = c_3 \\
-\lambda_4 & \geq c_4 \\
-c_2 + c_4 & > 0 \\
-0 & \geq c_2
-\end{align*}
-
-Because the unsatisfiable core contains constraints originating from
-the disjunctions to be expressed in a single term template.  This let
-us determine to relax the template to allow a enlarged solution.
-
-Now we slightly change the problem and the template.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A(x) \wedge B_1(x) & \implies 1 \leq 0 \\
-A(x) \wedge B_2(x) & \implies 1 \leq 0 \\
-\\
-\text{Templates:} \\
-A(x) \text{ : } & c_1 x + c_2 \leq 0 \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-When we construct the constraints, as same as previous, we again
-obtain an unsatisfiable core:
-
-\begin{align*}
-\text{Constraint:} \\
-TODO \\
-\\
-\text{Unsatisfiable core:} \\
-TODO
-\end{align*}
-
-This time, the unsatisfiable core contains the constraints that
-originate from
-
-TODO
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A_i(x) \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A_i(x) \wedge B_j(x) & \implies 1 \leq 0 \quad (i,j \in \left\lbrace 1,2 \right\rbrace ) \\
-\\
-\text{Templates:} \\
-A_i(x) \text{ : } & c_{1,i} x + c_{2,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-Finally, we obtain the following constraint:
-
-TODO
-
-The process also generates a predicate template for every predicate
-variable that the solution will become.  under certain solution
-template.
-
-\section{Preliminaries}
-
-For ease of discussion, our algorithm handles an input set of Horn
-clauses by a \emph{Horn graph} that express the equivalent problem.
-The graph is then used throughout the algorithm to represent the
-problem.  A Horn graph is a labelled directed acyclic graph
-$G=(V,E,\varphi)$.
-\begin{itemize}
-\item $V$ is a union of two disjoint sets of vertices; Horn term
-  vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex $u
-  \in V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
-  denoted as $succ(u)$.
-\item $E$ is a set of labelled edges, which are expressed in the form
-  $(u,v,\theta)$ where $\theta$ is a finite map from variables to
-  variables.
-\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
-  term vertices, where L is a set of elements of a predicate variable
-  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
-  inequality $a_1 x_1 + \cdots + a_n x_n + b \leq 0$.
-\end{itemize}
-There are two kinds of edges in a Horn graph;
-\begin{itemize}
-\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
-  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
-\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
-  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
-  empty.
-\end{itemize}
-
-We define the relation $\leadsto$ over $V_T$ as follows.
-\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow \exists a \in V_\Rightarrow;
-\edgel{u}{\theta}{a} \wedge \edge{a}{v} \] We may omit $\theta$ and
-write $u \leadsto v$ for convenience in later discussions. By the
-assumption that $G$ is acyclic, both $\rightarrow^+$ and $\leadsto^+$ are
-well-founded relations. $v_\bot$ is the greatest element with respect
-to both $\rightarrow^+$ and $\leadsto^+$.
-
-The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
-Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$.
-\begin{align*}
-\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(succ(v)) \\
-\llbracket G \rrbracket & = \bigwedge_{v \in V_\Rightarrow} \mathcal{HC}_v
-\end{align*}
-We implicitly assume each $\mathcal{HC}_v$ is universally quantified
-by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.
-
-Horn term vertices with predicate variable labels are unique each
-other that any of them do not have the same variable on their labels.
-From now on, we may denote a Horn term vertex with a predicate
-variable label by its predicate variable name if no ambiguity is
-present.
-
-Without loss of generality, we assume the existence of the \emph{root
-  vertex} $v_\bot \in V_T$ with linear expression labels $1 \leq 0$,
-meaning $\bot$.  This is because if we have a Horn clause $P \implies
-e$ which $e$ is a linear expression, we can convert it to $P \wedge
-\neg e \implies 1 \leq 0$.  We also assume that other vertices with
-linear expression labels have no incoming edges to them.
-
-For simplicity of discussion, we define \emph{term arity} $k_v$ for
-every Horn term vertex $v \in V_T$ with incoming edges.
-\begin{align*}
-k_v =
-\begin{cases}
-\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
-0 & \mbox{if } \varphi(v) = \bot
-\end{cases}
-\end{align*}
-We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.
-
-A variable mapping label $\theta$ on edges is used to map the index of
-one predicate variable to another. (TODO?)
-
-(Example?)
-
-The algorithm assigns a \emph{predicate template} for all predicate
-variables in $G$. A \emph{predicate templateset} is a set of mappings from
-a predicate variable to a predicate template.  A template for a
-predicate variable $P(x_1, \ldots, x_n)$ is a formula $\psi$
-constructed from the following definition.
-\begin{align*}
-\psi ::= & a_1 x_1 + \cdots + a_n x_n + b \leq 0 \mid \\
-& \psi \wedge \psi \mid \psi \vee \psi
-\end{align*}
-A linear expression in the definition above may also be represented in
-a equivalent vector form $\mathbf{a} \mathbf{x} + b \leq 0$ for short.
-Parameters $\mathbf{a}$ and $b$ are specified by a linear constraint.
-
-The solution of $G$ is a map $\rho$ from predicate variables to linear
-predicates of the form $\lambda x_1, \ldots ,x_n. \psi$ where
-parameters in $\psi$ are replaced with constant and $\rho[G]$ becomes
-a tautology.  A solution is \textit{simple} if $\rho[P]$ is of the
-form $\mathbf{a} \mathbf{x} + b \leq 0$ for every $P$ in
-$\mathrm{dom}(\rho)$.
-
-The problem of solving Horn clauses (represented by $G$) is to find a
-solution of $G$ (and a simple solution if possible).
-
-\section{Simple solution}
-
-We first propose an algorithm to compute a simple solution for the
-given Horn graph $G$ if one exists.  Later we add modifications to the
-algorithm for the case that a simple solution does not exist for $G$.
-In such a case, we transform the input graph $G$ to have a larger
-solution template.
-
-For each vertex $v \in V$, we define
-$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v + b_v \leq 0, C_v)$
-, where
-\begin{itemize}
-\item $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ is a linear inequality that
-  contains coefficient variables $\mathbf{a}_v$ and a constant $b_v$,
-  and
-\item $C_v$ is a set of linear expressions on coefficient variables.
-\end{itemize}
-Intuitively, it denotes the set of inequalities:
-\begin{align*}
-\left\lbrace
- \mathbf{a}_v \mathbf{x}_v + b_v \leq 0 \mid
- \exists \left( \textsc{FV}(C_v)
-  \setminus (\mathbf{a}_v \cup \left\lbrace b_v \right\rbrace
- \right); C_v
-\right\rbrace
-\end{align*}
-For a predicate variable vertex, $\mathbf{a}_v \mathbf{x}_v + b_v \leq
-0$ serves as a predicate template and $C_v$ describes the parameters
-$\mathbf{a}_v$ and $b_v$.
-
-$C_v$ is defined by well-founded induction on $v$ with respect to the
-relation $\rightarrow^+$. Then, $C_v$ can be shown as follows.
-\begin{itemize}
-\item For $v \in V_\Rightarrow$;
-\begin{align*}
-C_v & =
- \bigcup_{\edgel{u}{\theta}{v}} C_u \cup
- \bigcup_{0 < i \leq K}
- \left\lbrace
-  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{v}} \mathbf{a}_{u, \theta^{-1} (i)}
- \right\rbrace \\
- & \hspace{2cm} \cup
- \left\lbrace
-  b_v \leq \sum_{\edgel{u}{\theta}{v}} b_u
- \right\rbrace \cup
- \bigcup_{k_{succ(v)} < i \leq K}
- \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that
-  $v \neq v_{\bot} \wedge \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0$
-  where $\mathbf{a}_0$ and $b_0$ are constants;
-\begin{align*}
-C_v =
-\left\lbrace
- \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
- b_v = \lambda_v b_0
-\right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that $\varphi(v) = P(\mathbf{x}_v)$;
-\begin{align*}
-C_v = & \bigcup_{\edge{u}{v}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{a}_v, b_u = b_v \right\rbrace
-\right)
-\end{align*}
-\item For $v_\bot$;
-\begin{align*}
-C_{v_\bot} = \bigcup_{\edge{u}{v_\bot}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{0}, b_u > 0 \right\rbrace \right)
-\end{align*}
-\end{itemize}
-
-Any model $\sigma$ to the constraint $C_{v_\bot}$ gives a simple
-solution for $G$ if $C_{v_\bot}$ is satisfiable, i.e., we should have:
-
-\begin{quote}
-If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
-\begin{align*}
- \rho = \left\lbrace
-  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \mid
-  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
- \right\rbrace
-\end{align*}
-is a solution of $G$.
-\end{quote}
-
-If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for $G$.
-There are two cases that a simple solution does not exist for a
-predicate variable.
-\begin{itemize}
-\item A predicate variable appears multiple times on left-hand of Horn
-  clauses, and those appearances require different expressions as an
-  assumption.
-\item A predicate variable is implied by a disjunctive Horn clause,
-  which has disjunctions on its left-hand, and a common premise that
-  the disjunctions induce is not strong enough for solving a problem.
-\end{itemize}
-
-The first case occurs in a case like follows.
-\begin{align*}
-% conj.txt ... originated from uc/01.txt
-\left( x \leq 0 \right) \wedge \left( y \leq 0 \right) & \implies P(x,y) \\
-P(x,a) \wedge P(b,y) & \implies x+y \leq 0
-\end{align*}
-In this example, the predicate variable $P$ appears twice on the
-second clause and they require different premises; $x \leq 0$ for the
-first and $y \leq 0$ for the second.  They cannot be represented by
-$\mathbf{ax} + b \leq 0$ template at the same time.
-
-The second case occurs in a case like follows.
-\begin{align*}
-% disj-simp.txt ... originated from 23simp_neg.txt
-\left( x+1 \leq 0 \right) \vee \left( -x+1 \leq 0 \right) & \implies P(x) \\
-P(x,y) \wedge \left( x \leq 0 \right) \wedge \left( -x \leq 0 \right) & \implies \bot
-\end{align*}
-In this example, the predicate variable $P$ is implied by the first
-clause, which is a disjunctive Horn clause.  As a simple solution, $P$
-can take a common consequence that both of $x+1 \leq 0$ and $-x+1 \leq
-0$ induce.  However, such a simple linear expression does not exist
-except for $0 \leq 0$, which is equivalent to $\top$, and it is not
-sufficient to satisfy the second clause.
-
-Both cases can be fixed by enlarging the solution template.
-
-\section {Graph transformation}
-
-If a simple solution does not exist for $G$, we try to give a less
-simpler solution by transforming $G$ and relaxing the constraint
-$C_{v_\bot}$.  This transformation takes place by considering the
-unsatisfiable core $\mathcal{U}$.  An unsatisfiable core $\mathcal{U}$
-is a subset of $C_{v_\bot}$ which is inconsistent itself.  Depending
-on the reason for the absence of a simple solution, the way of the
-transformation differs.
-
-Algorithm~\ref{alg:solveHorn} shows the updated pseudo-code of our
-Horn clause solving algorithm with the graph transformation.  It takes
-an input Horn graph and returns a mapping between predicate variables
-and linear arithmetic formulas.
-
-\newcommand{\solveHorn}{\ensuremath{\mbox{\sc SolveHornGraph}}}
-\newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
-\newcommand{\transGraph}{\ensuremath{\mbox{\sc TransformGraph}}}
-
-\begin{algorithm}
-\caption{$ \solveHorn (G) $}\label{alg:solveHorn}
-\begin{algorithmic}
-\STATE {$v_\bot \gets $ the root node of $G$}
-\STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-\STATE {$T \gets $ initial predicate templateset for $G$}
-\REPEAT
-  \STATE {$C_{v_\bot} \gets \genConstr (G, v_\bot, V_\star)$}
-  \IF {$C_{v_\bot}$ is satisfiable}
-    \STATE {$\sigma \gets $ one model of $C_{v_\bot}$}
-    \STATE {$V_\star \gets V_\star \cup \left\lbrace u \in V_T \mid \forall v \in V_\star; u \leadsto v \right\rbrace$}
-  \ELSE
-    \STATE {$\mathcal{U} \gets $ unsatisfiable core for $C_{v_\bot}$}
-    \STATE {$(G, T) \gets \transGraph (G, T, \mathcal{U})$}
-    \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-  \ENDIF
-\UNTIL {$V_\star = V_T$}
-\RETURN {$\sigma \left( T \right)$}
-\end{algorithmic}
-\end{algorithm}
-\transGraph~procedure determines the unsatisfiable core of the
-constraint and transform the Horn graph $G$ and predicate templates.
-As mentioned earlier, two ways of graph transformation exist.
-
-\paragraph{Conjunction split}
-For the first case previously mentioned, the unsatisfiable core
-$\mathcal{U}$ contains coeffcient constraint expressions with a
-coefficient parameter of certain predicate variable.  If the predicate
-variable $P$ is the cause of unsatisfiability, some constraint
-expressions in the unsatisfiable core have a parameter $a_{P,i}$
-and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is copied to
-$P_1, \ldots, P_n$ so that the different appearances of $P$ on
-left-hand of Horn clauses can take different premises.  The solution
-template of $P$ becomes the conjunction of $P_1, \ldots, P_n$.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-Note that after this duplication of the predicate vertex $P$, in order
-to induce different linear expressions for $P_1, \ldots, P_n$, it may
-be necessary for preceeding predicate variable vertices in $G$ to be
-duplicated in the same manner as $P$. It is because that different
-derivation of $P$'s implication relation may be required to get
-different consequences.
-
-However, succeeding predicate vertices of $P$ in $G$ and other
-vertices than predecessors are not affected because those implications
-that have $P$ in their assumption can selectively choose the premise
-that $P_1, \ldots, P_n$ leads.
-
-Therefore, this graph transformation only affects the predecessors of
-$P$.
-
-\paragraph{Disjunction split}
-For the second case, the unsatisfiable core $\mathcal{U}$ contains
-coefficient constraints that unifies induced linear expressions from
-each term in disjunctions to be the same expression.  Similarly as
-previous, if the predicate variable $P$ is the cause of
-unsatisfiability, some constraint expressions in the unsatisfiable
-core have a parameter $a_{P,i}$ and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is separated to
-$P_1, \ldots, P_n$ each of which is implied by a term, which was
-originally a part of a disjunction to imply $P$.  This intuitively
-enables us to avoid generating a common linear expression that
-disjunctive terms induces, and to propagate each term's premises to
-$P$'s successors instead.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-After this duplication of the predicate vertex $P$, unlike the
-conjunction split, this results in producing a disjunction at
-succeeding predicate variables vertices ($R$ in this case).
-\begin{align*}
-(P_1 \wedge Q) \vee (P_2 \wedge Q) \implies R
-\end{align*}
-Also, predicate variables that appeared with $P$ in the same
-assumption ($Q$ in this case) needs to be duplicated by conjunction,
-so as to obtain different premises that go along with $P_i$.
-Therefore, the effect of disjunction split may reach to neighboring
-predicate vertices.
-
-Consider these two ways of transformation, we show a pseudo-code for
-\transGraph~procedure.  The procedure is given the graph $G$, the
-predicate templateset $T$, and the unsatisfiable core
-$\mathcal{U}$. It returns the transformed graph $G'$ and the new
-templateset $T'$.
-
-\begin{algorithm}
-\caption{$ \transGraph (G, T, \mathcal{U}) $}\label{alg:transGraph}
-\begin{algorithmic}
-% TODO:
-\RETURN {$\left( G', T' \right)$}
-\end{algorithmic}
-\end{algorithm}
-
-\paragraph{Combinatorial explosion in disjunctions}
-Virtually, if multiple disjunctions exists in an input set of Horn
-clauses, a na\"{i}ve way of building a solution is solve all
-combinations of choices out of every disjunction and concatenate each
-expression with conjunctions and/or disjunctions appropriately.  This
-may cause a combinatorial explosion.
-
-Conside example below. $e_i$ represents some conjunctions of linear
-expressions.
-% TODO: Can't we replace e_i with actual expressions?
-\begin{align*}
-(e_1 \vee e_2) \implies P \\ (e_3 \vee e_4) \implies Q \\ P \wedge Q \implies \bot
-\end{align*}
-In this example, a na\"{i}ve algorithm may solve following four
-problems separately.
-\begin{itemize}
-\item $e_1 \implies P_1 \qquad e_3 \implies Q_1 \qquad P_1 \wedge Q_1 \implies \bot$
-\item $e_2 \implies P_2 \qquad e_3 \implies Q_2 \qquad P_2 \wedge Q_2 \implies \bot$
-\item $e_1 \implies P_3 \qquad e_4 \implies Q_3 \qquad P_3 \wedge Q_3 \implies \bot$
-\item $e_2 \implies P_4 \qquad e_4 \implies Q_4 \qquad P_4 \wedge Q_4 \implies \bot$
-\end{itemize}
-Then the solution is built by follows.
-\begin{itemize}
-\item $P = ( P_1 \vee P_2 ) \wedge ( P_3 \vee P_4 )$
-\item $Q = ( P_1 \vee P_3 ) \wedge ( P_2 \vee P_4 )$
-\end{itemize}
-However, this method is not efficient.
-
-\paragraph{Algorithm}
-To avoid this, our algorithm lets a predicate variable $P$ which is
-disjunctively implied on the Horn graph $G$ to have a linear
-expression that is consequences of all terms in the disjunction.  This
-prevents the possibility of choices for disjunctions to be propagated
-to succeeding vertices. The simple linear expression is propagated
-instead.
-
-On the other hand, the algorithm allows a predicate variable $P$ that
-appears on multiple implications or appears multiple times in a single
-implication to take different linear expressions as a premise at
-different appearances.  This is for the algorithm to determine the
-cause of unsatisfiability efficiently.  It is done by copying a
-predicate templates and its parameters' constraint with appropriate
-renaming.
-
-The algorithm first execute this duplication on all predicate
-variables with multiple appearances.  It is the weakest constraint
-that the algorithm makes for the certain $G$.  Then, the algorithm
-gradually reduces the number of vertices that are duplicated by each
-time testing the satisfiability of the constraint.  When the
-constraint bebcome unsatisfiable at the certain iteration, the
-algorithm can determine a predicate vertex which has to be split in a
-conjunctive template.  If all the predicate variable constraints are
-no longer duplicated and the constraint is the strongest, any model
-for such constraint gives a simple solution under the Horn graph $G$
-and the templateset $T$.
-
-Note that duplicating the constraint in a na\"{i}ve way may cause the
-constraint to exponentially grow in size.  Our algorithm prevent it by
-simplifying a predicate vertex's constraint $C_v$ by quantifier
-elimination when $v$ is duplicated in generating linear constraint.
-For a predicate variable vertex $v_P$, variables except for the
-coefficient parameter $\mathbf{a}_P$ and the constant parameter $b_P$
-are eliminated.  Due to the constraints generation process, the
-simplified constraint is also linear and can be efficiently check the
-satisfiability.
-
-We finally show the modified version of pseudo-code for $\genConstr$
-procedure.  It produces the constraint $C_v$ for a given vertex $v$,
-considering constraint duplication and simplification.
-
-\begin{algorithm}
-\caption{$ \genConstr (G, v, V_\star) $}\label{alg:genConstr}
-\begin{algorithmic}
-\STATE {$C \gets \emptyset$}
-\FORALL {$v_\Rightarrow \in preds_G(v)$}
-  \FORALL {$v_p \in preds_G(v_\Rightarrow)$}
-    \IF {$\varphi (v_p)$ is a predicate variable}
-      \STATE {$C \gets C \cup \genConstr (G, v_p, visited)$}
-    \ELSE
-      \STATE { $id$ }
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-traverses the input graph in the topological order on
-$V_\Rightarrow$, and concatenate the constraints from predecessors to
-build the problem constraints.  Each process to build a constraint on
-$\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$ is equivalent
-to generate a linear constraint in interpolation.
-
-\chapter{Experiment}
-\label{chap:experiment}
-
-We incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It
-is specifically designed to verify the correctness of higher-orde
-programs written in a small subset of ML.  It infers dependent types
-for input higher-order programs \cite{conf/ppdp/UnnoK09}.
-
-Our implementation make use of Z3 \cite{conf/tacas/MouraB08} as
-an SMT solver.
-
-For simplicity of the implementation, we assume that the linear
-inequalities in the integer space.  Because of this, our algorithm
-lost the completeness.  By using Motzkin's transpose theorem
-\cite{journals/networks/Rajan90}, it is possible to handle linear
-inequalities in both $\mathbf{ax} + b \leq 0$ and $\mathbf{ax} + b <
-0$.
-
-\section{Interpolating theorem prover}
-
-Our interpolating theorem prover is used as a sub-procedure to
-solve Horn clauses.
-
-\begin{itemize}
-\item CSIsat \cite{conf/cav/BeyerZM08}
-\item FOCI \cite{website/foci}
-\item CLP-Prover \cite{website/clp}
-\end{itemize}
-
-\section{Horn clause solver}
-
-\chapter{Related work}
-\label{chap:related}
-\input{related.tex}
-
-\chapter{Future Work}
-\label{chap:future}
-
-\paragraph{Non-linear extension}
-\cite{conf/cav/DaiXZ13}
-
-\chapter{Conclusion}
-\label{chap:conclusion}
-
-\bibliographystyle{plain}
-\bibliography{./biblio}
-
-\end{document}

### 2013-08-26 article/experiment.tex
Other chapters also separated

--- 2013-08-26 article/thesis.tex
@@
-\documentclass[master,final,12pt]{iscs-thesis}
-
-\etitle{Simple Solutions of Acyclic Horn Clauses for Program Verification}
-\eauthor{Yuto Takei}
-\esupervisor{Naoki Kobayashi}
-
-\jtitle{Lipsum}
-\jauthor{Lipsum}
-\jsupervisor{Lipsum}
-
-\supervisortitle{Professor}
-\date{August 28, 2013}
-
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{stmaryrd}
-\usepackage{syntax}
-\usepackage{url}
-
-\newcommand{\sembrack}[1]{[\![#1]\!}
-\newcommand{\edge}[2]{#1\rightarrow#2}
-\newcommand{\edgel}[3]{#1\xrightarrow{#2}#3}
-
-\begin{document}
-
-\begin{eabstract}
-Automated theorem proving plays a crucial role in modern program
-verification techniques.  In particular, in the context of software
-model checking, the interpolation problem and its generalization
-called \emph{Horn clause solving} have attracted special attention.
-The latter problem asks for a solution for a set of acyclic Horn
-clauses containing unknown predicate variables.  The solution, i.e., a
-substitution for predicate variables, then serves as a candidate of
-program invariants.  However, existing methods are unsatisfactory in
-that they tend to generate too specific a solution, which does not
-correspond to program invariants.
-
-To address the problem above, we propose a new algorithm for solving
-Horn clauses consisting of linear arithmetics and unknown predicate
-variables.  The new algorithm tries to find small predicates, based on
-the hypothesis that program's invariants tend to be simple formulas.
-To this end, our algorithm prepares a template of linear arithmetic
-formulas of a fixed size as a candidate of each unknown predicate, so
-that Horn clause solving is reduced to linear programming.  The
-algorithm iteratively increases the size of the templates until a
-solution is found.  It enables us to construct relatively small
-solutions in a reasonable time compared to existing algorithms.  We
-have implemented and incorporated our Horn clause solver into MoCHi, a
-software model checker for functional programs.  We have evaluated the
-effectiveness of our approach through experiments.
-\end{eabstract}
-\begin{jabstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.  Morbi
-vulputate ornare arcu et malesuada.  Proin sem urna, feugiat et velit
-ac, elementum blandit orci.  Etiam ut neque ac felis laoreet venenatis
-eget quis enim.  Ut aliquam dapibus sem egestas molestie.  Ut sit amet
-porttitor orci, id fermentum dui.  In non dui at ante sagittis
-malesuada at in neque.  Fusce euismod velit quis ante ullamcorper
-qsodales.
-\end{jabstract}
-\maketitle
-
-\begin{acknowledge}
-Thank you.
-\end{acknowledge}
-
-\frontmatter
-\tableofcontents
-
-\mainmatter
-
-\chapter{Introduction}
-
-Automated theorem proving plays an important role in the modern
-program verification.  Particularly in the context of verification
-with model checking techniques, an automated theorem prover is used to
-abstract an infinite set of program states into a finite one so that
-the model checker can virtually explore all possible program execution
-paths.
-
-One of the most common methods of such abstraction is predicate
-abstraction\cite{conf/cav/GrafS97} within CounterExample-Guided
-Abstraction Refinement (CEGAR) framework
-\cite{conf/cav/ClarkeGJLV00,conf/popl/BallR02,conf/popl/HenzingerJMS02}.
-An automated theorem prover finds appropriate predicates for every
-program location.  An execution states at those locations is then
-expressed by a Boolean valuation of the given predicates.  If the
-abstraction is too coarse and a model checker discovers an infeasible
-error execution path in an abstract system, the prover computes an
-additional predicate at those program locations to refine the
-abstraction predicates.
-
-In order to obtain an additional abstraction predicate at program
-locations along the infeasible counterexample, one class of automated
-theorem provers, called an interpolating theorem prover
-\cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07},
-computes a separating predicate called an interpolant between two sets
-of constraints; the one consists of the constraints from the program
-entry point to the location, and the other consists of the constraints
-from the location to the program failure.  Those constraints include
-variable assignments and assertions to be satisfied originated from
-conditional branches.
-
-The computed predicate by an interpolating theorem prover is satisfied
-at the location under the execution in the concrete model along the
-discovered counterexample, and any execution paths which satisfies the
-predicate does not lead to the failure state.  After the refinement,
-the infeasible path is ruled out and no longer found by the model
-checker.
-
-However, existing interpolating theorem provers may return too complex
-a solution which is heavily affected by the constraints of the
-specific path.  It is not favorable in terms of the program
-verification because a complex predicate may not be sufficient to rule
-out similar spurious paths in programs with loops and recursion calls.
-This causes the model checker to discover spurious paths infinitely
-which pass the same program location by loops or recursion calls at
-different numbers of iteration, and the verification process does not
-to terminate.
-
-To avoid this, it is desired to discover an invariant formula for the
-predicate abstraction.  Based on the hypothesis that program's
-invariants tend to be simple formulas, we propose a new interpolating
-algorithm which tries to minimize the number of conjunctions and
-disjunctions in a solution.  The predicate obtained by this algorithm
-may become similar to the invariant, and the number of trials in model
-checking may be reduced.
-
-Additionally, we extend our interpolating algorithm to solve a
-symmetric interpolation problem, in which the algorithm computes
-predicates at multiple program locations along a path at once.  This
-enables the abstraction predicates at multiple locations to be updated
-at the same computation.
-
-Finally, we extend our algorithm to solve Horn clause solving
-problems.  It allows us to obtain predicates which are highly likely
-to be program invariants, by computing the same abstraction predicates
-inside loops and recursion calls on different iterations, and by
-unifying predicates at the same location over multiple paths.  The
-problems are generated by encoding constraints over execution paths
-with unknown predicates which describe functions' constraints between
-parameters and return values.
-
-The rest of the paper is structured as follows.
-Chapter~\ref{chap:interpolation} proposes a new interpolating
-algorithm to obtain simple solutions, and extends it to solve
-symmetric interpolation problems.  Chapter~\ref{chap:horn} describes a
-new Horn clause solving algorithm by extending the previous
-interpolating algorithm.  Chapter~\ref{chap:experiment} describes
-experiments and illustrates the result to confirm the effect of our
-proposed algorithm.  Chapter~\ref{chap:future} gives a discussion of
-our improvements and proposes further improvements to make.
-Chapter~\ref{chap:related} mentions related work done by others.
-Finally, Chapter~\ref{chap:conclusion} reviews the impact of our
-research and brings the conclusion.
-
-\chapter{Interpolation}
-\label{chap:interpolation}
-
-Among various program verification techniques, when software model
-checking with predicate abstraction is adopted, the interpolation is
-used to compute the abstraction predicate along a spurious
-counterexample that a model checker discovered.  It is done by
-computing a Craig interpolant between the strongest postcondition at
-the location and the weakest precondition to the failure point at
-every program location.
-
-However, by the existing interpolating theorem prover, we may not be
-able to obtain a simple solution.  Consider the example below.
-\begin{align*}
-\varphi_1 : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
-\varphi_2 : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
-\end{align*}
-Although one may be able to obtain $x = y$ as the simplest
-interpolation between $\varphi_1$ and $\varphi_2$,
-CSIsat\cite{conf/cav/BeyerZM08} returns
-\[ TODO \]
-as a solution.
-
-We propose an interpolation algorithm which returns as simple a
-solution as possible in terms of the number of conjunctions and
-disjunctions.  Our algorithm focuses on interpolating problems on
-linear arithmetic and it works in a constructive manner.
-
-\section{Example}
-
-Consider the following interpolation problem.
-
-\begin{align*}
-\frac
-{A: (x \leq a) \wedge (a + 1 \leq y)}
-{B: (y \leq b) \wedge (b + 1 \leq x)}
-\end{align*}
-
-Here, the linear arithmetic formulas A and B are inconsistent.  The
-interpolant I for this problem is $x-y+1 \leq 0$, which is implied by
-$A$ and inconsistent with $B$.  The vocabulary of $I$ is $\left\lbrace
-x,y \right\rbrace$ and is over $A$'s vocabulary $\left\lbrace x,y,a
-\right\rbrace$ and $B$'s $\left\lbrace x,y,b \right\rbrace$.
-
-In computing the interpolant, we make a linear constraint of a
-interpolant by applying Farkas's lemma.  First, we assing weight
-parameters to every expression in conjunction form.
-
-\begin{align*}
-\begin{array}{r r r r r r l}
-\lambda_1 : & -a & & +x & & & \leq 0 \\
-\lambda_2 : & a & & & -y & +1 & \leq 0 \\
-\hline
-\lambda_3 : & & -b & & +y & & \leq 0 \\
-\lambda_4 : & & b & -x & & +1 & \leq 0
-\end{array}
-\end{align*}
-
-Because all expressions induce inconsistent as a whole, the sum of
-weighted expressions should satisfy the following constraint.
-
-\begin{align*}
-\begin{array}{r l l}
-- \lambda_1 + \lambda_2 & = 0 \quad & (a) \\
-- \lambda_3 + \lambda_4 & = 0 & (b) \\
-  \lambda_1 - \lambda_4 & = 0 & (x) \\
-- \lambda_2 + \lambda_3 & = 0 & (y) \\
-  \lambda_2 + \lambda_4 & > 0 & (\text{constant}) \\
-\lambda_1, \ldots, \lambda_4 & \geq 0 & (\text{non-negative weight})
-\end{array}
-\end{align*}
-
-With the linear constraint among $\lambda_i$ above, the interpolant is
-represented as follows by the sum of weighted expressions from the
-first formula groups.
-\[ (- \lambda_1 + \lambda_2) a + \lambda_1 x - \lambda_2 y + \lambda_2 \leq 0 \]
-One of the model of the linear constraint is $\lambda_1 = \lambda_2 =
-\lambda_3 = \lambda_4 = 1$, and we obtain $x-y+1 \leq 0$ as a
-solution.
-
-% TODO: Better to mention multiple interpolants possible for one
-% interpolating problem?
-
-\section{Preliminaries}
-
-Our algorithm aims to compute a Craig interpolant over linear
-arithmetic formulas.
-
-\paragraph{Craig Interpolation}
-Given two logical formulas $A$ and $B$ that are inconsistent each
-other, we call a new preposition $I$ a \emph{Craig interpolant}
-\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$
-is implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
-only free variables that appears in both $A$ and $B$.  If $A \wedge B$
-is unsatisfiable, an interpolant $I$ always exists.
-
-Our algorithm handles a linear expression $e$ which is in a form:
-\[ a_1 x_1 + \cdots + a_n x_n + b \leq 0 \]
-This expression contains coefficient parameters $a_1, \ldots, a_n$ and
-a constant parameter $b$.  We treat a special expressions $\top$ as $0
-\leq 0$ and $\bot$ as $1 \leq 0$.  The algorithm input and output is a
-formula $\psi$ that consists of conjunctions and/or disjunctions of
-linear expressions.
-\[ \psi = e | e \wedge e | e \vee e | \top | \bot \]
-
-In the simple interpolation, the algorithm receives two formulas in
-$\psi$ and returns an interpolant in $\psi$.  For the later extension
-to handle a symmetric interpolation problem, the algorithm receives
-multiple formulas in $\psi$ and returns intermediate interpolants in
-$\psi$.  For computing those interpolants over linear arithmetics, we
-use Farkas's Lemma.
-
-\paragraph{Farkas's Lemma on linear inequalities}
-Let a linear inequality $e_i$ be represented as $a_i1 x_1 + \cdots +
-a_im x_m <= a_i0$.  Assuming that $e_1,\cdots,e_n$ implies $e_0$,
-there exists $\lambda_1,\cdots,\lambda_n$ that satisfy $a_0j
-=\sum_(i=1)^n \lambda_i * a_ij (j=0, \ldots, m)$.
-
-For the special case of Farkas's Lemma, we can state that:
-
-Assuming that $e_1,\cdots,e_n$ implies $\bot$, there exists
-$\lambda_1,\cdots,\lambda_n$ that satisfy $???? TODO $.
-
-Our interpolating algorithm make use of the latter version of Farkas's
-Lemma.
-
-\section{Algorithm}
-
-Our algorithm aims to construct a relatively small interpolant in a
-reasonable computation time for any input.  It also aims to obtain a
-common solution among multiple interpolating problems.  When the
-algorithm is used as a sub-routine for solving Horn clause satisfying
-problems that are made while discovering predicates for abstraction
-during the program verification.  These advantages of our algorithm
-make it possible to return a relatively small predicates for those
-problems. To accomplish these aims, the algorithm preserves a set of
-interpolants during the computation, and define and execute operations
-over interpolant sets.
-
-\paragraph{Interpolation between conjunctions}
-We first present an interpolating algorithm between two conjunctive
-sets of linear inequalities $A = \left\lbrace e_1,\ldots,e_m
-\right\rbrace$ and $B = \left\lbrace e_{m+1},\ldots,e_{m+n}
-\right\rbrace$.  We assume that conjunctions of $A$ and $B$ are
-inconsistent.  For later convenience, we call this interpolating
-problem $\left( A, B \right)$.  Then there exists an interpolating
-linear expression $e_\star$ which satisfies
-\begin{align*}
-\left\lbrace e_1,\ldots,e_m \right\rbrace & \vdash e_\star \\
-e_\star & \nvdash \left\lbrace e_{m+1},\ldots,e_{m+n} \right\rbrace
-\end{align*}
-For convenience, we assume that any expression $e_i$ have variables
-$x_1, \ldots, x_a$.
-
-According to the Farkas's lemma, there exists a set of assignmentqs to
-$\lambda_i$, to conclude $\bot$ from a set of linear expressions $A
-\cup B$, namely,
-\begin{align*}
-\exists \lambda_1, \ldots, \lambda_m. \\
-a % TODO: coefficient, constant constraints
-\end{align*}
-
-Then, a linear combination of $A$'s formula with a weight $\lambda_1,
-\ldots, \lambda_m$ becomes an interpolant $I$ between $A$ and $B$.
-Note that any assignments to $\lambda_i$ that satisfy the expressions
-above will let an expression $I$ to be an interpolant.  Considering
-this, it is possible to obtain different interpolants for a
-interpolating problem by computing different assignments to
-$\lambda_i$.
-
-This enables the algorithm to preserve a set of interpolants by a
-linear expression template $a_1 x_1 + \cdots + a_n x_n + b \leq 0$
-with a constraint to describe a possible space of coefficient
-parameters $a_i$ and the constant parameter $b$.
-
-In practical, we may use linear programming solvers to obtain a model
-for $\lambda_i$.  By assigning concrete values to them, we can obtain
-a concrete interpolant.
-
-\paragraph{Computing a common interpolant}
-Next we consider a method to compute a common interpolant across two
-different interpolating problems.  That is, when given two
-interpolating problems $\left(A_1, B_1 \right)$ and $\left(A_2, B_2
-\right)$, the problem is about to compute the common interpolant I,
-which is an interpolant for the first problem and the second one at
-the same time.
-
-This problem is used when solving a following Horn clause solving
-problem, for instance.
-\[
-\implies
-% TODO
-\]
-
-To accomplish this, we simply need to compute the intersection of
-interpolant sets.  Let us call the interpolant space for the problems
-$\left(A_1, B_1 \right)$ and $\left(A_2, B_2 \right)$,
-$\mathcal{I}_1$, and $\mathcal{I}_2$.  In the previous method, a set
-of interpolants was represented by a parameterized linear expression.
-\begin{align*}
-\mathcal{I}_1 = \left\lbrace \mathbf{a}_1 \mathbf{x} \leq b_1 \mid
-C_1 (\mathbf{a}_1, b_1 ) \right\rbrace \\
-\mathcal{I}_2 = \left\lbrace \mathbf{a}_2 \mathbf{x} \leq b_2 \mid
-C_2 (\mathbf{a}_2, b_2 ) \right\rbrace
-\end{align*}
-To compute the intersection between them, it is sufficient to choose
-the same assignments of parameters and let them satisfy the both
-constraints $C_1$ and $C_2$.
-\[ \mathcal{I} = \left\lbrace \mathbf{a} \mathbf{x} \leq b \mid
-C_1(\mathbf{a}, b) \wedge C_2(\mathbf{a}, b) \right\rbrace \]
-
-However, if such an assignment of $\mathbf{a}$ and $b$ does not exist,
-no common solution exists across two problems because $\mathcal{I} =
-\emptyset$.
-
-By applying this approach, it is also possible to compute a common
-interpolant among more than three problems.
-
-\paragraph{Interpolation between logical formulas}
-We now extend the algorithm so that it can handle general logical
-formulas over linear arithmetic including disjunctions of linear
-expressions.  Here the algorithm takes two logical formulas over
-linear expressions as a problem input.  For convenience, we assume
-that they are represented in disjunctive normal form (DNF).  The
-solution for this problem is the interpolant that is a logical formula
-of linear expressions.
-
-First, consider the case of an interpolation problem between $A_1 \vee
-A_2$ and $B$.  When $(A_1 \vee A_2) \wedge B$ is inconsistent, $A_1
-\wedge B$ is necessarily inconsistent.  Let $\mathcal{I}_1$ denote the
-set of interpolants for $(A_1, B)$.  In the same manner, let
-$\mathcal{I}_2$ denote the set of interpolants for $(A_2, B)$.  We may
-naturally obtain an interpolant set by the intersection $\mathcal{I} =
-\mathcal{I}_1 \cup \mathcal{I}_2$ as a solution for $(A_1 \vee A_2,
-B)$.  When the algorithm computes $\mathcal{I}$ by the previous
-method, and it is $\emptyset$, we need to build a less simple solution
-by
-\[ \mathcal{I}' = \left\lbrace \left( I_1 \vee I_2 \right) \mid
-I_1 \in \mathcal{I}_1 \wedge I_2 \in \mathcal{I}_2 \right\rbrace \]
-Any formulas in the set $\mathcal{I}'$ are the interpolant for $(A_1
-\vee A_2, B)$ because
-\begin{align*}
-\forall I \in \mathcal{I}. & A_1 \vdash I \wedge A_2 \vdash I \wedge I \nvdash B \wedge \\
-& FV(I) = FV(A_1) \cup FV(A_2) \cup FV(B)
-\end{align*}
-
-Second, consider the case of an interpolation problem between $A$ and
-$B_1 \wedge B_2$.
-% TODO:
-
-By combining these two ways of interpolant construction, we may be
-able to build a small solution for more general cases like $(A_1,
-\ldots, A_n)$ and $(B_1, \ldots, B_n)$.  In order to obtain a small
-solution, we need to wisely choose the order of combining an
-interpolant set.
-
-\paragraph{Symmetric Interpolation}
-TODO
-
-\chapter{Horn clause solving}
-\label{chap:horn}
-
-We propose a new algorithm for solving recursion-less Horn clauses to
-generate a solution which is as simple as possible.  This enables us
-to obtain appropriate predicates that are suitable for predicate
-abstraction in the program verification techniques.
-
-The construction of the algorithm roughly consists of two parts.
-\begin{itemize}
-\item \textbf{Iterative Constraint Generation} is a process to build a
-  constraint to compute a solution.  It starts from generating the
-  weakest constraint in the sense that the resulting solution would
-  not be simple yet.  The process iteratively strengthen a constraint
-  until the strongest one while preserving the constraint
-  satisfiability.  The strongest constraint gives a simple solution.
-\item \textbf{Problem Restructuring} is a process to restructure the
-  problem to obtain a less simpler solution, if the constraint become
-  unsatisfiable during the iterative constraint generation.  This
-  process considers the unsatisfiable core of the constraint, and
-  wisely enlarge the solution size.
-\end{itemize}
-
-All predicate variables in an input problem are assigned simple
-templates.  Those predicate templates are parameterized.  In the
-iterative constraint generation process, algorithm builds a linear
-constraint that parameters in templates must satisfy.  This is done in
-a constructive manner, and constraints are build in the order of
-implication dependency in Horn clauses.  That is, when an input
-problem has a Horn clause thath has a predicate variable $P$ on its
-left-hand and $Q$ on its right-hand, a constraint to describe
-parameters in $P$'s template are first build before those in $Q$'s
-template.  When all parameters in the whole problem are described, the
-built constraint are evaluated to obtain a model.  Its instanciation
-becomes an appropriate solution for the input problem.
-
-The novelty of our algorithm lies in the incremental constraint
-refinement.  The algorithm tries to solve from the weakest constraint
-that an input problem allows that does not give a simple solution, and
-gradually strengthen the constraint until it reaches to the simplest
-solution that the problem has.  When a constraint becomes
-unsatisfiable during this step, the algorithm analyzes the
-unsatisfiable core and transform the problem to obtain a less simple
-solution.
-
-This construction of the algorithm enables us to compute a relatively
-simpler solution compared to existing algorithms while preserving the
-computation cost in a reasonable time.
-
-\section{Example}
-
-\paragraph {Tree interpolation}
-First, consider the following example.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-y \leq 2 & \implies B(y) \\
-A(x) \wedge B(y) & \implies 3x+2y \leq 10
-\end{align*}
-
-In this example, we would like to compute linear arithmetic formulas
-for the predicate variables $A$ and $B$.  They appear only once on
-both left-hand and right-hand globally in the problem.
-
-We try to obtain a simple solution that is expressed by linear
-expression templates, whose coefficients and constants are
-parameterized.  Assume that they have following predicate templates.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\
-B(y) : c_3 y + c_4 \leq 0
-\end{align*}
-
-By naturally encoding the first Horn clause with Farkas's lemma, we
-obtain a linear constraint
-$ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
-Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will
-make $A(x)$ valid for the first clause.  In the same manner, we obtain
-$ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
-from the second clause, and
-$ c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0 $
-from the third.
-
-One of the models satisfying all the constraints above is
-$( c_1, c_2, c_3, c_4, \lambda_1, \lambda_2, \lambda_3 ) = (3, -3, 2, -4, 3, 2, 1)$.
-By assigning them to the template, we obtain the following solution.
-
-\begin{align*}
-A(x) : 3 x - 3 \leq 0 \\
-B(y) : 2 y - 4 \leq 0
-\end{align*}
-
-\paragraph {DAG problem solving}
-Second, we explain a case that some predicate variables appear
-multiple times on left-hand side in different Horn clauses.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-A(x) \wedge A(y) & \implies x+y \leq 2
-\end{align*}
-
-Assigning a template $A(x) : c_1 x + c_2 \leq 0$, we obtain the same
-constraint for the first clause as the previous example.
-
-\begin{align} \label{eq:constr1}
-\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0
-\end{align}
-
-Sinve we have $A$'s appearance twice on the second clause, it is
-possible for $A$ to have two expressions each of which is used for
-$A(x)$ and $A(y)$.  Although we need to use the same template for
-multiple occurrences in order to compute the simplest solution, it may
-make the whole constraints unsatisfiable if the simplest solution does
-not exist.
-
-Therefore, we generate a weaker constraint for the second clause that
-allows different expressions of $A$ for multiple occurrence.  If the
-whole constraint is satisfiable in the case, we try to solve in a
-stronger constraint to generate a simple solution.
-
-In this specific case, it is virtually equivalent to solve the
-following problem in a weaker constraint.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 1 & \implies A_1(x) \\
-x \leq 1 & \implies A_2(x) \\
-A_1(x) \wedge A_2(y) & \implies x+y \leq 2 \\
-\\
-\text{Templates:} \\
-A_1(x) \text{ : } & c_{1,1} x + c_{2,1} \leq 0 \\
-A_2(x) \text{ : } & c_{1,2} x + c_{2,2} \leq 0 \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-\end{align*}
-
-By duplicating the constraint~\ref{eq:constr1} of $A$ for newly
-created $A_1$ and $A_2$, we obtain the whole constraint as follows.
-
-\begin{align*}
-& \lambda_{1,1} = c_{1,1} \wedge - \lambda_{1,1} \geq c_{2,1} \wedge \lambda_{1,1} \geq 0 \wedge \\
-& \lambda_{1,2} = c_{1,2} \wedge - \lambda_{1,2} \geq c_{2,2} \wedge \lambda_{1,2} \geq 0 \wedge \\
-& c_{1,1} = \lambda_2 \wedge c_{1,2} = \lambda_2 \wedge c_{2,1} + c_{2,2} \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-This constraint is satisfiable and one model is
-\[ ( c_{1,1}, c_{1,2}, c_{2,1}, c_{2,2}, \lambda_{1,1}, \lambda_{1,2}, \lambda_2 ) =
-( 1, -1, 1, -1, 1, 1, 1 ) \] and the solution then becomes
-$A(x) : (x -1 \leq 0) \wedge (x -1 \leq 0)$.
-
-Since the weak constraint is satisfiable, we now refine the constraint
-to obtain the better solution.  That is, simply eliminate the
-duplication to build the following constraint for the original
-template.
-
-\begin{align*}
-& \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \\
-& c_1 = \lambda_2 \wedge c_1 = \lambda_2 \wedge 2 c_2 \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-Solving this constraint simply gives us a solution $A(x) : x -1 \leq 0$.
-
-\paragraph {Disjunctive Horn clause solving}
-Finally, we explain a case that an input set of Horn clauses contain
-disjunctions on left-hand side.
-
-\begin{align*}
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 \vee -x+1 \leq 0 & \implies B(x) \\
-A(x) \wedge B(x) & \implies 1 \leq 0
-\end{align*}
-
-Same as the previous examples, templates for predicate variables are
-prepared.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\ B(x) : c_3 x + c_4 \leq 0
-\end{align*}
-
-The constraints generated are follows.
-
-\begin{align*}
-& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
-& \lambda_3 = c_3 \wedge \lambda_3 \geq c_4 \wedge \lambda_3 \geq 0 \wedge \\
-& - \lambda_4 = c_3 \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
-& c_1 + c_3 = 0 \wedge c_2 + c_4 > 0
-\end{align*}
-
-The constraints here become unsatisfiable, and one possible
-unsatisfiable core is:
-
-\begin{align*}
-\lambda_3 & = c_3 \\
-\lambda_3 & \geq c_4 \\
-- \lambda_4 & = c_3 \\
-\lambda_4 & \geq c_4 \\
-c_2 + c_4 & > 0 \\
-0 & \geq c_2
-\end{align*}
-
-Because the unsatisfiable core contains constraints originating from
-the disjunctions to be expressed in a single term template.  This let
-us determine to relax the template to allow a enlarged solution.
-
-Now we slightly change the problem and the template.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A(x) \wedge B_1(x) & \implies 1 \leq 0 \\
-A(x) \wedge B_2(x) & \implies 1 \leq 0 \\
-\\
-\text{Templates:} \\
-A(x) \text{ : } & c_1 x + c_2 \leq 0 \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-When we construct the constraints, as same as previous, we again
-obtain an unsatisfiable core:
-
-\begin{align*}
-\text{Constraint:} \\
-TODO \\
-\\
-\text{Unsatisfiable core:} \\
-TODO
-\end{align*}
-
-This time, the unsatisfiable core contains the constraints that
-originate from
-
-TODO
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A_i(x) \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A_i(x) \wedge B_j(x) & \implies 1 \leq 0 \quad (i,j \in \left\lbrace 1,2 \right\rbrace ) \\
-\\
-\text{Templates:} \\
-A_i(x) \text{ : } & c_{1,i} x + c_{2,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-Finally, we obtain the following constraint:
-
-TODO
-
-The process also generates a predicate template for every predicate
-variable that the solution will become.  under certain solution
-template.
-
-\section{Preliminaries}
-
-For ease of discussion, our algorithm handles an input set of Horn
-clauses by a \emph{Horn graph} that express the equivalent problem.
-The graph is then used throughout the algorithm to represent the
-problem.  A Horn graph is a labelled directed acyclic graph
-$G=(V,E,\varphi)$.
-\begin{itemize}
-\item $V$ is a union of two disjoint sets of vertices; Horn term
-  vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex $u
-  \in V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
-  denoted as $succ(u)$.
-\item $E$ is a set of labelled edges, which are expressed in the form
-  $(u,v,\theta)$ where $\theta$ is a finite map from variables to
-  variables.
-\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
-  term vertices, where L is a set of elements of a predicate variable
-  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
-  inequality $a_1 x_1 + \cdots + a_n x_n + b \leq 0$.
-\end{itemize}
-There are two kinds of edges in a Horn graph;
-\begin{itemize}
-\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
-  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
-\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
-  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
-  empty.
-\end{itemize}
-
-We define the relation $\leadsto$ over $V_T$ as follows.
-\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow \exists a \in V_\Rightarrow;
-\edgel{u}{\theta}{a} \wedge \edge{a}{v} \] We may omit $\theta$ and
-write $u \leadsto v$ for convenience in later discussions. By the
-assumption that $G$ is acyclic, both $\rightarrow^+$ and $\leadsto^+$ are
-well-founded relations. $v_\bot$ is the greatest element with respect
-to both $\rightarrow^+$ and $\leadsto^+$.
-
-The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
-Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$.
-\begin{align*}
-\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(succ(v)) \\
-\llbracket G \rrbracket & = \bigwedge_{v \in V_\Rightarrow} \mathcal{HC}_v
-\end{align*}
-We implicitly assume each $\mathcal{HC}_v$ is universally quantified
-by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.
-
-Horn term vertices with predicate variable labels are unique each
-other that any of them do not have the same variable on their labels.
-From now on, we may denote a Horn term vertex with a predicate
-variable label by its predicate variable name if no ambiguity is
-present.
-
-Without loss of generality, we assume the existence of the \emph{root
-  vertex} $v_\bot \in V_T$ with linear expression labels $1 \leq 0$,
-meaning $\bot$.  This is because if we have a Horn clause $P \implies
-e$ which $e$ is a linear expression, we can convert it to $P \wedge
-\neg e \implies 1 \leq 0$.  We also assume that other vertices with
-linear expression labels have no incoming edges to them.
-
-For simplicity of discussion, we define \emph{term arity} $k_v$ for
-every Horn term vertex $v \in V_T$ with incoming edges.
-\begin{align*}
-k_v =
-\begin{cases}
-\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
-0 & \mbox{if } \varphi(v) = \bot
-\end{cases}
-\end{align*}
-We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.
-
-A variable mapping label $\theta$ on edges is used to map the index of
-one predicate variable to another. (TODO?)
-
-(Example?)
-
-The algorithm assigns a \emph{predicate template} for all predicate
-variables in $G$. A \emph{predicate templateset} is a set of mappings from
-a predicate variable to a predicate template.  A template for a
-predicate variable $P(x_1, \ldots, x_n)$ is a formula $\psi$
-constructed from the following definition.
-\begin{align*}
-\psi ::= & a_1 x_1 + \cdots + a_n x_n + b \leq 0 \mid \\
-& \psi \wedge \psi \mid \psi \vee \psi
-\end{align*}
-A linear expression in the definition above may also be represented in
-a equivalent vector form $\mathbf{a} \mathbf{x} + b \leq 0$ for short.
-Parameters $\mathbf{a}$ and $b$ are specified by a linear constraint.
-
-The solution of $G$ is a map $\rho$ from predicate variables to linear
-predicates of the form $\lambda x_1, \ldots ,x_n. \psi$ where
-parameters in $\psi$ are replaced with constant and $\rho[G]$ becomes
-a tautology.  A solution is \textit{simple} if $\rho[P]$ is of the
-form $\mathbf{a} \mathbf{x} + b \leq 0$ for every $P$ in
-$\mathrm{dom}(\rho)$.
-
-The problem of solving Horn clauses (represented by $G$) is to find a
-solution of $G$ (and a simple solution if possible).
-
-\section{Simple solution}
-
-We first propose an algorithm to compute a simple solution for the
-given Horn graph $G$ if one exists.  Later we add modifications to the
-algorithm for the case that a simple solution does not exist for $G$.
-In such a case, we transform the input graph $G$ to have a larger
-solution template.
-
-For each vertex $v \in V$, we define
-$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v + b_v \leq 0, C_v)$
-, where
-\begin{itemize}
-\item $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ is a linear inequality that
-  contains coefficient variables $\mathbf{a}_v$ and a constant $b_v$,
-  and
-\item $C_v$ is a set of linear expressions on coefficient variables.
-\end{itemize}
-Intuitively, it denotes the set of inequalities:
-\begin{align*}
-\left\lbrace
- \mathbf{a}_v \mathbf{x}_v + b_v \leq 0 \mid
- \exists \left( \textsc{FV}(C_v)
-  \setminus (\mathbf{a}_v \cup \left\lbrace b_v \right\rbrace
- \right); C_v
-\right\rbrace
-\end{align*}
-For a predicate variable vertex, $\mathbf{a}_v \mathbf{x}_v + b_v \leq
-0$ serves as a predicate template and $C_v$ describes the parameters
-$\mathbf{a}_v$ and $b_v$.
-
-$C_v$ is defined by well-founded induction on $v$ with respect to the
-relation $\rightarrow^+$. Then, $C_v$ can be shown as follows.
-\begin{itemize}
-\item For $v \in V_\Rightarrow$;
-\begin{align*}
-C_v & =
- \bigcup_{\edgel{u}{\theta}{v}} C_u \cup
- \bigcup_{0 < i \leq K}
- \left\lbrace
-  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{v}} \mathbf{a}_{u, \theta^{-1} (i)}
- \right\rbrace \\
- & \hspace{2cm} \cup
- \left\lbrace
-  b_v \leq \sum_{\edgel{u}{\theta}{v}} b_u
- \right\rbrace \cup
- \bigcup_{k_{succ(v)} < i \leq K}
- \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that
-  $v \neq v_{\bot} \wedge \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0$
-  where $\mathbf{a}_0$ and $b_0$ are constants;
-\begin{align*}
-C_v =
-\left\lbrace
- \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
- b_v = \lambda_v b_0
-\right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that $\varphi(v) = P(\mathbf{x}_v)$;
-\begin{align*}
-C_v = & \bigcup_{\edge{u}{v}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{a}_v, b_u = b_v \right\rbrace
-\right)
-\end{align*}
-\item For $v_\bot$;
-\begin{align*}
-C_{v_\bot} = \bigcup_{\edge{u}{v_\bot}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{0}, b_u > 0 \right\rbrace \right)
-\end{align*}
-\end{itemize}
-
-Any model $\sigma$ to the constraint $C_{v_\bot}$ gives a simple
-solution for $G$ if $C_{v_\bot}$ is satisfiable, i.e., we should have:
-
-\begin{quote}
-If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
-\begin{align*}
- \rho = \left\lbrace
-  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \mid
-  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
- \right\rbrace
-\end{align*}
-is a solution of $G$.
-\end{quote}
-
-If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for $G$.
-There are two cases that a simple solution does not exist for a
-predicate variable.
-\begin{itemize}
-\item A predicate variable appears multiple times on left-hand of Horn
-  clauses, and those appearances require different expressions as an
-  assumption.
-\item A predicate variable is implied by a disjunctive Horn clause,
-  which has disjunctions on its left-hand, and a common premise that
-  the disjunctions induce is not strong enough for solving a problem.
-\end{itemize}
-
-The first case occurs in a case like follows.
-\begin{align*}
-% conj.txt ... originated from uc/01.txt
-\left( x \leq 0 \right) \wedge \left( y \leq 0 \right) & \implies P(x,y) \\
-P(x,a) \wedge P(b,y) & \implies x+y \leq 0
-\end{align*}
-In this example, the predicate variable $P$ appears twice on the
-second clause and they require different premises; $x \leq 0$ for the
-first and $y \leq 0$ for the second.  They cannot be represented by
-$\mathbf{ax} + b \leq 0$ template at the same time.
-
-The second case occurs in a case like follows.
-\begin{align*}
-% disj-simp.txt ... originated from 23simp_neg.txt
-\left( x+1 \leq 0 \right) \vee \left( -x+1 \leq 0 \right) & \implies P(x) \\
-P(x,y) \wedge \left( x \leq 0 \right) \wedge \left( -x \leq 0 \right) & \implies \bot
-\end{align*}
-In this example, the predicate variable $P$ is implied by the first
-clause, which is a disjunctive Horn clause.  As a simple solution, $P$
-can take a common consequence that both of $x+1 \leq 0$ and $-x+1 \leq
-0$ induce.  However, such a simple linear expression does not exist
-except for $0 \leq 0$, which is equivalent to $\top$, and it is not
-sufficient to satisfy the second clause.
-
-Both cases can be fixed by enlarging the solution template.
-
-\section {Graph transformation}
-
-If a simple solution does not exist for $G$, we try to give a less
-simpler solution by transforming $G$ and relaxing the constraint
-$C_{v_\bot}$.  This transformation takes place by considering the
-unsatisfiable core $\mathcal{U}$.  An unsatisfiable core $\mathcal{U}$
-is a subset of $C_{v_\bot}$ which is inconsistent itself.  Depending
-on the reason for the absence of a simple solution, the way of the
-transformation differs.
-
-Algorithm~\ref{alg:solveHorn} shows the updated pseudo-code of our
-Horn clause solving algorithm with the graph transformation.  It takes
-an input Horn graph and returns a mapping between predicate variables
-and linear arithmetic formulas.
-
-\newcommand{\solveHorn}{\ensuremath{\mbox{\sc SolveHornGraph}}}
-\newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
-\newcommand{\transGraph}{\ensuremath{\mbox{\sc TransformGraph}}}
-
-\begin{algorithm}
-\caption{$ \solveHorn (G) $}\label{alg:solveHorn}
-\begin{algorithmic}
-\STATE {$v_\bot \gets $ the root node of $G$}
-\STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-\STATE {$T \gets $ initial predicate templateset for $G$}
-\REPEAT
-  \STATE {$C_{v_\bot} \gets \genConstr (G, v_\bot, V_\star)$}
-  \IF {$C_{v_\bot}$ is satisfiable}
-    \STATE {$\sigma \gets $ one model of $C_{v_\bot}$}
-    \STATE {$V_\star \gets V_\star \cup \left\lbrace u \in V_T \mid \forall v \in V_\star; u \leadsto v \right\rbrace$}
-  \ELSE
-    \STATE {$\mathcal{U} \gets $ unsatisfiable core for $C_{v_\bot}$}
-    \STATE {$(G, T) \gets \transGraph (G, T, \mathcal{U})$}
-    \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-  \ENDIF
-\UNTIL {$V_\star = V_T$}
-\RETURN {$\sigma \left( T \right)$}
-\end{algorithmic}
-\end{algorithm}
-\transGraph~procedure determines the unsatisfiable core of the
-constraint and transform the Horn graph $G$ and predicate templates.
-As mentioned earlier, two ways of graph transformation exist.
-
-\paragraph{Conjunction split}
-For the first case previously mentioned, the unsatisfiable core
-$\mathcal{U}$ contains coeffcient constraint expressions with a
-coefficient parameter of certain predicate variable.  If the predicate
-variable $P$ is the cause of unsatisfiability, some constraint
-expressions in the unsatisfiable core have a parameter $a_{P,i}$
-and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is copied to
-$P_1, \ldots, P_n$ so that the different appearances of $P$ on
-left-hand of Horn clauses can take different premises.  The solution
-template of $P$ becomes the conjunction of $P_1, \ldots, P_n$.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-Note that after this duplication of the predicate vertex $P$, in order
-to induce different linear expressions for $P_1, \ldots, P_n$, it may
-be necessary for preceeding predicate variable vertices in $G$ to be
-duplicated in the same manner as $P$. It is because that different
-derivation of $P$'s implication relation may be required to get
-different consequences.
-
-However, succeeding predicate vertices of $P$ in $G$ and other
-vertices than predecessors are not affected because those implications
-that have $P$ in their assumption can selectively choose the premise
-that $P_1, \ldots, P_n$ leads.
-
-Therefore, this graph transformation only affects the predecessors of
-$P$.
-
-\paragraph{Disjunction split}
-For the second case, the unsatisfiable core $\mathcal{U}$ contains
-coefficient constraints that unifies induced linear expressions from
-each term in disjunctions to be the same expression.  Similarly as
-previous, if the predicate variable $P$ is the cause of
-unsatisfiability, some constraint expressions in the unsatisfiable
-core have a parameter $a_{P,i}$ and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is separated to
-$P_1, \ldots, P_n$ each of which is implied by a term, which was
-originally a part of a disjunction to imply $P$.  This intuitively
-enables us to avoid generating a common linear expression that
-disjunctive terms induces, and to propagate each term's premises to
-$P$'s successors instead.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-After this duplication of the predicate vertex $P$, unlike the
-conjunction split, this results in producing a disjunction at
-succeeding predicate variables vertices ($R$ in this case).
-\begin{align*}
-(P_1 \wedge Q) \vee (P_2 \wedge Q) \implies R
-\end{align*}
-Also, predicate variables that appeared with $P$ in the same
-assumption ($Q$ in this case) needs to be duplicated by conjunction,
-so as to obtain different premises that go along with $P_i$.
-Therefore, the effect of disjunction split may reach to neighboring
-predicate vertices.
-
-Consider these two ways of transformation, we show a pseudo-code for
-\transGraph~procedure.  The procedure is given the graph $G$, the
-predicate templateset $T$, and the unsatisfiable core
-$\mathcal{U}$. It returns the transformed graph $G'$ and the new
-templateset $T'$.
-
-\begin{algorithm}
-\caption{$ \transGraph (G, T, \mathcal{U}) $}\label{alg:transGraph}
-\begin{algorithmic}
-% TODO:
-\RETURN {$\left( G', T' \right)$}
-\end{algorithmic}
-\end{algorithm}
-
-\paragraph{Combinatorial explosion in disjunctions}
-Virtually, if multiple disjunctions exists in an input set of Horn
-clauses, a na\"{i}ve way of building a solution is solve all
-combinations of choices out of every disjunction and concatenate each
-expression with conjunctions and/or disjunctions appropriately.  This
-may cause a combinatorial explosion.
-
-Conside example below. $e_i$ represents some conjunctions of linear
-expressions.
-% TODO: Can't we replace e_i with actual expressions?
-\begin{align*}
-(e_1 \vee e_2) \implies P \\ (e_3 \vee e_4) \implies Q \\ P \wedge Q \implies \bot
-\end{align*}
-In this example, a na\"{i}ve algorithm may solve following four
-problems separately.
-\begin{itemize}
-\item $e_1 \implies P_1 \qquad e_3 \implies Q_1 \qquad P_1 \wedge Q_1 \implies \bot$
-\item $e_2 \implies P_2 \qquad e_3 \implies Q_2 \qquad P_2 \wedge Q_2 \implies \bot$
-\item $e_1 \implies P_3 \qquad e_4 \implies Q_3 \qquad P_3 \wedge Q_3 \implies \bot$
-\item $e_2 \implies P_4 \qquad e_4 \implies Q_4 \qquad P_4 \wedge Q_4 \implies \bot$
-\end{itemize}
-Then the solution is built by follows.
-\begin{itemize}
-\item $P = ( P_1 \vee P_2 ) \wedge ( P_3 \vee P_4 )$
-\item $Q = ( P_1 \vee P_3 ) \wedge ( P_2 \vee P_4 )$
-\end{itemize}
-However, this method is not efficient.
-
-\paragraph{Algorithm}
-To avoid this, our algorithm lets a predicate variable $P$ which is
-disjunctively implied on the Horn graph $G$ to have a linear
-expression that is consequences of all terms in the disjunction.  This
-prevents the possibility of choices for disjunctions to be propagated
-to succeeding vertices. The simple linear expression is propagated
-instead.
-
-On the other hand, the algorithm allows a predicate variable $P$ that
-appears on multiple implications or appears multiple times in a single
-implication to take different linear expressions as a premise at
-different appearances.  This is for the algorithm to determine the
-cause of unsatisfiability efficiently.  It is done by copying a
-predicate templates and its parameters' constraint with appropriate
-renaming.
-
-The algorithm first execute this duplication on all predicate
-variables with multiple appearances.  It is the weakest constraint
-that the algorithm makes for the certain $G$.  Then, the algorithm
-gradually reduces the number of vertices that are duplicated by each
-time testing the satisfiability of the constraint.  When the
-constraint bebcome unsatisfiable at the certain iteration, the
-algorithm can determine a predicate vertex which has to be split in a
-conjunctive template.  If all the predicate variable constraints are
-no longer duplicated and the constraint is the strongest, any model
-for such constraint gives a simple solution under the Horn graph $G$
-and the templateset $T$.
-
-Note that duplicating the constraint in a na\"{i}ve way may cause the
-constraint to exponentially grow in size.  Our algorithm prevent it by
-simplifying a predicate vertex's constraint $C_v$ by quantifier
-elimination when $v$ is duplicated in generating linear constraint.
-For a predicate variable vertex $v_P$, variables except for the
-coefficient parameter $\mathbf{a}_P$ and the constant parameter $b_P$
-are eliminated.  Due to the constraints generation process, the
-simplified constraint is also linear and can be efficiently check the
-satisfiability.
-
-We finally show the modified version of pseudo-code for $\genConstr$
-procedure.  It produces the constraint $C_v$ for a given vertex $v$,
-considering constraint duplication and simplification.
-
-\begin{algorithm}
-\caption{$ \genConstr (G, v, V_\star) $}\label{alg:genConstr}
-\begin{algorithmic}
-\STATE {$C \gets \emptyset$}
-\FORALL {$v_\Rightarrow \in preds_G(v)$}
-  \FORALL {$v_p \in preds_G(v_\Rightarrow)$}
-    \IF {$\varphi (v_p)$ is a predicate variable}
-      \STATE {$C \gets C \cup \genConstr (G, v_p, visited)$}
-    \ELSE
-      \STATE { $id$ }
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-traverses the input graph in the topological order on
-$V_\Rightarrow$, and concatenate the constraints from predecessors to
-build the problem constraints.  Each process to build a constraint on
-$\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$ is equivalent
-to generate a linear constraint in interpolation.
-
 \chapter{Experiment}
@@
 
-Our interpolating theorem prover is used as a sub-procedure to
-solve Horn clauses.
+Our interpolating theorem prover is used as a sub-procedure to solve
+Horn clauses.
 
@@
 \section{Horn clause solver}
-
-\chapter{Related work}
-\label{chap:related}
-\input{related.tex}
-
-\chapter{Future Work}
-\label{chap:future}
-
-\paragraph{Non-linear extension}
-\cite{conf/cav/DaiXZ13}
-
-\chapter{Conclusion}
-\label{chap:conclusion}
-
-\bibliographystyle{plain}
-\bibliography{./biblio}
-
-\end{document}

### 2013-08-26 article/future.tex
Other chapters also separated

--- 2013-08-26 article/thesis.tex
@@
-\documentclass[master,final,12pt]{iscs-thesis}
-
-\etitle{Simple Solutions of Acyclic Horn Clauses for Program Verification}
-\eauthor{Yuto Takei}
-\esupervisor{Naoki Kobayashi}
-
-\jtitle{Lipsum}
-\jauthor{Lipsum}
-\jsupervisor{Lipsum}
-
-\supervisortitle{Professor}
-\date{August 28, 2013}
-
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{stmaryrd}
-\usepackage{syntax}
-\usepackage{url}
-
-\newcommand{\sembrack}[1]{[\![#1]\!}
-\newcommand{\edge}[2]{#1\rightarrow#2}
-\newcommand{\edgel}[3]{#1\xrightarrow{#2}#3}
-
-\begin{document}
-
-\begin{eabstract}
-Automated theorem proving plays a crucial role in modern program
-verification techniques.  In particular, in the context of software
-model checking, the interpolation problem and its generalization
-called \emph{Horn clause solving} have attracted special attention.
-The latter problem asks for a solution for a set of acyclic Horn
-clauses containing unknown predicate variables.  The solution, i.e., a
-substitution for predicate variables, then serves as a candidate of
-program invariants.  However, existing methods are unsatisfactory in
-that they tend to generate too specific a solution, which does not
-correspond to program invariants.
-
-To address the problem above, we propose a new algorithm for solving
-Horn clauses consisting of linear arithmetics and unknown predicate
-variables.  The new algorithm tries to find small predicates, based on
-the hypothesis that program's invariants tend to be simple formulas.
-To this end, our algorithm prepares a template of linear arithmetic
-formulas of a fixed size as a candidate of each unknown predicate, so
-that Horn clause solving is reduced to linear programming.  The
-algorithm iteratively increases the size of the templates until a
-solution is found.  It enables us to construct relatively small
-solutions in a reasonable time compared to existing algorithms.  We
-have implemented and incorporated our Horn clause solver into MoCHi, a
-software model checker for functional programs.  We have evaluated the
-effectiveness of our approach through experiments.
-\end{eabstract}
-\begin{jabstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.  Morbi
-vulputate ornare arcu et malesuada.  Proin sem urna, feugiat et velit
-ac, elementum blandit orci.  Etiam ut neque ac felis laoreet venenatis
-eget quis enim.  Ut aliquam dapibus sem egestas molestie.  Ut sit amet
-porttitor orci, id fermentum dui.  In non dui at ante sagittis
-malesuada at in neque.  Fusce euismod velit quis ante ullamcorper
-qsodales.
-\end{jabstract}
-\maketitle
-
-\begin{acknowledge}
-Thank you.
-\end{acknowledge}
-
-\frontmatter
-\tableofcontents
-
-\mainmatter
-
-\chapter{Introduction}
-
-Automated theorem proving plays an important role in the modern
-program verification.  Particularly in the context of verification
-with model checking techniques, an automated theorem prover is used to
-abstract an infinite set of program states into a finite one so that
-the model checker can virtually explore all possible program execution
-paths.
-
-One of the most common methods of such abstraction is predicate
-abstraction\cite{conf/cav/GrafS97} within CounterExample-Guided
-Abstraction Refinement (CEGAR) framework
-\cite{conf/cav/ClarkeGJLV00,conf/popl/BallR02,conf/popl/HenzingerJMS02}.
-An automated theorem prover finds appropriate predicates for every
-program location.  An execution states at those locations is then
-expressed by a Boolean valuation of the given predicates.  If the
-abstraction is too coarse and a model checker discovers an infeasible
-error execution path in an abstract system, the prover computes an
-additional predicate at those program locations to refine the
-abstraction predicates.
-
-In order to obtain an additional abstraction predicate at program
-locations along the infeasible counterexample, one class of automated
-theorem provers, called an interpolating theorem prover
-\cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07},
-computes a separating predicate called an interpolant between two sets
-of constraints; the one consists of the constraints from the program
-entry point to the location, and the other consists of the constraints
-from the location to the program failure.  Those constraints include
-variable assignments and assertions to be satisfied originated from
-conditional branches.
-
-The computed predicate by an interpolating theorem prover is satisfied
-at the location under the execution in the concrete model along the
-discovered counterexample, and any execution paths which satisfies the
-predicate does not lead to the failure state.  After the refinement,
-the infeasible path is ruled out and no longer found by the model
-checker.
-
-However, existing interpolating theorem provers may return too complex
-a solution which is heavily affected by the constraints of the
-specific path.  It is not favorable in terms of the program
-verification because a complex predicate may not be sufficient to rule
-out similar spurious paths in programs with loops and recursion calls.
-This causes the model checker to discover spurious paths infinitely
-which pass the same program location by loops or recursion calls at
-different numbers of iteration, and the verification process does not
-to terminate.
-
-To avoid this, it is desired to discover an invariant formula for the
-predicate abstraction.  Based on the hypothesis that program's
-invariants tend to be simple formulas, we propose a new interpolating
-algorithm which tries to minimize the number of conjunctions and
-disjunctions in a solution.  The predicate obtained by this algorithm
-may become similar to the invariant, and the number of trials in model
-checking may be reduced.
-
-Additionally, we extend our interpolating algorithm to solve a
-symmetric interpolation problem, in which the algorithm computes
-predicates at multiple program locations along a path at once.  This
-enables the abstraction predicates at multiple locations to be updated
-at the same computation.
-
-Finally, we extend our algorithm to solve Horn clause solving
-problems.  It allows us to obtain predicates which are highly likely
-to be program invariants, by computing the same abstraction predicates
-inside loops and recursion calls on different iterations, and by
-unifying predicates at the same location over multiple paths.  The
-problems are generated by encoding constraints over execution paths
-with unknown predicates which describe functions' constraints between
-parameters and return values.
-
-The rest of the paper is structured as follows.
-Chapter~\ref{chap:interpolation} proposes a new interpolating
-algorithm to obtain simple solutions, and extends it to solve
-symmetric interpolation problems.  Chapter~\ref{chap:horn} describes a
-new Horn clause solving algorithm by extending the previous
-interpolating algorithm.  Chapter~\ref{chap:experiment} describes
-experiments and illustrates the result to confirm the effect of our
-proposed algorithm.  Chapter~\ref{chap:future} gives a discussion of
-our improvements and proposes further improvements to make.
-Chapter~\ref{chap:related} mentions related work done by others.
-Finally, Chapter~\ref{chap:conclusion} reviews the impact of our
-research and brings the conclusion.
-
-\chapter{Interpolation}
-\label{chap:interpolation}
-
-Among various program verification techniques, when software model
-checking with predicate abstraction is adopted, the interpolation is
-used to compute the abstraction predicate along a spurious
-counterexample that a model checker discovered.  It is done by
-computing a Craig interpolant between the strongest postcondition at
-the location and the weakest precondition to the failure point at
-every program location.
-
-However, by the existing interpolating theorem prover, we may not be
-able to obtain a simple solution.  Consider the example below.
-\begin{align*}
-\varphi_1 : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
-\varphi_2 : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
-\end{align*}
-Although one may be able to obtain $x = y$ as the simplest
-interpolation between $\varphi_1$ and $\varphi_2$,
-CSIsat\cite{conf/cav/BeyerZM08} returns
-\[ TODO \]
-as a solution.
-
-We propose an interpolation algorithm which returns as simple a
-solution as possible in terms of the number of conjunctions and
-disjunctions.  Our algorithm focuses on interpolating problems on
-linear arithmetic and it works in a constructive manner.
-
-\section{Example}
-
-Consider the following interpolation problem.
-
-\begin{align*}
-\frac
-{A: (x \leq a) \wedge (a + 1 \leq y)}
-{B: (y \leq b) \wedge (b + 1 \leq x)}
-\end{align*}
-
-Here, the linear arithmetic formulas A and B are inconsistent.  The
-interpolant I for this problem is $x-y+1 \leq 0$, which is implied by
-$A$ and inconsistent with $B$.  The vocabulary of $I$ is $\left\lbrace
-x,y \right\rbrace$ and is over $A$'s vocabulary $\left\lbrace x,y,a
-\right\rbrace$ and $B$'s $\left\lbrace x,y,b \right\rbrace$.
-
-In computing the interpolant, we make a linear constraint of a
-interpolant by applying Farkas's lemma.  First, we assing weight
-parameters to every expression in conjunction form.
-
-\begin{align*}
-\begin{array}{r r r r r r l}
-\lambda_1 : & -a & & +x & & & \leq 0 \\
-\lambda_2 : & a & & & -y & +1 & \leq 0 \\
-\hline
-\lambda_3 : & & -b & & +y & & \leq 0 \\
-\lambda_4 : & & b & -x & & +1 & \leq 0
-\end{array}
-\end{align*}
-
-Because all expressions induce inconsistent as a whole, the sum of
-weighted expressions should satisfy the following constraint.
-
-\begin{align*}
-\begin{array}{r l l}
-- \lambda_1 + \lambda_2 & = 0 \quad & (a) \\
-- \lambda_3 + \lambda_4 & = 0 & (b) \\
-  \lambda_1 - \lambda_4 & = 0 & (x) \\
-- \lambda_2 + \lambda_3 & = 0 & (y) \\
-  \lambda_2 + \lambda_4 & > 0 & (\text{constant}) \\
-\lambda_1, \ldots, \lambda_4 & \geq 0 & (\text{non-negative weight})
-\end{array}
-\end{align*}
-
-With the linear constraint among $\lambda_i$ above, the interpolant is
-represented as follows by the sum of weighted expressions from the
-first formula groups.
-\[ (- \lambda_1 + \lambda_2) a + \lambda_1 x - \lambda_2 y + \lambda_2 \leq 0 \]
-One of the model of the linear constraint is $\lambda_1 = \lambda_2 =
-\lambda_3 = \lambda_4 = 1$, and we obtain $x-y+1 \leq 0$ as a
-solution.
-
-% TODO: Better to mention multiple interpolants possible for one
-% interpolating problem?
-
-\section{Preliminaries}
-
-Our algorithm aims to compute a Craig interpolant over linear
-arithmetic formulas.
-
-\paragraph{Craig Interpolation}
-Given two logical formulas $A$ and $B$ that are inconsistent each
-other, we call a new preposition $I$ a \emph{Craig interpolant}
-\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$
-is implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
-only free variables that appears in both $A$ and $B$.  If $A \wedge B$
-is unsatisfiable, an interpolant $I$ always exists.
-
-Our algorithm handles a linear expression $e$ which is in a form:
-\[ a_1 x_1 + \cdots + a_n x_n + b \leq 0 \]
-This expression contains coefficient parameters $a_1, \ldots, a_n$ and
-a constant parameter $b$.  We treat a special expressions $\top$ as $0
-\leq 0$ and $\bot$ as $1 \leq 0$.  The algorithm input and output is a
-formula $\psi$ that consists of conjunctions and/or disjunctions of
-linear expressions.
-\[ \psi = e | e \wedge e | e \vee e | \top | \bot \]
-
-In the simple interpolation, the algorithm receives two formulas in
-$\psi$ and returns an interpolant in $\psi$.  For the later extension
-to handle a symmetric interpolation problem, the algorithm receives
-multiple formulas in $\psi$ and returns intermediate interpolants in
-$\psi$.  For computing those interpolants over linear arithmetics, we
-use Farkas's Lemma.
-
-\paragraph{Farkas's Lemma on linear inequalities}
-Let a linear inequality $e_i$ be represented as $a_i1 x_1 + \cdots +
-a_im x_m <= a_i0$.  Assuming that $e_1,\cdots,e_n$ implies $e_0$,
-there exists $\lambda_1,\cdots,\lambda_n$ that satisfy $a_0j
-=\sum_(i=1)^n \lambda_i * a_ij (j=0, \ldots, m)$.
-
-For the special case of Farkas's Lemma, we can state that:
-
-Assuming that $e_1,\cdots,e_n$ implies $\bot$, there exists
-$\lambda_1,\cdots,\lambda_n$ that satisfy $???? TODO $.
-
-Our interpolating algorithm make use of the latter version of Farkas's
-Lemma.
-
-\section{Algorithm}
-
-Our algorithm aims to construct a relatively small interpolant in a
-reasonable computation time for any input.  It also aims to obtain a
-common solution among multiple interpolating problems.  When the
-algorithm is used as a sub-routine for solving Horn clause satisfying
-problems that are made while discovering predicates for abstraction
-during the program verification.  These advantages of our algorithm
-make it possible to return a relatively small predicates for those
-problems. To accomplish these aims, the algorithm preserves a set of
-interpolants during the computation, and define and execute operations
-over interpolant sets.
-
-\paragraph{Interpolation between conjunctions}
-We first present an interpolating algorithm between two conjunctive
-sets of linear inequalities $A = \left\lbrace e_1,\ldots,e_m
-\right\rbrace$ and $B = \left\lbrace e_{m+1},\ldots,e_{m+n}
-\right\rbrace$.  We assume that conjunctions of $A$ and $B$ are
-inconsistent.  For later convenience, we call this interpolating
-problem $\left( A, B \right)$.  Then there exists an interpolating
-linear expression $e_\star$ which satisfies
-\begin{align*}
-\left\lbrace e_1,\ldots,e_m \right\rbrace & \vdash e_\star \\
-e_\star & \nvdash \left\lbrace e_{m+1},\ldots,e_{m+n} \right\rbrace
-\end{align*}
-For convenience, we assume that any expression $e_i$ have variables
-$x_1, \ldots, x_a$.
-
-According to the Farkas's lemma, there exists a set of assignmentqs to
-$\lambda_i$, to conclude $\bot$ from a set of linear expressions $A
-\cup B$, namely,
-\begin{align*}
-\exists \lambda_1, \ldots, \lambda_m. \\
-a % TODO: coefficient, constant constraints
-\end{align*}
-
-Then, a linear combination of $A$'s formula with a weight $\lambda_1,
-\ldots, \lambda_m$ becomes an interpolant $I$ between $A$ and $B$.
-Note that any assignments to $\lambda_i$ that satisfy the expressions
-above will let an expression $I$ to be an interpolant.  Considering
-this, it is possible to obtain different interpolants for a
-interpolating problem by computing different assignments to
-$\lambda_i$.
-
-This enables the algorithm to preserve a set of interpolants by a
-linear expression template $a_1 x_1 + \cdots + a_n x_n + b \leq 0$
-with a constraint to describe a possible space of coefficient
-parameters $a_i$ and the constant parameter $b$.
-
-In practical, we may use linear programming solvers to obtain a model
-for $\lambda_i$.  By assigning concrete values to them, we can obtain
-a concrete interpolant.
-
-\paragraph{Computing a common interpolant}
-Next we consider a method to compute a common interpolant across two
-different interpolating problems.  That is, when given two
-interpolating problems $\left(A_1, B_1 \right)$ and $\left(A_2, B_2
-\right)$, the problem is about to compute the common interpolant I,
-which is an interpolant for the first problem and the second one at
-the same time.
-
-This problem is used when solving a following Horn clause solving
-problem, for instance.
-\[
-\implies
-% TODO
-\]
-
-To accomplish this, we simply need to compute the intersection of
-interpolant sets.  Let us call the interpolant space for the problems
-$\left(A_1, B_1 \right)$ and $\left(A_2, B_2 \right)$,
-$\mathcal{I}_1$, and $\mathcal{I}_2$.  In the previous method, a set
-of interpolants was represented by a parameterized linear expression.
-\begin{align*}
-\mathcal{I}_1 = \left\lbrace \mathbf{a}_1 \mathbf{x} \leq b_1 \mid
-C_1 (\mathbf{a}_1, b_1 ) \right\rbrace \\
-\mathcal{I}_2 = \left\lbrace \mathbf{a}_2 \mathbf{x} \leq b_2 \mid
-C_2 (\mathbf{a}_2, b_2 ) \right\rbrace
-\end{align*}
-To compute the intersection between them, it is sufficient to choose
-the same assignments of parameters and let them satisfy the both
-constraints $C_1$ and $C_2$.
-\[ \mathcal{I} = \left\lbrace \mathbf{a} \mathbf{x} \leq b \mid
-C_1(\mathbf{a}, b) \wedge C_2(\mathbf{a}, b) \right\rbrace \]
-
-However, if such an assignment of $\mathbf{a}$ and $b$ does not exist,
-no common solution exists across two problems because $\mathcal{I} =
-\emptyset$.
-
-By applying this approach, it is also possible to compute a common
-interpolant among more than three problems.
-
-\paragraph{Interpolation between logical formulas}
-We now extend the algorithm so that it can handle general logical
-formulas over linear arithmetic including disjunctions of linear
-expressions.  Here the algorithm takes two logical formulas over
-linear expressions as a problem input.  For convenience, we assume
-that they are represented in disjunctive normal form (DNF).  The
-solution for this problem is the interpolant that is a logical formula
-of linear expressions.
-
-First, consider the case of an interpolation problem between $A_1 \vee
-A_2$ and $B$.  When $(A_1 \vee A_2) \wedge B$ is inconsistent, $A_1
-\wedge B$ is necessarily inconsistent.  Let $\mathcal{I}_1$ denote the
-set of interpolants for $(A_1, B)$.  In the same manner, let
-$\mathcal{I}_2$ denote the set of interpolants for $(A_2, B)$.  We may
-naturally obtain an interpolant set by the intersection $\mathcal{I} =
-\mathcal{I}_1 \cup \mathcal{I}_2$ as a solution for $(A_1 \vee A_2,
-B)$.  When the algorithm computes $\mathcal{I}$ by the previous
-method, and it is $\emptyset$, we need to build a less simple solution
-by
-\[ \mathcal{I}' = \left\lbrace \left( I_1 \vee I_2 \right) \mid
-I_1 \in \mathcal{I}_1 \wedge I_2 \in \mathcal{I}_2 \right\rbrace \]
-Any formulas in the set $\mathcal{I}'$ are the interpolant for $(A_1
-\vee A_2, B)$ because
-\begin{align*}
-\forall I \in \mathcal{I}. & A_1 \vdash I \wedge A_2 \vdash I \wedge I \nvdash B \wedge \\
-& FV(I) = FV(A_1) \cup FV(A_2) \cup FV(B)
-\end{align*}
-
-Second, consider the case of an interpolation problem between $A$ and
-$B_1 \wedge B_2$.
-% TODO:
-
-By combining these two ways of interpolant construction, we may be
-able to build a small solution for more general cases like $(A_1,
-\ldots, A_n)$ and $(B_1, \ldots, B_n)$.  In order to obtain a small
-solution, we need to wisely choose the order of combining an
-interpolant set.
-
-\paragraph{Symmetric Interpolation}
-TODO
-
-\chapter{Horn clause solving}
-\label{chap:horn}
-
-We propose a new algorithm for solving recursion-less Horn clauses to
-generate a solution which is as simple as possible.  This enables us
-to obtain appropriate predicates that are suitable for predicate
-abstraction in the program verification techniques.
-
-The construction of the algorithm roughly consists of two parts.
-\begin{itemize}
-\item \textbf{Iterative Constraint Generation} is a process to build a
-  constraint to compute a solution.  It starts from generating the
-  weakest constraint in the sense that the resulting solution would
-  not be simple yet.  The process iteratively strengthen a constraint
-  until the strongest one while preserving the constraint
-  satisfiability.  The strongest constraint gives a simple solution.
-\item \textbf{Problem Restructuring} is a process to restructure the
-  problem to obtain a less simpler solution, if the constraint become
-  unsatisfiable during the iterative constraint generation.  This
-  process considers the unsatisfiable core of the constraint, and
-  wisely enlarge the solution size.
-\end{itemize}
-
-All predicate variables in an input problem are assigned simple
-templates.  Those predicate templates are parameterized.  In the
-iterative constraint generation process, algorithm builds a linear
-constraint that parameters in templates must satisfy.  This is done in
-a constructive manner, and constraints are build in the order of
-implication dependency in Horn clauses.  That is, when an input
-problem has a Horn clause thath has a predicate variable $P$ on its
-left-hand and $Q$ on its right-hand, a constraint to describe
-parameters in $P$'s template are first build before those in $Q$'s
-template.  When all parameters in the whole problem are described, the
-built constraint are evaluated to obtain a model.  Its instanciation
-becomes an appropriate solution for the input problem.
-
-The novelty of our algorithm lies in the incremental constraint
-refinement.  The algorithm tries to solve from the weakest constraint
-that an input problem allows that does not give a simple solution, and
-gradually strengthen the constraint until it reaches to the simplest
-solution that the problem has.  When a constraint becomes
-unsatisfiable during this step, the algorithm analyzes the
-unsatisfiable core and transform the problem to obtain a less simple
-solution.
-
-This construction of the algorithm enables us to compute a relatively
-simpler solution compared to existing algorithms while preserving the
-computation cost in a reasonable time.
-
-\section{Example}
-
-\paragraph {Tree interpolation}
-First, consider the following example.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-y \leq 2 & \implies B(y) \\
-A(x) \wedge B(y) & \implies 3x+2y \leq 10
-\end{align*}
-
-In this example, we would like to compute linear arithmetic formulas
-for the predicate variables $A$ and $B$.  They appear only once on
-both left-hand and right-hand globally in the problem.
-
-We try to obtain a simple solution that is expressed by linear
-expression templates, whose coefficients and constants are
-parameterized.  Assume that they have following predicate templates.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\
-B(y) : c_3 y + c_4 \leq 0
-\end{align*}
-
-By naturally encoding the first Horn clause with Farkas's lemma, we
-obtain a linear constraint
-$ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
-Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will
-make $A(x)$ valid for the first clause.  In the same manner, we obtain
-$ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
-from the second clause, and
-$ c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0 $
-from the third.
-
-One of the models satisfying all the constraints above is
-$( c_1, c_2, c_3, c_4, \lambda_1, \lambda_2, \lambda_3 ) = (3, -3, 2, -4, 3, 2, 1)$.
-By assigning them to the template, we obtain the following solution.
-
-\begin{align*}
-A(x) : 3 x - 3 \leq 0 \\
-B(y) : 2 y - 4 \leq 0
-\end{align*}
-
-\paragraph {DAG problem solving}
-Second, we explain a case that some predicate variables appear
-multiple times on left-hand side in different Horn clauses.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-A(x) \wedge A(y) & \implies x+y \leq 2
-\end{align*}
-
-Assigning a template $A(x) : c_1 x + c_2 \leq 0$, we obtain the same
-constraint for the first clause as the previous example.
-
-\begin{align} \label{eq:constr1}
-\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0
-\end{align}
-
-Sinve we have $A$'s appearance twice on the second clause, it is
-possible for $A$ to have two expressions each of which is used for
-$A(x)$ and $A(y)$.  Although we need to use the same template for
-multiple occurrences in order to compute the simplest solution, it may
-make the whole constraints unsatisfiable if the simplest solution does
-not exist.
-
-Therefore, we generate a weaker constraint for the second clause that
-allows different expressions of $A$ for multiple occurrence.  If the
-whole constraint is satisfiable in the case, we try to solve in a
-stronger constraint to generate a simple solution.
-
-In this specific case, it is virtually equivalent to solve the
-following problem in a weaker constraint.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 1 & \implies A_1(x) \\
-x \leq 1 & \implies A_2(x) \\
-A_1(x) \wedge A_2(y) & \implies x+y \leq 2 \\
-\\
-\text{Templates:} \\
-A_1(x) \text{ : } & c_{1,1} x + c_{2,1} \leq 0 \\
-A_2(x) \text{ : } & c_{1,2} x + c_{2,2} \leq 0 \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-\end{align*}
-
-By duplicating the constraint~\ref{eq:constr1} of $A$ for newly
-created $A_1$ and $A_2$, we obtain the whole constraint as follows.
-
-\begin{align*}
-& \lambda_{1,1} = c_{1,1} \wedge - \lambda_{1,1} \geq c_{2,1} \wedge \lambda_{1,1} \geq 0 \wedge \\
-& \lambda_{1,2} = c_{1,2} \wedge - \lambda_{1,2} \geq c_{2,2} \wedge \lambda_{1,2} \geq 0 \wedge \\
-& c_{1,1} = \lambda_2 \wedge c_{1,2} = \lambda_2 \wedge c_{2,1} + c_{2,2} \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-This constraint is satisfiable and one model is
-\[ ( c_{1,1}, c_{1,2}, c_{2,1}, c_{2,2}, \lambda_{1,1}, \lambda_{1,2}, \lambda_2 ) =
-( 1, -1, 1, -1, 1, 1, 1 ) \] and the solution then becomes
-$A(x) : (x -1 \leq 0) \wedge (x -1 \leq 0)$.
-
-Since the weak constraint is satisfiable, we now refine the constraint
-to obtain the better solution.  That is, simply eliminate the
-duplication to build the following constraint for the original
-template.
-
-\begin{align*}
-& \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \\
-& c_1 = \lambda_2 \wedge c_1 = \lambda_2 \wedge 2 c_2 \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-Solving this constraint simply gives us a solution $A(x) : x -1 \leq 0$.
-
-\paragraph {Disjunctive Horn clause solving}
-Finally, we explain a case that an input set of Horn clauses contain
-disjunctions on left-hand side.
-
-\begin{align*}
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 \vee -x+1 \leq 0 & \implies B(x) \\
-A(x) \wedge B(x) & \implies 1 \leq 0
-\end{align*}
-
-Same as the previous examples, templates for predicate variables are
-prepared.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\ B(x) : c_3 x + c_4 \leq 0
-\end{align*}
-
-The constraints generated are follows.
-
-\begin{align*}
-& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
-& \lambda_3 = c_3 \wedge \lambda_3 \geq c_4 \wedge \lambda_3 \geq 0 \wedge \\
-& - \lambda_4 = c_3 \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
-& c_1 + c_3 = 0 \wedge c_2 + c_4 > 0
-\end{align*}
-
-The constraints here become unsatisfiable, and one possible
-unsatisfiable core is:
-
-\begin{align*}
-\lambda_3 & = c_3 \\
-\lambda_3 & \geq c_4 \\
-- \lambda_4 & = c_3 \\
-\lambda_4 & \geq c_4 \\
-c_2 + c_4 & > 0 \\
-0 & \geq c_2
-\end{align*}
-
-Because the unsatisfiable core contains constraints originating from
-the disjunctions to be expressed in a single term template.  This let
-us determine to relax the template to allow a enlarged solution.
-
-Now we slightly change the problem and the template.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A(x) \wedge B_1(x) & \implies 1 \leq 0 \\
-A(x) \wedge B_2(x) & \implies 1 \leq 0 \\
-\\
-\text{Templates:} \\
-A(x) \text{ : } & c_1 x + c_2 \leq 0 \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-When we construct the constraints, as same as previous, we again
-obtain an unsatisfiable core:
-
-\begin{align*}
-\text{Constraint:} \\
-TODO \\
-\\
-\text{Unsatisfiable core:} \\
-TODO
-\end{align*}
-
-This time, the unsatisfiable core contains the constraints that
-originate from
-
-TODO
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A_i(x) \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A_i(x) \wedge B_j(x) & \implies 1 \leq 0 \quad (i,j \in \left\lbrace 1,2 \right\rbrace ) \\
-\\
-\text{Templates:} \\
-A_i(x) \text{ : } & c_{1,i} x + c_{2,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-Finally, we obtain the following constraint:
-
-TODO
-
-The process also generates a predicate template for every predicate
-variable that the solution will become.  under certain solution
-template.
-
-\section{Preliminaries}
-
-For ease of discussion, our algorithm handles an input set of Horn
-clauses by a \emph{Horn graph} that express the equivalent problem.
-The graph is then used throughout the algorithm to represent the
-problem.  A Horn graph is a labelled directed acyclic graph
-$G=(V,E,\varphi)$.
-\begin{itemize}
-\item $V$ is a union of two disjoint sets of vertices; Horn term
-  vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex $u
-  \in V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
-  denoted as $succ(u)$.
-\item $E$ is a set of labelled edges, which are expressed in the form
-  $(u,v,\theta)$ where $\theta$ is a finite map from variables to
-  variables.
-\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
-  term vertices, where L is a set of elements of a predicate variable
-  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
-  inequality $a_1 x_1 + \cdots + a_n x_n + b \leq 0$.
-\end{itemize}
-There are two kinds of edges in a Horn graph;
-\begin{itemize}
-\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
-  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
-\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
-  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
-  empty.
-\end{itemize}
-
-We define the relation $\leadsto$ over $V_T$ as follows.
-\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow \exists a \in V_\Rightarrow;
-\edgel{u}{\theta}{a} \wedge \edge{a}{v} \] We may omit $\theta$ and
-write $u \leadsto v$ for convenience in later discussions. By the
-assumption that $G$ is acyclic, both $\rightarrow^+$ and $\leadsto^+$ are
-well-founded relations. $v_\bot$ is the greatest element with respect
-to both $\rightarrow^+$ and $\leadsto^+$.
-
-The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
-Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$.
-\begin{align*}
-\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(succ(v)) \\
-\llbracket G \rrbracket & = \bigwedge_{v \in V_\Rightarrow} \mathcal{HC}_v
-\end{align*}
-We implicitly assume each $\mathcal{HC}_v$ is universally quantified
-by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.
-
-Horn term vertices with predicate variable labels are unique each
-other that any of them do not have the same variable on their labels.
-From now on, we may denote a Horn term vertex with a predicate
-variable label by its predicate variable name if no ambiguity is
-present.
-
-Without loss of generality, we assume the existence of the \emph{root
-  vertex} $v_\bot \in V_T$ with linear expression labels $1 \leq 0$,
-meaning $\bot$.  This is because if we have a Horn clause $P \implies
-e$ which $e$ is a linear expression, we can convert it to $P \wedge
-\neg e \implies 1 \leq 0$.  We also assume that other vertices with
-linear expression labels have no incoming edges to them.
-
-For simplicity of discussion, we define \emph{term arity} $k_v$ for
-every Horn term vertex $v \in V_T$ with incoming edges.
-\begin{align*}
-k_v =
-\begin{cases}
-\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
-0 & \mbox{if } \varphi(v) = \bot
-\end{cases}
-\end{align*}
-We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.
-
-A variable mapping label $\theta$ on edges is used to map the index of
-one predicate variable to another. (TODO?)
-
-(Example?)
-
-The algorithm assigns a \emph{predicate template} for all predicate
-variables in $G$. A \emph{predicate templateset} is a set of mappings from
-a predicate variable to a predicate template.  A template for a
-predicate variable $P(x_1, \ldots, x_n)$ is a formula $\psi$
-constructed from the following definition.
-\begin{align*}
-\psi ::= & a_1 x_1 + \cdots + a_n x_n + b \leq 0 \mid \\
-& \psi \wedge \psi \mid \psi \vee \psi
-\end{align*}
-A linear expression in the definition above may also be represented in
-a equivalent vector form $\mathbf{a} \mathbf{x} + b \leq 0$ for short.
-Parameters $\mathbf{a}$ and $b$ are specified by a linear constraint.
-
-The solution of $G$ is a map $\rho$ from predicate variables to linear
-predicates of the form $\lambda x_1, \ldots ,x_n. \psi$ where
-parameters in $\psi$ are replaced with constant and $\rho[G]$ becomes
-a tautology.  A solution is \textit{simple} if $\rho[P]$ is of the
-form $\mathbf{a} \mathbf{x} + b \leq 0$ for every $P$ in
-$\mathrm{dom}(\rho)$.
-
-The problem of solving Horn clauses (represented by $G$) is to find a
-solution of $G$ (and a simple solution if possible).
-
-\section{Simple solution}
-
-We first propose an algorithm to compute a simple solution for the
-given Horn graph $G$ if one exists.  Later we add modifications to the
-algorithm for the case that a simple solution does not exist for $G$.
-In such a case, we transform the input graph $G$ to have a larger
-solution template.
-
-For each vertex $v \in V$, we define
-$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v + b_v \leq 0, C_v)$
-, where
-\begin{itemize}
-\item $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ is a linear inequality that
-  contains coefficient variables $\mathbf{a}_v$ and a constant $b_v$,
-  and
-\item $C_v$ is a set of linear expressions on coefficient variables.
-\end{itemize}
-Intuitively, it denotes the set of inequalities:
-\begin{align*}
-\left\lbrace
- \mathbf{a}_v \mathbf{x}_v + b_v \leq 0 \mid
- \exists \left( \textsc{FV}(C_v)
-  \setminus (\mathbf{a}_v \cup \left\lbrace b_v \right\rbrace
- \right); C_v
-\right\rbrace
-\end{align*}
-For a predicate variable vertex, $\mathbf{a}_v \mathbf{x}_v + b_v \leq
-0$ serves as a predicate template and $C_v$ describes the parameters
-$\mathbf{a}_v$ and $b_v$.
-
-$C_v$ is defined by well-founded induction on $v$ with respect to the
-relation $\rightarrow^+$. Then, $C_v$ can be shown as follows.
-\begin{itemize}
-\item For $v \in V_\Rightarrow$;
-\begin{align*}
-C_v & =
- \bigcup_{\edgel{u}{\theta}{v}} C_u \cup
- \bigcup_{0 < i \leq K}
- \left\lbrace
-  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{v}} \mathbf{a}_{u, \theta^{-1} (i)}
- \right\rbrace \\
- & \hspace{2cm} \cup
- \left\lbrace
-  b_v \leq \sum_{\edgel{u}{\theta}{v}} b_u
- \right\rbrace \cup
- \bigcup_{k_{succ(v)} < i \leq K}
- \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that
-  $v \neq v_{\bot} \wedge \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0$
-  where $\mathbf{a}_0$ and $b_0$ are constants;
-\begin{align*}
-C_v =
-\left\lbrace
- \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
- b_v = \lambda_v b_0
-\right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that $\varphi(v) = P(\mathbf{x}_v)$;
-\begin{align*}
-C_v = & \bigcup_{\edge{u}{v}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{a}_v, b_u = b_v \right\rbrace
-\right)
-\end{align*}
-\item For $v_\bot$;
-\begin{align*}
-C_{v_\bot} = \bigcup_{\edge{u}{v_\bot}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{0}, b_u > 0 \right\rbrace \right)
-\end{align*}
-\end{itemize}
-
-Any model $\sigma$ to the constraint $C_{v_\bot}$ gives a simple
-solution for $G$ if $C_{v_\bot}$ is satisfiable, i.e., we should have:
-
-\begin{quote}
-If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
-\begin{align*}
- \rho = \left\lbrace
-  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \mid
-  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
- \right\rbrace
-\end{align*}
-is a solution of $G$.
-\end{quote}
-
-If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for $G$.
-There are two cases that a simple solution does not exist for a
-predicate variable.
-\begin{itemize}
-\item A predicate variable appears multiple times on left-hand of Horn
-  clauses, and those appearances require different expressions as an
-  assumption.
-\item A predicate variable is implied by a disjunctive Horn clause,
-  which has disjunctions on its left-hand, and a common premise that
-  the disjunctions induce is not strong enough for solving a problem.
-\end{itemize}
-
-The first case occurs in a case like follows.
-\begin{align*}
-% conj.txt ... originated from uc/01.txt
-\left( x \leq 0 \right) \wedge \left( y \leq 0 \right) & \implies P(x,y) \\
-P(x,a) \wedge P(b,y) & \implies x+y \leq 0
-\end{align*}
-In this example, the predicate variable $P$ appears twice on the
-second clause and they require different premises; $x \leq 0$ for the
-first and $y \leq 0$ for the second.  They cannot be represented by
-$\mathbf{ax} + b \leq 0$ template at the same time.
-
-The second case occurs in a case like follows.
-\begin{align*}
-% disj-simp.txt ... originated from 23simp_neg.txt
-\left( x+1 \leq 0 \right) \vee \left( -x+1 \leq 0 \right) & \implies P(x) \\
-P(x,y) \wedge \left( x \leq 0 \right) \wedge \left( -x \leq 0 \right) & \implies \bot
-\end{align*}
-In this example, the predicate variable $P$ is implied by the first
-clause, which is a disjunctive Horn clause.  As a simple solution, $P$
-can take a common consequence that both of $x+1 \leq 0$ and $-x+1 \leq
-0$ induce.  However, such a simple linear expression does not exist
-except for $0 \leq 0$, which is equivalent to $\top$, and it is not
-sufficient to satisfy the second clause.
-
-Both cases can be fixed by enlarging the solution template.
-
-\section {Graph transformation}
-
-If a simple solution does not exist for $G$, we try to give a less
-simpler solution by transforming $G$ and relaxing the constraint
-$C_{v_\bot}$.  This transformation takes place by considering the
-unsatisfiable core $\mathcal{U}$.  An unsatisfiable core $\mathcal{U}$
-is a subset of $C_{v_\bot}$ which is inconsistent itself.  Depending
-on the reason for the absence of a simple solution, the way of the
-transformation differs.
-
-Algorithm~\ref{alg:solveHorn} shows the updated pseudo-code of our
-Horn clause solving algorithm with the graph transformation.  It takes
-an input Horn graph and returns a mapping between predicate variables
-and linear arithmetic formulas.
-
-\newcommand{\solveHorn}{\ensuremath{\mbox{\sc SolveHornGraph}}}
-\newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
-\newcommand{\transGraph}{\ensuremath{\mbox{\sc TransformGraph}}}
-
-\begin{algorithm}
-\caption{$ \solveHorn (G) $}\label{alg:solveHorn}
-\begin{algorithmic}
-\STATE {$v_\bot \gets $ the root node of $G$}
-\STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-\STATE {$T \gets $ initial predicate templateset for $G$}
-\REPEAT
-  \STATE {$C_{v_\bot} \gets \genConstr (G, v_\bot, V_\star)$}
-  \IF {$C_{v_\bot}$ is satisfiable}
-    \STATE {$\sigma \gets $ one model of $C_{v_\bot}$}
-    \STATE {$V_\star \gets V_\star \cup \left\lbrace u \in V_T \mid \forall v \in V_\star; u \leadsto v \right\rbrace$}
-  \ELSE
-    \STATE {$\mathcal{U} \gets $ unsatisfiable core for $C_{v_\bot}$}
-    \STATE {$(G, T) \gets \transGraph (G, T, \mathcal{U})$}
-    \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-  \ENDIF
-\UNTIL {$V_\star = V_T$}
-\RETURN {$\sigma \left( T \right)$}
-\end{algorithmic}
-\end{algorithm}
-\transGraph~procedure determines the unsatisfiable core of the
-constraint and transform the Horn graph $G$ and predicate templates.
-As mentioned earlier, two ways of graph transformation exist.
-
-\paragraph{Conjunction split}
-For the first case previously mentioned, the unsatisfiable core
-$\mathcal{U}$ contains coeffcient constraint expressions with a
-coefficient parameter of certain predicate variable.  If the predicate
-variable $P$ is the cause of unsatisfiability, some constraint
-expressions in the unsatisfiable core have a parameter $a_{P,i}$
-and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is copied to
-$P_1, \ldots, P_n$ so that the different appearances of $P$ on
-left-hand of Horn clauses can take different premises.  The solution
-template of $P$ becomes the conjunction of $P_1, \ldots, P_n$.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-Note that after this duplication of the predicate vertex $P$, in order
-to induce different linear expressions for $P_1, \ldots, P_n$, it may
-be necessary for preceeding predicate variable vertices in $G$ to be
-duplicated in the same manner as $P$. It is because that different
-derivation of $P$'s implication relation may be required to get
-different consequences.
-
-However, succeeding predicate vertices of $P$ in $G$ and other
-vertices than predecessors are not affected because those implications
-that have $P$ in their assumption can selectively choose the premise
-that $P_1, \ldots, P_n$ leads.
-
-Therefore, this graph transformation only affects the predecessors of
-$P$.
-
-\paragraph{Disjunction split}
-For the second case, the unsatisfiable core $\mathcal{U}$ contains
-coefficient constraints that unifies induced linear expressions from
-each term in disjunctions to be the same expression.  Similarly as
-previous, if the predicate variable $P$ is the cause of
-unsatisfiability, some constraint expressions in the unsatisfiable
-core have a parameter $a_{P,i}$ and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is separated to
-$P_1, \ldots, P_n$ each of which is implied by a term, which was
-originally a part of a disjunction to imply $P$.  This intuitively
-enables us to avoid generating a common linear expression that
-disjunctive terms induces, and to propagate each term's premises to
-$P$'s successors instead.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-After this duplication of the predicate vertex $P$, unlike the
-conjunction split, this results in producing a disjunction at
-succeeding predicate variables vertices ($R$ in this case).
-\begin{align*}
-(P_1 \wedge Q) \vee (P_2 \wedge Q) \implies R
-\end{align*}
-Also, predicate variables that appeared with $P$ in the same
-assumption ($Q$ in this case) needs to be duplicated by conjunction,
-so as to obtain different premises that go along with $P_i$.
-Therefore, the effect of disjunction split may reach to neighboring
-predicate vertices.
-
-Consider these two ways of transformation, we show a pseudo-code for
-\transGraph~procedure.  The procedure is given the graph $G$, the
-predicate templateset $T$, and the unsatisfiable core
-$\mathcal{U}$. It returns the transformed graph $G'$ and the new
-templateset $T'$.
-
-\begin{algorithm}
-\caption{$ \transGraph (G, T, \mathcal{U}) $}\label{alg:transGraph}
-\begin{algorithmic}
-% TODO:
-\RETURN {$\left( G', T' \right)$}
-\end{algorithmic}
-\end{algorithm}
-
-\paragraph{Combinatorial explosion in disjunctions}
-Virtually, if multiple disjunctions exists in an input set of Horn
-clauses, a na\"{i}ve way of building a solution is solve all
-combinations of choices out of every disjunction and concatenate each
-expression with conjunctions and/or disjunctions appropriately.  This
-may cause a combinatorial explosion.
-
-Conside example below. $e_i$ represents some conjunctions of linear
-expressions.
-% TODO: Can't we replace e_i with actual expressions?
-\begin{align*}
-(e_1 \vee e_2) \implies P \\ (e_3 \vee e_4) \implies Q \\ P \wedge Q \implies \bot
-\end{align*}
-In this example, a na\"{i}ve algorithm may solve following four
-problems separately.
-\begin{itemize}
-\item $e_1 \implies P_1 \qquad e_3 \implies Q_1 \qquad P_1 \wedge Q_1 \implies \bot$
-\item $e_2 \implies P_2 \qquad e_3 \implies Q_2 \qquad P_2 \wedge Q_2 \implies \bot$
-\item $e_1 \implies P_3 \qquad e_4 \implies Q_3 \qquad P_3 \wedge Q_3 \implies \bot$
-\item $e_2 \implies P_4 \qquad e_4 \implies Q_4 \qquad P_4 \wedge Q_4 \implies \bot$
-\end{itemize}
-Then the solution is built by follows.
-\begin{itemize}
-\item $P = ( P_1 \vee P_2 ) \wedge ( P_3 \vee P_4 )$
-\item $Q = ( P_1 \vee P_3 ) \wedge ( P_2 \vee P_4 )$
-\end{itemize}
-However, this method is not efficient.
-
-\paragraph{Algorithm}
-To avoid this, our algorithm lets a predicate variable $P$ which is
-disjunctively implied on the Horn graph $G$ to have a linear
-expression that is consequences of all terms in the disjunction.  This
-prevents the possibility of choices for disjunctions to be propagated
-to succeeding vertices. The simple linear expression is propagated
-instead.
-
-On the other hand, the algorithm allows a predicate variable $P$ that
-appears on multiple implications or appears multiple times in a single
-implication to take different linear expressions as a premise at
-different appearances.  This is for the algorithm to determine the
-cause of unsatisfiability efficiently.  It is done by copying a
-predicate templates and its parameters' constraint with appropriate
-renaming.
-
-The algorithm first execute this duplication on all predicate
-variables with multiple appearances.  It is the weakest constraint
-that the algorithm makes for the certain $G$.  Then, the algorithm
-gradually reduces the number of vertices that are duplicated by each
-time testing the satisfiability of the constraint.  When the
-constraint bebcome unsatisfiable at the certain iteration, the
-algorithm can determine a predicate vertex which has to be split in a
-conjunctive template.  If all the predicate variable constraints are
-no longer duplicated and the constraint is the strongest, any model
-for such constraint gives a simple solution under the Horn graph $G$
-and the templateset $T$.
-
-Note that duplicating the constraint in a na\"{i}ve way may cause the
-constraint to exponentially grow in size.  Our algorithm prevent it by
-simplifying a predicate vertex's constraint $C_v$ by quantifier
-elimination when $v$ is duplicated in generating linear constraint.
-For a predicate variable vertex $v_P$, variables except for the
-coefficient parameter $\mathbf{a}_P$ and the constant parameter $b_P$
-are eliminated.  Due to the constraints generation process, the
-simplified constraint is also linear and can be efficiently check the
-satisfiability.
-
-We finally show the modified version of pseudo-code for $\genConstr$
-procedure.  It produces the constraint $C_v$ for a given vertex $v$,
-considering constraint duplication and simplification.
-
-\begin{algorithm}
-\caption{$ \genConstr (G, v, V_\star) $}\label{alg:genConstr}
-\begin{algorithmic}
-\STATE {$C \gets \emptyset$}
-\FORALL {$v_\Rightarrow \in preds_G(v)$}
-  \FORALL {$v_p \in preds_G(v_\Rightarrow)$}
-    \IF {$\varphi (v_p)$ is a predicate variable}
-      \STATE {$C \gets C \cup \genConstr (G, v_p, visited)$}
-    \ELSE
-      \STATE { $id$ }
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-traverses the input graph in the topological order on
-$V_\Rightarrow$, and concatenate the constraints from predecessors to
-build the problem constraints.  Each process to build a constraint on
-$\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$ is equivalent
-to generate a linear constraint in interpolation.
-
-\chapter{Experiment}
-\label{chap:experiment}
-
-We incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It
-is specifically designed to verify the correctness of higher-orde
-programs written in a small subset of ML.  It infers dependent types
-for input higher-order programs \cite{conf/ppdp/UnnoK09}.
-
-Our implementation make use of Z3 \cite{conf/tacas/MouraB08} as
-an SMT solver.
-
-For simplicity of the implementation, we assume that the linear
-inequalities in the integer space.  Because of this, our algorithm
-lost the completeness.  By using Motzkin's transpose theorem
-\cite{journals/networks/Rajan90}, it is possible to handle linear
-inequalities in both $\mathbf{ax} + b \leq 0$ and $\mathbf{ax} + b <
-0$.
-
-\section{Interpolating theorem prover}
-
-Our interpolating theorem prover is used as a sub-procedure to
-solve Horn clauses.
-
-\begin{itemize}
-\item CSIsat \cite{conf/cav/BeyerZM08}
-\item FOCI \cite{website/foci}
-\item CLP-Prover \cite{website/clp}
-\end{itemize}
-
-\section{Horn clause solver}
-
-\chapter{Related work}
-\label{chap:related}
-\input{related.tex}
-
 \chapter{Future Work}
@@
 \cite{conf/cav/DaiXZ13}
-
-\chapter{Conclusion}
-\label{chap:conclusion}
-
-\bibliographystyle{plain}
-\bibliography{./biblio}
-
-\end{document}

### 2013-08-26 article/horn.tex
Other chapters also separated

--- 2013-08-26 article/thesis.tex
@@
-\documentclass[master,final,12pt]{iscs-thesis}
-
-\etitle{Simple Solutions of Acyclic Horn Clauses for Program Verification}
-\eauthor{Yuto Takei}
-\esupervisor{Naoki Kobayashi}
-
-\jtitle{Lipsum}
-\jauthor{Lipsum}
-\jsupervisor{Lipsum}
-
-\supervisortitle{Professor}
-\date{August 28, 2013}
-
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{stmaryrd}
-\usepackage{syntax}
-\usepackage{url}
-
-\newcommand{\sembrack}[1]{[\![#1]\!}
-\newcommand{\edge}[2]{#1\rightarrow#2}
-\newcommand{\edgel}[3]{#1\xrightarrow{#2}#3}
-
-\begin{document}
-
-\begin{eabstract}
-Automated theorem proving plays a crucial role in modern program
-verification techniques.  In particular, in the context of software
-model checking, the interpolation problem and its generalization
-called \emph{Horn clause solving} have attracted special attention.
-The latter problem asks for a solution for a set of acyclic Horn
-clauses containing unknown predicate variables.  The solution, i.e., a
-substitution for predicate variables, then serves as a candidate of
-program invariants.  However, existing methods are unsatisfactory in
-that they tend to generate too specific a solution, which does not
-correspond to program invariants.
-
-To address the problem above, we propose a new algorithm for solving
-Horn clauses consisting of linear arithmetics and unknown predicate
-variables.  The new algorithm tries to find small predicates, based on
-the hypothesis that program's invariants tend to be simple formulas.
-To this end, our algorithm prepares a template of linear arithmetic
-formulas of a fixed size as a candidate of each unknown predicate, so
-that Horn clause solving is reduced to linear programming.  The
-algorithm iteratively increases the size of the templates until a
-solution is found.  It enables us to construct relatively small
-solutions in a reasonable time compared to existing algorithms.  We
-have implemented and incorporated our Horn clause solver into MoCHi, a
-software model checker for functional programs.  We have evaluated the
-effectiveness of our approach through experiments.
-\end{eabstract}
-\begin{jabstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.  Morbi
-vulputate ornare arcu et malesuada.  Proin sem urna, feugiat et velit
-ac, elementum blandit orci.  Etiam ut neque ac felis laoreet venenatis
-eget quis enim.  Ut aliquam dapibus sem egestas molestie.  Ut sit amet
-porttitor orci, id fermentum dui.  In non dui at ante sagittis
-malesuada at in neque.  Fusce euismod velit quis ante ullamcorper
-qsodales.
-\end{jabstract}
-\maketitle
-
-\begin{acknowledge}
-Thank you.
-\end{acknowledge}
-
-\frontmatter
-\tableofcontents
-
-\mainmatter
-
-\chapter{Introduction}
-
-Automated theorem proving plays an important role in the modern
-program verification.  Particularly in the context of verification
-with model checking techniques, an automated theorem prover is used to
-abstract an infinite set of program states into a finite one so that
-the model checker can virtually explore all possible program execution
-paths.
-
-One of the most common methods of such abstraction is predicate
-abstraction\cite{conf/cav/GrafS97} within CounterExample-Guided
-Abstraction Refinement (CEGAR) framework
-\cite{conf/cav/ClarkeGJLV00,conf/popl/BallR02,conf/popl/HenzingerJMS02}.
-An automated theorem prover finds appropriate predicates for every
-program location.  An execution states at those locations is then
-expressed by a Boolean valuation of the given predicates.  If the
-abstraction is too coarse and a model checker discovers an infeasible
-error execution path in an abstract system, the prover computes an
-additional predicate at those program locations to refine the
-abstraction predicates.
-
-In order to obtain an additional abstraction predicate at program
-locations along the infeasible counterexample, one class of automated
-theorem provers, called an interpolating theorem prover
-\cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07},
-computes a separating predicate called an interpolant between two sets
-of constraints; the one consists of the constraints from the program
-entry point to the location, and the other consists of the constraints
-from the location to the program failure.  Those constraints include
-variable assignments and assertions to be satisfied originated from
-conditional branches.
-
-The computed predicate by an interpolating theorem prover is satisfied
-at the location under the execution in the concrete model along the
-discovered counterexample, and any execution paths which satisfies the
-predicate does not lead to the failure state.  After the refinement,
-the infeasible path is ruled out and no longer found by the model
-checker.
-
-However, existing interpolating theorem provers may return too complex
-a solution which is heavily affected by the constraints of the
-specific path.  It is not favorable in terms of the program
-verification because a complex predicate may not be sufficient to rule
-out similar spurious paths in programs with loops and recursion calls.
-This causes the model checker to discover spurious paths infinitely
-which pass the same program location by loops or recursion calls at
-different numbers of iteration, and the verification process does not
-to terminate.
-
-To avoid this, it is desired to discover an invariant formula for the
-predicate abstraction.  Based on the hypothesis that program's
-invariants tend to be simple formulas, we propose a new interpolating
-algorithm which tries to minimize the number of conjunctions and
-disjunctions in a solution.  The predicate obtained by this algorithm
-may become similar to the invariant, and the number of trials in model
-checking may be reduced.
-
-Additionally, we extend our interpolating algorithm to solve a
-symmetric interpolation problem, in which the algorithm computes
-predicates at multiple program locations along a path at once.  This
-enables the abstraction predicates at multiple locations to be updated
-at the same computation.
-
-Finally, we extend our algorithm to solve Horn clause solving
-problems.  It allows us to obtain predicates which are highly likely
-to be program invariants, by computing the same abstraction predicates
-inside loops and recursion calls on different iterations, and by
-unifying predicates at the same location over multiple paths.  The
-problems are generated by encoding constraints over execution paths
-with unknown predicates which describe functions' constraints between
-parameters and return values.
-
-The rest of the paper is structured as follows.
-Chapter~\ref{chap:interpolation} proposes a new interpolating
-algorithm to obtain simple solutions, and extends it to solve
-symmetric interpolation problems.  Chapter~\ref{chap:horn} describes a
-new Horn clause solving algorithm by extending the previous
-interpolating algorithm.  Chapter~\ref{chap:experiment} describes
-experiments and illustrates the result to confirm the effect of our
-proposed algorithm.  Chapter~\ref{chap:future} gives a discussion of
-our improvements and proposes further improvements to make.
-Chapter~\ref{chap:related} mentions related work done by others.
-Finally, Chapter~\ref{chap:conclusion} reviews the impact of our
-research and brings the conclusion.
-
-\chapter{Interpolation}
-\label{chap:interpolation}
-
-Among various program verification techniques, when software model
-checking with predicate abstraction is adopted, the interpolation is
-used to compute the abstraction predicate along a spurious
-counterexample that a model checker discovered.  It is done by
-computing a Craig interpolant between the strongest postcondition at
-the location and the weakest precondition to the failure point at
-every program location.
-
-However, by the existing interpolating theorem prover, we may not be
-able to obtain a simple solution.  Consider the example below.
-\begin{align*}
-\varphi_1 : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
-\varphi_2 : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
-\end{align*}
-Although one may be able to obtain $x = y$ as the simplest
-interpolation between $\varphi_1$ and $\varphi_2$,
-CSIsat\cite{conf/cav/BeyerZM08} returns
-\[ TODO \]
-as a solution.
-
-We propose an interpolation algorithm which returns as simple a
-solution as possible in terms of the number of conjunctions and
-disjunctions.  Our algorithm focuses on interpolating problems on
-linear arithmetic and it works in a constructive manner.
-
-\section{Example}
-
-Consider the following interpolation problem.
-
-\begin{align*}
-\frac
-{A: (x \leq a) \wedge (a + 1 \leq y)}
-{B: (y \leq b) \wedge (b + 1 \leq x)}
-\end{align*}
-
-Here, the linear arithmetic formulas A and B are inconsistent.  The
-interpolant I for this problem is $x-y+1 \leq 0$, which is implied by
-$A$ and inconsistent with $B$.  The vocabulary of $I$ is $\left\lbrace
-x,y \right\rbrace$ and is over $A$'s vocabulary $\left\lbrace x,y,a
-\right\rbrace$ and $B$'s $\left\lbrace x,y,b \right\rbrace$.
-
-In computing the interpolant, we make a linear constraint of a
-interpolant by applying Farkas's lemma.  First, we assing weight
-parameters to every expression in conjunction form.
-
-\begin{align*}
-\begin{array}{r r r r r r l}
-\lambda_1 : & -a & & +x & & & \leq 0 \\
-\lambda_2 : & a & & & -y & +1 & \leq 0 \\
-\hline
-\lambda_3 : & & -b & & +y & & \leq 0 \\
-\lambda_4 : & & b & -x & & +1 & \leq 0
-\end{array}
-\end{align*}
-
-Because all expressions induce inconsistent as a whole, the sum of
-weighted expressions should satisfy the following constraint.
-
-\begin{align*}
-\begin{array}{r l l}
-- \lambda_1 + \lambda_2 & = 0 \quad & (a) \\
-- \lambda_3 + \lambda_4 & = 0 & (b) \\
-  \lambda_1 - \lambda_4 & = 0 & (x) \\
-- \lambda_2 + \lambda_3 & = 0 & (y) \\
-  \lambda_2 + \lambda_4 & > 0 & (\text{constant}) \\
-\lambda_1, \ldots, \lambda_4 & \geq 0 & (\text{non-negative weight})
-\end{array}
-\end{align*}
-
-With the linear constraint among $\lambda_i$ above, the interpolant is
-represented as follows by the sum of weighted expressions from the
-first formula groups.
-\[ (- \lambda_1 + \lambda_2) a + \lambda_1 x - \lambda_2 y + \lambda_2 \leq 0 \]
-One of the model of the linear constraint is $\lambda_1 = \lambda_2 =
-\lambda_3 = \lambda_4 = 1$, and we obtain $x-y+1 \leq 0$ as a
-solution.
-
-% TODO: Better to mention multiple interpolants possible for one
-% interpolating problem?
-
-\section{Preliminaries}
-
-Our algorithm aims to compute a Craig interpolant over linear
-arithmetic formulas.
-
-\paragraph{Craig Interpolation}
-Given two logical formulas $A$ and $B$ that are inconsistent each
-other, we call a new preposition $I$ a \emph{Craig interpolant}
-\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$
-is implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
-only free variables that appears in both $A$ and $B$.  If $A \wedge B$
-is unsatisfiable, an interpolant $I$ always exists.
-
-Our algorithm handles a linear expression $e$ which is in a form:
-\[ a_1 x_1 + \cdots + a_n x_n + b \leq 0 \]
-This expression contains coefficient parameters $a_1, \ldots, a_n$ and
-a constant parameter $b$.  We treat a special expressions $\top$ as $0
-\leq 0$ and $\bot$ as $1 \leq 0$.  The algorithm input and output is a
-formula $\psi$ that consists of conjunctions and/or disjunctions of
-linear expressions.
-\[ \psi = e | e \wedge e | e \vee e | \top | \bot \]
-
-In the simple interpolation, the algorithm receives two formulas in
-$\psi$ and returns an interpolant in $\psi$.  For the later extension
-to handle a symmetric interpolation problem, the algorithm receives
-multiple formulas in $\psi$ and returns intermediate interpolants in
-$\psi$.  For computing those interpolants over linear arithmetics, we
-use Farkas's Lemma.
-
-\paragraph{Farkas's Lemma on linear inequalities}
-Let a linear inequality $e_i$ be represented as $a_i1 x_1 + \cdots +
-a_im x_m <= a_i0$.  Assuming that $e_1,\cdots,e_n$ implies $e_0$,
-there exists $\lambda_1,\cdots,\lambda_n$ that satisfy $a_0j
-=\sum_(i=1)^n \lambda_i * a_ij (j=0, \ldots, m)$.
-
-For the special case of Farkas's Lemma, we can state that:
-
-Assuming that $e_1,\cdots,e_n$ implies $\bot$, there exists
-$\lambda_1,\cdots,\lambda_n$ that satisfy $???? TODO $.
-
-Our interpolating algorithm make use of the latter version of Farkas's
-Lemma.
-
-\section{Algorithm}
-
-Our algorithm aims to construct a relatively small interpolant in a
-reasonable computation time for any input.  It also aims to obtain a
-common solution among multiple interpolating problems.  When the
-algorithm is used as a sub-routine for solving Horn clause satisfying
-problems that are made while discovering predicates for abstraction
-during the program verification.  These advantages of our algorithm
-make it possible to return a relatively small predicates for those
-problems. To accomplish these aims, the algorithm preserves a set of
-interpolants during the computation, and define and execute operations
-over interpolant sets.
-
-\paragraph{Interpolation between conjunctions}
-We first present an interpolating algorithm between two conjunctive
-sets of linear inequalities $A = \left\lbrace e_1,\ldots,e_m
-\right\rbrace$ and $B = \left\lbrace e_{m+1},\ldots,e_{m+n}
-\right\rbrace$.  We assume that conjunctions of $A$ and $B$ are
-inconsistent.  For later convenience, we call this interpolating
-problem $\left( A, B \right)$.  Then there exists an interpolating
-linear expression $e_\star$ which satisfies
-\begin{align*}
-\left\lbrace e_1,\ldots,e_m \right\rbrace & \vdash e_\star \\
-e_\star & \nvdash \left\lbrace e_{m+1},\ldots,e_{m+n} \right\rbrace
-\end{align*}
-For convenience, we assume that any expression $e_i$ have variables
-$x_1, \ldots, x_a$.
-
-According to the Farkas's lemma, there exists a set of assignmentqs to
-$\lambda_i$, to conclude $\bot$ from a set of linear expressions $A
-\cup B$, namely,
-\begin{align*}
-\exists \lambda_1, \ldots, \lambda_m. \\
-a % TODO: coefficient, constant constraints
-\end{align*}
-
-Then, a linear combination of $A$'s formula with a weight $\lambda_1,
-\ldots, \lambda_m$ becomes an interpolant $I$ between $A$ and $B$.
-Note that any assignments to $\lambda_i$ that satisfy the expressions
-above will let an expression $I$ to be an interpolant.  Considering
-this, it is possible to obtain different interpolants for a
-interpolating problem by computing different assignments to
-$\lambda_i$.
-
-This enables the algorithm to preserve a set of interpolants by a
-linear expression template $a_1 x_1 + \cdots + a_n x_n + b \leq 0$
-with a constraint to describe a possible space of coefficient
-parameters $a_i$ and the constant parameter $b$.
-
-In practical, we may use linear programming solvers to obtain a model
-for $\lambda_i$.  By assigning concrete values to them, we can obtain
-a concrete interpolant.
-
-\paragraph{Computing a common interpolant}
-Next we consider a method to compute a common interpolant across two
-different interpolating problems.  That is, when given two
-interpolating problems $\left(A_1, B_1 \right)$ and $\left(A_2, B_2
-\right)$, the problem is about to compute the common interpolant I,
-which is an interpolant for the first problem and the second one at
-the same time.
-
-This problem is used when solving a following Horn clause solving
-problem, for instance.
-\[
-\implies
-% TODO
-\]
-
-To accomplish this, we simply need to compute the intersection of
-interpolant sets.  Let us call the interpolant space for the problems
-$\left(A_1, B_1 \right)$ and $\left(A_2, B_2 \right)$,
-$\mathcal{I}_1$, and $\mathcal{I}_2$.  In the previous method, a set
-of interpolants was represented by a parameterized linear expression.
-\begin{align*}
-\mathcal{I}_1 = \left\lbrace \mathbf{a}_1 \mathbf{x} \leq b_1 \mid
-C_1 (\mathbf{a}_1, b_1 ) \right\rbrace \\
-\mathcal{I}_2 = \left\lbrace \mathbf{a}_2 \mathbf{x} \leq b_2 \mid
-C_2 (\mathbf{a}_2, b_2 ) \right\rbrace
-\end{align*}
-To compute the intersection between them, it is sufficient to choose
-the same assignments of parameters and let them satisfy the both
-constraints $C_1$ and $C_2$.
-\[ \mathcal{I} = \left\lbrace \mathbf{a} \mathbf{x} \leq b \mid
-C_1(\mathbf{a}, b) \wedge C_2(\mathbf{a}, b) \right\rbrace \]
-
-However, if such an assignment of $\mathbf{a}$ and $b$ does not exist,
-no common solution exists across two problems because $\mathcal{I} =
-\emptyset$.
-
-By applying this approach, it is also possible to compute a common
-interpolant among more than three problems.
-
-\paragraph{Interpolation between logical formulas}
-We now extend the algorithm so that it can handle general logical
-formulas over linear arithmetic including disjunctions of linear
-expressions.  Here the algorithm takes two logical formulas over
-linear expressions as a problem input.  For convenience, we assume
-that they are represented in disjunctive normal form (DNF).  The
-solution for this problem is the interpolant that is a logical formula
-of linear expressions.
-
-First, consider the case of an interpolation problem between $A_1 \vee
-A_2$ and $B$.  When $(A_1 \vee A_2) \wedge B$ is inconsistent, $A_1
-\wedge B$ is necessarily inconsistent.  Let $\mathcal{I}_1$ denote the
-set of interpolants for $(A_1, B)$.  In the same manner, let
-$\mathcal{I}_2$ denote the set of interpolants for $(A_2, B)$.  We may
-naturally obtain an interpolant set by the intersection $\mathcal{I} =
-\mathcal{I}_1 \cup \mathcal{I}_2$ as a solution for $(A_1 \vee A_2,
-B)$.  When the algorithm computes $\mathcal{I}$ by the previous
-method, and it is $\emptyset$, we need to build a less simple solution
-by
-\[ \mathcal{I}' = \left\lbrace \left( I_1 \vee I_2 \right) \mid
-I_1 \in \mathcal{I}_1 \wedge I_2 \in \mathcal{I}_2 \right\rbrace \]
-Any formulas in the set $\mathcal{I}'$ are the interpolant for $(A_1
-\vee A_2, B)$ because
-\begin{align*}
-\forall I \in \mathcal{I}. & A_1 \vdash I \wedge A_2 \vdash I \wedge I \nvdash B \wedge \\
-& FV(I) = FV(A_1) \cup FV(A_2) \cup FV(B)
-\end{align*}
-
-Second, consider the case of an interpolation problem between $A$ and
-$B_1 \wedge B_2$.
-% TODO:
-
-By combining these two ways of interpolant construction, we may be
-able to build a small solution for more general cases like $(A_1,
-\ldots, A_n)$ and $(B_1, \ldots, B_n)$.  In order to obtain a small
-solution, we need to wisely choose the order of combining an
-interpolant set.
-
-\paragraph{Symmetric Interpolation}
-TODO
-
 \chapter{Horn clause solving}
@@
 obtain a linear constraint
-$ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
+$\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0$.
 Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will
 make $A(x)$ valid for the first clause.  In the same manner, we obtain
-$ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
+$\lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0$
 from the second clause, and 
-$ c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0 $
+$c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0$
 from the third.
@@
 
-Solving this constraint simply gives us a solution $A(x) : x -1 \leq 0$.
+Solving this constraint simply gives us a solution
+$A(x) : x -1 \leq 0$.
 
@@
 \item $V$ is a union of two disjoint sets of vertices; Horn term
-  vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex $u
-  \in V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
-  denoted as $succ(u)$.
+  vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex
+  $u \in V_\Rightarrow$ has eactly one outgoing edge to some
+  $v \in V_T$, denoted as $succ(u)$.
 \item $E$ is a set of labelled edges, which are expressed in the form
@@
 \[ u \mathop{\leadsto}^\theta v \Longleftrightarrow \exists a \in V_\Rightarrow;
-\edgel{u}{\theta}{a} \wedge \edge{a}{v} \] We may omit $\theta$ and
-write $u \leadsto v$ for convenience in later discussions. By the
-assumption that $G$ is acyclic, both $\rightarrow^+$ and $\leadsto^+$ are
-well-founded relations. $v_\bot$ is the greatest element with respect
-to both $\rightarrow^+$ and $\leadsto^+$.
+\edgel{u}{\theta}{a} \wedge \edge{a}{v} \]
+We may omit $\theta$ and write $u \leadsto v$ for convenience in later
+discussions. By the assumption that $G$ is acyclic, both
+$\rightarrow^+$ and $\leadsto^+$ are well-founded relations. $v_\bot$
+is the greatest element with respect to both $\rightarrow^+$ and
+$\leadsto^+$.
 
@@
 
-Without loss of generality, we assume the existence of the \emph{root
-  vertex} $v_\bot \in V_T$ with linear expression labels $1 \leq 0$,
-meaning $\bot$.  This is because if we have a Horn clause $P \implies
-e$ which $e$ is a linear expression, we can convert it to $P \wedge
-\neg e \implies 1 \leq 0$.  We also assume that other vertices with
-linear expression labels have no incoming edges to them.
+Without loss of generality, we assume the existence of the
+\emph{root vertex} $v_\bot \in V_T$ with linear expression labels
+$1 \leq 0$, meaning $\bot$.  This is because if we have a Horn clause
+$P \implies e$ which $e$ is a linear expression, we can convert it to
+$P \wedge \neg e \implies 1 \leq 0$.  We also assume that other
+vertices with linear expression labels have no incoming edges to them.
 
@@
 The algorithm assigns a \emph{predicate template} for all predicate
-variables in $G$. A \emph{predicate templateset} is a set of mappings from
-a predicate variable to a predicate template.  A template for a
+variables in $G$. A \emph{predicate templateset} is a set of mappings
+from a predicate variable to a predicate template.  A template for a
 predicate variable $P(x_1, \ldots, x_n)$ is a formula $\psi$
@@
 \begin{itemize}
-\item $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ is a linear inequality that
-  contains coefficient variables $\mathbf{a}_v$ and a constant $b_v$,
-  and
+\item $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ is a linear inequality
+  that contains coefficient variables $\mathbf{a}_v$ and a constant
+  $b_v$, and
 \item $C_v$ is a set of linear expressions on coefficient variables.
@@
 \end{align*}
-For a predicate variable vertex, $\mathbf{a}_v \mathbf{x}_v + b_v \leq
-0$ serves as a predicate template and $C_v$ describes the parameters
-$\mathbf{a}_v$ and $b_v$.
+For a predicate variable vertex,
+$\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ serves as a predicate
+template and $C_v$ describes the parameters $\mathbf{a}_v$ and $b_v$.
 
@@
 clause, which is a disjunctive Horn clause.  As a simple solution, $P$
-can take a common consequence that both of $x+1 \leq 0$ and $-x+1 \leq
-0$ induce.  However, such a simple linear expression does not exist
-except for $0 \leq 0$, which is equivalent to $\top$, and it is not
-sufficient to satisfy the second clause.
+can take a common consequence that both of $x+1 \leq 0$ and
+$-x+1 \leq 0$ induce.  However, such a simple linear expression does
+not exist except for $0 \leq 0$, which is equivalent to $\top$, and it
+is not sufficient to satisfy the second clause.
 
@@
 
-traverses the input graph in the topological order on
-$V_\Rightarrow$, and concatenate the constraints from predecessors to
-build the problem constraints.  Each process to build a constraint on
-$\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$ is equivalent
-to generate a linear constraint in interpolation.
-
-\chapter{Experiment}
-\label{chap:experiment}
-
-We incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It
-is specifically designed to verify the correctness of higher-orde
-programs written in a small subset of ML.  It infers dependent types
-for input higher-order programs \cite{conf/ppdp/UnnoK09}.
-
-Our implementation make use of Z3 \cite{conf/tacas/MouraB08} as
-an SMT solver.
-
-For simplicity of the implementation, we assume that the linear
-inequalities in the integer space.  Because of this, our algorithm
-lost the completeness.  By using Motzkin's transpose theorem
-\cite{journals/networks/Rajan90}, it is possible to handle linear
-inequalities in both $\mathbf{ax} + b \leq 0$ and $\mathbf{ax} + b <
-0$.
-
-\section{Interpolating theorem prover}
-
-Our interpolating theorem prover is used as a sub-procedure to
-solve Horn clauses.
-
-\begin{itemize}
-\item CSIsat \cite{conf/cav/BeyerZM08}
-\item FOCI \cite{website/foci}
-\item CLP-Prover \cite{website/clp}
-\end{itemize}
-
-\section{Horn clause solver}
-
-\chapter{Related work}
-\label{chap:related}
-\input{related.tex}
-
-\chapter{Future Work}
-\label{chap:future}
-
-\paragraph{Non-linear extension}
-\cite{conf/cav/DaiXZ13}
-
-\chapter{Conclusion}
-\label{chap:conclusion}
-
-\bibliographystyle{plain}
-\bibliography{./biblio}
-
-\end{document}
+% TODO: What's this paragraph?
+traverses the input graph in the topological order on $V_\Rightarrow$,
+and concatenate the constraints from predecessors to build the problem
+constraints.  Each process to build a constraint on $\mathcal{HC}_v$
+for all vertices $v \in V_\Rightarrow$ is equivalent to generate a
+linear constraint in interpolation.

### 2013-08-26 article/interpolation.tex
Other chapters also separated

--- 2013-08-26 article/thesis.tex
@@
-\documentclass[master,final,12pt]{iscs-thesis}
-
-\etitle{Simple Solutions of Acyclic Horn Clauses for Program Verification}
-\eauthor{Yuto Takei}
-\esupervisor{Naoki Kobayashi}
-
-\jtitle{Lipsum}
-\jauthor{Lipsum}
-\jsupervisor{Lipsum}
-
-\supervisortitle{Professor}
-\date{August 28, 2013}
-
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{stmaryrd}
-\usepackage{syntax}
-\usepackage{url}
-
-\newcommand{\sembrack}[1]{[\![#1]\!}
-\newcommand{\edge}[2]{#1\rightarrow#2}
-\newcommand{\edgel}[3]{#1\xrightarrow{#2}#3}
-
-\begin{document}
-
-\begin{eabstract}
-Automated theorem proving plays a crucial role in modern program
-verification techniques.  In particular, in the context of software
-model checking, the interpolation problem and its generalization
-called \emph{Horn clause solving} have attracted special attention.
-The latter problem asks for a solution for a set of acyclic Horn
-clauses containing unknown predicate variables.  The solution, i.e., a
-substitution for predicate variables, then serves as a candidate of
-program invariants.  However, existing methods are unsatisfactory in
-that they tend to generate too specific a solution, which does not
-correspond to program invariants.
-
-To address the problem above, we propose a new algorithm for solving
-Horn clauses consisting of linear arithmetics and unknown predicate
-variables.  The new algorithm tries to find small predicates, based on
-the hypothesis that program's invariants tend to be simple formulas.
-To this end, our algorithm prepares a template of linear arithmetic
-formulas of a fixed size as a candidate of each unknown predicate, so
-that Horn clause solving is reduced to linear programming.  The
-algorithm iteratively increases the size of the templates until a
-solution is found.  It enables us to construct relatively small
-solutions in a reasonable time compared to existing algorithms.  We
-have implemented and incorporated our Horn clause solver into MoCHi, a
-software model checker for functional programs.  We have evaluated the
-effectiveness of our approach through experiments.
-\end{eabstract}
-\begin{jabstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.  Morbi
-vulputate ornare arcu et malesuada.  Proin sem urna, feugiat et velit
-ac, elementum blandit orci.  Etiam ut neque ac felis laoreet venenatis
-eget quis enim.  Ut aliquam dapibus sem egestas molestie.  Ut sit amet
-porttitor orci, id fermentum dui.  In non dui at ante sagittis
-malesuada at in neque.  Fusce euismod velit quis ante ullamcorper
-qsodales.
-\end{jabstract}
-\maketitle
-
-\begin{acknowledge}
-Thank you.
-\end{acknowledge}
-
-\frontmatter
-\tableofcontents
-
-\mainmatter
-
-\chapter{Introduction}
-
-Automated theorem proving plays an important role in the modern
-program verification.  Particularly in the context of verification
-with model checking techniques, an automated theorem prover is used to
-abstract an infinite set of program states into a finite one so that
-the model checker can virtually explore all possible program execution
-paths.
-
-One of the most common methods of such abstraction is predicate
-abstraction\cite{conf/cav/GrafS97} within CounterExample-Guided
-Abstraction Refinement (CEGAR) framework
-\cite{conf/cav/ClarkeGJLV00,conf/popl/BallR02,conf/popl/HenzingerJMS02}.
-An automated theorem prover finds appropriate predicates for every
-program location.  An execution states at those locations is then
-expressed by a Boolean valuation of the given predicates.  If the
-abstraction is too coarse and a model checker discovers an infeasible
-error execution path in an abstract system, the prover computes an
-additional predicate at those program locations to refine the
-abstraction predicates.
-
-In order to obtain an additional abstraction predicate at program
-locations along the infeasible counterexample, one class of automated
-theorem provers, called an interpolating theorem prover
-\cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07},
-computes a separating predicate called an interpolant between two sets
-of constraints; the one consists of the constraints from the program
-entry point to the location, and the other consists of the constraints
-from the location to the program failure.  Those constraints include
-variable assignments and assertions to be satisfied originated from
-conditional branches.
-
-The computed predicate by an interpolating theorem prover is satisfied
-at the location under the execution in the concrete model along the
-discovered counterexample, and any execution paths which satisfies the
-predicate does not lead to the failure state.  After the refinement,
-the infeasible path is ruled out and no longer found by the model
-checker.
-
-However, existing interpolating theorem provers may return too complex
-a solution which is heavily affected by the constraints of the
-specific path.  It is not favorable in terms of the program
-verification because a complex predicate may not be sufficient to rule
-out similar spurious paths in programs with loops and recursion calls.
-This causes the model checker to discover spurious paths infinitely
-which pass the same program location by loops or recursion calls at
-different numbers of iteration, and the verification process does not
-to terminate.
-
-To avoid this, it is desired to discover an invariant formula for the
-predicate abstraction.  Based on the hypothesis that program's
-invariants tend to be simple formulas, we propose a new interpolating
-algorithm which tries to minimize the number of conjunctions and
-disjunctions in a solution.  The predicate obtained by this algorithm
-may become similar to the invariant, and the number of trials in model
-checking may be reduced.
-
-Additionally, we extend our interpolating algorithm to solve a
-symmetric interpolation problem, in which the algorithm computes
-predicates at multiple program locations along a path at once.  This
-enables the abstraction predicates at multiple locations to be updated
-at the same computation.
-
-Finally, we extend our algorithm to solve Horn clause solving
-problems.  It allows us to obtain predicates which are highly likely
-to be program invariants, by computing the same abstraction predicates
-inside loops and recursion calls on different iterations, and by
-unifying predicates at the same location over multiple paths.  The
-problems are generated by encoding constraints over execution paths
-with unknown predicates which describe functions' constraints between
-parameters and return values.
-
-The rest of the paper is structured as follows.
-Chapter~\ref{chap:interpolation} proposes a new interpolating
-algorithm to obtain simple solutions, and extends it to solve
-symmetric interpolation problems.  Chapter~\ref{chap:horn} describes a
-new Horn clause solving algorithm by extending the previous
-interpolating algorithm.  Chapter~\ref{chap:experiment} describes
-experiments and illustrates the result to confirm the effect of our
-proposed algorithm.  Chapter~\ref{chap:future} gives a discussion of
-our improvements and proposes further improvements to make.
-Chapter~\ref{chap:related} mentions related work done by others.
-Finally, Chapter~\ref{chap:conclusion} reviews the impact of our
-research and brings the conclusion.
-
 \chapter{Interpolation}
@@
 other, we call a new preposition $I$ a \emph{Craig interpolant}
-\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$
-is implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
+\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$ is
+implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
 only free variables that appears in both $A$ and $B$.  If $A \wedge B$
@@
 TODO
-
-\chapter{Horn clause solving}
-\label{chap:horn}
-
-We propose a new algorithm for solving recursion-less Horn clauses to
-generate a solution which is as simple as possible.  This enables us
-to obtain appropriate predicates that are suitable for predicate
-abstraction in the program verification techniques.
-
-The construction of the algorithm roughly consists of two parts.
-\begin{itemize}
-\item \textbf{Iterative Constraint Generation} is a process to build a
-  constraint to compute a solution.  It starts from generating the
-  weakest constraint in the sense that the resulting solution would
-  not be simple yet.  The process iteratively strengthen a constraint
-  until the strongest one while preserving the constraint
-  satisfiability.  The strongest constraint gives a simple solution.
-\item \textbf{Problem Restructuring} is a process to restructure the
-  problem to obtain a less simpler solution, if the constraint become
-  unsatisfiable during the iterative constraint generation.  This
-  process considers the unsatisfiable core of the constraint, and
-  wisely enlarge the solution size.
-\end{itemize}
-
-All predicate variables in an input problem are assigned simple
-templates.  Those predicate templates are parameterized.  In the
-iterative constraint generation process, algorithm builds a linear
-constraint that parameters in templates must satisfy.  This is done in
-a constructive manner, and constraints are build in the order of
-implication dependency in Horn clauses.  That is, when an input
-problem has a Horn clause thath has a predicate variable $P$ on its
-left-hand and $Q$ on its right-hand, a constraint to describe
-parameters in $P$'s template are first build before those in $Q$'s
-template.  When all parameters in the whole problem are described, the
-built constraint are evaluated to obtain a model.  Its instanciation
-becomes an appropriate solution for the input problem.
-
-The novelty of our algorithm lies in the incremental constraint
-refinement.  The algorithm tries to solve from the weakest constraint
-that an input problem allows that does not give a simple solution, and
-gradually strengthen the constraint until it reaches to the simplest
-solution that the problem has.  When a constraint becomes
-unsatisfiable during this step, the algorithm analyzes the
-unsatisfiable core and transform the problem to obtain a less simple
-solution.
-
-This construction of the algorithm enables us to compute a relatively
-simpler solution compared to existing algorithms while preserving the
-computation cost in a reasonable time.
-
-\section{Example}
-
-\paragraph {Tree interpolation}
-First, consider the following example.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-y \leq 2 & \implies B(y) \\
-A(x) \wedge B(y) & \implies 3x+2y \leq 10
-\end{align*}
-
-In this example, we would like to compute linear arithmetic formulas
-for the predicate variables $A$ and $B$.  They appear only once on
-both left-hand and right-hand globally in the problem.
-
-We try to obtain a simple solution that is expressed by linear
-expression templates, whose coefficients and constants are
-parameterized.  Assume that they have following predicate templates.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\
-B(y) : c_3 y + c_4 \leq 0
-\end{align*}
-
-By naturally encoding the first Horn clause with Farkas's lemma, we
-obtain a linear constraint
-$ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
-Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will
-make $A(x)$ valid for the first clause.  In the same manner, we obtain
-$ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
-from the second clause, and
-$ c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0 $
-from the third.
-
-One of the models satisfying all the constraints above is
-$( c_1, c_2, c_3, c_4, \lambda_1, \lambda_2, \lambda_3 ) = (3, -3, 2, -4, 3, 2, 1)$.
-By assigning them to the template, we obtain the following solution.
-
-\begin{align*}
-A(x) : 3 x - 3 \leq 0 \\
-B(y) : 2 y - 4 \leq 0
-\end{align*}
-
-\paragraph {DAG problem solving}
-Second, we explain a case that some predicate variables appear
-multiple times on left-hand side in different Horn clauses.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-A(x) \wedge A(y) & \implies x+y \leq 2
-\end{align*}
-
-Assigning a template $A(x) : c_1 x + c_2 \leq 0$, we obtain the same
-constraint for the first clause as the previous example.
-
-\begin{align} \label{eq:constr1}
-\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0
-\end{align}
-
-Sinve we have $A$'s appearance twice on the second clause, it is
-possible for $A$ to have two expressions each of which is used for
-$A(x)$ and $A(y)$.  Although we need to use the same template for
-multiple occurrences in order to compute the simplest solution, it may
-make the whole constraints unsatisfiable if the simplest solution does
-not exist.
-
-Therefore, we generate a weaker constraint for the second clause that
-allows different expressions of $A$ for multiple occurrence.  If the
-whole constraint is satisfiable in the case, we try to solve in a
-stronger constraint to generate a simple solution.
-
-In this specific case, it is virtually equivalent to solve the
-following problem in a weaker constraint.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 1 & \implies A_1(x) \\
-x \leq 1 & \implies A_2(x) \\
-A_1(x) \wedge A_2(y) & \implies x+y \leq 2 \\
-\\
-\text{Templates:} \\
-A_1(x) \text{ : } & c_{1,1} x + c_{2,1} \leq 0 \\
-A_2(x) \text{ : } & c_{1,2} x + c_{2,2} \leq 0 \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-\end{align*}
-
-By duplicating the constraint~\ref{eq:constr1} of $A$ for newly
-created $A_1$ and $A_2$, we obtain the whole constraint as follows.
-
-\begin{align*}
-& \lambda_{1,1} = c_{1,1} \wedge - \lambda_{1,1} \geq c_{2,1} \wedge \lambda_{1,1} \geq 0 \wedge \\
-& \lambda_{1,2} = c_{1,2} \wedge - \lambda_{1,2} \geq c_{2,2} \wedge \lambda_{1,2} \geq 0 \wedge \\
-& c_{1,1} = \lambda_2 \wedge c_{1,2} = \lambda_2 \wedge c_{2,1} + c_{2,2} \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-This constraint is satisfiable and one model is
-\[ ( c_{1,1}, c_{1,2}, c_{2,1}, c_{2,2}, \lambda_{1,1}, \lambda_{1,2}, \lambda_2 ) =
-( 1, -1, 1, -1, 1, 1, 1 ) \] and the solution then becomes
-$A(x) : (x -1 \leq 0) \wedge (x -1 \leq 0)$.
-
-Since the weak constraint is satisfiable, we now refine the constraint
-to obtain the better solution.  That is, simply eliminate the
-duplication to build the following constraint for the original
-template.
-
-\begin{align*}
-& \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \\
-& c_1 = \lambda_2 \wedge c_1 = \lambda_2 \wedge 2 c_2 \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-Solving this constraint simply gives us a solution $A(x) : x -1 \leq 0$.
-
-\paragraph {Disjunctive Horn clause solving}
-Finally, we explain a case that an input set of Horn clauses contain
-disjunctions on left-hand side.
-
-\begin{align*}
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 \vee -x+1 \leq 0 & \implies B(x) \\
-A(x) \wedge B(x) & \implies 1 \leq 0
-\end{align*}
-
-Same as the previous examples, templates for predicate variables are
-prepared.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\ B(x) : c_3 x + c_4 \leq 0
-\end{align*}
-
-The constraints generated are follows.
-
-\begin{align*}
-& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
-& \lambda_3 = c_3 \wedge \lambda_3 \geq c_4 \wedge \lambda_3 \geq 0 \wedge \\
-& - \lambda_4 = c_3 \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
-& c_1 + c_3 = 0 \wedge c_2 + c_4 > 0
-\end{align*}
-
-The constraints here become unsatisfiable, and one possible
-unsatisfiable core is:
-
-\begin{align*}
-\lambda_3 & = c_3 \\
-\lambda_3 & \geq c_4 \\
-- \lambda_4 & = c_3 \\
-\lambda_4 & \geq c_4 \\
-c_2 + c_4 & > 0 \\
-0 & \geq c_2
-\end{align*}
-
-Because the unsatisfiable core contains constraints originating from
-the disjunctions to be expressed in a single term template.  This let
-us determine to relax the template to allow a enlarged solution.
-
-Now we slightly change the problem and the template.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A(x) \wedge B_1(x) & \implies 1 \leq 0 \\
-A(x) \wedge B_2(x) & \implies 1 \leq 0 \\
-\\
-\text{Templates:} \\
-A(x) \text{ : } & c_1 x + c_2 \leq 0 \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-When we construct the constraints, as same as previous, we again
-obtain an unsatisfiable core:
-
-\begin{align*}
-\text{Constraint:} \\
-TODO \\
-\\
-\text{Unsatisfiable core:} \\
-TODO
-\end{align*}
-
-This time, the unsatisfiable core contains the constraints that
-originate from
-
-TODO
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A_i(x) \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A_i(x) \wedge B_j(x) & \implies 1 \leq 0 \quad (i,j \in \left\lbrace 1,2 \right\rbrace ) \\
-\\
-\text{Templates:} \\
-A_i(x) \text{ : } & c_{1,i} x + c_{2,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-Finally, we obtain the following constraint:
-
-TODO
-
-The process also generates a predicate template for every predicate
-variable that the solution will become.  under certain solution
-template.
-
-\section{Preliminaries}
-
-For ease of discussion, our algorithm handles an input set of Horn
-clauses by a \emph{Horn graph} that express the equivalent problem.
-The graph is then used throughout the algorithm to represent the
-problem.  A Horn graph is a labelled directed acyclic graph
-$G=(V,E,\varphi)$.
-\begin{itemize}
-\item $V$ is a union of two disjoint sets of vertices; Horn term
-  vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex $u
-  \in V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
-  denoted as $succ(u)$.
-\item $E$ is a set of labelled edges, which are expressed in the form
-  $(u,v,\theta)$ where $\theta$ is a finite map from variables to
-  variables.
-\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
-  term vertices, where L is a set of elements of a predicate variable
-  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
-  inequality $a_1 x_1 + \cdots + a_n x_n + b \leq 0$.
-\end{itemize}
-There are two kinds of edges in a Horn graph;
-\begin{itemize}
-\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
-  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
-\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
-  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
-  empty.
-\end{itemize}
-
-We define the relation $\leadsto$ over $V_T$ as follows.
-\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow \exists a \in V_\Rightarrow;
-\edgel{u}{\theta}{a} \wedge \edge{a}{v} \] We may omit $\theta$ and
-write $u \leadsto v$ for convenience in later discussions. By the
-assumption that $G$ is acyclic, both $\rightarrow^+$ and $\leadsto^+$ are
-well-founded relations. $v_\bot$ is the greatest element with respect
-to both $\rightarrow^+$ and $\leadsto^+$.
-
-The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
-Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$.
-\begin{align*}
-\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(succ(v)) \\
-\llbracket G \rrbracket & = \bigwedge_{v \in V_\Rightarrow} \mathcal{HC}_v
-\end{align*}
-We implicitly assume each $\mathcal{HC}_v$ is universally quantified
-by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.
-
-Horn term vertices with predicate variable labels are unique each
-other that any of them do not have the same variable on their labels.
-From now on, we may denote a Horn term vertex with a predicate
-variable label by its predicate variable name if no ambiguity is
-present.
-
-Without loss of generality, we assume the existence of the \emph{root
-  vertex} $v_\bot \in V_T$ with linear expression labels $1 \leq 0$,
-meaning $\bot$.  This is because if we have a Horn clause $P \implies
-e$ which $e$ is a linear expression, we can convert it to $P \wedge
-\neg e \implies 1 \leq 0$.  We also assume that other vertices with
-linear expression labels have no incoming edges to them.
-
-For simplicity of discussion, we define \emph{term arity} $k_v$ for
-every Horn term vertex $v \in V_T$ with incoming edges.
-\begin{align*}
-k_v =
-\begin{cases}
-\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
-0 & \mbox{if } \varphi(v) = \bot
-\end{cases}
-\end{align*}
-We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.
-
-A variable mapping label $\theta$ on edges is used to map the index of
-one predicate variable to another. (TODO?)
-
-(Example?)
-
-The algorithm assigns a \emph{predicate template} for all predicate
-variables in $G$. A \emph{predicate templateset} is a set of mappings from
-a predicate variable to a predicate template.  A template for a
-predicate variable $P(x_1, \ldots, x_n)$ is a formula $\psi$
-constructed from the following definition.
-\begin{align*}
-\psi ::= & a_1 x_1 + \cdots + a_n x_n + b \leq 0 \mid \\
-& \psi \wedge \psi \mid \psi \vee \psi
-\end{align*}
-A linear expression in the definition above may also be represented in
-a equivalent vector form $\mathbf{a} \mathbf{x} + b \leq 0$ for short.
-Parameters $\mathbf{a}$ and $b$ are specified by a linear constraint.
-
-The solution of $G$ is a map $\rho$ from predicate variables to linear
-predicates of the form $\lambda x_1, \ldots ,x_n. \psi$ where
-parameters in $\psi$ are replaced with constant and $\rho[G]$ becomes
-a tautology.  A solution is \textit{simple} if $\rho[P]$ is of the
-form $\mathbf{a} \mathbf{x} + b \leq 0$ for every $P$ in
-$\mathrm{dom}(\rho)$.
-
-The problem of solving Horn clauses (represented by $G$) is to find a
-solution of $G$ (and a simple solution if possible).
-
-\section{Simple solution}
-
-We first propose an algorithm to compute a simple solution for the
-given Horn graph $G$ if one exists.  Later we add modifications to the
-algorithm for the case that a simple solution does not exist for $G$.
-In such a case, we transform the input graph $G$ to have a larger
-solution template.
-
-For each vertex $v \in V$, we define
-$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v + b_v \leq 0, C_v)$
-, where
-\begin{itemize}
-\item $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ is a linear inequality that
-  contains coefficient variables $\mathbf{a}_v$ and a constant $b_v$,
-  and
-\item $C_v$ is a set of linear expressions on coefficient variables.
-\end{itemize}
-Intuitively, it denotes the set of inequalities:
-\begin{align*}
-\left\lbrace
- \mathbf{a}_v \mathbf{x}_v + b_v \leq 0 \mid
- \exists \left( \textsc{FV}(C_v)
-  \setminus (\mathbf{a}_v \cup \left\lbrace b_v \right\rbrace
- \right); C_v
-\right\rbrace
-\end{align*}
-For a predicate variable vertex, $\mathbf{a}_v \mathbf{x}_v + b_v \leq
-0$ serves as a predicate template and $C_v$ describes the parameters
-$\mathbf{a}_v$ and $b_v$.
-
-$C_v$ is defined by well-founded induction on $v$ with respect to the
-relation $\rightarrow^+$. Then, $C_v$ can be shown as follows.
-\begin{itemize}
-\item For $v \in V_\Rightarrow$;
-\begin{align*}
-C_v & =
- \bigcup_{\edgel{u}{\theta}{v}} C_u \cup
- \bigcup_{0 < i \leq K}
- \left\lbrace
-  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{v}} \mathbf{a}_{u, \theta^{-1} (i)}
- \right\rbrace \\
- & \hspace{2cm} \cup
- \left\lbrace
-  b_v \leq \sum_{\edgel{u}{\theta}{v}} b_u
- \right\rbrace \cup
- \bigcup_{k_{succ(v)} < i \leq K}
- \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that
-  $v \neq v_{\bot} \wedge \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0$
-  where $\mathbf{a}_0$ and $b_0$ are constants;
-\begin{align*}
-C_v =
-\left\lbrace
- \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
- b_v = \lambda_v b_0
-\right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that $\varphi(v) = P(\mathbf{x}_v)$;
-\begin{align*}
-C_v = & \bigcup_{\edge{u}{v}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{a}_v, b_u = b_v \right\rbrace
-\right)
-\end{align*}
-\item For $v_\bot$;
-\begin{align*}
-C_{v_\bot} = \bigcup_{\edge{u}{v_\bot}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{0}, b_u > 0 \right\rbrace \right)
-\end{align*}
-\end{itemize}
-
-Any model $\sigma$ to the constraint $C_{v_\bot}$ gives a simple
-solution for $G$ if $C_{v_\bot}$ is satisfiable, i.e., we should have:
-
-\begin{quote}
-If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
-\begin{align*}
- \rho = \left\lbrace
-  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \mid
-  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
- \right\rbrace
-\end{align*}
-is a solution of $G$.
-\end{quote}
-
-If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for $G$.
-There are two cases that a simple solution does not exist for a
-predicate variable.
-\begin{itemize}
-\item A predicate variable appears multiple times on left-hand of Horn
-  clauses, and those appearances require different expressions as an
-  assumption.
-\item A predicate variable is implied by a disjunctive Horn clause,
-  which has disjunctions on its left-hand, and a common premise that
-  the disjunctions induce is not strong enough for solving a problem.
-\end{itemize}
-
-The first case occurs in a case like follows.
-\begin{align*}
-% conj.txt ... originated from uc/01.txt
-\left( x \leq 0 \right) \wedge \left( y \leq 0 \right) & \implies P(x,y) \\
-P(x,a) \wedge P(b,y) & \implies x+y \leq 0
-\end{align*}
-In this example, the predicate variable $P$ appears twice on the
-second clause and they require different premises; $x \leq 0$ for the
-first and $y \leq 0$ for the second.  They cannot be represented by
-$\mathbf{ax} + b \leq 0$ template at the same time.
-
-The second case occurs in a case like follows.
-\begin{align*}
-% disj-simp.txt ... originated from 23simp_neg.txt
-\left( x+1 \leq 0 \right) \vee \left( -x+1 \leq 0 \right) & \implies P(x) \\
-P(x,y) \wedge \left( x \leq 0 \right) \wedge \left( -x \leq 0 \right) & \implies \bot
-\end{align*}
-In this example, the predicate variable $P$ is implied by the first
-clause, which is a disjunctive Horn clause.  As a simple solution, $P$
-can take a common consequence that both of $x+1 \leq 0$ and $-x+1 \leq
-0$ induce.  However, such a simple linear expression does not exist
-except for $0 \leq 0$, which is equivalent to $\top$, and it is not
-sufficient to satisfy the second clause.
-
-Both cases can be fixed by enlarging the solution template.
-
-\section {Graph transformation}
-
-If a simple solution does not exist for $G$, we try to give a less
-simpler solution by transforming $G$ and relaxing the constraint
-$C_{v_\bot}$.  This transformation takes place by considering the
-unsatisfiable core $\mathcal{U}$.  An unsatisfiable core $\mathcal{U}$
-is a subset of $C_{v_\bot}$ which is inconsistent itself.  Depending
-on the reason for the absence of a simple solution, the way of the
-transformation differs.
-
-Algorithm~\ref{alg:solveHorn} shows the updated pseudo-code of our
-Horn clause solving algorithm with the graph transformation.  It takes
-an input Horn graph and returns a mapping between predicate variables
-and linear arithmetic formulas.
-
-\newcommand{\solveHorn}{\ensuremath{\mbox{\sc SolveHornGraph}}}
-\newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
-\newcommand{\transGraph}{\ensuremath{\mbox{\sc TransformGraph}}}
-
-\begin{algorithm}
-\caption{$ \solveHorn (G) $}\label{alg:solveHorn}
-\begin{algorithmic}
-\STATE {$v_\bot \gets $ the root node of $G$}
-\STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-\STATE {$T \gets $ initial predicate templateset for $G$}
-\REPEAT
-  \STATE {$C_{v_\bot} \gets \genConstr (G, v_\bot, V_\star)$}
-  \IF {$C_{v_\bot}$ is satisfiable}
-    \STATE {$\sigma \gets $ one model of $C_{v_\bot}$}
-    \STATE {$V_\star \gets V_\star \cup \left\lbrace u \in V_T \mid \forall v \in V_\star; u \leadsto v \right\rbrace$}
-  \ELSE
-    \STATE {$\mathcal{U} \gets $ unsatisfiable core for $C_{v_\bot}$}
-    \STATE {$(G, T) \gets \transGraph (G, T, \mathcal{U})$}
-    \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-  \ENDIF
-\UNTIL {$V_\star = V_T$}
-\RETURN {$\sigma \left( T \right)$}
-\end{algorithmic}
-\end{algorithm}
-\transGraph~procedure determines the unsatisfiable core of the
-constraint and transform the Horn graph $G$ and predicate templates.
-As mentioned earlier, two ways of graph transformation exist.
-
-\paragraph{Conjunction split}
-For the first case previously mentioned, the unsatisfiable core
-$\mathcal{U}$ contains coeffcient constraint expressions with a
-coefficient parameter of certain predicate variable.  If the predicate
-variable $P$ is the cause of unsatisfiability, some constraint
-expressions in the unsatisfiable core have a parameter $a_{P,i}$
-and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is copied to
-$P_1, \ldots, P_n$ so that the different appearances of $P$ on
-left-hand of Horn clauses can take different premises.  The solution
-template of $P$ becomes the conjunction of $P_1, \ldots, P_n$.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-Note that after this duplication of the predicate vertex $P$, in order
-to induce different linear expressions for $P_1, \ldots, P_n$, it may
-be necessary for preceeding predicate variable vertices in $G$ to be
-duplicated in the same manner as $P$. It is because that different
-derivation of $P$'s implication relation may be required to get
-different consequences.
-
-However, succeeding predicate vertices of $P$ in $G$ and other
-vertices than predecessors are not affected because those implications
-that have $P$ in their assumption can selectively choose the premise
-that $P_1, \ldots, P_n$ leads.
-
-Therefore, this graph transformation only affects the predecessors of
-$P$.
-
-\paragraph{Disjunction split}
-For the second case, the unsatisfiable core $\mathcal{U}$ contains
-coefficient constraints that unifies induced linear expressions from
-each term in disjunctions to be the same expression.  Similarly as
-previous, if the predicate variable $P$ is the cause of
-unsatisfiability, some constraint expressions in the unsatisfiable
-core have a parameter $a_{P,i}$ and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is separated to
-$P_1, \ldots, P_n$ each of which is implied by a term, which was
-originally a part of a disjunction to imply $P$.  This intuitively
-enables us to avoid generating a common linear expression that
-disjunctive terms induces, and to propagate each term's premises to
-$P$'s successors instead.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-After this duplication of the predicate vertex $P$, unlike the
-conjunction split, this results in producing a disjunction at
-succeeding predicate variables vertices ($R$ in this case).
-\begin{align*}
-(P_1 \wedge Q) \vee (P_2 \wedge Q) \implies R
-\end{align*}
-Also, predicate variables that appeared with $P$ in the same
-assumption ($Q$ in this case) needs to be duplicated by conjunction,
-so as to obtain different premises that go along with $P_i$.
-Therefore, the effect of disjunction split may reach to neighboring
-predicate vertices.
-
-Consider these two ways of transformation, we show a pseudo-code for
-\transGraph~procedure.  The procedure is given the graph $G$, the
-predicate templateset $T$, and the unsatisfiable core
-$\mathcal{U}$. It returns the transformed graph $G'$ and the new
-templateset $T'$.
-
-\begin{algorithm}
-\caption{$ \transGraph (G, T, \mathcal{U}) $}\label{alg:transGraph}
-\begin{algorithmic}
-% TODO:
-\RETURN {$\left( G', T' \right)$}
-\end{algorithmic}
-\end{algorithm}
-
-\paragraph{Combinatorial explosion in disjunctions}
-Virtually, if multiple disjunctions exists in an input set of Horn
-clauses, a na\"{i}ve way of building a solution is solve all
-combinations of choices out of every disjunction and concatenate each
-expression with conjunctions and/or disjunctions appropriately.  This
-may cause a combinatorial explosion.
-
-Conside example below. $e_i$ represents some conjunctions of linear
-expressions.
-% TODO: Can't we replace e_i with actual expressions?
-\begin{align*}
-(e_1 \vee e_2) \implies P \\ (e_3 \vee e_4) \implies Q \\ P \wedge Q \implies \bot
-\end{align*}
-In this example, a na\"{i}ve algorithm may solve following four
-problems separately.
-\begin{itemize}
-\item $e_1 \implies P_1 \qquad e_3 \implies Q_1 \qquad P_1 \wedge Q_1 \implies \bot$
-\item $e_2 \implies P_2 \qquad e_3 \implies Q_2 \qquad P_2 \wedge Q_2 \implies \bot$
-\item $e_1 \implies P_3 \qquad e_4 \implies Q_3 \qquad P_3 \wedge Q_3 \implies \bot$
-\item $e_2 \implies P_4 \qquad e_4 \implies Q_4 \qquad P_4 \wedge Q_4 \implies \bot$
-\end{itemize}
-Then the solution is built by follows.
-\begin{itemize}
-\item $P = ( P_1 \vee P_2 ) \wedge ( P_3 \vee P_4 )$
-\item $Q = ( P_1 \vee P_3 ) \wedge ( P_2 \vee P_4 )$
-\end{itemize}
-However, this method is not efficient.
-
-\paragraph{Algorithm}
-To avoid this, our algorithm lets a predicate variable $P$ which is
-disjunctively implied on the Horn graph $G$ to have a linear
-expression that is consequences of all terms in the disjunction.  This
-prevents the possibility of choices for disjunctions to be propagated
-to succeeding vertices. The simple linear expression is propagated
-instead.
-
-On the other hand, the algorithm allows a predicate variable $P$ that
-appears on multiple implications or appears multiple times in a single
-implication to take different linear expressions as a premise at
-different appearances.  This is for the algorithm to determine the
-cause of unsatisfiability efficiently.  It is done by copying a
-predicate templates and its parameters' constraint with appropriate
-renaming.
-
-The algorithm first execute this duplication on all predicate
-variables with multiple appearances.  It is the weakest constraint
-that the algorithm makes for the certain $G$.  Then, the algorithm
-gradually reduces the number of vertices that are duplicated by each
-time testing the satisfiability of the constraint.  When the
-constraint bebcome unsatisfiable at the certain iteration, the
-algorithm can determine a predicate vertex which has to be split in a
-conjunctive template.  If all the predicate variable constraints are
-no longer duplicated and the constraint is the strongest, any model
-for such constraint gives a simple solution under the Horn graph $G$
-and the templateset $T$.
-
-Note that duplicating the constraint in a na\"{i}ve way may cause the
-constraint to exponentially grow in size.  Our algorithm prevent it by
-simplifying a predicate vertex's constraint $C_v$ by quantifier
-elimination when $v$ is duplicated in generating linear constraint.
-For a predicate variable vertex $v_P$, variables except for the
-coefficient parameter $\mathbf{a}_P$ and the constant parameter $b_P$
-are eliminated.  Due to the constraints generation process, the
-simplified constraint is also linear and can be efficiently check the
-satisfiability.
-
-We finally show the modified version of pseudo-code for $\genConstr$
-procedure.  It produces the constraint $C_v$ for a given vertex $v$,
-considering constraint duplication and simplification.
-
-\begin{algorithm}
-\caption{$ \genConstr (G, v, V_\star) $}\label{alg:genConstr}
-\begin{algorithmic}
-\STATE {$C \gets \emptyset$}
-\FORALL {$v_\Rightarrow \in preds_G(v)$}
-  \FORALL {$v_p \in preds_G(v_\Rightarrow)$}
-    \IF {$\varphi (v_p)$ is a predicate variable}
-      \STATE {$C \gets C \cup \genConstr (G, v_p, visited)$}
-    \ELSE
-      \STATE { $id$ }
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-traverses the input graph in the topological order on
-$V_\Rightarrow$, and concatenate the constraints from predecessors to
-build the problem constraints.  Each process to build a constraint on
-$\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$ is equivalent
-to generate a linear constraint in interpolation.
-
-\chapter{Experiment}
-\label{chap:experiment}
-
-We incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It
-is specifically designed to verify the correctness of higher-orde
-programs written in a small subset of ML.  It infers dependent types
-for input higher-order programs \cite{conf/ppdp/UnnoK09}.
-
-Our implementation make use of Z3 \cite{conf/tacas/MouraB08} as
-an SMT solver.
-
-For simplicity of the implementation, we assume that the linear
-inequalities in the integer space.  Because of this, our algorithm
-lost the completeness.  By using Motzkin's transpose theorem
-\cite{journals/networks/Rajan90}, it is possible to handle linear
-inequalities in both $\mathbf{ax} + b \leq 0$ and $\mathbf{ax} + b <
-0$.
-
-\section{Interpolating theorem prover}
-
-Our interpolating theorem prover is used as a sub-procedure to
-solve Horn clauses.
-
-\begin{itemize}
-\item CSIsat \cite{conf/cav/BeyerZM08}
-\item FOCI \cite{website/foci}
-\item CLP-Prover \cite{website/clp}
-\end{itemize}
-
-\section{Horn clause solver}
-
-\chapter{Related work}
-\label{chap:related}
-\input{related.tex}
-
-\chapter{Future Work}
-\label{chap:future}
-
-\paragraph{Non-linear extension}
-\cite{conf/cav/DaiXZ13}
-
-\chapter{Conclusion}
-\label{chap:conclusion}
-
-\bibliographystyle{plain}
-\bibliography{./biblio}
-
-\end{document}

### 2013-08-26 article/intro.tex
Other chapters also separated

--- 2013-08-26 article/thesis.tex
@@
-\documentclass[master,final,12pt]{iscs-thesis}
-
-\etitle{Simple Solutions of Acyclic Horn Clauses for Program Verification}
-\eauthor{Yuto Takei}
-\esupervisor{Naoki Kobayashi}
-
-\jtitle{Lipsum}
-\jauthor{Lipsum}
-\jsupervisor{Lipsum}
-
-\supervisortitle{Professor}
-\date{August 28, 2013}
-
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage{stmaryrd}
-\usepackage{syntax}
-\usepackage{url}
-
-\newcommand{\sembrack}[1]{[\![#1]\!}
-\newcommand{\edge}[2]{#1\rightarrow#2}
-\newcommand{\edgel}[3]{#1\xrightarrow{#2}#3}
-
-\begin{document}
-
-\begin{eabstract}
-Automated theorem proving plays a crucial role in modern program
-verification techniques.  In particular, in the context of software
-model checking, the interpolation problem and its generalization
-called \emph{Horn clause solving} have attracted special attention.
-The latter problem asks for a solution for a set of acyclic Horn
-clauses containing unknown predicate variables.  The solution, i.e., a
-substitution for predicate variables, then serves as a candidate of
-program invariants.  However, existing methods are unsatisfactory in
-that they tend to generate too specific a solution, which does not
-correspond to program invariants.
-
-To address the problem above, we propose a new algorithm for solving
-Horn clauses consisting of linear arithmetics and unknown predicate
-variables.  The new algorithm tries to find small predicates, based on
-the hypothesis that program's invariants tend to be simple formulas.
-To this end, our algorithm prepares a template of linear arithmetic
-formulas of a fixed size as a candidate of each unknown predicate, so
-that Horn clause solving is reduced to linear programming.  The
-algorithm iteratively increases the size of the templates until a
-solution is found.  It enables us to construct relatively small
-solutions in a reasonable time compared to existing algorithms.  We
-have implemented and incorporated our Horn clause solver into MoCHi, a
-software model checker for functional programs.  We have evaluated the
-effectiveness of our approach through experiments.
-\end{eabstract}
-\begin{jabstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.  Morbi
-vulputate ornare arcu et malesuada.  Proin sem urna, feugiat et velit
-ac, elementum blandit orci.  Etiam ut neque ac felis laoreet venenatis
-eget quis enim.  Ut aliquam dapibus sem egestas molestie.  Ut sit amet
-porttitor orci, id fermentum dui.  In non dui at ante sagittis
-malesuada at in neque.  Fusce euismod velit quis ante ullamcorper
-qsodales.
-\end{jabstract}
-\maketitle
-
-\begin{acknowledge}
-Thank you.
-\end{acknowledge}
-
-\frontmatter
-\tableofcontents
-
-\mainmatter
-
 \chapter{Introduction}
@@
 theorem provers, called an interpolating theorem prover
-\cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07},
-computes a separating predicate called an interpolant between two sets
-of constraints; the one consists of the constraints from the program
+\cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07}, computes a
+separating predicate called an interpolant between two sets of
+constraints; the one consists of the constraints from the program
 entry point to the location, and the other consists of the constraints
@@
 research and brings the conclusion.
-
-\chapter{Interpolation}
-\label{chap:interpolation}
-
-Among various program verification techniques, when software model
-checking with predicate abstraction is adopted, the interpolation is
-used to compute the abstraction predicate along a spurious
-counterexample that a model checker discovered.  It is done by
-computing a Craig interpolant between the strongest postcondition at
-the location and the weakest precondition to the failure point at
-every program location.
-
-However, by the existing interpolating theorem prover, we may not be
-able to obtain a simple solution.  Consider the example below.
-\begin{align*}
-\varphi_1 : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
-\varphi_2 : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
-\end{align*}
-Although one may be able to obtain $x = y$ as the simplest
-interpolation between $\varphi_1$ and $\varphi_2$,
-CSIsat\cite{conf/cav/BeyerZM08} returns
-\[ TODO \]
-as a solution.
-
-We propose an interpolation algorithm which returns as simple a
-solution as possible in terms of the number of conjunctions and
-disjunctions.  Our algorithm focuses on interpolating problems on
-linear arithmetic and it works in a constructive manner.
-
-\section{Example}
-
-Consider the following interpolation problem.
-
-\begin{align*}
-\frac
-{A: (x \leq a) \wedge (a + 1 \leq y)}
-{B: (y \leq b) \wedge (b + 1 \leq x)}
-\end{align*}
-
-Here, the linear arithmetic formulas A and B are inconsistent.  The
-interpolant I for this problem is $x-y+1 \leq 0$, which is implied by
-$A$ and inconsistent with $B$.  The vocabulary of $I$ is $\left\lbrace
-x,y \right\rbrace$ and is over $A$'s vocabulary $\left\lbrace x,y,a
-\right\rbrace$ and $B$'s $\left\lbrace x,y,b \right\rbrace$.
-
-In computing the interpolant, we make a linear constraint of a
-interpolant by applying Farkas's lemma.  First, we assing weight
-parameters to every expression in conjunction form.
-
-\begin{align*}
-\begin{array}{r r r r r r l}
-\lambda_1 : & -a & & +x & & & \leq 0 \\
-\lambda_2 : & a & & & -y & +1 & \leq 0 \\
-\hline
-\lambda_3 : & & -b & & +y & & \leq 0 \\
-\lambda_4 : & & b & -x & & +1 & \leq 0
-\end{array}
-\end{align*}
-
-Because all expressions induce inconsistent as a whole, the sum of
-weighted expressions should satisfy the following constraint.
-
-\begin{align*}
-\begin{array}{r l l}
-- \lambda_1 + \lambda_2 & = 0 \quad & (a) \\
-- \lambda_3 + \lambda_4 & = 0 & (b) \\
-  \lambda_1 - \lambda_4 & = 0 & (x) \\
-- \lambda_2 + \lambda_3 & = 0 & (y) \\
-  \lambda_2 + \lambda_4 & > 0 & (\text{constant}) \\
-\lambda_1, \ldots, \lambda_4 & \geq 0 & (\text{non-negative weight})
-\end{array}
-\end{align*}
-
-With the linear constraint among $\lambda_i$ above, the interpolant is
-represented as follows by the sum of weighted expressions from the
-first formula groups.
-\[ (- \lambda_1 + \lambda_2) a + \lambda_1 x - \lambda_2 y + \lambda_2 \leq 0 \]
-One of the model of the linear constraint is $\lambda_1 = \lambda_2 =
-\lambda_3 = \lambda_4 = 1$, and we obtain $x-y+1 \leq 0$ as a
-solution.
-
-% TODO: Better to mention multiple interpolants possible for one
-% interpolating problem?
-
-\section{Preliminaries}
-
-Our algorithm aims to compute a Craig interpolant over linear
-arithmetic formulas.
-
-\paragraph{Craig Interpolation}
-Given two logical formulas $A$ and $B$ that are inconsistent each
-other, we call a new preposition $I$ a \emph{Craig interpolant}
-\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$
-is implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
-only free variables that appears in both $A$ and $B$.  If $A \wedge B$
-is unsatisfiable, an interpolant $I$ always exists.
-
-Our algorithm handles a linear expression $e$ which is in a form:
-\[ a_1 x_1 + \cdots + a_n x_n + b \leq 0 \]
-This expression contains coefficient parameters $a_1, \ldots, a_n$ and
-a constant parameter $b$.  We treat a special expressions $\top$ as $0
-\leq 0$ and $\bot$ as $1 \leq 0$.  The algorithm input and output is a
-formula $\psi$ that consists of conjunctions and/or disjunctions of
-linear expressions.
-\[ \psi = e | e \wedge e | e \vee e | \top | \bot \]
-
-In the simple interpolation, the algorithm receives two formulas in
-$\psi$ and returns an interpolant in $\psi$.  For the later extension
-to handle a symmetric interpolation problem, the algorithm receives
-multiple formulas in $\psi$ and returns intermediate interpolants in
-$\psi$.  For computing those interpolants over linear arithmetics, we
-use Farkas's Lemma.
-
-\paragraph{Farkas's Lemma on linear inequalities}
-Let a linear inequality $e_i$ be represented as $a_i1 x_1 + \cdots +
-a_im x_m <= a_i0$.  Assuming that $e_1,\cdots,e_n$ implies $e_0$,
-there exists $\lambda_1,\cdots,\lambda_n$ that satisfy $a_0j
-=\sum_(i=1)^n \lambda_i * a_ij (j=0, \ldots, m)$.
-
-For the special case of Farkas's Lemma, we can state that:
-
-Assuming that $e_1,\cdots,e_n$ implies $\bot$, there exists
-$\lambda_1,\cdots,\lambda_n$ that satisfy $???? TODO $.
-
-Our interpolating algorithm make use of the latter version of Farkas's
-Lemma.
-
-\section{Algorithm}
-
-Our algorithm aims to construct a relatively small interpolant in a
-reasonable computation time for any input.  It also aims to obtain a
-common solution among multiple interpolating problems.  When the
-algorithm is used as a sub-routine for solving Horn clause satisfying
-problems that are made while discovering predicates for abstraction
-during the program verification.  These advantages of our algorithm
-make it possible to return a relatively small predicates for those
-problems. To accomplish these aims, the algorithm preserves a set of
-interpolants during the computation, and define and execute operations
-over interpolant sets.
-
-\paragraph{Interpolation between conjunctions}
-We first present an interpolating algorithm between two conjunctive
-sets of linear inequalities $A = \left\lbrace e_1,\ldots,e_m
-\right\rbrace$ and $B = \left\lbrace e_{m+1},\ldots,e_{m+n}
-\right\rbrace$.  We assume that conjunctions of $A$ and $B$ are
-inconsistent.  For later convenience, we call this interpolating
-problem $\left( A, B \right)$.  Then there exists an interpolating
-linear expression $e_\star$ which satisfies
-\begin{align*}
-\left\lbrace e_1,\ldots,e_m \right\rbrace & \vdash e_\star \\
-e_\star & \nvdash \left\lbrace e_{m+1},\ldots,e_{m+n} \right\rbrace
-\end{align*}
-For convenience, we assume that any expression $e_i$ have variables
-$x_1, \ldots, x_a$.
-
-According to the Farkas's lemma, there exists a set of assignmentqs to
-$\lambda_i$, to conclude $\bot$ from a set of linear expressions $A
-\cup B$, namely,
-\begin{align*}
-\exists \lambda_1, \ldots, \lambda_m. \\
-a % TODO: coefficient, constant constraints
-\end{align*}
-
-Then, a linear combination of $A$'s formula with a weight $\lambda_1,
-\ldots, \lambda_m$ becomes an interpolant $I$ between $A$ and $B$.
-Note that any assignments to $\lambda_i$ that satisfy the expressions
-above will let an expression $I$ to be an interpolant.  Considering
-this, it is possible to obtain different interpolants for a
-interpolating problem by computing different assignments to
-$\lambda_i$.
-
-This enables the algorithm to preserve a set of interpolants by a
-linear expression template $a_1 x_1 + \cdots + a_n x_n + b \leq 0$
-with a constraint to describe a possible space of coefficient
-parameters $a_i$ and the constant parameter $b$.
-
-In practical, we may use linear programming solvers to obtain a model
-for $\lambda_i$.  By assigning concrete values to them, we can obtain
-a concrete interpolant.
-
-\paragraph{Computing a common interpolant}
-Next we consider a method to compute a common interpolant across two
-different interpolating problems.  That is, when given two
-interpolating problems $\left(A_1, B_1 \right)$ and $\left(A_2, B_2
-\right)$, the problem is about to compute the common interpolant I,
-which is an interpolant for the first problem and the second one at
-the same time.
-
-This problem is used when solving a following Horn clause solving
-problem, for instance.
-\[
-\implies
-% TODO
-\]
-
-To accomplish this, we simply need to compute the intersection of
-interpolant sets.  Let us call the interpolant space for the problems
-$\left(A_1, B_1 \right)$ and $\left(A_2, B_2 \right)$,
-$\mathcal{I}_1$, and $\mathcal{I}_2$.  In the previous method, a set
-of interpolants was represented by a parameterized linear expression.
-\begin{align*}
-\mathcal{I}_1 = \left\lbrace \mathbf{a}_1 \mathbf{x} \leq b_1 \mid
-C_1 (\mathbf{a}_1, b_1 ) \right\rbrace \\
-\mathcal{I}_2 = \left\lbrace \mathbf{a}_2 \mathbf{x} \leq b_2 \mid
-C_2 (\mathbf{a}_2, b_2 ) \right\rbrace
-\end{align*}
-To compute the intersection between them, it is sufficient to choose
-the same assignments of parameters and let them satisfy the both
-constraints $C_1$ and $C_2$.
-\[ \mathcal{I} = \left\lbrace \mathbf{a} \mathbf{x} \leq b \mid
-C_1(\mathbf{a}, b) \wedge C_2(\mathbf{a}, b) \right\rbrace \]
-
-However, if such an assignment of $\mathbf{a}$ and $b$ does not exist,
-no common solution exists across two problems because $\mathcal{I} =
-\emptyset$.
-
-By applying this approach, it is also possible to compute a common
-interpolant among more than three problems.
-
-\paragraph{Interpolation between logical formulas}
-We now extend the algorithm so that it can handle general logical
-formulas over linear arithmetic including disjunctions of linear
-expressions.  Here the algorithm takes two logical formulas over
-linear expressions as a problem input.  For convenience, we assume
-that they are represented in disjunctive normal form (DNF).  The
-solution for this problem is the interpolant that is a logical formula
-of linear expressions.
-
-First, consider the case of an interpolation problem between $A_1 \vee
-A_2$ and $B$.  When $(A_1 \vee A_2) \wedge B$ is inconsistent, $A_1
-\wedge B$ is necessarily inconsistent.  Let $\mathcal{I}_1$ denote the
-set of interpolants for $(A_1, B)$.  In the same manner, let
-$\mathcal{I}_2$ denote the set of interpolants for $(A_2, B)$.  We may
-naturally obtain an interpolant set by the intersection $\mathcal{I} =
-\mathcal{I}_1 \cup \mathcal{I}_2$ as a solution for $(A_1 \vee A_2,
-B)$.  When the algorithm computes $\mathcal{I}$ by the previous
-method, and it is $\emptyset$, we need to build a less simple solution
-by
-\[ \mathcal{I}' = \left\lbrace \left( I_1 \vee I_2 \right) \mid
-I_1 \in \mathcal{I}_1 \wedge I_2 \in \mathcal{I}_2 \right\rbrace \]
-Any formulas in the set $\mathcal{I}'$ are the interpolant for $(A_1
-\vee A_2, B)$ because
-\begin{align*}
-\forall I \in \mathcal{I}. & A_1 \vdash I \wedge A_2 \vdash I \wedge I \nvdash B \wedge \\
-& FV(I) = FV(A_1) \cup FV(A_2) \cup FV(B)
-\end{align*}
-
-Second, consider the case of an interpolation problem between $A$ and
-$B_1 \wedge B_2$.
-% TODO:
-
-By combining these two ways of interpolant construction, we may be
-able to build a small solution for more general cases like $(A_1,
-\ldots, A_n)$ and $(B_1, \ldots, B_n)$.  In order to obtain a small
-solution, we need to wisely choose the order of combining an
-interpolant set.
-
-\paragraph{Symmetric Interpolation}
-TODO
-
-\chapter{Horn clause solving}
-\label{chap:horn}
-
-We propose a new algorithm for solving recursion-less Horn clauses to
-generate a solution which is as simple as possible.  This enables us
-to obtain appropriate predicates that are suitable for predicate
-abstraction in the program verification techniques.
-
-The construction of the algorithm roughly consists of two parts.
-\begin{itemize}
-\item \textbf{Iterative Constraint Generation} is a process to build a
-  constraint to compute a solution.  It starts from generating the
-  weakest constraint in the sense that the resulting solution would
-  not be simple yet.  The process iteratively strengthen a constraint
-  until the strongest one while preserving the constraint
-  satisfiability.  The strongest constraint gives a simple solution.
-\item \textbf{Problem Restructuring} is a process to restructure the
-  problem to obtain a less simpler solution, if the constraint become
-  unsatisfiable during the iterative constraint generation.  This
-  process considers the unsatisfiable core of the constraint, and
-  wisely enlarge the solution size.
-\end{itemize}
-
-All predicate variables in an input problem are assigned simple
-templates.  Those predicate templates are parameterized.  In the
-iterative constraint generation process, algorithm builds a linear
-constraint that parameters in templates must satisfy.  This is done in
-a constructive manner, and constraints are build in the order of
-implication dependency in Horn clauses.  That is, when an input
-problem has a Horn clause thath has a predicate variable $P$ on its
-left-hand and $Q$ on its right-hand, a constraint to describe
-parameters in $P$'s template are first build before those in $Q$'s
-template.  When all parameters in the whole problem are described, the
-built constraint are evaluated to obtain a model.  Its instanciation
-becomes an appropriate solution for the input problem.
-
-The novelty of our algorithm lies in the incremental constraint
-refinement.  The algorithm tries to solve from the weakest constraint
-that an input problem allows that does not give a simple solution, and
-gradually strengthen the constraint until it reaches to the simplest
-solution that the problem has.  When a constraint becomes
-unsatisfiable during this step, the algorithm analyzes the
-unsatisfiable core and transform the problem to obtain a less simple
-solution.
-
-This construction of the algorithm enables us to compute a relatively
-simpler solution compared to existing algorithms while preserving the
-computation cost in a reasonable time.
-
-\section{Example}
-
-\paragraph {Tree interpolation}
-First, consider the following example.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-y \leq 2 & \implies B(y) \\
-A(x) \wedge B(y) & \implies 3x+2y \leq 10
-\end{align*}
-
-In this example, we would like to compute linear arithmetic formulas
-for the predicate variables $A$ and $B$.  They appear only once on
-both left-hand and right-hand globally in the problem.
-
-We try to obtain a simple solution that is expressed by linear
-expression templates, whose coefficients and constants are
-parameterized.  Assume that they have following predicate templates.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\
-B(y) : c_3 y + c_4 \leq 0
-\end{align*}
-
-By naturally encoding the first Horn clause with Farkas's lemma, we
-obtain a linear constraint
-$ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
-Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will
-make $A(x)$ valid for the first clause.  In the same manner, we obtain
-$ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
-from the second clause, and
-$ c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0 $
-from the third.
-
-One of the models satisfying all the constraints above is
-$( c_1, c_2, c_3, c_4, \lambda_1, \lambda_2, \lambda_3 ) = (3, -3, 2, -4, 3, 2, 1)$.
-By assigning them to the template, we obtain the following solution.
-
-\begin{align*}
-A(x) : 3 x - 3 \leq 0 \\
-B(y) : 2 y - 4 \leq 0
-\end{align*}
-
-\paragraph {DAG problem solving}
-Second, we explain a case that some predicate variables appear
-multiple times on left-hand side in different Horn clauses.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-A(x) \wedge A(y) & \implies x+y \leq 2
-\end{align*}
-
-Assigning a template $A(x) : c_1 x + c_2 \leq 0$, we obtain the same
-constraint for the first clause as the previous example.
-
-\begin{align} \label{eq:constr1}
-\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0
-\end{align}
-
-Sinve we have $A$'s appearance twice on the second clause, it is
-possible for $A$ to have two expressions each of which is used for
-$A(x)$ and $A(y)$.  Although we need to use the same template for
-multiple occurrences in order to compute the simplest solution, it may
-make the whole constraints unsatisfiable if the simplest solution does
-not exist.
-
-Therefore, we generate a weaker constraint for the second clause that
-allows different expressions of $A$ for multiple occurrence.  If the
-whole constraint is satisfiable in the case, we try to solve in a
-stronger constraint to generate a simple solution.
-
-In this specific case, it is virtually equivalent to solve the
-following problem in a weaker constraint.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 1 & \implies A_1(x) \\
-x \leq 1 & \implies A_2(x) \\
-A_1(x) \wedge A_2(y) & \implies x+y \leq 2 \\
-\\
-\text{Templates:} \\
-A_1(x) \text{ : } & c_{1,1} x + c_{2,1} \leq 0 \\
-A_2(x) \text{ : } & c_{1,2} x + c_{2,2} \leq 0 \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-\end{align*}
-
-By duplicating the constraint~\ref{eq:constr1} of $A$ for newly
-created $A_1$ and $A_2$, we obtain the whole constraint as follows.
-
-\begin{align*}
-& \lambda_{1,1} = c_{1,1} \wedge - \lambda_{1,1} \geq c_{2,1} \wedge \lambda_{1,1} \geq 0 \wedge \\
-& \lambda_{1,2} = c_{1,2} \wedge - \lambda_{1,2} \geq c_{2,2} \wedge \lambda_{1,2} \geq 0 \wedge \\
-& c_{1,1} = \lambda_2 \wedge c_{1,2} = \lambda_2 \wedge c_{2,1} + c_{2,2} \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-This constraint is satisfiable and one model is
-\[ ( c_{1,1}, c_{1,2}, c_{2,1}, c_{2,2}, \lambda_{1,1}, \lambda_{1,2}, \lambda_2 ) =
-( 1, -1, 1, -1, 1, 1, 1 ) \] and the solution then becomes
-$A(x) : (x -1 \leq 0) \wedge (x -1 \leq 0)$.
-
-Since the weak constraint is satisfiable, we now refine the constraint
-to obtain the better solution.  That is, simply eliminate the
-duplication to build the following constraint for the original
-template.
-
-\begin{align*}
-& \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \\
-& c_1 = \lambda_2 \wedge c_1 = \lambda_2 \wedge 2 c_2 \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-Solving this constraint simply gives us a solution $A(x) : x -1 \leq 0$.
-
-\paragraph {Disjunctive Horn clause solving}
-Finally, we explain a case that an input set of Horn clauses contain
-disjunctions on left-hand side.
-
-\begin{align*}
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 \vee -x+1 \leq 0 & \implies B(x) \\
-A(x) \wedge B(x) & \implies 1 \leq 0
-\end{align*}
-
-Same as the previous examples, templates for predicate variables are
-prepared.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\ B(x) : c_3 x + c_4 \leq 0
-\end{align*}
-
-The constraints generated are follows.
-
-\begin{align*}
-& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
-& \lambda_3 = c_3 \wedge \lambda_3 \geq c_4 \wedge \lambda_3 \geq 0 \wedge \\
-& - \lambda_4 = c_3 \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
-& c_1 + c_3 = 0 \wedge c_2 + c_4 > 0
-\end{align*}
-
-The constraints here become unsatisfiable, and one possible
-unsatisfiable core is:
-
-\begin{align*}
-\lambda_3 & = c_3 \\
-\lambda_3 & \geq c_4 \\
-- \lambda_4 & = c_3 \\
-\lambda_4 & \geq c_4 \\
-c_2 + c_4 & > 0 \\
-0 & \geq c_2
-\end{align*}
-
-Because the unsatisfiable core contains constraints originating from
-the disjunctions to be expressed in a single term template.  This let
-us determine to relax the template to allow a enlarged solution.
-
-Now we slightly change the problem and the template.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A(x) \wedge B_1(x) & \implies 1 \leq 0 \\
-A(x) \wedge B_2(x) & \implies 1 \leq 0 \\
-\\
-\text{Templates:} \\
-A(x) \text{ : } & c_1 x + c_2 \leq 0 \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-When we construct the constraints, as same as previous, we again
-obtain an unsatisfiable core:
-
-\begin{align*}
-\text{Constraint:} \\
-TODO \\
-\\
-\text{Unsatisfiable core:} \\
-TODO
-\end{align*}
-
-This time, the unsatisfiable core contains the constraints that
-originate from
-
-TODO
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A_i(x) \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A_i(x) \wedge B_j(x) & \implies 1 \leq 0 \quad (i,j \in \left\lbrace 1,2 \right\rbrace ) \\
-\\
-\text{Templates:} \\
-A_i(x) \text{ : } & c_{1,i} x + c_{2,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-Finally, we obtain the following constraint:
-
-TODO
-
-The process also generates a predicate template for every predicate
-variable that the solution will become.  under certain solution
-template.
-
-\section{Preliminaries}
-
-For ease of discussion, our algorithm handles an input set of Horn
-clauses by a \emph{Horn graph} that express the equivalent problem.
-The graph is then used throughout the algorithm to represent the
-problem.  A Horn graph is a labelled directed acyclic graph
-$G=(V,E,\varphi)$.
-\begin{itemize}
-\item $V$ is a union of two disjoint sets of vertices; Horn term
-  vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex $u
-  \in V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
-  denoted as $succ(u)$.
-\item $E$ is a set of labelled edges, which are expressed in the form
-  $(u,v,\theta)$ where $\theta$ is a finite map from variables to
-  variables.
-\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
-  term vertices, where L is a set of elements of a predicate variable
-  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
-  inequality $a_1 x_1 + \cdots + a_n x_n + b \leq 0$.
-\end{itemize}
-There are two kinds of edges in a Horn graph;
-\begin{itemize}
-\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
-  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
-\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
-  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
-  empty.
-\end{itemize}
-
-We define the relation $\leadsto$ over $V_T$ as follows.
-\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow \exists a \in V_\Rightarrow;
-\edgel{u}{\theta}{a} \wedge \edge{a}{v} \] We may omit $\theta$ and
-write $u \leadsto v$ for convenience in later discussions. By the
-assumption that $G$ is acyclic, both $\rightarrow^+$ and $\leadsto^+$ are
-well-founded relations. $v_\bot$ is the greatest element with respect
-to both $\rightarrow^+$ and $\leadsto^+$.
-
-The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
-Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$.
-\begin{align*}
-\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(succ(v)) \\
-\llbracket G \rrbracket & = \bigwedge_{v \in V_\Rightarrow} \mathcal{HC}_v
-\end{align*}
-We implicitly assume each $\mathcal{HC}_v$ is universally quantified
-by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.
-
-Horn term vertices with predicate variable labels are unique each
-other that any of them do not have the same variable on their labels.
-From now on, we may denote a Horn term vertex with a predicate
-variable label by its predicate variable name if no ambiguity is
-present.
-
-Without loss of generality, we assume the existence of the \emph{root
-  vertex} $v_\bot \in V_T$ with linear expression labels $1 \leq 0$,
-meaning $\bot$.  This is because if we have a Horn clause $P \implies
-e$ which $e$ is a linear expression, we can convert it to $P \wedge
-\neg e \implies 1 \leq 0$.  We also assume that other vertices with
-linear expression labels have no incoming edges to them.
-
-For simplicity of discussion, we define \emph{term arity} $k_v$ for
-every Horn term vertex $v \in V_T$ with incoming edges.
-\begin{align*}
-k_v =
-\begin{cases}
-\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
-0 & \mbox{if } \varphi(v) = \bot
-\end{cases}
-\end{align*}
-We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.
-
-A variable mapping label $\theta$ on edges is used to map the index of
-one predicate variable to another. (TODO?)
-
-(Example?)
-
-The algorithm assigns a \emph{predicate template} for all predicate
-variables in $G$. A \emph{predicate templateset} is a set of mappings from
-a predicate variable to a predicate template.  A template for a
-predicate variable $P(x_1, \ldots, x_n)$ is a formula $\psi$
-constructed from the following definition.
-\begin{align*}
-\psi ::= & a_1 x_1 + \cdots + a_n x_n + b \leq 0 \mid \\
-& \psi \wedge \psi \mid \psi \vee \psi
-\end{align*}
-A linear expression in the definition above may also be represented in
-a equivalent vector form $\mathbf{a} \mathbf{x} + b \leq 0$ for short.
-Parameters $\mathbf{a}$ and $b$ are specified by a linear constraint.
-
-The solution of $G$ is a map $\rho$ from predicate variables to linear
-predicates of the form $\lambda x_1, \ldots ,x_n. \psi$ where
-parameters in $\psi$ are replaced with constant and $\rho[G]$ becomes
-a tautology.  A solution is \textit{simple} if $\rho[P]$ is of the
-form $\mathbf{a} \mathbf{x} + b \leq 0$ for every $P$ in
-$\mathrm{dom}(\rho)$.
-
-The problem of solving Horn clauses (represented by $G$) is to find a
-solution of $G$ (and a simple solution if possible).
-
-\section{Simple solution}
-
-We first propose an algorithm to compute a simple solution for the
-given Horn graph $G$ if one exists.  Later we add modifications to the
-algorithm for the case that a simple solution does not exist for $G$.
-In such a case, we transform the input graph $G$ to have a larger
-solution template.
-
-For each vertex $v \in V$, we define
-$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v + b_v \leq 0, C_v)$
-, where
-\begin{itemize}
-\item $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ is a linear inequality that
-  contains coefficient variables $\mathbf{a}_v$ and a constant $b_v$,
-  and
-\item $C_v$ is a set of linear expressions on coefficient variables.
-\end{itemize}
-Intuitively, it denotes the set of inequalities:
-\begin{align*}
-\left\lbrace
- \mathbf{a}_v \mathbf{x}_v + b_v \leq 0 \mid
- \exists \left( \textsc{FV}(C_v)
-  \setminus (\mathbf{a}_v \cup \left\lbrace b_v \right\rbrace
- \right); C_v
-\right\rbrace
-\end{align*}
-For a predicate variable vertex, $\mathbf{a}_v \mathbf{x}_v + b_v \leq
-0$ serves as a predicate template and $C_v$ describes the parameters
-$\mathbf{a}_v$ and $b_v$.
-
-$C_v$ is defined by well-founded induction on $v$ with respect to the
-relation $\rightarrow^+$. Then, $C_v$ can be shown as follows.
-\begin{itemize}
-\item For $v \in V_\Rightarrow$;
-\begin{align*}
-C_v & =
- \bigcup_{\edgel{u}{\theta}{v}} C_u \cup
- \bigcup_{0 < i \leq K}
- \left\lbrace
-  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{v}} \mathbf{a}_{u, \theta^{-1} (i)}
- \right\rbrace \\
- & \hspace{2cm} \cup
- \left\lbrace
-  b_v \leq \sum_{\edgel{u}{\theta}{v}} b_u
- \right\rbrace \cup
- \bigcup_{k_{succ(v)} < i \leq K}
- \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that
-  $v \neq v_{\bot} \wedge \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0$
-  where $\mathbf{a}_0$ and $b_0$ are constants;
-\begin{align*}
-C_v =
-\left\lbrace
- \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
- b_v = \lambda_v b_0
-\right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that $\varphi(v) = P(\mathbf{x}_v)$;
-\begin{align*}
-C_v = & \bigcup_{\edge{u}{v}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{a}_v, b_u = b_v \right\rbrace
-\right)
-\end{align*}
-\item For $v_\bot$;
-\begin{align*}
-C_{v_\bot} = \bigcup_{\edge{u}{v_\bot}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{0}, b_u > 0 \right\rbrace \right)
-\end{align*}
-\end{itemize}
-
-Any model $\sigma$ to the constraint $C_{v_\bot}$ gives a simple
-solution for $G$ if $C_{v_\bot}$ is satisfiable, i.e., we should have:
-
-\begin{quote}
-If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
-\begin{align*}
- \rho = \left\lbrace
-  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \mid
-  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
- \right\rbrace
-\end{align*}
-is a solution of $G$.
-\end{quote}
-
-If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for $G$.
-There are two cases that a simple solution does not exist for a
-predicate variable.
-\begin{itemize}
-\item A predicate variable appears multiple times on left-hand of Horn
-  clauses, and those appearances require different expressions as an
-  assumption.
-\item A predicate variable is implied by a disjunctive Horn clause,
-  which has disjunctions on its left-hand, and a common premise that
-  the disjunctions induce is not strong enough for solving a problem.
-\end{itemize}
-
-The first case occurs in a case like follows.
-\begin{align*}
-% conj.txt ... originated from uc/01.txt
-\left( x \leq 0 \right) \wedge \left( y \leq 0 \right) & \implies P(x,y) \\
-P(x,a) \wedge P(b,y) & \implies x+y \leq 0
-\end{align*}
-In this example, the predicate variable $P$ appears twice on the
-second clause and they require different premises; $x \leq 0$ for the
-first and $y \leq 0$ for the second.  They cannot be represented by
-$\mathbf{ax} + b \leq 0$ template at the same time.
-
-The second case occurs in a case like follows.
-\begin{align*}
-% disj-simp.txt ... originated from 23simp_neg.txt
-\left( x+1 \leq 0 \right) \vee \left( -x+1 \leq 0 \right) & \implies P(x) \\
-P(x,y) \wedge \left( x \leq 0 \right) \wedge \left( -x \leq 0 \right) & \implies \bot
-\end{align*}
-In this example, the predicate variable $P$ is implied by the first
-clause, which is a disjunctive Horn clause.  As a simple solution, $P$
-can take a common consequence that both of $x+1 \leq 0$ and $-x+1 \leq
-0$ induce.  However, such a simple linear expression does not exist
-except for $0 \leq 0$, which is equivalent to $\top$, and it is not
-sufficient to satisfy the second clause.
-
-Both cases can be fixed by enlarging the solution template.
-
-\section {Graph transformation}
-
-If a simple solution does not exist for $G$, we try to give a less
-simpler solution by transforming $G$ and relaxing the constraint
-$C_{v_\bot}$.  This transformation takes place by considering the
-unsatisfiable core $\mathcal{U}$.  An unsatisfiable core $\mathcal{U}$
-is a subset of $C_{v_\bot}$ which is inconsistent itself.  Depending
-on the reason for the absence of a simple solution, the way of the
-transformation differs.
-
-Algorithm~\ref{alg:solveHorn} shows the updated pseudo-code of our
-Horn clause solving algorithm with the graph transformation.  It takes
-an input Horn graph and returns a mapping between predicate variables
-and linear arithmetic formulas.
-
-\newcommand{\solveHorn}{\ensuremath{\mbox{\sc SolveHornGraph}}}
-\newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
-\newcommand{\transGraph}{\ensuremath{\mbox{\sc TransformGraph}}}
-
-\begin{algorithm}
-\caption{$ \solveHorn (G) $}\label{alg:solveHorn}
-\begin{algorithmic}
-\STATE {$v_\bot \gets $ the root node of $G$}
-\STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-\STATE {$T \gets $ initial predicate templateset for $G$}
-\REPEAT
-  \STATE {$C_{v_\bot} \gets \genConstr (G, v_\bot, V_\star)$}
-  \IF {$C_{v_\bot}$ is satisfiable}
-    \STATE {$\sigma \gets $ one model of $C_{v_\bot}$}
-    \STATE {$V_\star \gets V_\star \cup \left\lbrace u \in V_T \mid \forall v \in V_\star; u \leadsto v \right\rbrace$}
-  \ELSE
-    \STATE {$\mathcal{U} \gets $ unsatisfiable core for $C_{v_\bot}$}
-    \STATE {$(G, T) \gets \transGraph (G, T, \mathcal{U})$}
-    \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-  \ENDIF
-\UNTIL {$V_\star = V_T$}
-\RETURN {$\sigma \left( T \right)$}
-\end{algorithmic}
-\end{algorithm}
-\transGraph~procedure determines the unsatisfiable core of the
-constraint and transform the Horn graph $G$ and predicate templates.
-As mentioned earlier, two ways of graph transformation exist.
-
-\paragraph{Conjunction split}
-For the first case previously mentioned, the unsatisfiable core
-$\mathcal{U}$ contains coeffcient constraint expressions with a
-coefficient parameter of certain predicate variable.  If the predicate
-variable $P$ is the cause of unsatisfiability, some constraint
-expressions in the unsatisfiable core have a parameter $a_{P,i}$
-and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is copied to
-$P_1, \ldots, P_n$ so that the different appearances of $P$ on
-left-hand of Horn clauses can take different premises.  The solution
-template of $P$ becomes the conjunction of $P_1, \ldots, P_n$.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-Note that after this duplication of the predicate vertex $P$, in order
-to induce different linear expressions for $P_1, \ldots, P_n$, it may
-be necessary for preceeding predicate variable vertices in $G$ to be
-duplicated in the same manner as $P$. It is because that different
-derivation of $P$'s implication relation may be required to get
-different consequences.
-
-However, succeeding predicate vertices of $P$ in $G$ and other
-vertices than predecessors are not affected because those implications
-that have $P$ in their assumption can selectively choose the premise
-that $P_1, \ldots, P_n$ leads.
-
-Therefore, this graph transformation only affects the predecessors of
-$P$.
-
-\paragraph{Disjunction split}
-For the second case, the unsatisfiable core $\mathcal{U}$ contains
-coefficient constraints that unifies induced linear expressions from
-each term in disjunctions to be the same expression.  Similarly as
-previous, if the predicate variable $P$ is the cause of
-unsatisfiability, some constraint expressions in the unsatisfiable
-core have a parameter $a_{P,i}$ and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is separated to
-$P_1, \ldots, P_n$ each of which is implied by a term, which was
-originally a part of a disjunction to imply $P$.  This intuitively
-enables us to avoid generating a common linear expression that
-disjunctive terms induces, and to propagate each term's premises to
-$P$'s successors instead.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-After this duplication of the predicate vertex $P$, unlike the
-conjunction split, this results in producing a disjunction at
-succeeding predicate variables vertices ($R$ in this case).
-\begin{align*}
-(P_1 \wedge Q) \vee (P_2 \wedge Q) \implies R
-\end{align*}
-Also, predicate variables that appeared with $P$ in the same
-assumption ($Q$ in this case) needs to be duplicated by conjunction,
-so as to obtain different premises that go along with $P_i$.
-Therefore, the effect of disjunction split may reach to neighboring
-predicate vertices.
-
-Consider these two ways of transformation, we show a pseudo-code for
-\transGraph~procedure.  The procedure is given the graph $G$, the
-predicate templateset $T$, and the unsatisfiable core
-$\mathcal{U}$. It returns the transformed graph $G'$ and the new
-templateset $T'$.
-
-\begin{algorithm}
-\caption{$ \transGraph (G, T, \mathcal{U}) $}\label{alg:transGraph}
-\begin{algorithmic}
-% TODO:
-\RETURN {$\left( G', T' \right)$}
-\end{algorithmic}
-\end{algorithm}
-
-\paragraph{Combinatorial explosion in disjunctions}
-Virtually, if multiple disjunctions exists in an input set of Horn
-clauses, a na\"{i}ve way of building a solution is solve all
-combinations of choices out of every disjunction and concatenate each
-expression with conjunctions and/or disjunctions appropriately.  This
-may cause a combinatorial explosion.
-
-Conside example below. $e_i$ represents some conjunctions of linear
-expressions.
-% TODO: Can't we replace e_i with actual expressions?
-\begin{align*}
-(e_1 \vee e_2) \implies P \\ (e_3 \vee e_4) \implies Q \\ P \wedge Q \implies \bot
-\end{align*}
-In this example, a na\"{i}ve algorithm may solve following four
-problems separately.
-\begin{itemize}
-\item $e_1 \implies P_1 \qquad e_3 \implies Q_1 \qquad P_1 \wedge Q_1 \implies \bot$
-\item $e_2 \implies P_2 \qquad e_3 \implies Q_2 \qquad P_2 \wedge Q_2 \implies \bot$
-\item $e_1 \implies P_3 \qquad e_4 \implies Q_3 \qquad P_3 \wedge Q_3 \implies \bot$
-\item $e_2 \implies P_4 \qquad e_4 \implies Q_4 \qquad P_4 \wedge Q_4 \implies \bot$
-\end{itemize}
-Then the solution is built by follows.
-\begin{itemize}
-\item $P = ( P_1 \vee P_2 ) \wedge ( P_3 \vee P_4 )$
-\item $Q = ( P_1 \vee P_3 ) \wedge ( P_2 \vee P_4 )$
-\end{itemize}
-However, this method is not efficient.
-
-\paragraph{Algorithm}
-To avoid this, our algorithm lets a predicate variable $P$ which is
-disjunctively implied on the Horn graph $G$ to have a linear
-expression that is consequences of all terms in the disjunction.  This
-prevents the possibility of choices for disjunctions to be propagated
-to succeeding vertices. The simple linear expression is propagated
-instead.
-
-On the other hand, the algorithm allows a predicate variable $P$ that
-appears on multiple implications or appears multiple times in a single
-implication to take different linear expressions as a premise at
-different appearances.  This is for the algorithm to determine the
-cause of unsatisfiability efficiently.  It is done by copying a
-predicate templates and its parameters' constraint with appropriate
-renaming.
-
-The algorithm first execute this duplication on all predicate
-variables with multiple appearances.  It is the weakest constraint
-that the algorithm makes for the certain $G$.  Then, the algorithm
-gradually reduces the number of vertices that are duplicated by each
-time testing the satisfiability of the constraint.  When the
-constraint bebcome unsatisfiable at the certain iteration, the
-algorithm can determine a predicate vertex which has to be split in a
-conjunctive template.  If all the predicate variable constraints are
-no longer duplicated and the constraint is the strongest, any model
-for such constraint gives a simple solution under the Horn graph $G$
-and the templateset $T$.
-
-Note that duplicating the constraint in a na\"{i}ve way may cause the
-constraint to exponentially grow in size.  Our algorithm prevent it by
-simplifying a predicate vertex's constraint $C_v$ by quantifier
-elimination when $v$ is duplicated in generating linear constraint.
-For a predicate variable vertex $v_P$, variables except for the
-coefficient parameter $\mathbf{a}_P$ and the constant parameter $b_P$
-are eliminated.  Due to the constraints generation process, the
-simplified constraint is also linear and can be efficiently check the
-satisfiability.
-
-We finally show the modified version of pseudo-code for $\genConstr$
-procedure.  It produces the constraint $C_v$ for a given vertex $v$,
-considering constraint duplication and simplification.
-
-\begin{algorithm}
-\caption{$ \genConstr (G, v, V_\star) $}\label{alg:genConstr}
-\begin{algorithmic}
-\STATE {$C \gets \emptyset$}
-\FORALL {$v_\Rightarrow \in preds_G(v)$}
-  \FORALL {$v_p \in preds_G(v_\Rightarrow)$}
-    \IF {$\varphi (v_p)$ is a predicate variable}
-      \STATE {$C \gets C \cup \genConstr (G, v_p, visited)$}
-    \ELSE
-      \STATE { $id$ }
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-traverses the input graph in the topological order on
-$V_\Rightarrow$, and concatenate the constraints from predecessors to
-build the problem constraints.  Each process to build a constraint on
-$\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$ is equivalent
-to generate a linear constraint in interpolation.
-
-\chapter{Experiment}
-\label{chap:experiment}
-
-We incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It
-is specifically designed to verify the correctness of higher-orde
-programs written in a small subset of ML.  It infers dependent types
-for input higher-order programs \cite{conf/ppdp/UnnoK09}.
-
-Our implementation make use of Z3 \cite{conf/tacas/MouraB08} as
-an SMT solver.
-
-For simplicity of the implementation, we assume that the linear
-inequalities in the integer space.  Because of this, our algorithm
-lost the completeness.  By using Motzkin's transpose theorem
-\cite{journals/networks/Rajan90}, it is possible to handle linear
-inequalities in both $\mathbf{ax} + b \leq 0$ and $\mathbf{ax} + b <
-0$.
-
-\section{Interpolating theorem prover}
-
-Our interpolating theorem prover is used as a sub-procedure to
-solve Horn clauses.
-
-\begin{itemize}
-\item CSIsat \cite{conf/cav/BeyerZM08}
-\item FOCI \cite{website/foci}
-\item CLP-Prover \cite{website/clp}
-\end{itemize}
-
-\section{Horn clause solver}
-
-\chapter{Related work}
-\label{chap:related}
-\input{related.tex}
-
-\chapter{Future Work}
-\label{chap:future}
-
-\paragraph{Non-linear extension}
-\cite{conf/cav/DaiXZ13}
-
-\chapter{Conclusion}
-\label{chap:conclusion}
-
-\bibliographystyle{plain}
-\bibliography{./biblio}
-
-\end{document}

### 2013-08-26 article/thesis.tex
Other chapters also separated

@@
 
-\begin{eabstract}
-Automated theorem proving plays a crucial role in modern program
-verification techniques.  In particular, in the context of software
-model checking, the interpolation problem and its generalization
-called \emph{Horn clause solving} have attracted special attention.
-The latter problem asks for a solution for a set of acyclic Horn
-clauses containing unknown predicate variables.  The solution, i.e., a
-substitution for predicate variables, then serves as a candidate of
-program invariants.  However, existing methods are unsatisfactory in
-that they tend to generate too specific a solution, which does not
-correspond to program invariants.
-
-To address the problem above, we propose a new algorithm for solving
-Horn clauses consisting of linear arithmetics and unknown predicate
-variables.  The new algorithm tries to find small predicates, based on
-the hypothesis that program's invariants tend to be simple formulas.
-To this end, our algorithm prepares a template of linear arithmetic
-formulas of a fixed size as a candidate of each unknown predicate, so
-that Horn clause solving is reduced to linear programming.  The
-algorithm iteratively increases the size of the templates until a
-solution is found.  It enables us to construct relatively small
-solutions in a reasonable time compared to existing algorithms.  We
-have implemented and incorporated our Horn clause solver into MoCHi, a
-software model checker for functional programs.  We have evaluated the
-effectiveness of our approach through experiments.
-\end{eabstract}
+\input{eabstract.tex}
 \begin{jabstract}
@@
 
-\begin{acknowledge}
-Thank you.
-\end{acknowledge}
+\input{acknowledge.tex}
 
@@
 
-\chapter{Introduction}
-
-Automated theorem proving plays an important role in the modern
-program verification.  Particularly in the context of verification
-with model checking techniques, an automated theorem prover is used to
-abstract an infinite set of program states into a finite one so that
-the model checker can virtually explore all possible program execution
-paths.
-
-One of the most common methods of such abstraction is predicate
-abstraction\cite{conf/cav/GrafS97} within CounterExample-Guided
-Abstraction Refinement (CEGAR) framework
-\cite{conf/cav/ClarkeGJLV00,conf/popl/BallR02,conf/popl/HenzingerJMS02}.
-An automated theorem prover finds appropriate predicates for every
-program location.  An execution states at those locations is then
-expressed by a Boolean valuation of the given predicates.  If the
-abstraction is too coarse and a model checker discovers an infeasible
-error execution path in an abstract system, the prover computes an
-additional predicate at those program locations to refine the
-abstraction predicates.
-
-In order to obtain an additional abstraction predicate at program
-locations along the infeasible counterexample, one class of automated
-theorem provers, called an interpolating theorem prover
-\cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07},
-computes a separating predicate called an interpolant between two sets
-of constraints; the one consists of the constraints from the program
-entry point to the location, and the other consists of the constraints
-from the location to the program failure.  Those constraints include
-variable assignments and assertions to be satisfied originated from
-conditional branches.
-
-The computed predicate by an interpolating theorem prover is satisfied
-at the location under the execution in the concrete model along the
-discovered counterexample, and any execution paths which satisfies the
-predicate does not lead to the failure state.  After the refinement,
-the infeasible path is ruled out and no longer found by the model
-checker.
-
-However, existing interpolating theorem provers may return too complex
-a solution which is heavily affected by the constraints of the
-specific path.  It is not favorable in terms of the program
-verification because a complex predicate may not be sufficient to rule
-out similar spurious paths in programs with loops and recursion calls.
-This causes the model checker to discover spurious paths infinitely
-which pass the same program location by loops or recursion calls at
-different numbers of iteration, and the verification process does not
-to terminate.
-
-To avoid this, it is desired to discover an invariant formula for the
-predicate abstraction.  Based on the hypothesis that program's
-invariants tend to be simple formulas, we propose a new interpolating
-algorithm which tries to minimize the number of conjunctions and
-disjunctions in a solution.  The predicate obtained by this algorithm
-may become similar to the invariant, and the number of trials in model
-checking may be reduced.
-
-Additionally, we extend our interpolating algorithm to solve a
-symmetric interpolation problem, in which the algorithm computes
-predicates at multiple program locations along a path at once.  This
-enables the abstraction predicates at multiple locations to be updated
-at the same computation.
-
-Finally, we extend our algorithm to solve Horn clause solving
-problems.  It allows us to obtain predicates which are highly likely
-to be program invariants, by computing the same abstraction predicates
-inside loops and recursion calls on different iterations, and by
-unifying predicates at the same location over multiple paths.  The
-problems are generated by encoding constraints over execution paths
-with unknown predicates which describe functions' constraints between
-parameters and return values.
-
-The rest of the paper is structured as follows.
-Chapter~\ref{chap:interpolation} proposes a new interpolating
-algorithm to obtain simple solutions, and extends it to solve
-symmetric interpolation problems.  Chapter~\ref{chap:horn} describes a
-new Horn clause solving algorithm by extending the previous
-interpolating algorithm.  Chapter~\ref{chap:experiment} describes
-experiments and illustrates the result to confirm the effect of our
-proposed algorithm.  Chapter~\ref{chap:future} gives a discussion of
-our improvements and proposes further improvements to make.
-Chapter~\ref{chap:related} mentions related work done by others.
-Finally, Chapter~\ref{chap:conclusion} reviews the impact of our
-research and brings the conclusion.
-
-\chapter{Interpolation}
-\label{chap:interpolation}
-
-Among various program verification techniques, when software model
-checking with predicate abstraction is adopted, the interpolation is
-used to compute the abstraction predicate along a spurious
-counterexample that a model checker discovered.  It is done by
-computing a Craig interpolant between the strongest postcondition at
-the location and the weakest precondition to the failure point at
-every program location.
-
-However, by the existing interpolating theorem prover, we may not be
-able to obtain a simple solution.  Consider the example below.
-\begin{align*}
-\varphi_1 : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
-\varphi_2 : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
-\end{align*}
-Although one may be able to obtain $x = y$ as the simplest
-interpolation between $\varphi_1$ and $\varphi_2$,
-CSIsat\cite{conf/cav/BeyerZM08} returns
-\[ TODO \]
-as a solution.
-
-We propose an interpolation algorithm which returns as simple a
-solution as possible in terms of the number of conjunctions and
-disjunctions.  Our algorithm focuses on interpolating problems on
-linear arithmetic and it works in a constructive manner.
-
-\section{Example}
-
-Consider the following interpolation problem.
-
-\begin{align*}
-\frac
-{A: (x \leq a) \wedge (a + 1 \leq y)}
-{B: (y \leq b) \wedge (b + 1 \leq x)}
-\end{align*}
-
-Here, the linear arithmetic formulas A and B are inconsistent.  The
-interpolant I for this problem is $x-y+1 \leq 0$, which is implied by
-$A$ and inconsistent with $B$.  The vocabulary of $I$ is $\left\lbrace
-x,y \right\rbrace$ and is over $A$'s vocabulary $\left\lbrace x,y,a
-\right\rbrace$ and $B$'s $\left\lbrace x,y,b \right\rbrace$.
-
-In computing the interpolant, we make a linear constraint of a
-interpolant by applying Farkas's lemma.  First, we assing weight
-parameters to every expression in conjunction form.
-
-\begin{align*}
-\begin{array}{r r r r r r l}
-\lambda_1 : & -a & & +x & & & \leq 0 \\
-\lambda_2 : & a & & & -y & +1 & \leq 0 \\
-\hline
-\lambda_3 : & & -b & & +y & & \leq 0 \\
-\lambda_4 : & & b & -x & & +1 & \leq 0
-\end{array}
-\end{align*}
-
-Because all expressions induce inconsistent as a whole, the sum of
-weighted expressions should satisfy the following constraint.
-
-\begin{align*}
-\begin{array}{r l l}
-- \lambda_1 + \lambda_2 & = 0 \quad & (a) \\
-- \lambda_3 + \lambda_4 & = 0 & (b) \\
-  \lambda_1 - \lambda_4 & = 0 & (x) \\
-- \lambda_2 + \lambda_3 & = 0 & (y) \\
-  \lambda_2 + \lambda_4 & > 0 & (\text{constant}) \\
-\lambda_1, \ldots, \lambda_4 & \geq 0 & (\text{non-negative weight})
-\end{array}
-\end{align*}
-
-With the linear constraint among $\lambda_i$ above, the interpolant is
-represented as follows by the sum of weighted expressions from the
-first formula groups.
-\[ (- \lambda_1 + \lambda_2) a + \lambda_1 x - \lambda_2 y + \lambda_2 \leq 0 \]
-One of the model of the linear constraint is $\lambda_1 = \lambda_2 =
-\lambda_3 = \lambda_4 = 1$, and we obtain $x-y+1 \leq 0$ as a
-solution.
-
-% TODO: Better to mention multiple interpolants possible for one
-% interpolating problem?
-
-\section{Preliminaries}
-
-Our algorithm aims to compute a Craig interpolant over linear
-arithmetic formulas.
-
-\paragraph{Craig Interpolation}
-Given two logical formulas $A$ and $B$ that are inconsistent each
-other, we call a new preposition $I$ a \emph{Craig interpolant}
-\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$
-is implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
-only free variables that appears in both $A$ and $B$.  If $A \wedge B$
-is unsatisfiable, an interpolant $I$ always exists.
-
-Our algorithm handles a linear expression $e$ which is in a form:
-\[ a_1 x_1 + \cdots + a_n x_n + b \leq 0 \]
-This expression contains coefficient parameters $a_1, \ldots, a_n$ and
-a constant parameter $b$.  We treat a special expressions $\top$ as $0
-\leq 0$ and $\bot$ as $1 \leq 0$.  The algorithm input and output is a
-formula $\psi$ that consists of conjunctions and/or disjunctions of
-linear expressions.
-\[ \psi = e | e \wedge e | e \vee e | \top | \bot \]
-
-In the simple interpolation, the algorithm receives two formulas in
-$\psi$ and returns an interpolant in $\psi$.  For the later extension
-to handle a symmetric interpolation problem, the algorithm receives
-multiple formulas in $\psi$ and returns intermediate interpolants in
-$\psi$.  For computing those interpolants over linear arithmetics, we
-use Farkas's Lemma.
-
-\paragraph{Farkas's Lemma on linear inequalities}
-Let a linear inequality $e_i$ be represented as $a_i1 x_1 + \cdots +
-a_im x_m <= a_i0$.  Assuming that $e_1,\cdots,e_n$ implies $e_0$,
-there exists $\lambda_1,\cdots,\lambda_n$ that satisfy $a_0j
-=\sum_(i=1)^n \lambda_i * a_ij (j=0, \ldots, m)$.
-
-For the special case of Farkas's Lemma, we can state that:
-
-Assuming that $e_1,\cdots,e_n$ implies $\bot$, there exists
-$\lambda_1,\cdots,\lambda_n$ that satisfy $???? TODO $.
-
-Our interpolating algorithm make use of the latter version of Farkas's
-Lemma.
-
-\section{Algorithm}
-
-Our algorithm aims to construct a relatively small interpolant in a
-reasonable computation time for any input.  It also aims to obtain a
-common solution among multiple interpolating problems.  When the
-algorithm is used as a sub-routine for solving Horn clause satisfying
-problems that are made while discovering predicates for abstraction
-during the program verification.  These advantages of our algorithm
-make it possible to return a relatively small predicates for those
-problems. To accomplish these aims, the algorithm preserves a set of
-interpolants during the computation, and define and execute operations
-over interpolant sets.
-
-\paragraph{Interpolation between conjunctions}
-We first present an interpolating algorithm between two conjunctive
-sets of linear inequalities $A = \left\lbrace e_1,\ldots,e_m
-\right\rbrace$ and $B = \left\lbrace e_{m+1},\ldots,e_{m+n}
-\right\rbrace$.  We assume that conjunctions of $A$ and $B$ are
-inconsistent.  For later convenience, we call this interpolating
-problem $\left( A, B \right)$.  Then there exists an interpolating
-linear expression $e_\star$ which satisfies
-\begin{align*}
-\left\lbrace e_1,\ldots,e_m \right\rbrace & \vdash e_\star \\
-e_\star & \nvdash \left\lbrace e_{m+1},\ldots,e_{m+n} \right\rbrace
-\end{align*}
-For convenience, we assume that any expression $e_i$ have variables
-$x_1, \ldots, x_a$.
-
-According to the Farkas's lemma, there exists a set of assignmentqs to
-$\lambda_i$, to conclude $\bot$ from a set of linear expressions $A
-\cup B$, namely,
-\begin{align*}
-\exists \lambda_1, \ldots, \lambda_m. \\
-a % TODO: coefficient, constant constraints
-\end{align*}
-
-Then, a linear combination of $A$'s formula with a weight $\lambda_1,
-\ldots, \lambda_m$ becomes an interpolant $I$ between $A$ and $B$.
-Note that any assignments to $\lambda_i$ that satisfy the expressions
-above will let an expression $I$ to be an interpolant.  Considering
-this, it is possible to obtain different interpolants for a
-interpolating problem by computing different assignments to
-$\lambda_i$.
-
-This enables the algorithm to preserve a set of interpolants by a
-linear expression template $a_1 x_1 + \cdots + a_n x_n + b \leq 0$
-with a constraint to describe a possible space of coefficient
-parameters $a_i$ and the constant parameter $b$.
-
-In practical, we may use linear programming solvers to obtain a model
-for $\lambda_i$.  By assigning concrete values to them, we can obtain
-a concrete interpolant.
-
-\paragraph{Computing a common interpolant}
-Next we consider a method to compute a common interpolant across two
-different interpolating problems.  That is, when given two
-interpolating problems $\left(A_1, B_1 \right)$ and $\left(A_2, B_2
-\right)$, the problem is about to compute the common interpolant I,
-which is an interpolant for the first problem and the second one at
-the same time.
-
-This problem is used when solving a following Horn clause solving
-problem, for instance.
-\[
-\implies
-% TODO
-\]
-
-To accomplish this, we simply need to compute the intersection of
-interpolant sets.  Let us call the interpolant space for the problems
-$\left(A_1, B_1 \right)$ and $\left(A_2, B_2 \right)$,
-$\mathcal{I}_1$, and $\mathcal{I}_2$.  In the previous method, a set
-of interpolants was represented by a parameterized linear expression.
-\begin{align*}
-\mathcal{I}_1 = \left\lbrace \mathbf{a}_1 \mathbf{x} \leq b_1 \mid
-C_1 (\mathbf{a}_1, b_1 ) \right\rbrace \\
-\mathcal{I}_2 = \left\lbrace \mathbf{a}_2 \mathbf{x} \leq b_2 \mid
-C_2 (\mathbf{a}_2, b_2 ) \right\rbrace
-\end{align*}
-To compute the intersection between them, it is sufficient to choose
-the same assignments of parameters and let them satisfy the both
-constraints $C_1$ and $C_2$.
-\[ \mathcal{I} = \left\lbrace \mathbf{a} \mathbf{x} \leq b \mid
-C_1(\mathbf{a}, b) \wedge C_2(\mathbf{a}, b) \right\rbrace \]
-
-However, if such an assignment of $\mathbf{a}$ and $b$ does not exist,
-no common solution exists across two problems because $\mathcal{I} =
-\emptyset$.
-
-By applying this approach, it is also possible to compute a common
-interpolant among more than three problems.
-
-\paragraph{Interpolation between logical formulas}
-We now extend the algorithm so that it can handle general logical
-formulas over linear arithmetic including disjunctions of linear
-expressions.  Here the algorithm takes two logical formulas over
-linear expressions as a problem input.  For convenience, we assume
-that they are represented in disjunctive normal form (DNF).  The
-solution for this problem is the interpolant that is a logical formula
-of linear expressions.
-
-First, consider the case of an interpolation problem between $A_1 \vee
-A_2$ and $B$.  When $(A_1 \vee A_2) \wedge B$ is inconsistent, $A_1
-\wedge B$ is necessarily inconsistent.  Let $\mathcal{I}_1$ denote the
-set of interpolants for $(A_1, B)$.  In the same manner, let
-$\mathcal{I}_2$ denote the set of interpolants for $(A_2, B)$.  We may
-naturally obtain an interpolant set by the intersection $\mathcal{I} =
-\mathcal{I}_1 \cup \mathcal{I}_2$ as a solution for $(A_1 \vee A_2,
-B)$.  When the algorithm computes $\mathcal{I}$ by the previous
-method, and it is $\emptyset$, we need to build a less simple solution
-by
-\[ \mathcal{I}' = \left\lbrace \left( I_1 \vee I_2 \right) \mid
-I_1 \in \mathcal{I}_1 \wedge I_2 \in \mathcal{I}_2 \right\rbrace \]
-Any formulas in the set $\mathcal{I}'$ are the interpolant for $(A_1
-\vee A_2, B)$ because
-\begin{align*}
-\forall I \in \mathcal{I}. & A_1 \vdash I \wedge A_2 \vdash I \wedge I \nvdash B \wedge \\
-& FV(I) = FV(A_1) \cup FV(A_2) \cup FV(B)
-\end{align*}
-
-Second, consider the case of an interpolation problem between $A$ and
-$B_1 \wedge B_2$.
-% TODO:
-
-By combining these two ways of interpolant construction, we may be
-able to build a small solution for more general cases like $(A_1,
-\ldots, A_n)$ and $(B_1, \ldots, B_n)$.  In order to obtain a small
-solution, we need to wisely choose the order of combining an
-interpolant set.
-
-\paragraph{Symmetric Interpolation}
-TODO
-
-\chapter{Horn clause solving}
-\label{chap:horn}
-
-We propose a new algorithm for solving recursion-less Horn clauses to
-generate a solution which is as simple as possible.  This enables us
-to obtain appropriate predicates that are suitable for predicate
-abstraction in the program verification techniques.
-
-The construction of the algorithm roughly consists of two parts.
-\begin{itemize}
-\item \textbf{Iterative Constraint Generation} is a process to build a
-  constraint to compute a solution.  It starts from generating the
-  weakest constraint in the sense that the resulting solution would
-  not be simple yet.  The process iteratively strengthen a constraint
-  until the strongest one while preserving the constraint
-  satisfiability.  The strongest constraint gives a simple solution.
-\item \textbf{Problem Restructuring} is a process to restructure the
-  problem to obtain a less simpler solution, if the constraint become
-  unsatisfiable during the iterative constraint generation.  This
-  process considers the unsatisfiable core of the constraint, and
-  wisely enlarge the solution size.
-\end{itemize}
-
-All predicate variables in an input problem are assigned simple
-templates.  Those predicate templates are parameterized.  In the
-iterative constraint generation process, algorithm builds a linear
-constraint that parameters in templates must satisfy.  This is done in
-a constructive manner, and constraints are build in the order of
-implication dependency in Horn clauses.  That is, when an input
-problem has a Horn clause thath has a predicate variable $P$ on its
-left-hand and $Q$ on its right-hand, a constraint to describe
-parameters in $P$'s template are first build before those in $Q$'s
-template.  When all parameters in the whole problem are described, the
-built constraint are evaluated to obtain a model.  Its instanciation
-becomes an appropriate solution for the input problem.
-
-The novelty of our algorithm lies in the incremental constraint
-refinement.  The algorithm tries to solve from the weakest constraint
-that an input problem allows that does not give a simple solution, and
-gradually strengthen the constraint until it reaches to the simplest
-solution that the problem has.  When a constraint becomes
-unsatisfiable during this step, the algorithm analyzes the
-unsatisfiable core and transform the problem to obtain a less simple
-solution.
-
-This construction of the algorithm enables us to compute a relatively
-simpler solution compared to existing algorithms while preserving the
-computation cost in a reasonable time.
-
-\section{Example}
-
-\paragraph {Tree interpolation}
-First, consider the following example.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-y \leq 2 & \implies B(y) \\
-A(x) \wedge B(y) & \implies 3x+2y \leq 10
-\end{align*}
-
-In this example, we would like to compute linear arithmetic formulas
-for the predicate variables $A$ and $B$.  They appear only once on
-both left-hand and right-hand globally in the problem.
-
-We try to obtain a simple solution that is expressed by linear
-expression templates, whose coefficients and constants are
-parameterized.  Assume that they have following predicate templates.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\
-B(y) : c_3 y + c_4 \leq 0
-\end{align*}
-
-By naturally encoding the first Horn clause with Farkas's lemma, we
-obtain a linear constraint
-$ \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 $.
-Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will
-make $A(x)$ valid for the first clause.  In the same manner, we obtain
-$ \lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0 $
-from the second clause, and
-$ c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0 $
-from the third.
-
-One of the models satisfying all the constraints above is
-$( c_1, c_2, c_3, c_4, \lambda_1, \lambda_2, \lambda_3 ) = (3, -3, 2, -4, 3, 2, 1)$.
-By assigning them to the template, we obtain the following solution.
-
-\begin{align*}
-A(x) : 3 x - 3 \leq 0 \\
-B(y) : 2 y - 4 \leq 0
-\end{align*}
-
-\paragraph {DAG problem solving}
-Second, we explain a case that some predicate variables appear
-multiple times on left-hand side in different Horn clauses.
-
-\begin{align*}
-x \leq 1 & \implies A(x) \\
-A(x) \wedge A(y) & \implies x+y \leq 2
-\end{align*}
-
-Assigning a template $A(x) : c_1 x + c_2 \leq 0$, we obtain the same
-constraint for the first clause as the previous example.
-
-\begin{align} \label{eq:constr1}
-\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0
-\end{align}
-
-Sinve we have $A$'s appearance twice on the second clause, it is
-possible for $A$ to have two expressions each of which is used for
-$A(x)$ and $A(y)$.  Although we need to use the same template for
-multiple occurrences in order to compute the simplest solution, it may
-make the whole constraints unsatisfiable if the simplest solution does
-not exist.
-
-Therefore, we generate a weaker constraint for the second clause that
-allows different expressions of $A$ for multiple occurrence.  If the
-whole constraint is satisfiable in the case, we try to solve in a
-stronger constraint to generate a simple solution.
-
-In this specific case, it is virtually equivalent to solve the
-following problem in a weaker constraint.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 1 & \implies A_1(x) \\
-x \leq 1 & \implies A_2(x) \\
-A_1(x) \wedge A_2(y) & \implies x+y \leq 2 \\
-\\
-\text{Templates:} \\
-A_1(x) \text{ : } & c_{1,1} x + c_{2,1} \leq 0 \\
-A_2(x) \text{ : } & c_{1,2} x + c_{2,2} \leq 0 \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-\end{align*}
-
-By duplicating the constraint~\ref{eq:constr1} of $A$ for newly
-created $A_1$ and $A_2$, we obtain the whole constraint as follows.
-
-\begin{align*}
-& \lambda_{1,1} = c_{1,1} \wedge - \lambda_{1,1} \geq c_{2,1} \wedge \lambda_{1,1} \geq 0 \wedge \\
-& \lambda_{1,2} = c_{1,2} \wedge - \lambda_{1,2} \geq c_{2,2} \wedge \lambda_{1,2} \geq 0 \wedge \\
-& c_{1,1} = \lambda_2 \wedge c_{1,2} = \lambda_2 \wedge c_{2,1} + c_{2,2} \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-This constraint is satisfiable and one model is
-\[ ( c_{1,1}, c_{1,2}, c_{2,1}, c_{2,2}, \lambda_{1,1}, \lambda_{1,2}, \lambda_2 ) =
-( 1, -1, 1, -1, 1, 1, 1 ) \] and the solution then becomes
-$A(x) : (x -1 \leq 0) \wedge (x -1 \leq 0)$.
-
-Since the weak constraint is satisfiable, we now refine the constraint
-to obtain the better solution.  That is, simply eliminate the
-duplication to build the following constraint for the original
-template.
-
-\begin{align*}
-& \lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \\
-& c_1 = \lambda_2 \wedge c_1 = \lambda_2 \wedge 2 c_2 \geq 2 \lambda_2 \wedge \lambda_2 > 0
-\end{align*}
-
-Solving this constraint simply gives us a solution $A(x) : x -1 \leq 0$.
-
-\paragraph {Disjunctive Horn clause solving}
-Finally, we explain a case that an input set of Horn clauses contain
-disjunctions on left-hand side.
-
-\begin{align*}
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 \vee -x+1 \leq 0 & \implies B(x) \\
-A(x) \wedge B(x) & \implies 1 \leq 0
-\end{align*}
-
-Same as the previous examples, templates for predicate variables are
-prepared.
-
-\begin{align*}
-A(x) : c_1 x + c_2 \leq 0 \\ B(x) : c_3 x + c_4 \leq 0
-\end{align*}
-
-The constraints generated are follows.
-
-\begin{align*}
-& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
-& \lambda_3 = c_3 \wedge \lambda_3 \geq c_4 \wedge \lambda_3 \geq 0 \wedge \\
-& - \lambda_4 = c_3 \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
-& c_1 + c_3 = 0 \wedge c_2 + c_4 > 0
-\end{align*}
-
-The constraints here become unsatisfiable, and one possible
-unsatisfiable core is:
-
-\begin{align*}
-\lambda_3 & = c_3 \\
-\lambda_3 & \geq c_4 \\
-- \lambda_4 & = c_3 \\
-\lambda_4 & \geq c_4 \\
-c_2 + c_4 & > 0 \\
-0 & \geq c_2
-\end{align*}
-
-Because the unsatisfiable core contains constraints originating from
-the disjunctions to be expressed in a single term template.  This let
-us determine to relax the template to allow a enlarged solution.
-
-Now we slightly change the problem and the template.
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A(x) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A(x) \wedge B_1(x) & \implies 1 \leq 0 \\
-A(x) \wedge B_2(x) & \implies 1 \leq 0 \\
-\\
-\text{Templates:} \\
-A(x) \text{ : } & c_1 x + c_2 \leq 0 \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-When we construct the constraints, as same as previous, we again
-obtain an unsatisfiable core:
-
-\begin{align*}
-\text{Constraint:} \\
-TODO \\
-\\
-\text{Unsatisfiable core:} \\
-TODO
-\end{align*}
-
-This time, the unsatisfiable core contains the constraints that
-originate from
-
-TODO
-
-\begin{align*}
-\text{Problem:} \\
-x \leq 0 \wedge -x \leq 0 & \implies A_i(x) \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-x+1 \leq 0 & \implies B_1(x) \\
--x+1 \leq 0 & \implies B_2(x) \\
-A_i(x) \wedge B_j(x) & \implies 1 \leq 0 \quad (i,j \in \left\lbrace 1,2 \right\rbrace ) \\
-\\
-\text{Templates:} \\
-A_i(x) \text{ : } & c_{1,i} x + c_{2,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-A(x) \text{ : } & A_1(x) \wedge A_2(x) \\
-B_i(x) \text{ : } & c_{3,i} x + c_{4,i} \leq 0 \quad (i \in \left\lbrace 1,2 \right\rbrace ) \\
-B(x) \text{ : } & B_1(x) \vee B_2(x)
-\end{align*}
-
-Finally, we obtain the following constraint:
-
-TODO
-
-The process also generates a predicate template for every predicate
-variable that the solution will become.  under certain solution
-template.
-
-\section{Preliminaries}
-
-For ease of discussion, our algorithm handles an input set of Horn
-clauses by a \emph{Horn graph} that express the equivalent problem.
-The graph is then used throughout the algorithm to represent the
-problem.  A Horn graph is a labelled directed acyclic graph
-$G=(V,E,\varphi)$.
-\begin{itemize}
-\item $V$ is a union of two disjoint sets of vertices; Horn term
-  vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex $u
-  \in V_\Rightarrow$ has eactly one outgoing edge to some $v \in V_T$,
-  denoted as $succ(u)$.
-\item $E$ is a set of labelled edges, which are expressed in the form
-  $(u,v,\theta)$ where $\theta$ is a finite map from variables to
-  variables.
-\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
-  term vertices, where L is a set of elements of a predicate variable
-  in the form $P(x_1, \ldots, x_{\mathrm{arity}(P)})$ or a linear
-  inequality $a_1 x_1 + \cdots + a_n x_n + b \leq 0$.
-\end{itemize}
-There are two kinds of edges in a Horn graph;
-\begin{itemize}
-\item for $u \in V_T, v \in V_\Rightarrow$, an edge takes a form
-  $(u,v,\theta)$, written $\edgel{u}{\theta}{v}$, or,
-\item for $u \in V_\Rightarrow, v \in V_T$, an edge takes a form
-  $(u,v,\emptyset)$, written $\edge{u}{v}$, where the mapping is
-  empty.
-\end{itemize}
-
-We define the relation $\leadsto$ over $V_T$ as follows.
-\[ u \mathop{\leadsto}^\theta v \Longleftrightarrow \exists a \in V_\Rightarrow;
-\edgel{u}{\theta}{a} \wedge \edge{a}{v} \] We may omit $\theta$ and
-write $u \leadsto v$ for convenience in later discussions. By the
-assumption that $G$ is acyclic, both $\rightarrow^+$ and $\leadsto^+$ are
-well-founded relations. $v_\bot$ is the greatest element with respect
-to both $\rightarrow^+$ and $\leadsto^+$.
-
-The meaning $\llbracket G \rrbracket $ of a Horn graph $G$ is a set of
-Horn clauses $\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$.
-\begin{align*}
-\mathcal{HC}_v & = \left( \bigwedge_{(u,v,\theta) \in E} \theta \varphi(u) \right) \Longrightarrow \varphi(succ(v)) \\
-\llbracket G \rrbracket & = \bigwedge_{v \in V_\Rightarrow} \mathcal{HC}_v
-\end{align*}
-We implicitly assume each $\mathcal{HC}_v$ is universally quantified
-by its all free variables $\textsc{FV}(\mathcal{HC}_v)$.
-
-Horn term vertices with predicate variable labels are unique each
-other that any of them do not have the same variable on their labels.
-From now on, we may denote a Horn term vertex with a predicate
-variable label by its predicate variable name if no ambiguity is
-present.
-
-Without loss of generality, we assume the existence of the \emph{root
-  vertex} $v_\bot \in V_T$ with linear expression labels $1 \leq 0$,
-meaning $\bot$.  This is because if we have a Horn clause $P \implies
-e$ which $e$ is a linear expression, we can convert it to $P \wedge
-\neg e \implies 1 \leq 0$.  We also assume that other vertices with
-linear expression labels have no incoming edges to them.
-
-For simplicity of discussion, we define \emph{term arity} $k_v$ for
-every Horn term vertex $v \in V_T$ with incoming edges.
-\begin{align*}
-k_v =
-\begin{cases}
-\mathrm{arity}(P) & \mbox{if } \varphi(v) = P(x_1,...,x_{\mathrm{arity}(P)}) \\
-0 & \mbox{if } \varphi(v) = \bot
-\end{cases}
-\end{align*}
-We let $K$ denote the maximum of $k_v$ over all $v \in V_T$.
-
-A variable mapping label $\theta$ on edges is used to map the index of
-one predicate variable to another. (TODO?)
-
-(Example?)
-
-The algorithm assigns a \emph{predicate template} for all predicate
-variables in $G$. A \emph{predicate templateset} is a set of mappings from
-a predicate variable to a predicate template.  A template for a
-predicate variable $P(x_1, \ldots, x_n)$ is a formula $\psi$
-constructed from the following definition.
-\begin{align*}
-\psi ::= & a_1 x_1 + \cdots + a_n x_n + b \leq 0 \mid \\
-& \psi \wedge \psi \mid \psi \vee \psi
-\end{align*}
-A linear expression in the definition above may also be represented in
-a equivalent vector form $\mathbf{a} \mathbf{x} + b \leq 0$ for short.
-Parameters $\mathbf{a}$ and $b$ are specified by a linear constraint.
-
-The solution of $G$ is a map $\rho$ from predicate variables to linear
-predicates of the form $\lambda x_1, \ldots ,x_n. \psi$ where
-parameters in $\psi$ are replaced with constant and $\rho[G]$ becomes
-a tautology.  A solution is \textit{simple} if $\rho[P]$ is of the
-form $\mathbf{a} \mathbf{x} + b \leq 0$ for every $P$ in
-$\mathrm{dom}(\rho)$.
-
-The problem of solving Horn clauses (represented by $G$) is to find a
-solution of $G$ (and a simple solution if possible).
-
-\section{Simple solution}
-
-We first propose an algorithm to compute a simple solution for the
-given Horn graph $G$ if one exists.  Later we add modifications to the
-algorithm for the case that a simple solution does not exist for $G$.
-In such a case, we transform the input graph $G$ to have a larger
-solution template.
-
-For each vertex $v \in V$, we define
-$\Delta(v) = (\mathbf{a}_v \mathbf{x}_v + b_v \leq 0, C_v)$
-, where
-\begin{itemize}
-\item $\mathbf{a}_v \mathbf{x}_v + b_v \leq 0$ is a linear inequality that
-  contains coefficient variables $\mathbf{a}_v$ and a constant $b_v$,
-  and
-\item $C_v$ is a set of linear expressions on coefficient variables.
-\end{itemize}
-Intuitively, it denotes the set of inequalities:
-\begin{align*}
-\left\lbrace
- \mathbf{a}_v \mathbf{x}_v + b_v \leq 0 \mid
- \exists \left( \textsc{FV}(C_v)
-  \setminus (\mathbf{a}_v \cup \left\lbrace b_v \right\rbrace
- \right); C_v
-\right\rbrace
-\end{align*}
-For a predicate variable vertex, $\mathbf{a}_v \mathbf{x}_v + b_v \leq
-0$ serves as a predicate template and $C_v$ describes the parameters
-$\mathbf{a}_v$ and $b_v$.
-
-$C_v$ is defined by well-founded induction on $v$ with respect to the
-relation $\rightarrow^+$. Then, $C_v$ can be shown as follows.
-\begin{itemize}
-\item For $v \in V_\Rightarrow$;
-\begin{align*}
-C_v & =
- \bigcup_{\edgel{u}{\theta}{v}} C_u \cup
- \bigcup_{0 < i \leq K}
- \left\lbrace
-  \mathbf{a}_{v,i} = \sum_{\edgel{u}{\theta}{v}} \mathbf{a}_{u, \theta^{-1} (i)}
- \right\rbrace \\
- & \hspace{2cm} \cup
- \left\lbrace
-  b_v \leq \sum_{\edgel{u}{\theta}{v}} b_u
- \right\rbrace \cup
- \bigcup_{k_{succ(v)} < i \leq K}
- \left\lbrace \mathbf{a}_{v,i} = 0 \right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that
-  $v \neq v_{\bot} \wedge \varphi(v) = \mathbf{a}_0 \mathbf{x}_v \leq b_0$
-  where $\mathbf{a}_0$ and $b_0$ are constants;
-\begin{align*}
-C_v =
-\left\lbrace
- \lambda_v \geq 0, \mathbf{a}_v = \lambda_v \mathbf{a}_0,
- b_v = \lambda_v b_0
-\right\rbrace
-\end{align*}
-\item For $v \in V_T$ such that $\varphi(v) = P(\mathbf{x}_v)$;
-\begin{align*}
-C_v = & \bigcup_{\edge{u}{v}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{a}_v, b_u = b_v \right\rbrace
-\right)
-\end{align*}
-\item For $v_\bot$;
-\begin{align*}
-C_{v_\bot} = \bigcup_{\edge{u}{v_\bot}} \left( C_{u} \cup
-\left\lbrace \mathbf{a}_u = \mathbf{0}, b_u > 0 \right\rbrace \right)
-\end{align*}
-\end{itemize}
-
-Any model $\sigma$ to the constraint $C_{v_\bot}$ gives a simple
-solution for $G$ if $C_{v_\bot}$ is satisfiable, i.e., we should have:
-
-\begin{quote}
-If $\models \sigma(C_{v_\bot})$, then $\rho$ defined by
-\begin{align*}
- \rho = \left\lbrace
-  \left( P, \sigma(\mathbf{a}_v \mathbf{x}_v \leq b_v) \right) \mid
-  \forall v \in V_T; \varphi(v) = P(\mathbf{x}_v)
- \right\rbrace
-\end{align*}
-is a solution of $G$.
-\end{quote}
-
-If $C_{v_\bot}$ is unsatisfiable, no simple solution exists for $G$.
-There are two cases that a simple solution does not exist for a
-predicate variable.
-\begin{itemize}
-\item A predicate variable appears multiple times on left-hand of Horn
-  clauses, and those appearances require different expressions as an
-  assumption.
-\item A predicate variable is implied by a disjunctive Horn clause,
-  which has disjunctions on its left-hand, and a common premise that
-  the disjunctions induce is not strong enough for solving a problem.
-\end{itemize}
-
-The first case occurs in a case like follows.
-\begin{align*}
-% conj.txt ... originated from uc/01.txt
-\left( x \leq 0 \right) \wedge \left( y \leq 0 \right) & \implies P(x,y) \\
-P(x,a) \wedge P(b,y) & \implies x+y \leq 0
-\end{align*}
-In this example, the predicate variable $P$ appears twice on the
-second clause and they require different premises; $x \leq 0$ for the
-first and $y \leq 0$ for the second.  They cannot be represented by
-$\mathbf{ax} + b \leq 0$ template at the same time.
-
-The second case occurs in a case like follows.
-\begin{align*}
-% disj-simp.txt ... originated from 23simp_neg.txt
-\left( x+1 \leq 0 \right) \vee \left( -x+1 \leq 0 \right) & \implies P(x) \\
-P(x,y) \wedge \left( x \leq 0 \right) \wedge \left( -x \leq 0 \right) & \implies \bot
-\end{align*}
-In this example, the predicate variable $P$ is implied by the first
-clause, which is a disjunctive Horn clause.  As a simple solution, $P$
-can take a common consequence that both of $x+1 \leq 0$ and $-x+1 \leq
-0$ induce.  However, such a simple linear expression does not exist
-except for $0 \leq 0$, which is equivalent to $\top$, and it is not
-sufficient to satisfy the second clause.
-
-Both cases can be fixed by enlarging the solution template.
-
-\section {Graph transformation}
-
-If a simple solution does not exist for $G$, we try to give a less
-simpler solution by transforming $G$ and relaxing the constraint
-$C_{v_\bot}$.  This transformation takes place by considering the
-unsatisfiable core $\mathcal{U}$.  An unsatisfiable core $\mathcal{U}$
-is a subset of $C_{v_\bot}$ which is inconsistent itself.  Depending
-on the reason for the absence of a simple solution, the way of the
-transformation differs.
-
-Algorithm~\ref{alg:solveHorn} shows the updated pseudo-code of our
-Horn clause solving algorithm with the graph transformation.  It takes
-an input Horn graph and returns a mapping between predicate variables
-and linear arithmetic formulas.
-
-\newcommand{\solveHorn}{\ensuremath{\mbox{\sc SolveHornGraph}}}
-\newcommand{\genConstr}{\ensuremath{\mbox{\sc GenerateConstraint}}}
-\newcommand{\transGraph}{\ensuremath{\mbox{\sc TransformGraph}}}
-
-\begin{algorithm}
-\caption{$ \solveHorn (G) $}\label{alg:solveHorn}
-\begin{algorithmic}
-\STATE {$v_\bot \gets $ the root node of $G$}
-\STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-\STATE {$T \gets $ initial predicate templateset for $G$}
-\REPEAT
-  \STATE {$C_{v_\bot} \gets \genConstr (G, v_\bot, V_\star)$}
-  \IF {$C_{v_\bot}$ is satisfiable}
-    \STATE {$\sigma \gets $ one model of $C_{v_\bot}$}
-    \STATE {$V_\star \gets V_\star \cup \left\lbrace u \in V_T \mid \forall v \in V_\star; u \leadsto v \right\rbrace$}
-  \ELSE
-    \STATE {$\mathcal{U} \gets $ unsatisfiable core for $C_{v_\bot}$}
-    \STATE {$(G, T) \gets \transGraph (G, T, \mathcal{U})$}
-    \STATE {$V_\star \gets \left\lbrace v_\bot \right\rbrace$}
-  \ENDIF
-\UNTIL {$V_\star = V_T$}
-\RETURN {$\sigma \left( T \right)$}
-\end{algorithmic}
-\end{algorithm}
-\transGraph~procedure determines the unsatisfiable core of the
-constraint and transform the Horn graph $G$ and predicate templates.
-As mentioned earlier, two ways of graph transformation exist.
-
-\paragraph{Conjunction split}
-For the first case previously mentioned, the unsatisfiable core
-$\mathcal{U}$ contains coeffcient constraint expressions with a
-coefficient parameter of certain predicate variable.  If the predicate
-variable $P$ is the cause of unsatisfiability, some constraint
-expressions in the unsatisfiable core have a parameter $a_{P,i}$
-and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is copied to
-$P_1, \ldots, P_n$ so that the different appearances of $P$ on
-left-hand of Horn clauses can take different premises.  The solution
-template of $P$ becomes the conjunction of $P_1, \ldots, P_n$.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-Note that after this duplication of the predicate vertex $P$, in order
-to induce different linear expressions for $P_1, \ldots, P_n$, it may
-be necessary for preceeding predicate variable vertices in $G$ to be
-duplicated in the same manner as $P$. It is because that different
-derivation of $P$'s implication relation may be required to get
-different consequences.
-
-However, succeeding predicate vertices of $P$ in $G$ and other
-vertices than predecessors are not affected because those implications
-that have $P$ in their assumption can selectively choose the premise
-that $P_1, \ldots, P_n$ leads.
-
-Therefore, this graph transformation only affects the predecessors of
-$P$.
-
-\paragraph{Disjunction split}
-For the second case, the unsatisfiable core $\mathcal{U}$ contains
-coefficient constraints that unifies induced linear expressions from
-each term in disjunctions to be the same expression.  Similarly as
-previous, if the predicate variable $P$ is the cause of
-unsatisfiability, some constraint expressions in the unsatisfiable
-core have a parameter $a_{P,i}$ and/or $b_P$ in the following way.
-\begin{align*}
-% TODO:
-& \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
-& \left( \exists i, j, k;
-\left\lbrace \left( \mathbf{a}_{v_a,j} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right),
-\left( \mathbf{a}_{v_b,k} = \ldots + \mathbf{a}_{v_P,i} + \ldots \right)
-\right\rbrace \subseteq \mathcal{U} \right)
-\end{align*}
-
-In such a case, the predicate variable $P$ in question is separated to
-$P_1, \ldots, P_n$ each of which is implied by a term, which was
-originally a part of a disjunction to imply $P$.  This intuitively
-enables us to avoid generating a common linear expression that
-disjunctive terms induces, and to propagate each term's premises to
-$P$'s successors instead.
-
-<fig...>
-
-Here, the graph $G=(V,E,\varphi)$ will be transformed into
-$G'=(V',E',\varphi')$ such that:
-\begin{align*}
-V' & = V \setminus \left\lbrace v_P \right\rbrace \cup
-  \left\lbrace v_{P_i} \mid \emptyset \right\rbrace \\
-E' & = E \\
-\varphi' & = \varphi
-\end{align*}
-
-After this duplication of the predicate vertex $P$, unlike the
-conjunction split, this results in producing a disjunction at
-succeeding predicate variables vertices ($R$ in this case).
-\begin{align*}
-(P_1 \wedge Q) \vee (P_2 \wedge Q) \implies R
-\end{align*}
-Also, predicate variables that appeared with $P$ in the same
-assumption ($Q$ in this case) needs to be duplicated by conjunction,
-so as to obtain different premises that go along with $P_i$.
-Therefore, the effect of disjunction split may reach to neighboring
-predicate vertices.
-
-Consider these two ways of transformation, we show a pseudo-code for
-\transGraph~procedure.  The procedure is given the graph $G$, the
-predicate templateset $T$, and the unsatisfiable core
-$\mathcal{U}$. It returns the transformed graph $G'$ and the new
-templateset $T'$.
-
-\begin{algorithm}
-\caption{$ \transGraph (G, T, \mathcal{U}) $}\label{alg:transGraph}
-\begin{algorithmic}
-% TODO:
-\RETURN {$\left( G', T' \right)$}
-\end{algorithmic}
-\end{algorithm}
-
-\paragraph{Combinatorial explosion in disjunctions}
-Virtually, if multiple disjunctions exists in an input set of Horn
-clauses, a na\"{i}ve way of building a solution is solve all
-combinations of choices out of every disjunction and concatenate each
-expression with conjunctions and/or disjunctions appropriately.  This
-may cause a combinatorial explosion.
-
-Conside example below. $e_i$ represents some conjunctions of linear
-expressions.
-% TODO: Can't we replace e_i with actual expressions?
-\begin{align*}
-(e_1 \vee e_2) \implies P \\ (e_3 \vee e_4) \implies Q \\ P \wedge Q \implies \bot
-\end{align*}
-In this example, a na\"{i}ve algorithm may solve following four
-problems separately.
-\begin{itemize}
-\item $e_1 \implies P_1 \qquad e_3 \implies Q_1 \qquad P_1 \wedge Q_1 \implies \bot$
-\item $e_2 \implies P_2 \qquad e_3 \implies Q_2 \qquad P_2 \wedge Q_2 \implies \bot$
-\item $e_1 \implies P_3 \qquad e_4 \implies Q_3 \qquad P_3 \wedge Q_3 \implies \bot$
-\item $e_2 \implies P_4 \qquad e_4 \implies Q_4 \qquad P_4 \wedge Q_4 \implies \bot$
-\end{itemize}
-Then the solution is built by follows.
-\begin{itemize}
-\item $P = ( P_1 \vee P_2 ) \wedge ( P_3 \vee P_4 )$
-\item $Q = ( P_1 \vee P_3 ) \wedge ( P_2 \vee P_4 )$
-\end{itemize}
-However, this method is not efficient.
-
-\paragraph{Algorithm}
-To avoid this, our algorithm lets a predicate variable $P$ which is
-disjunctively implied on the Horn graph $G$ to have a linear
-expression that is consequences of all terms in the disjunction.  This
-prevents the possibility of choices for disjunctions to be propagated
-to succeeding vertices. The simple linear expression is propagated
-instead.
-
-On the other hand, the algorithm allows a predicate variable $P$ that
-appears on multiple implications or appears multiple times in a single
-implication to take different linear expressions as a premise at
-different appearances.  This is for the algorithm to determine the
-cause of unsatisfiability efficiently.  It is done by copying a
-predicate templates and its parameters' constraint with appropriate
-renaming.
-
-The algorithm first execute this duplication on all predicate
-variables with multiple appearances.  It is the weakest constraint
-that the algorithm makes for the certain $G$.  Then, the algorithm
-gradually reduces the number of vertices that are duplicated by each
-time testing the satisfiability of the constraint.  When the
-constraint bebcome unsatisfiable at the certain iteration, the
-algorithm can determine a predicate vertex which has to be split in a
-conjunctive template.  If all the predicate variable constraints are
-no longer duplicated and the constraint is the strongest, any model
-for such constraint gives a simple solution under the Horn graph $G$
-and the templateset $T$.
-
-Note that duplicating the constraint in a na\"{i}ve way may cause the
-constraint to exponentially grow in size.  Our algorithm prevent it by
-simplifying a predicate vertex's constraint $C_v$ by quantifier
-elimination when $v$ is duplicated in generating linear constraint.
-For a predicate variable vertex $v_P$, variables except for the
-coefficient parameter $\mathbf{a}_P$ and the constant parameter $b_P$
-are eliminated.  Due to the constraints generation process, the
-simplified constraint is also linear and can be efficiently check the
-satisfiability.
-
-We finally show the modified version of pseudo-code for $\genConstr$
-procedure.  It produces the constraint $C_v$ for a given vertex $v$,
-considering constraint duplication and simplification.
-
-\begin{algorithm}
-\caption{$ \genConstr (G, v, V_\star) $}\label{alg:genConstr}
-\begin{algorithmic}
-\STATE {$C \gets \emptyset$}
-\FORALL {$v_\Rightarrow \in preds_G(v)$}
-  \FORALL {$v_p \in preds_G(v_\Rightarrow)$}
-    \IF {$\varphi (v_p)$ is a predicate variable}
-      \STATE {$C \gets C \cup \genConstr (G, v_p, visited)$}
-    \ELSE
-      \STATE { $id$ }
-    \ENDIF
-  \ENDFOR
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-traverses the input graph in the topological order on
-$V_\Rightarrow$, and concatenate the constraints from predecessors to
-build the problem constraints.  Each process to build a constraint on
-$\mathcal{HC}_v$ for all vertices $v \in V_\Rightarrow$ is equivalent
-to generate a linear constraint in interpolation.
-
-\chapter{Experiment}
-\label{chap:experiment}
-
-We incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It
-is specifically designed to verify the correctness of higher-orde
-programs written in a small subset of ML.  It infers dependent types
-for input higher-order programs \cite{conf/ppdp/UnnoK09}.
-
-Our implementation make use of Z3 \cite{conf/tacas/MouraB08} as
-an SMT solver.
-
-For simplicity of the implementation, we assume that the linear
-inequalities in the integer space.  Because of this, our algorithm
-lost the completeness.  By using Motzkin's transpose theorem
-\cite{journals/networks/Rajan90}, it is possible to handle linear
-inequalities in both $\mathbf{ax} + b \leq 0$ and $\mathbf{ax} + b <
-0$.
-
-\section{Interpolating theorem prover}
-
-Our interpolating theorem prover is used as a sub-procedure to
-solve Horn clauses.
-
-\begin{itemize}
-\item CSIsat \cite{conf/cav/BeyerZM08}
-\item FOCI \cite{website/foci}
-\item CLP-Prover \cite{website/clp}
-\end{itemize}
-
-\section{Horn clause solver}
-
-\chapter{Related work}
-\label{chap:related}
+\input{intro.tex}
+\input{interpolation.tex}
+\input{horn.tex}
+\input{experiment.tex}
 \input{related.tex}
-
-\chapter{Future Work}
-\label{chap:future}
-
-\paragraph{Non-linear extension}
-\cite{conf/cav/DaiXZ13}
-
-\chapter{Conclusion}
-\label{chap:conclusion}
+\input{future.tex}
+\input{conclusion.tex}
 

### 2013-08-26 article/conclusion.tex
Conclusion

\chapter{Conclusion}
\label{chap:conclusion}

In this thesis, we have suggested a new algorithm to discover
predicates that are suitable for program verification.  We put a
hypothesis that the program invariants are prone to be expressed in
simple expressions.  Based on this hypothesis, our algorithm tries to
obtain a small predicate with a fewer number of terms.

To accomplish this goal on Craig interpolation over linear
arithmetics, we first explained a method to reduce interpolation into
linear constraint generation by using Farkas's Lemma.  This method
preserves a set of solutions during the computation process so to make
it possible to choose a small solution later. This enabled a Horn
clause solver to build small solutions when it uses the method as a
subroutine due to the ability to choose a common predicate among
different interpolating problems. Additionally, we mentioned the
possibility of chosing small solutions in interpolating general
formulas by wisely choose the order to combine set of solutions during
the computation. In the end we extended the method to solve symmetric
interpolation problems to perform predicate discovery at multiple
locations in the program at the same time.

Next we extended our interpolating method to solve recursion-less
disjunctive Horn clauses.  Same as previous, this method also
describes a set of predicates for each predicate variable in linear
constraints.  The algorithm builds a solution in constructive manner
and is thus scalable for the size of input problems. If the algorithm
fails to find a simple solution, it tries to find a larger solution in
error-guided manner.  By observing an unsatisfiable core, the
algorithm enlarges a size of solution for certain predicates.  We have
also proposed to simplify linear constraints by quantifier elimination
for preventing exponential growth of them.

We have incorporated our algorithm into MoCHi and confirmed the
effectiveness to a certain level.

We would like to extend the algorithm further to achieve better
performance in constraint generation.  It is also desired to support
uninterpreted function symbols in our theory.

### 2013-08-27 article/future.tex
Future work ongoing

@@
 
-\paragraph{Non-linear extension}
-\cite{conf/cav/DaiXZ13}
+Our proposed algorithms have still numbers of rooms for further
+improvement.  Based on the experiment result from the previous
+chapter, here we give raise some analaysis and discussion for them.
+
+\paragraph{Space explosion for constraint duplication}
+Our experimental implementation did not simplify the linear constraint
+for every predicate variable and it caused a linear constraint
+unsolvable due to the shortage of space resources by exponential
+growth of constraints.  As we described in the algorithm explanation,
+it may be possible to perform quantifier elimination for reduce the
+size of linear constraints.  On the other hand without the
+optimization, the constraint duplication may occur every time a
+constraint is generated for the predicate vertices with multiple
+successors.
+
+We think that the algorithm with such optimization would not cause a
+memory shortage because constraints gently grow according to roughly
+the number of vertices.  Although the worst case can be produced by
+causing a conjunction split frequently, we consider that it is not the
+common case in program verification because of our hypothesis.
+
+In practice, the size of constraint may be limited by simplification,
+and additionally, normally the algorithm does not need to preserve the
+constraint until the time to build a concrete solution in the end.
+
+\paragraph{Quantifier elimination}
+It may be useful to prepare a customized algorithm for performing
+quantifier elimination for linear expressions because linear
+constraints that are generated from our algorithm have certain pattern
+in their format.
+
+In the preliminary expriment before incorporating into MoCHi, although
+we have tried the quantifier elimination over integer theory, the
+result was unsatisfactory in the sense that simplified formulas had
+disjunctions.  It caused the loss of efficiency in solving constraints
+and generating unsatisfiable cores.
+
+We have also tried the quantifier elimination over real space, and the
+result was better than the integer version.  However, because our algorithm
+first treats all linear inequalities over integer space, more study on
+the gap betweem these two are required when we consider the completeness.
+
+\paragraph{Solving linear constraints}
+
+Although we solve the constraints by a general SMT solver Z3, GNU
+Linear Programming Kit (GLPK) may be sufficient for solving
+constraints.  The linear programming solver should support sparse
+matrix because our algorithm generates constraints that are sparse.
+
+\paragraph{Generation and choice of unsatisfiable core}
+Our Horn clause solver make use of Z3 to obtain an unsatisfiable core
+from the linear constraints.  However the extraction of unsatisfiable
+core from the proof may cost certain level of computation.
+Originally, it is better that the theorem prover customizes to extract
+
+\paragraph{Applications for other work}
+In our research, we put a special focus on solving Horn clauses for
+dependent type inference in higher-order program verification.
+However, solving Horn clauses are becoming common to describe various
+kinds of programs' behavior. For instance in verifying multi-threaded
+programs, Horn clauses can describe the conditions of data exchange
+between threads, and confirm the validity of semaphores for mutural
+exclusion.  We expect our algorithm achieve better performance in
+other work when customized into different contexts.
+
+By puting such effort for further improvement on our algorithm, we can
+conclude that the algorithm may be able to achive better performance
+in Horn clause solving, and become a necessary tool for program verification.

### 2013-08-28 article/future.tex
Future work

@@
 We have also tried the quantifier elimination over real space, and the
-result was better than the integer version.  However, because our algorithm
-first treats all linear inequalities over integer space, more study on
-the gap betweem these two are required when we consider the completeness.
+result was better than the integer version.  However, because our
+algorithm first treats all linear inequalities over integer space,
+more study on the gap betweem these two are required when we consider
+the completeness.
 
 \paragraph{Solving linear constraints}
-
 Although we solve the constraints by a general SMT solver Z3, GNU
@@
 \paragraph{Generation and choice of unsatisfiable core}
-Our Horn clause solver make use of Z3 to obtain an unsatisfiable core
-from the linear constraints.  However the extraction of unsatisfiable
-core from the proof may cost certain level of computation.
-Originally, it is better that the theorem prover customizes to extract
+We can efficiently build linear constraints for the given $G$ after
+the transformation in $\transGraph$~procedure by appropriate inference
+from the previous constraint and the SMT query result.  Our Horn
+clause solver make use of Z3 to obtain an unsatisfiable core from the
+linear constraints.  However the extraction of an unsatisfiable core
+from the proof sums up to a large portion of the whole computation
+time in our algorithm.
+
+Although our implementation in this work did not consider such the way
+of optimization to reduce the cost of linear constraints queries, it
+is necessary to reduce the number of them for reduce the computation
+cost.  For instance, the optimization is possible by appropriatealy
+updating $V_\star$ in $\solveHorn$~procedure since the vertices that
+are already examined for split does not need to be considered again.
+
+The choice of subset of constraints for an unsatisfiable core is also
+important.  In some cases, there exists different multiple cores and
+the algorithm returns a different shape of the solution in relation to
+the choice of them.  Thus, we need further study to optimize the order
+of choice of multiple unsatisfiable cores if ones exists.
+
+\paragraph{Separation of edges}
+To transform the graph by $\transGraph$~procedure, there may be
+multiple transformations possible.  For example, consider a predicate
+variable $P$ and edges $\left\lbrace e_1, e_2, e_3, e_4 \right\rbrace$
+that go out from the vertex.  If an unsatisfiable core contains a set
+of edges $\left\lbrace e_1, e_2 \right\rbrace$,
+$\transGraph$~procedure determines to split those edges by duplicating
+$P$.  However at this point, $e_3$ and $e_4$ are free to be connected
+to any duplication of $P$.  Alghough our current algorithm split
+$\left\lbrace e_1, e_2, e_3, e_4 \right\rbrace$ into four duplications
+$\left\lbrace e_1 \right\rbrace$ to $\left\lbrace e_4 \right\rbrace$,
+it is possible to split into $\left\lbrace e_1, e_3\right\rbrace$ and
+$\left\lbrace e_2, e_4 \right\rbrace$ for example.  By intellilgently
+choose this, we may be able to obtain much simpler solution than the
+current algorithm.
+
+We suggest to reduce this edge split problem into a graph coloring
+problem.  By generating a graph which a set of vertices have an edge
+of Horn graph as a label and solve a coloring problem on it, the
+algorithm can wisely determine the minuminm number of duplications.
+
+\paragraph{Support for Horn clauses with Recursions}
+Considering usages of Horn clause solving in program verification, it
+is useful to update our algorithm so that it can solve Horn clauses
+with recursions as well. For instance, this makes possible to generate
+invariants upon the loop programs on the control flow graphs, rather
+than to generate a set of disjunctive constraints
+\cite{conf/pldi/BeyerHMR07}.
+
+One known method to let the algorithm support Horn clauses with
+recursions is to unroll a looping predicate in the problem
+\cite{conf/ppdp/UnnoK09}.  By computing a fixedpoint of the predicate
+solution over unrolled programs, it is possible to virtually solve
+Horn clauses with recursions as well.
 
@@
 conclude that the algorithm may be able to achive better performance
-in Horn clause solving, and become a necessary tool for program verification.
+in Horn clause solving, and become a necessary tool for program
+verification.

### 2013-08-28 article/experiment.tex
Experiment chapter ongoing

@@
 
-We incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It
-is specifically designed to verify the correctness of higher-orde
-programs written in a small subset of ML.  It infers dependent types
-for input higher-order programs \cite{conf/ppdp/UnnoK09}.
+We have incorporated our algorithm to the software model checker for
+higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It is
+specifically designed to verify the correctness of higher-orde
+programs written in a small subset of ML.  This verification tool
+abstracts a given input program by inferring dependent types for
+higher-order programs \cite{conf/ppdp/UnnoK09}.
 
-Our implementation make use of Z3 \cite{conf/tacas/MouraB08} as
-an SMT solver.
+We implemented our proposed algorithm into MoCHi by two steps.  First,
+we replace an existing theorem prover with our interpolating theorem
+prover as a sub-routine for solving Horn clauses in dependend typing.
+We used a method to obtain a common small interpolant across different
+interpolating problems, while reducing a Horn clause problem into
+interpolation \cite{conf/ppdp/UnnoK09}.  Second, we completely
+replaced MoCHi's Horn solving algorithm with our algorithm.
 
@@
 
+While solving interpolating problems, our algorithm treated linear
+equalities in the form $\mathbf{ax} + b = 0$ specially without
+normalizing into a conjunction of linear equalities so that the
+algorithm obtain a linear equality in a set of interpolants.  It is
+because we observed some cases that requires a equality relation
+between two variables but not an inequality as an program invariant.
+
+Note that, however, in Horn clauses solving problems, when a predicate
+variable $P$ needs to take an equality $=$ or negation $\neq$, the
+predicate template grows during the solving process as
+$\mathbf{ax}-b \leq 0 \wedge -\mathbf{ax}+b \leq 0$ in the former
+case, and $\mathbf{ax}-b+1 \leq 0 \vee -\mathbf{ax}+b+1 \leq 0$ in the
+latter case.  We also thought it possible to obtain predicates that
+are similar to program invariants by considering multiple paths.
+
+For linear constraint satisfaction, we used GNU Linear Programming Kit
+(GLPK) first for its light computation load.  For Horn clause solving
+however, Our implementation make use of Z3 \cite{conf/tacas/MouraB08}
+as an SMT solver to obtain unsatisfiable core of the constraints.
+
 \section{Interpolating theorem prover}
 
-Our interpolating theorem prover is used as a sub-procedure to solve
-Horn clauses.
-
+There are some interpolating theorem provers known.
 \begin{itemize}
-\item CSIsat \cite{conf/cav/BeyerZM08}
-\item FOCI \cite{website/foci}
-\item CLP-Prover \cite{website/clp}
+\item \textsc{CSIsat} \cite{conf/cav/BeyerZM08} combines linear
+  arithmetic reasonings and SAT-based logical reasonings.
+\item \textsc{FOCI} \cite{website/foci} extracts an interpolant by the
+  proof-based method.
+\item \textsc{CLP-Prover} \cite{website/clp} generates an interpolant
+ from constraints by linear arithmetic reasonings. It only supports
+ conjunctive problems.
 \end{itemize}
 
-\section{Horn clause solver}
+In our experiment, we compared the performance of our algorithm with
+CSIsat, that is mainly used as a subprocedure of MoCHi.  Note that
+CSIsat may return different interpolants for an interpolating problem
+because of the randomized algorithm.

### 2013-08-28 article/horn.tex
Conclusion
Formatting for final

@@
 Consider these two ways of transformation, we show a pseudo-code for
-\transGraph~procedure.  The procedure is given the graph $G$, the
-predicate templateset $T$, and the unsatisfiable core
+\linebreak \transGraph~procedure.  The procedure is given the graph
+$G$, the predicate templateset $T$, and the unsatisfiable core
 $\mathcal{U}$. It returns the transformed graph $G'$ and the new
@@
 \end{algorithm}
-
-% TODO: What's this paragraph?
-traverses the input graph in the topological order on $V_\Rightarrow$,
-and concatenate the constraints from predecessors to build the problem
-constraints.  Each process to build a constraint on $\mathcal{HC}_v$
-for all vertices $v \in V_\Rightarrow$ is equivalent to generate a
-linear constraint in interpolation.

### 2013-08-28 article/jabstract.tex
Formatting for final

\begin{jabstract}
近年のプログラム検証技術においては、自動定理証明が重要な役割を担っている。特にソフトウェア モデル検査の分野では、補間問題およびその一般化である Horn 節充足問題が注目を集めてきた。後者の問題は、循環した依存関係を持たない、未知の述語変数を複数含む Horn 節の集合の解を求める問題である。Horn 節充足問題の解、すなわち Horn 節集合を恒真とするような述語変数への代入は、プログラムの不変式の候補として用いられる。しかしながら、Horn 節充足問題の既存解法が出力する解は複雑かつ特殊すぎて不変式にならないことが多い。
上述の問題を解決するために、われわれは線形代数上の Horn 節充足問題を解く新しいアルゴリズムを提案する。われわれのアルゴリズムは、プログラム不変式が単純な式で表されるという仮定のもとで、小さな述語の生成を試みる。この目的を達成するため、このアルゴリズムはある大きさの線形不等式の雛形を未知の述語変数それぞれに対して生成し、Horn 節充足問題を線形計画法に帰着させる。そして、解が見つかるまで雛形の拡大を繰り返す。この提案手法では、合理的な時間内に比較的小さな解を構成することができる。われわれは、このアルゴリズムを実装し関数型言語向けのソフトウェア モデル検査器 MoCHi に組み込んだ。実験により、提案手法について一定の有効性を確認した。
\end{jabstract}

### 2013-08-28 article/thesis.tex
Include graphics
Formatting for final

@@
 
-\jtitle{Lipsum}
-\jauthor{Lipsum}
-\jsupervisor{Lipsum}
+\jtitle{プログラム検証のための非循環 Horn 節充足問題における小さな解の生成}
+\jauthor{竹井 悠人}
+\jsupervisor{小林 直樹}
 
@@
 \usepackage{algorithmic}
-\usepackage{graphicx}
+\usepackage[dvipdfmx]{graphicx}
 \usepackage{listings}
@@
 \input{eabstract.tex}
-\begin{jabstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.  Morbi
-vulputate ornare arcu et malesuada.  Proin sem urna, feugiat et velit
-ac, elementum blandit orci.  Etiam ut neque ac felis laoreet venenatis
-eget quis enim.  Ut aliquam dapibus sem egestas molestie.  Ut sit amet
-porttitor orci, id fermentum dui.  In non dui at ante sagittis
-malesuada at in neque.  Fusce euismod velit quis ante ullamcorper
-qsodales.
-\end{jabstract}
+\input{jabstract.tex}
 \maketitle

### 2013-08-28 article/experiment.tex
Experiment

@@
 higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It is
-specifically designed to verify the correctness of higher-orde
+specifically designed to verify the correctness of higher-order
 programs written in a small subset of ML.  This verification tool
@@
 we replace an existing theorem prover with our interpolating theorem
-prover as a sub-routine for solving Horn clauses in dependend typing.
+prover as a sub-routine for solving Horn clauses in dependent typing.
 We used a method to obtain a common small interpolant across different
@@
 algorithm obtain a linear equality in a set of interpolants.  It is
-because we observed some cases that requires a equality relation
+because we observed some cases that require an equality relation
 between two variables but not an inequality as an program invariant.
@@
 (GLPK) first for its light computation load.  For Horn clause solving
-however, Our implementation make use of Z3 \cite{conf/tacas/MouraB08}
+however, our implementation make use of Z3 \cite{conf/tacas/MouraB08}
 as an SMT solver to obtain unsatisfiable core of the constraints.
@@
 \item \textsc{CSIsat} \cite{conf/cav/BeyerZM08} combines linear
-  arithmetic reasonings and SAT-based logical reasonings.
+  arithmetic reasoning and SAT-based logical reasoning.
 \item \textsc{FOCI} \cite{website/foci} extracts an interpolant by the
@@
 \item \textsc{CLP-Prover} \cite{website/clp} generates an interpolant
- from constraints by linear arithmetic reasonings. It only supports
+ from constraints by linear arithmetic reasoning. It only supports
  conjunctive problems.
@@
 In our experiment, we compared the performance of our algorithm with
-CSIsat, that is mainly used as a subprocedure of MoCHi.  Note that
+CSIsat, that is mainly used as a sub procedure of MoCHi.  Note that
 CSIsat may return different interpolants for an interpolating problem
 because of the randomized algorithm.
+
+\paragraph{Comparison over simple interpolation}
+First, we compare by simply generating an interpolant without
+incorporating into MoCHi.  \textsc{Yint} represents our interpolating
+theorem prover.
+
+\begin{itemize}
+\item Disjunctive interpolation in $A$.
+\[\begin{array}{r l}
+\varphi_A & (x-1 \geq 0) \vee (x+1 \leq 0) \\
+\hline
+\varphi_B & x=0
+\end{array}\]
+The result for this:
+\[\begin{array}{l c}
+\text{\textsc{Yint}} & (x+1 \leq 0) \vee (-x+1 \leq 0) \\
+\text{\textsc{CSIsat}} & (-x \leq -1 \vee x \leq -1)
+\end{array}\]
+\item Disjunctive interpolation in $B$.
+\[\begin{array}{r l}
+\varphi_A & x=0 \\
+\hline
+\varphi_B & (x-1 \geq 0) \vee (x+1 \leq 0)
+\end{array}\]
+\[\begin{array}{l c}
+\text{\textsc{Yint}} & -x=0 \\
+\text{\textsc{CSIsat}} & (-x \leq 0 \wedge x \leq 0)
+\end{array}\]
+\item Complex formula in $A$.
+% complicated disj
+\[\begin{array}{r l}
+\varphi_A & ((x-5<0) \wedge (-x<0)) \vee ((x<0) \wedge (-x-5<0)) \\
+\hline
+\varphi_B & x=0
+\end{array}\]
+\[\begin{array}{l c}
+\text{\textsc{Yint}} & (x+1 \leq 0) \vee (-x+1 \leq 0) \\
+\text{\textsc{CSIsat}} & (x < 0 \vee -x < 0)
+\end{array}\]
+\end{itemize}
+For those problems that are simple enough, Yint and CSIsat returns the
+same results.
+
+Next we tried the problem that an equality interpolant exists.
+% equality
+\[\begin{array}{r l}
+\varphi_A & x=0 \\
+\hline
+\varphi_B & x-1=0
+\end{array}\]
+\[\begin{array}{l c}
+\text{\textsc{Yint}} & x=0 \\
+\text{\textsc{CSIsat}} & x \leq 0
+\end{array}\]
+
+\bigskip\centerline{}
+
+% quantifier
+\[\begin{array}{r l}
+\varphi_A & (y-z=0) \wedge (x-z=0) \\
+\hline
+\varphi_B & (-w+y+1 \leq 0) \wedge (-w+x \geq 0)
+\end{array}\]
+\[\begin{array}{l c}
+\text{\textsc{Yint}} & x-y=0 \\
+\text{\textsc{CSIsat}} & x-y \leq 0
+\end{array}\]
+Because our algorithm handles linear equalities special, our algorithm makes an effort to return an interpolant of linear equality if one exists.
+
+Additionally, as the follwing resuls of \textsc{Yint} shows, if a simple interpolant exists for a given problem, our algorithm tries such result.
+\[\begin{array}{r l}
+\varphi_A & ((-y+1 \leq 0) \wedge (-x \leq 0)) \vee ((-y \leq 0) \wedge (-x+1 \leq 0)) \\
+\hline
+\varphi_B & ((y+1<0) \wedge (x<0)) \vee ((y<0) \wedge (x+1<0))
+\end{array}\]
+\[\begin{array}{l c}
+\text{\textsc{Yint}} & -x \leq 0 \\
+\text{\textsc{CSIsat}} & (((x < 0 \vee -x \leq 0) \wedge 0 \leq x) \vee ((y < 0 \vee -y \leq 0) \wedge 0 \leq y))
+\end{array}\]
+
+\bigskip\centerline{}
+
+\[\begin{array}{r l}
+\varphi_A & (x-1=0) \vee (x+1=0) \\
+\hline
+\varphi_B & (x-2=0) \vee (x=0) \vee (x+2=0)
+\end{array}\]
+\[\begin{array}{l c}
+\text{\textsc{Yint}} & (-x-1=0) \vee (-x+1=0) \\
+\text{\textsc{CSIsat}} & ((-x \leq -1 \vee x \leq -1) \wedge (-x \leq 1 \vee - x \leq - 1) \wedge (x \leq - 1 \vee x \leq 1))
+\end{array}\]
+
+Finally, we showed some test cases that may occur in practical interpolation.
+
+\[\begin{array}{r l}
+\multirow{2}{*}{$\varphi_A$}
+& ((x-y-1 \leq 0) \wedge (-x+y-2 \leq 0) \wedge (x+y-1 \leq 0) \wedge (-x-y-2 \leq 0)) \vee \\
+& ((x-y-2 \leq 0) \wedge (-x+y-1 \leq 0) \wedge (x+y-2 \leq 0) \wedge (-x-y-1 \leq 0)) \\
+\hline
+\multirow{2}{*}{$\varphi_B$}
+& ((x-y+1<0) \wedge (-x+y-2 \leq 0) \wedge (x+y-2 \leq 0) \wedge (-x-y+1<0)) \vee \\
+& ((x-y-2 \leq 0) \wedge (-x+y+1<0) \wedge (x+y+1<0) \wedge (-x-y-2 \leq 0))
+\end{array}\]
+\[\begin{array}{l c}
+\text{\textsc{Yint}} & (-x+17y-33 \leq 0) \wedge (x-17y-33 \leq 0) \\
+\text{\textsc{CSIsat}} & ((-1 \leq (x + y) \vee (x + -y) \leq 1) \wedge (-1 \leq (x + -y) \vee (x + y) \leq 1))
+\end{array}\]
+
+\bigskip\centerline{}
+
+\[\begin{array}{r l}
+\multirow{2}{*}{$\varphi_A$}
+& ((y+1=0) \wedge (x-1=0)) \vee ((y+1=0) \wedge (x+1=0)) \vee \\
+& ((y-1=0) \wedge (x+1=0)) \vee ((y-1=0) \wedge (x-1=0)) \\
+\hline
+\multirow{2}{*}{$\varphi_B$}
+& ((y+2 \leq 0) \wedge (x-2 \geq 0)) \vee ((y+2 \leq 0) \wedge (x+2 \leq 0)) \vee \\
+& ((y-2 \geq 0) \wedge (x+2 \leq 0)) \vee ((y-2 \geq 0) \wedge (x-2 \geq 0))
+\end{array}\]
+\begin{tabular}{l c}
+\textsc{Yint} & $(-2y-3 \leq 0) \wedge (2y-3 \leq 0)$
+\end{tabular}
+
+\bigskip\centerline{}
+
+\textsc{CSIsat}
+\begin{align*}
+& ((-y \leq 1 \wedge y \leq -1) \vee (-y \leq -1 \wedge x \leq 1) \vee x \leq -1) \wedge \\
+& \qquad (-x \leq 1 \vee -x \leq -1 \vee -y \leq -1) \wedge \\
+& \qquad (-x \leq 1 \vee -x \leq -1 \vee y \leq -1) \wedge \\
+& \qquad (-x \leq -1 \vee x \leq -1)
+\end{align*}
+
+\[\begin{array}{r l}
+\varphi_A & ((y-1=0) \wedge (x-1=0)) \vee ((y=0) \wedge (x=0)) \\
+\hline
+\varphi_B & (((y-1>0) \vee (y-1<0)) \wedge (x-1=0)) \vee (((y>0) \vee (y<0)) \wedge (x=0))
+\end{array}\]
+\begin{tabular}{l c}
+\textsc{Yint} & $x-y=0$
+\end{tabular}
+
+\textsc{CSIsat}
+\begin{align*}
+& ((((y \leq 0 \wedge -y \leq 0) \vee 1 = x) \wedge 1 \neq x) \vee ((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 \neq x) \wedge \\
+& \qquad (((0 \neq x \vee y \leq 1) \wedge (0 = x \vee y \leq 1)) \vee 1 \neq x) \wedge \\
+& \qquad (((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 = x) \wedge \\
+& \qquad ((1 \neq x \wedge 1 = x) \vee (y \leq 0 \wedge -y \leq 0) \vee 0 \neq x) \wedge \\
+& \qquad (1 = x \vee y \leq 0)
+\end{align*}

### 2013-08-28 article/horn.tex
Experiment

@@
   weakest constraint in the sense that the resulting solution would
-  not be simple yet.  The process iteratively strengthen a constraint
+  not be simple yet.  The process iteratively strengthens a constraint
   until the strongest one while preserving the constraint
@@
 \item \textbf{Problem Restructuring} is a process to restructure the
-  problem to obtain a less simpler solution, if the constraint become
+  problem to obtain a less simple solution, if the constraint become
   unsatisfiable during the iterative constraint generation.  This
   process considers the unsatisfiable core of the constraint, and
-  wisely enlarge the solution size.
+  wisely enlarges the solution size.
 \end{itemize}
@@
 constraint that parameters in templates must satisfy.  This is done in
-a constructive manner, and constraints are build in the order of
+a constructive manner, and constraints are built in the order of
 implication dependency in Horn clauses.  That is, when an input
-problem has a Horn clause thath has a predicate variable $P$ on its
+problem has a Horn clause that has a predicate variable $P$ on its
 left-hand and $Q$ on its right-hand, a constraint to describe
@@
 template.  When all parameters in the whole problem are described, the
-built constraint are evaluated to obtain a model.  Its instanciation
+built constraint is evaluated to obtain a model.  Its instantiation
 becomes an appropriate solution for the input problem.
@@
 unsatisfiable during this step, the algorithm analyzes the
-unsatisfiable core and transform the problem to obtain a less simple
+unsatisfiable core and transforms the problem to obtain a less simple
 solution.
@@
 
-Sinve we have $A$'s appearance twice on the second clause, it is
+Since we have $A$'s appearance twice on the second clause, it is
 possible for $A$ to have two expressions each of which is used for
@@
 Because the unsatisfiable core contains constraints originating from
-the disjunctions to be expressed in a single term template.  This let
-us determine to relax the template to allow a enlarged solution.
+the disjunctions to be expressed in a single term template, we can
+determine which predicate to relax its template to allow an enlarged solution.
 
@@
 \text{Constraint:} \\
-TODO \\
+& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
+& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \lambda_3 \geq 0 \wedge \\
+& \lambda_3 = c_{3,2} \wedge \lambda_3 \geq c_{4,2} \wedge \lambda_3 \geq 0 \wedge \\
+& - \lambda_4 = c_{3,1} \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
+& - \lambda_4 = c_{3,2} \wedge \lambda_4 \geq c_{4,2} \wedge \lambda_4 \geq 0 \wedge \\
+& c_1 + c_{3,1} = 0 \wedge c_2 + c_{4,1} > 0 \wedge \\
+& c_1 + c_{3,2} = 0 \wedge c_2 + c_{4,2} > 0
 \\
 \text{Unsatisfiable core:} \\
-TODO
+0 \geq c_2 \\
+& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \\
+& - \lambda_4 = c_{3,2} \wedge \\
+& c_1 + c_{3,1} = 0 \wedge c_2 + c_{4,1} > 0 \wedge \\
+& c_1 + c_{3,2} = 0
 \end{align*}
@@
 originate from
-
-TODO
 
@@
 
-Finally, we obtain the following constraint:
+After this, we obtain the following constraint:
 
-TODO
+\begin{align*}
+& \lambda_1 - \lambda_2 = c_{1,1} \wedge 0 \geq c_{2,1} \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
+& \lambda_1 - \lambda_2 = c_{1,2} \wedge 0 \geq c_{2,2} \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
+& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \lambda_3 \geq 0 \wedge \\
+& \lambda_3 = c_{3,2} \wedge \lambda_3 \geq c_{4,2} \wedge \lambda_3 \geq 0 \wedge \\
+& - \lambda_4 = c_{3,1} \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
+& - \lambda_4 = c_{3,2} \wedge \lambda_4 \geq c_{4,2} \wedge \lambda_4 \geq 0 \wedge \\
+& c_{1,1} + c_{1,2} + c_{3,1} = 0 \wedge c_{2,1} + c_{2,2} + c_{4,1} > 0 \wedge \\
+& c_{1,1} + c_{1,2} + c_{3,2} = 0 \wedge c_{2,1} + c_{2,2} + c_{4,2} > 0
+\end{align*}
 
-The process also generates a predicate template for every predicate
-variable that the solution will become.  under certain solution
-template.
+By continuing this process, the process in the end generates a
+predicate template that the solution will become.  Eventually, we
+obtain the following solution.
+
+\begin{align*}
+A(x) \text{ : } & x \leq 0 \wedge -x \leq 0 \\
+B(x) \text{ : } & x+1 \leq 0 \vee -x+1 \leq 0
+\end{align*}
 
@@
 The graph is then used throughout the algorithm to represent the
-problem.  A Horn graph is a labelled directed acyclic graph
+problem.  A Horn graph is a labeled directed acyclic graph
 $G=(V,E,\varphi)$.
@@
   vertices $V_T$ and arrow vertices $V_\Rightarrow$.  Each vertex
-  $u \in V_\Rightarrow$ has eactly one outgoing edge to some
+  $u \in V_\Rightarrow$ has exactly one outgoing edge to some
   $v \in V_T$, denoted as $succ(u)$.
-\item $E$ is a set of labelled edges, which are expressed in the form
+\item $E$ is a set of labeled edges, which are expressed in the form
   $(u,v,\theta)$ where $\theta$ is a finite map from variables to
   variables.
-\item $\varphi: V_T \rightarrow L$ is a labelling function for Horn
+\item $\varphi: V_T \rightarrow L$ is a labeling function for Horn
   term vertices, where L is a set of elements of a predicate variable
@@
 A variable mapping label $\theta$ on edges is used to map the index of
-one predicate variable to another. (TODO?)
-
-(Example?)
+one predicate variable to another.
 
@@
 For the first case previously mentioned, the unsatisfiable core
-$\mathcal{U}$ contains coeffcient constraint expressions with a
+$\mathcal{U}$ contains coefficient constraint expressions with a
 coefficient parameter of certain predicate variable.  If the predicate
@@
 to induce different linear expressions for $P_1, \ldots, P_n$, it may
-be necessary for preceeding predicate variable vertices in $G$ to be
+be necessary for preceding predicate variable vertices in $G$ to be
 duplicated in the same manner as $P$. It is because that different
@@
 After this duplication of the predicate vertex $P$, unlike the
-conjunction split, this results in producing a disjunction at
-succeeding predicate variables vertices ($R$ in this case).
+conjunction split, a disjunction at succeeding predicate
+variable vertices ($R$ in this case) is produced.
 \begin{align*}
@@
 
-Conside example below. $e_i$ represents some conjunctions of linear
+Consider example below. $e_i$ represents some conjunctions of linear
 expressions.
@@
 time testing the satisfiability of the constraint.  When the
-constraint bebcome unsatisfiable at the certain iteration, the
+constraint becomes unsatisfiable at the certain iteration, the
 algorithm can determine a predicate vertex which has to be split in a

### 2013-08-28 article/conclusion.tex
Spell check

@@
 different interpolating problems. Additionally, we mentioned the
-possibility of chosing small solutions in interpolating general
+possibility of choosing small solutions in interpolating general
 formulas by wisely choose the order to combine set of solutions during

### 2013-08-28 article/experiment.tex
Spell check

--- 2013-08-28 article/experiment.tex
@@
 In our experiment, we compared the performance of our algorithm with
-CSIsat, that is mainly used as a subprocedure of MoCHi.  Note that
+CSIsat, that is mainly used as a sub procedure of MoCHi.  Note that
 CSIsat may return different interpolants for an interpolating problem

### 2013-08-28 article/future.tex
Formatting for final
Spell check

@@
 improvement.  Based on the experiment result from the previous
-chapter, here we give raise some analaysis and discussion for them.
+chapter, here we give raise some analysis and discussion for them.
 
@@
 
-In the preliminary expriment before incorporating into MoCHi, although
+In the preliminary experiment before incorporating into MoCHi, although
 we have tried the quantifier elimination over integer theory, the
@@
 algorithm first treats all linear inequalities over integer space,
-more study on the gap betweem these two are required when we consider
+more study on the gap between these two are required when we consider
 the completeness.
@@
 from the previous constraint and the SMT query result.  Our Horn
-clause solver make use of Z3 to obtain an unsatisfiable core from the
+clause solver makes use of Z3 to obtain an unsatisfiable core from the
 linear constraints.  However the extraction of an unsatisfiable core
@@
 is necessary to reduce the number of them for reduce the computation
-cost.  For instance, the optimization is possible by appropriatealy
+cost.  For instance, the optimization is possible by appropriately
 updating $V_\star$ in $\solveHorn$~procedure since the vertices that
@@
 the choice of them.  Thus, we need further study to optimize the order
-of choice of multiple unsatisfiable cores if ones exists.
+of choice of multiple unsatisfiable cores if ones exist.
 
@@
 $P$.  However at this point, $e_3$ and $e_4$ are free to be connected
-to any duplication of $P$.  Alghough our current algorithm split
+to any duplication of $P$.  Although our current algorithm split
 $\left\lbrace e_1, e_2, e_3, e_4 \right\rbrace$ into four duplications
@@
 it is possible to split into $\left\lbrace e_1, e_3\right\rbrace$ and
-$\left\lbrace e_2, e_4 \right\rbrace$ for example.  By intellilgently
+$\left\lbrace e_2, e_4 \right\rbrace$ for example.  By intelligently
 choose this, we may be able to obtain much simpler solution than the
@@
 of Horn graph as a label and solve a coloring problem on it, the
-algorithm can wisely determine the minuminm number of duplications.
+algorithm can wisely determine the minimum number of duplications.
 
@@
 programs, Horn clauses can describe the conditions of data exchange
-between threads, and confirm the validity of semaphores for mutural
+between threads, and confirm the validity of semaphores for mutual
 exclusion.  We expect our algorithm achieve better performance in
@@
 
-By puting such effort for further improvement on our algorithm, we can
-conclude that the algorithm may be able to achive better performance
+By putting such effort for further improvement on our algorithm, we can
+conclude that the algorithm may be able to achieve better performance
 in Horn clause solving, and become a necessary tool for program

### 2013-08-28 article/horn.tex
Spell check

@@
 \text{Constraint:} \\
-& \lambda_1 - \lambda_2 = c_1 \wedge 0 \geq c_2 \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
-& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \lambda_3 \geq 0 \wedge \\
-& \lambda_3 = c_{3,2} \wedge \lambda_3 \geq c_{4,2} \wedge \lambda_3 \geq 0 \wedge \\
-& - \lambda_4 = c_{3,1} \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
-& - \lambda_4 = c_{3,2} \wedge \lambda_4 \geq c_{4,2} \wedge \lambda_4 \geq 0 \wedge \\
-& c_1 + c_{3,1} = 0 \wedge c_2 + c_{4,1} > 0 \wedge \\
-& c_1 + c_{3,2} = 0 \wedge c_2 + c_{4,2} > 0
+TODO \\
 \\
 \text{Unsatisfiable core:} \\
-0 \geq c_2 \\
-& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \\
-& - \lambda_4 = c_{3,2} \wedge \\
-& c_1 + c_{3,1} = 0 \wedge c_2 + c_{4,1} > 0 \wedge \\
-& c_1 + c_{3,2} = 0
+TODO
 \end{align*}
@@
 originate from
+
+TODO
 
@@
 
-After this, we obtain the following constraint:
+Finally, we obtain the following constraint:
 
-\begin{align*}
-& \lambda_1 - \lambda_2 = c_{1,1} \wedge 0 \geq c_{2,1} \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
-& \lambda_1 - \lambda_2 = c_{1,2} \wedge 0 \geq c_{2,2} \wedge \lambda_1 \geq 0 \wedge \lambda_2 \geq 0 \wedge \\
-& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \lambda_3 \geq 0 \wedge \\
-& \lambda_3 = c_{3,2} \wedge \lambda_3 \geq c_{4,2} \wedge \lambda_3 \geq 0 \wedge \\
-& - \lambda_4 = c_{3,1} \wedge \lambda_4 \geq c_4 \wedge \lambda_4 \geq 0 \wedge \\
-& - \lambda_4 = c_{3,2} \wedge \lambda_4 \geq c_{4,2} \wedge \lambda_4 \geq 0 \wedge \\
-& c_{1,1} + c_{1,2} + c_{3,1} = 0 \wedge c_{2,1} + c_{2,2} + c_{4,1} > 0 \wedge \\
-& c_{1,1} + c_{1,2} + c_{3,2} = 0 \wedge c_{2,1} + c_{2,2} + c_{4,2} > 0
-\end{align*}
+TODO
 
-By continuing this process, the process in the end generates a
-predicate template that the solution will become.  Eventually, we
-obtain the following solution.
-
-\begin{align*}
-A(x) \text{ : } & x \leq 0 \wedge -x \leq 0 \\
-B(x) \text{ : } & x+1 \leq 0 \vee -x+1 \leq 0
-\end{align*}
+The process also generates a predicate template for every predicate
+variable that the solution will become.
 
@@
 A variable mapping label $\theta$ on edges is used to map the index of
-one predicate variable to another.
+one predicate variable to another. (TODO?)
+
+(Example?)
 

### 2013-08-28 article/interpolation.tex
Spell check

@@
 In computing the interpolant, we make a linear constraint of a
-interpolant by applying Farkas's lemma.  First, we assing weight
+interpolant by applying Farkas's lemma.  First, we assign weight
 parameters to every expression in conjunction form.
@@
 reasonable computation time for any input.  It also aims to obtain a
-common solution among multiple interpolating problems.  When the
-algorithm is used as a sub-routine for solving Horn clause satisfying
-problems that are made while discovering predicates for abstraction
+common solution among multiple interpolating problems.  These goals
+are set because the algorithm is designed to be used as a sub-routine
+for solving Horn clause satisfying problems
+that are made while discovering predicates for abstraction
 during the program verification.  These advantages of our algorithm
@@
 problems. To accomplish these aims, the algorithm preserves a set of
-interpolants during the computation, and define and execute operations
+interpolants during the computation, and executes operations
 over interpolant sets.
@@
 above will let an expression $I$ to be an interpolant.  Considering
-this, it is possible to obtain different interpolants for a
+this, it is possible to obtain different interpolants for an
 interpolating problem by computing different assignments to

### 2013-08-28 article/related.tex
Other chapters also separated
Spell check

@@
+\chapter{Related work}
+\label{chap:related}
+
 Interpolating theorem proving have played an important role in program
@@
 An interpolating prover \textsc{FOCI} \cite{website/foci} was
-developed under the comprehensive wrok on quantifier-free interpolant
+developed under the comprehensive work on quantifier-free interpolant
 generation over the theory of linear arithmetic and uninterpreted
@@
 over it.  Here, the control flow graph may contain loops, it is not
-suffcieint by the convential interpolation, and rather it required a
+sufficient by the conventional interpolation, and rather it required a
 disjunctive reasoning.
@@
 As another direction of extension, some works are done for solving
-Horn clauses with qunatified consequences \cite{conf/sas/BjornerMR13,
+Horn clauses with quantified consequences \cite{conf/sas/BjornerMR13,
 conf/cav/BeyenePR13}.  They are used to abstract more complex systems

### 2013-08-28 article/experiment.tex
Small fixes

--- 2013-08-28 article/experiment.tex
@@
 \end{array}\]
-Because our algorithm handles linear equalities special, our algorithm makes an effort to return an interpolant of linear equality if one exists.
+Because our algorithm handles linear equalities special, our algorithm
+makes an effort to return an interpolant of linear equality if one
+exists.
 
-Additionally, as the follwing resuls of \textsc{Yint} shows, if a simple interpolant exists for a given problem, our algorithm tries such result.
+Additionally, as the follwing results of \textsc{Yint} shows, if a
+simple interpolant exists for a given problem, our algorithm tries
+such result.
 \[\begin{array}{r l}
@@
 \[\begin{array}{r l}
-\varphi_A & ((y-1=0) \wedge (x-1=0)) \vee ((y=0) \wedge (x=0)) \\
+\varphi_A & (x=1 \wedge y=1) \vee (x=0 \wedge y=0) \\
 \hline
-\varphi_B & (((y-1>0) \vee (y-1<0)) \wedge (x-1=0)) \vee (((y>0) \vee (y<0)) \wedge (x=0))
+\varphi_B & (x=1 \wedge y \neq 1) \vee (x=0 \wedge y \neq 0)
 \end{array}\]
@@
 \end{align*}
+
+\section{Horn clause solving}
+
+As an Horn clause solver we incorporated our Horn clause solver to MoCHi.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=.8]{figures/chart.png}
+\end{center}
+\caption{Result obtained from MoCHi with Horn clause solver CSIsat and
+Yhorn}
+\end{figure}
+
+We obtained better result in some cases compared to the one we had in CSIsat.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=.5]{figures/time.eps}
+\end{center}
+\caption{Time consumption ration of Yhorn compared to CSIsat}
+\end{figure}
+
+However we take a longer time for solving the same Horn clauses, which
+means further optimization is required for our algorithm.

### 2013-08-28 article/interpolation.tex
Small fixes

@@
 CSIsat\cite{conf/cav/BeyerZM08} returns
-\[ TODO \]
+\begin{align*}
+& ((((y \leq 0 \wedge -y \leq 0) \vee 1 = x) \wedge 1 \neq x) \vee ((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 \neq x) \wedge \\
+& \qquad (((0 \neq x \vee y \leq 1) \wedge (0 = x \vee y \leq 1)) \vee 1 \neq x) \wedge \\
+& \qquad (((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 = x) \wedge \\
+& \qquad ((1 \neq x \wedge 1 = x) \vee (y \leq 0 \wedge -y \leq 0) \vee 0 \neq x) \wedge \\
+& \qquad (1 = x \vee y \leq 0)
+\end{align*}
 as a solution.
@@
 \paragraph{Farkas's Lemma on linear inequalities}
-Let a linear inequality $e_i$ be represented as $a_i1 x_1 + \cdots +
-a_im x_m <= a_i0$.  Assuming that $e_1,\cdots,e_n$ implies $e_0$,
-there exists $\lambda_1,\cdots,\lambda_n$ that satisfy $a_0j
-=\sum_(i=1)^n \lambda_i * a_ij (j=0, \ldots, m)$.
+Let a linear inequality $e_i$ be represented as
+$a_{i,1} x_1 + \cdots + a_{i,m} x_m <= a_{i,0}$.  Assuming that
+$e_1,\cdots,e_n$ implies $e_0$, there exists
+$\lambda_1,\cdots,\lambda_n$ that satisfy
+$a_{0,j} = \sum_(i=1)^n \lambda_i * a_{i,j} \quad (j=0, \ldots, m)$.
 
@@
 
-Assuming that $e_1,\cdots,e_n$ implies $\bot$, there exists
-$\lambda_1,\cdots,\lambda_n$ that satisfy $???? TODO $.
+Let a linear inequality $e_i$ be represented as
+$a_{i,1} x_1 + \cdots + a_{i,m} x_m <= a_{i,0}$.  Assuming that
+$e_1,\cdots,e_n$ implies $\bot$, there exists
+$\lambda_1,\cdots,\lambda_n$ that satisfy
+$0 = \sum_(i=1)^n \lambda_i * a_{i,j} \quad (j=1, \ldots, m) \wedge
+0 > \sum_(i=1)^n \lambda_i * a_{i,0}$.
 
@@
 \left\lbrace e_1,\ldots,e_m \right\rbrace & \vdash e_\star \\
-e_\star & \nvdash \left\lbrace e_{m+1},\ldots,e_{m+n} \right\rbrace
+e_\star \wedge \left\lbrace e_{m+1},\ldots,e_{m+n} \right\rbrace & \vdash \bot
 \end{align*}
@@
 \begin{align*}
-\forall I \in \mathcal{I}. & A_1 \vdash I \wedge A_2 \vdash I \wedge I \nvdash B \wedge \\
+\forall I \in \mathcal{I}. & A_1 \vdash I \wedge A_2 \vdash I \wedge I \wedge B \vdash \bot \wedge \\
 & FV(I) = FV(A_1) \cup FV(A_2) \cup FV(B)

### 2013-09-17 article/interpolation.tex
Minor fix

@@
 $\lambda_1,\cdots,\lambda_n$ that satisfy
-$a_{0,j} = \sum_(i=1)^n \lambda_i * a_{i,j} \quad (j=0, \ldots, m)$.
+\[a_{0,j} = \sum_{i=1}^n \lambda_i * a_{i,j} \quad (j=0, \ldots,m)\]
 
@@
 $\lambda_1,\cdots,\lambda_n$ that satisfy
-$0 = \sum_(i=1)^n \lambda_i * a_{i,j} \quad (j=1, \ldots, m) \wedge
-0 > \sum_(i=1)^n \lambda_i * a_{i,0}$.
+\[0 = \sum_{i=1}^n \lambda_i a_{i,j} \quad (j=1, \ldots, m) \wedge
+0 > \sum_{i=1}^n \lambda_i a_{i,0}\]
 
@@
 \right\rbrace$ and $B = \left\lbrace e_{m+1},\ldots,e_{m+n}
-\right\rbrace$.  We assume that conjunctions of $A$ and $B$ are
-inconsistent.  For later convenience, we call this interpolating
-problem $\left( A, B \right)$.  Then there exists an interpolating
-linear expression $e_\star$ which satisfies
+\right\rbrace$, where $e_i$ takes a form $\mathbf{a}_i \mathbf{x} + b
+\leq 0$.  We assume that conjunctions of $A$ and $B$ are inconsistent.
+For later convenience, we call this interpolating problem $\left( A, B
+\right)$.  Then there exists an interpolating linear expression
+$e_\star$ which satisfies
 \begin{align*}
@@
 
-According to the Farkas's lemma, there exists a set of assignmentqs to
-$\lambda_i$, to conclude $\bot$ from a set of linear expressions $A
+According to the Farkas's lemma, there exists a set of assignments to
+$\lambda_i$ to conclude $\bot$ from a set of linear expressions $A
 \cup B$, namely,
 \begin{align*}
-\exists \lambda_1, \ldots, \lambda_m. \\
-a % TODO: coefficient, constant constraints
+& \exists \lambda_1, \ldots, \lambda_m. \\
+& \sum_{i=1}^{m+n} \lambda_i \mathbf{a}_i = 0\wedge \sum_{i=1}^{m+n} \lambda_i b_i > 0
 \end{align*}

### 2013-09-25 article/intro.tex
Introduction amendment

@@
 with model checking techniques, an automated theorem prover is used to
-abstract an infinite set of program states into a finite one so that
-the model checker can virtually explore all possible program execution
+abstract a program with infinite states into a finite model so that a
+model checker can virtually explore all possible program execution
 paths.
@@
 An automated theorem prover finds appropriate predicates for every
-program location.  An execution states at those locations is then
-expressed by a Boolean valuation of the given predicates.  If the
-abstraction is too coarse and a model checker discovers an infeasible
-error execution path in an abstract system, the prover computes an
+program location.  Execution states at those locations are expressed
+by a Boolean valuation of the given predicates.  If the abstraction is
+too coarse and a model checker discovers an infeasible error execution
+path, also called a spurious counterexample, the prover computes an
 additional predicate at those program locations to refine the
-abstraction predicates.
+abstraction.
 
-In order to obtain an additional abstraction predicate at program
-locations along the infeasible counterexample, one class of automated
+In order to obtain an additional abstraction predicate at certain
+program location along the infeasible path, one class of automated
 theorem provers, called an interpolating theorem prover
 \cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07}, computes a
-separating predicate called an interpolant between two sets of
-constraints; the one consists of the constraints from the program
-entry point to the location, and the other consists of the constraints
-from the location to the program failure.  Those constraints include
-variable assignments and assertions to be satisfied originated from
-conditional branches.
+logical separation between two sets of formula; the one consists of
+constraints from the program entry point to the location, and the
+other consists of the constraints from the location to the program
+failure.  Those formulas are built from variable assignments and
+assertions to be satisfied originated from conditional branches.
 
-The computed predicate by an interpolating theorem prover is satisfied
-at the location under the execution in the concrete model along the
-discovered counterexample, and any execution paths which satisfies the
-predicate does not lead to the failure state.  After the refinement,
-the infeasible path is ruled out and no longer found by the model
-checker.
+The computed logical separator is satisfied at the location under the
+execution in the program along the discovered path, and any execution
+paths which satisfies the predicate does not lead to the failure
+state.  After the refinement, the counterexample is ruled out and the
+same infeasible path is no longer found by the model checker.
 
@@
 verification because a complex predicate may not be sufficient to rule
-out similar spurious paths in programs with loops and recursion calls.
-This causes the model checker to discover spurious paths infinitely
-which pass the same program location by loops or recursion calls at
-different numbers of iteration, and the verification process does not
-to terminate.
+out resembling spurious paths in programs.  This causes the
+verification process not to terminate because the model checker
+infinitely discovers infeasible paths which pass the same program
+location by loops or recursion calls on the different number of
+iterations.
 
-To avoid this, it is desired to discover an invariant formula for the
-predicate abstraction.  Based on the hypothesis that program's
-invariants tend to be simple formulas, we propose a new interpolating
-algorithm which tries to minimize the number of conjunctions and
-disjunctions in a solution.  The predicate obtained by this algorithm
-may become similar to the invariant, and the number of trials in model
-checking may be reduced.
+Ideally it is desired to discover an invariant formula for the
+predicate abstraction.  We put a hypothesis that program invariants
+tend to be simple formulas, and we propose a new interpolating
+algorithm based on the hypothesis.  The algorithm tries to minimize
+the number of conjunctions and disjunctions in a solution.  The
+predicate obtained by this algorithm may become similar to the
+invariant. We expect the number of iterations for model checking to
+decrease as a result.
 
-Additionally, we extend our interpolating algorithm to solve a
-symmetric interpolation problem, in which the algorithm computes
-predicates at multiple program locations along a path at once.  This
-enables the abstraction predicates at multiple locations to be updated
-at the same computation.
+We also focus on solving constraint satisfaction problems over Horn
+clauses with unknown predicate variables.  As well as interpolation,
+Horn clause solving is another technique often used in program
+verification for capturing program behaviors.  By computing the
+solution for Horn clause problems, we obtain predicates that are true
+at specific program locations along specific paths.  Furthermore it
+may be possible to obtain predicates which are highly likely to be
+program invariants by computing same predicates inside loops and
+recursion calls on different iterations, or at the same location over
+multiple paths.  By the previous hypothesis, we propose a Horn clause
+solving algorithm which constructs as simple a solution as possible.
 
-Finally, we extend our algorithm to solve Horn clause solving
-problems.  It allows us to obtain predicates which are highly likely
-to be program invariants, by computing the same abstraction predicates
-inside loops and recursion calls on different iterations, and by
-unifying predicates at the same location over multiple paths.  The
-problems are generated by encoding constraints over execution paths
-with unknown predicates which describe functions' constraints between
-parameters and return values.
+TODO: Mention experiments
 
-The rest of the paper is structured as follows.
-Chapter~\ref{chap:interpolation} proposes a new interpolating
-algorithm to obtain simple solutions, and extends it to solve
-symmetric interpolation problems.  Chapter~\ref{chap:horn} describes a
-new Horn clause solving algorithm by extending the previous
+The rest of this thesis is structured as follows.
+Chapter~\ref{chap:interpolation} shows detailed usage of interpolation
+problems in program verification and proposes a new interpolating
+algorithm to obtain simple solutions.  Chapter~\ref{chap:horn}
+describes a Horn clause solving algorithm by extending the previous
 interpolating algorithm.  Chapter~\ref{chap:experiment} describes
 experiments and illustrates the result to confirm the effect of our
-proposed algorithm.  Chapter~\ref{chap:future} gives a discussion of
-our improvements and proposes further improvements to make.
-Chapter~\ref{chap:related} mentions related work done by others.
-Finally, Chapter~\ref{chap:conclusion} reviews the impact of our
-research and brings the conclusion.
+algorithms.  Chapter~\ref{chap:future} gives a discussion on further
+possibilities for improvements. Chapter~\ref{chap:related} mentions
+related work done by others.  Finally, Chapter~\ref{chap:conclusion}
+reviews the impact of our research and brings the conclusion.

### 2013-09-25 article/interpolation.tex
Stash

@@
 
-Among various program verification techniques, when software model
-checking with predicate abstraction is adopted, the interpolation is
-used to compute the abstraction predicate along a spurious
-counterexample that a model checker discovered.  It is done by
-computing a Craig interpolant between the strongest postcondition at
-the location and the weakest precondition to the failure point at
-every program location.
+Among various program verification techniques, the interpolation has
+been used to extract a program behavior as a logical formula.
+Especially when one adopts software model checking technique with
+predicate abstraction, the interpolation computes the abstraction
+predicate along a spurious counterexample that a model checker
+discovered.
+
+It is done by computing a logical separation between the strongest
+postcondition at the location and the weakest precondition to the
+failure point at every program location.
 

### 2013-09-25 article/experiment.tex
Minor fix
Amendment

@@
 We have incorporated our algorithm to the software model checker for
-higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It is
+higher-order programs \textsc{MoCHi} \cite{conf/pldi/KobayashiSU11}.  It is
 specifically designed to verify the correctness of higher-order
@@
 
-We implemented our proposed algorithm into MoCHi by two steps.  First,
+We implemented our proposed algorithm into \textsc{MoCHi} by two steps.  First,
 we replace an existing theorem prover with our interpolating theorem
@@
 
-There are some interpolating theorem provers known.
-\begin{itemize}
-\item \textsc{CSIsat} \cite{conf/cav/BeyerZM08} combines linear
-  arithmetic reasoning and SAT-based logical reasoning.
-\item \textsc{FOCI} \cite{website/foci} extracts an interpolant by the
-  proof-based method.
-\item \textsc{CLP-Prover} \cite{website/clp} generates an interpolant
- from constraints by linear arithmetic reasoning. It only supports
- conjunctive problems.
-\end{itemize}
-
-In our experiment, we compared the performance of our algorithm with
-CSIsat, that is mainly used as a sub procedure of MoCHi.  Note that
-CSIsat may return different interpolants for an interpolating problem
-because of the randomized algorithm.
-
-\paragraph{Comparison over simple interpolation}
-First, we compare by simply generating an interpolant without
-incorporating into MoCHi.  \textsc{Yint} represents our interpolating
-theorem prover.
+As an experiment of our interpolating algorithm, we compared the
+quality of resulting interpolants with
+ones that \testsc{CSIsat} \cite{conf/cav/BeyerZM08} generates.
+It combines linear arithmetic reasoning and SAT-based logical reasoning.
+\textsc{CSIsat} is mainly used as a sub procedure of current \textsc{MoCHi}.  Note that
+it may return different interpolants for an interpolating problem
+because of the randomized algorithm.  Results shown below
+are ones that are randomly chosen.
+\textsc{Yint} represents our interpolating algorithm.
 
@@
 \item Disjunctive interpolation in $A$.
-\[\begin{array}{r l}
-\varphi_A & (x-1 \geq 0) \vee (x+1 \leq 0) \\
+\[\begin{array}{r c}
+\varphi_A : & (x-1 \geq 0) \vee (x+1 \leq 0) \\
 \hline
-\varphi_B & x=0
+\varphi_B : & x=0
 \end{array}\]
@@
 \item Disjunctive interpolation in $B$.
-\[\begin{array}{r l}
-\varphi_A & x=0 \\
+\[\begin{array}{r c}
+\varphi_A : & x=0 \\
 \hline
-\varphi_B & (x-1 \geq 0) \vee (x+1 \leq 0)
+\varphi_B : & (x-1 \geq 0) \vee (x+1 \leq 0)
 \end{array}\]
@@
 % complicated disj
-\[\begin{array}{r l}
-\varphi_A & ((x-5<0) \wedge (-x<0)) \vee ((x<0) \wedge (-x-5<0)) \\
+\[\begin{array}{r c}
+\varphi_A : & ((x-5<0) \wedge (-x<0)) \vee ((x<0) \wedge (-x-5<0)) \\
 \hline
-\varphi_B & x=0
+\varphi_B : & x=0
 \end{array}\]
@@
 \end{itemize}
-For those problems that are simple enough, Yint and CSIsat returns the
+For those problems that are simple enough, \textsc{Yint} and \textsc{CSIsat} returns the
 same results.
@@
 % equality
-\[\begin{array}{r l}
-\varphi_A & x=0 \\
+\[\begin{array}{r c}
+\varphi_A : & x=0 \\
 \hline
-\varphi_B & x-1=0
+\varphi_B : & x-1=0
 \end{array}\]
@@
 % quantifier
-\[\begin{array}{r l}
-\varphi_A & (y-z=0) \wedge (x-z=0) \\
+\[\begin{array}{r c}
+\varphi_A : & (y-z=0) \wedge (x-z=0) \\
 \hline
-\varphi_B & (-w+y+1 \leq 0) \wedge (-w+x \geq 0)
+\varphi_B : & (-w+y+1 \leq 0) \wedge (-w+x \geq 0)
 \end{array}\]
@@
 such result.
-\[\begin{array}{r l}
-\varphi_A & ((-y+1 \leq 0) \wedge (-x \leq 0)) \vee ((-y \leq 0) \wedge (-x+1 \leq 0)) \\
+\[\begin{array}{r c}
+\varphi_A : & ((-y+1 \leq 0) \wedge (-x \leq 0)) \vee ((-y \leq 0) \wedge (-x+1 \leq 0)) \\
 \hline
-\varphi_B & ((y+1<0) \wedge (x<0)) \vee ((y<0) \wedge (x+1<0))
+\varphi_B : & ((y+1<0) \wedge (x<0)) \vee ((y<0) \wedge (x+1<0))
 \end{array}\]
@@
 
-\[\begin{array}{r l}
-\varphi_A & (x-1=0) \vee (x+1=0) \\
+\[\begin{array}{r c}
+\varphi_A : & (x-1=0) \vee (x+1=0) \\
 \hline
-\varphi_B & (x-2=0) \vee (x=0) \vee (x+2=0)
+\varphi_B : & (x-2=0) \vee (x=0) \vee (x+2=0)
 \end{array}\]
@@
 
-\[\begin{array}{r l}
+\[\begin{array}{r c}
 \multirow{2}{*}{$\varphi_A$}
@@
 \hline
-\multirow{2}{*}{$\varphi_B$}
-& ((x-y+1<0) \wedge (-x+y-2 \leq 0) \wedge (x+y-2 \leq 0) \wedge (-x-y+1<0)) \vee \\
-& ((x-y-2 \leq 0) \wedge (-x+y+1<0) \wedge (x+y+1<0) \wedge (-x-y-2 \leq 0))
+\varphi_B : & (x=0 \wedge y=2) \vee (x=0 \wedge y=-2)
 \end{array}\]
@@
 
-\[\begin{array}{r l}
+\[\begin{array}{r c}
 \varphi_A & (x=1 \wedge y=1) \vee (x=0 \wedge y=0) \\
@@
 \begin{center}
-\includegraphics[scale=.8]{figures/chart.png}
+\includegraphics[scale=.6]{figures/chart.png}
 \end{center}
-\caption{Result obtained from MoCHi with Horn clause solver CSIsat and
-Yhorn}
+\caption{Result obtained from \textsc{MoCHi} with Horn clause solver \textsc{CSIsat} and
+\textsc{Yhorn}}
 \end{figure}
 
-We obtained better result in some cases compared to the one we had in CSIsat.
+We obtained better result in some cases compared to the one we had in \textsc{CSIsat}.
 
@@
 \begin{center}
-\includegraphics[scale=.5]{figures/time.eps}
+\includegraphics[scale=.8]{figures/time.eps}
 \end{center}
-\caption{Time consumption ration of Yhorn compared to CSIsat}
+\caption{Time consumption ration of Yhorn compared to \textsc{CSIsat}}
 \end{figure}

### 2013-09-25 article/future.tex
Small fixes
Amendment

@@
 
-Our proposed algorithms have still numbers of rooms for further
+Our proposed algorithms have still a number of rooms for further
 improvement.  Based on the experiment result from the previous
 chapter, here we give raise some analysis and discussion for them.
-
-\paragraph{Space explosion for constraint duplication}
-Our experimental implementation did not simplify the linear constraint
-for every predicate variable and it caused a linear constraint
-unsolvable due to the shortage of space resources by exponential
-growth of constraints.  As we described in the algorithm explanation,
-it may be possible to perform quantifier elimination for reduce the
-size of linear constraints.  On the other hand without the
-optimization, the constraint duplication may occur every time a
-constraint is generated for the predicate vertices with multiple
-successors.
-
-We think that the algorithm with such optimization would not cause a
-memory shortage because constraints gently grow according to roughly
-the number of vertices.  Although the worst case can be produced by
-causing a conjunction split frequently, we consider that it is not the
-common case in program verification because of our hypothesis.
-
-In practice, the size of constraint may be limited by simplification,
-and additionally, normally the algorithm does not need to preserve the
-constraint until the time to build a concrete solution in the end.
-
-\paragraph{Quantifier elimination}
-It may be useful to prepare a customized algorithm for performing
-quantifier elimination for linear expressions because linear
-constraints that are generated from our algorithm have certain pattern
-in their format.
-
-In the preliminary experiment before incorporating into MoCHi, although
-we have tried the quantifier elimination over integer theory, the
-result was unsatisfactory in the sense that simplified formulas had
-disjunctions.  It caused the loss of efficiency in solving constraints
-and generating unsatisfiable cores.
-
-We have also tried the quantifier elimination over real space, and the
-result was better than the integer version.  However, because our
-algorithm first treats all linear inequalities over integer space,
-more study on the gap between these two are required when we consider
-the completeness.
 

### 2013-09-25 article/interpolation.tex
Amendment

@@
 This expression contains coefficient parameters $a_1, \ldots, a_n$ and
-a constant parameter $b$.  We treat a special expressions $\top$ as $0
+a constant parameter $b$.  We treat special expressions $\top$ as $0
 \leq 0$ and $\bot$ as $1 \leq 0$.  The algorithm input and output is a
@@
 In the simple interpolation, the algorithm receives two formulas in
-$\psi$ and returns an interpolant in $\psi$.  For the later extension
-to handle a symmetric interpolation problem, the algorithm receives
-multiple formulas in $\psi$ and returns intermediate interpolants in
-$\psi$.  For computing those interpolants over linear arithmetics, we
-use Farkas's Lemma.
+$\psi$ and returns an interpolant in $\psi$.
 
@@
 $\lambda_1,\cdots,\lambda_n$ that satisfy
-\[a_{0,j} = \sum_{i=1}^n \lambda_i * a_{i,j} \quad (j=0, \ldots,m)\]
+\[a_{0,j} = \sum_{i=1}^n \lambda_i a_{i,j} \quad (j=0, \ldots,m)\]
 
@@
 reasonable computation time for any input.  It also aims to obtain a
-common solution among multiple interpolating problems.  These goals
-are set because the algorithm is designed to be used as a sub-routine
-for solving Horn clause satisfying problems
-that are made while discovering predicates for abstraction
-during the program verification.  These advantages of our algorithm
-make it possible to return a relatively small predicates for those
-problems. To accomplish these aims, the algorithm preserves a set of
-interpolants during the computation, and executes operations
-over interpolant sets.
+common solution among multiple interpolating problems.
+To accomplish these goals, the algorithm preserves a set of solution
+interpolants instead of one interpolant during the computation of sub
+problems, and executes operations over interpolant sets.
 
@@
 \end{align*}
-For convenience, we assume that any expression $e_i$ have variables
-$x_1, \ldots, x_a$.
 
@@
 interpolant set.
-
-\paragraph{Symmetric Interpolation}
-TODO

### 2013-09-26 article/conclusion.tex
Amendment
Interpolation

@@
 
-In this thesis, we have suggested a new algorithm to discover
+In this thesis, we have suggested new algorithms to discover
 predicates that are suitable for program verification.  We put a
-hypothesis that the program invariants are prone to be expressed in
-simple expressions.  Based on this hypothesis, our algorithm tries to
-obtain a small predicate with a fewer number of terms.
+hypothesis that program invariants are prone to be expressed in simple
+expressions.  Based on this hypothesis, our algorithm tries to obtain
+a small predicate with a fewer number of terms.
 

### 2013-09-26 article/experiment.tex
Interpolation

@@
 quality of resulting interpolants with
-ones that \testsc{CSIsat} \cite{conf/cav/BeyerZM08} generates.
+ones that \textsc{CSIsat} \cite{conf/cav/BeyerZM08} generates.
 It combines linear arithmetic reasoning and SAT-based logical reasoning.
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1]{figures/int1.eps}
+\end{center}
+\caption{Interpolation}
+\end{figure}
+
 \bigskip\centerline{}
@@
 \end{tabular}
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1]{figures/int2.eps}
+\quad
+\includegraphics[scale=1]{figures/int3.eps}
+\end{center}
+\caption{Interpolation}
+\end{figure}
 
@@
 \begin{center}
-\includegraphics[scale=.6]{figures/chart.png}
+\includegraphics[scale=.6]{figures/unsat2.eps}
+\quad
+\includegraphics[scale=.6]{figures/unsat3.eps}
 \end{center}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=.8]{figures/time.eps}
-\end{center}
-\caption{Time consumption ration of Yhorn compared to \textsc{CSIsat}}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\includegraphics[scale=.8]{figures/time.eps}
+%\end{center}
+%\caption{Time consumption ration of Yhorn compared to \textsc{CSIsat}}
+%\end{figure}
 

### 2013-09-26 article/interpolation.tex
Interpolation

@@
 Among various program verification techniques, the interpolation has
-been used to extract a program behavior as a logical formula.
-Especially when one adopts software model checking technique with
-predicate abstraction, the interpolation computes the abstraction
-predicate along a spurious counterexample that a model checker
-discovered.
+been used to extract program behavior as a logical formula.
+Especially when one adopts software model checking with
+\emph{predicate abstraction}, a technique to abstract a program by a
+Boolean valuation of given abstraction predicates, an interpolating
+theorem prover plays a role to refine the program abstraction by
+computing such predicates along a spurious counterexample that a model
+checker discovered.
 
-It is done by computing a logical separation between the strongest
-postcondition at the location and the weakest precondition to the
-failure point at every program location.
+This process is done by computing a logical separation between the
+strongest postcondition and the weakest precondition to the failure
+point at every program location.  We apply Craig interpolation for
+obtaining such logical separation.
 
-However, by the existing interpolating theorem prover, we may not be
-able to obtain a simple solution.  Consider the example below.
+\paragraph{Craig Interpolation}
+Given two logical formulas $A$ and $B$ that are inconsistent each
+other, we call a new preposition $I$ a \emph{Craig interpolant}
+\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$ is
+implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
+only free variables that appears in both $A$ and $B$.  If $A \wedge B$
+is unsatisfiable, an interpolant $I$ always exists.
+\vspace{10pt}
+
+However, by existing interpolating theorem provers, we may not be able
+to obtain a simple separation.  Consider the example below.
 \begin{align*}
-\varphi_1 : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
-\varphi_2 : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
+A : (x = 0 \wedge y = 0) \vee (x = 1 \wedge y = 1) \\
+B : (x \neq 0 \wedge y = 0) \vee (x \neq 1 \wedge y = 1)
 \end{align*}
 Although one may be able to obtain $x = y$ as the simplest
-interpolation between $\varphi_1$ and $\varphi_2$,
-CSIsat\cite{conf/cav/BeyerZM08} returns
+interpolation between $A$ and $B$, CSIsat\cite{conf/cav/BeyerZM08}
+returns
 \begin{align*}
-& ((((y \leq 0 \wedge -y \leq 0) \vee 1 = x) \wedge 1 \neq x) \vee ((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 \neq x) \wedge \\
-& \qquad (((0 \neq x \vee y \leq 1) \wedge (0 = x \vee y \leq 1)) \vee 1 \neq x) \wedge \\
-& \qquad (((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 = x) \wedge \\
-& \qquad ((1 \neq x \wedge 1 = x) \vee (y \leq 0 \wedge -y \leq 0) \vee 0 \neq x) \wedge \\
-& \qquad (1 = x \vee y \leq 0)
+I: \quad & ((((y \leq 0 \wedge -y \leq 0) \vee 1 = x) \wedge 1 \neq x) \vee ((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 \neq x) \wedge \\
+& (((0 \neq x \vee y \leq 1) \wedge (0 = x \vee y \leq 1)) \vee 1 \neq x) \wedge \\
+& (((1 \neq x \vee -y \leq -1) \wedge 1 = x) \vee 0 = x) \wedge \\
+& ((1 \neq x \wedge 1 = x) \vee (y \leq 0 \wedge -y \leq 0) \vee 0 \neq x) \wedge \\
+& (1 = x \vee y \leq 0)
 \end{align*}
-as a solution.
+as a solution.  By the complex solution, a program verification
+process may continue to abstract a program by complex predicate,
+causing the verification not to terminate.
 
-We propose an interpolation algorithm which returns as simple a
-solution as possible in terms of the number of conjunctions and
-disjunctions.  Our algorithm focuses on interpolating problems on
-linear arithmetic and it works in a constructive manner.
+We put a hypothesis that program invariants are often expressed in
+simple expressions. Under the hypothesis, we propose an interpolation
+algorithm which returns as simple a solution as possible in terms of
+the number of conjunctions and disjunctions.  Our algorithm focuses on
+interpolating problems on linear arithmetic.
+
+\section{Background}
+
+For ease of understanding, we review a method of software model
+checking with predicate abstraction\cite{conf/cav/GrafS97} in
+CounterExample-Guided Abstraction Refinement (CEGAR) framework.
+
+Consider an example program below.
+
+\begin{alltt}
+    a = 0;
+    b = 0;
+\(lbl\):
+    // equivalent to assert (a == b);
+    if (a != b) fail;
+\end{alltt}
+
+At first of the program verification, a predicate abstraction process
+translate the input program with empty set of abstraction predicates.
+Under such abstraction, the program is abstracted equivalent as a
+following listing and a model checker discovers a spurious
+counterexample which fails at the the last $fail$ statement.
+
+\begin{alltt}
+    \(\star\)
+    \(\star\)
+\(lbl\):
+    if (\(\star\)) fail;
+\end{alltt}
+
+In this example, it is sufficient to compute a predicate $I$ that is
+satisfied at the label $lbl$, and does not lead to the assertion
+failure.  Namely,
+
+\begin{align*}
+a = 0 \wedge b = 0 & \vdash I \\
+I \wedge (a \neq b) & \vdash \bot
+\end{align*}
+
+Here an interpolating theorem prover is used to compute a
+Craig interpolant.  A prover may return $I: (a = b)$ as a solution.
+By using this newly obtained predicate, the abstraction is refined as follows.
+
+\begin{alltt}
+    // p reprensents the state \(a=b\)
+    p = true;
+\(lbl\):
+    if (\(\neg\)p) fail;
+\end{alltt}
+
+After this refinement, the model checker does not discover the
+counterexample and program verificaction succeeds.
 
@@
 
-Consider the following interpolation problem.
+We briefly explain the mechanism of our algorithm. Consider the
+following interpolation problem between $A$ and $B$. Note that
+$a,b,x,y$ are variables.
 
@@
 
-Here, the linear arithmetic formulas A and B are inconsistent.  The
-interpolant I for this problem is $x-y+1 \leq 0$, which is implied by
+Here, the linear arithmetic formulas $A$ and $B$ are inconsistent.  The
+interpolant $I$ for this problem is $x-y+1 \leq 0$, which is implied by
 $A$ and inconsistent with $B$.  The vocabulary of $I$ is $\left\lbrace
@@
 Our algorithm aims to compute a Craig interpolant over linear
-arithmetic formulas.
-
-\paragraph{Craig Interpolation}
-Given two logical formulas $A$ and $B$ that are inconsistent each
-other, we call a new preposition $I$ a \emph{Craig interpolant}
-\cite{journals/jsyml/Craig57} between $A$ and $B$ such that $I$ is
-implied by $A$ and inconsistent with $B$.  $I$'s vocabulary must be
-only free variables that appears in both $A$ and $B$.  If $A \wedge B$
-is unsatisfiable, an interpolant $I$ always exists.
-
-Our algorithm handles a linear expression $e$ which is in a form:
+arithmetic formulas.  We handle a linear expression $e$ which is in a
+form:
 \[ a_1 x_1 + \cdots + a_n x_n + b \leq 0 \]
@@
 \[a_{0,j} = \sum_{i=1}^n \lambda_i a_{i,j} \quad (j=0, \ldots,m)\]
+\vspace{10pt}
 
-For the special case of Farkas's Lemma, we can state that:
+Especially, our algorithm often uses the special case of Farkas's Lemma.
 
+\paragraph{Farkas's Lemma to induce contradiction}
 Let a linear inequality $e_i$ be represented as
@@
 $\lambda_1,\cdots,\lambda_n$ that satisfy
-\[0 = \sum_{i=1}^n \lambda_i a_{i,j} \quad (j=1, \ldots, m) \wedge
-0 > \sum_{i=1}^n \lambda_i a_{i,0}\]
-
-Our interpolating algorithm make use of the latter version of Farkas's
-Lemma.
+\begin{align*}
+& 0 = \sum_{i=1}^n \lambda_i a_{i,j} \quad (j=1, \ldots, m) \wedge \\
+& 0 > \sum_{i=1}^n \lambda_i a_{i,0}\
+\end{align*}
 
@@
 \leq 0$.  We assume that conjunctions of $A$ and $B$ are inconsistent.
-For later convenience, we call this interpolating problem $\left( A, B
-\right)$.  Then there exists an interpolating linear expression
-$e_\star$ which satisfies
+For later convenience, we refer this interpolating problem as $\left(
+A, B \right)$.  Because $A \wedge B$ is contradiction, according to
+Craig's Interpolation Theorem, there exists an interpolating linear
+expression $e_\star$ which satisfies
 \begin{align*}
@@
 According to the Farkas's lemma, there exists a set of assignments to
-$\lambda_i$ to conclude $\bot$ from a set of linear expressions $A
-\cup B$, namely,
+$\lambda_i$ to build $\bot$ from a linear combination of expressions
+in $A$ and $B$, namely,
 \begin{align*}
@@
 above will let an expression $I$ to be an interpolant.  Considering
-this, it is possible to obtain different interpolants for an
+that, it is possible to obtain different interpolants for an
 interpolating problem by computing different assignments to
@@
 interpolating problems $\left(A_1, B_1 \right)$ and $\left(A_2, B_2
-\right)$, the problem is about to compute the common interpolant I,
+\right)$, the problem is about to compute the common interpolant $I$,
 which is an interpolant for the first problem and the second one at
@@
 
-This problem is used when solving a following Horn clause solving
-problem, for instance.
-\[
-\implies
-% TODO
-\]
-
 To accomplish this, we simply need to compute the intersection of
-interpolant sets.  Let us call the interpolant space for the problems
-$\left(A_1, B_1 \right)$ and $\left(A_2, B_2 \right)$,
-$\mathcal{I}_1$, and $\mathcal{I}_2$.  In the previous method, a set
-of interpolants was represented by a parameterized linear expression.
+interpolant sets.  We use $\mathcal{I}_1$ and $\mathcal{I}_2$ to
+denote the interpolant space for the problems $\left(A_1, B_1 \right)$
+and $\left(A_2, B_2 \right)$ respectively.  In the previous method, a
+set of interpolants was represented by a parameterized linear
+expression.
 \begin{align*}
@@
 
-However, if such an assignment of $\mathbf{a}$ and $b$ does not exist,
-no common solution exists across two problems because $\mathcal{I} =
-\emptyset$.
-
-By applying this approach, it is also possible to compute a common
-interpolant among more than three problems.
+However, if such an assignment of $\mathbf{a}$ and $b$ does not exist
+and $\mathcal{I}$ becomes $\emptyset$, no common solution exists
+across two problems.
 
 \paragraph{Interpolation between logical formulas}
-We now extend the algorithm so that it can handle general logical
-formulas over linear arithmetic including disjunctions of linear
-expressions.  Here the algorithm takes two logical formulas over
-linear expressions as a problem input.  For convenience, we assume
-that they are represented in disjunctive normal form (DNF).  The
-solution for this problem is the interpolant that is a logical formula
-of linear expressions.
+We now extend the algorithm so that it can handle more general
+problems.  We now treat the problem of general linear arithmetic
+formulas including disjunctions.
 
-First, consider the case of an interpolation problem between $A_1 \vee
-A_2$ and $B$.  When $(A_1 \vee A_2) \wedge B$ is inconsistent, $A_1
-\wedge B$ is necessarily inconsistent.  Let $\mathcal{I}_1$ denote the
-set of interpolants for $(A_1, B)$.  In the same manner, let
-$\mathcal{I}_2$ denote the set of interpolants for $(A_2, B)$.  We may
-naturally obtain an interpolant set by the intersection $\mathcal{I} =
-\mathcal{I}_1 \cup \mathcal{I}_2$ as a solution for $(A_1 \vee A_2,
-B)$.  When the algorithm computes $\mathcal{I}$ by the previous
-method, and it is $\emptyset$, we need to build a less simple solution
-by
+The algorithm takes two formulas over linear expressions as a problem
+input.  For convenience, we assume that they are represented in
+disjunctive normal form (DNF).  The solution interpolant for this
+problem is a logical formula of linear expressions.
+
+First, consider the case of an interpolation problem
+$(A_1 \vee A_2, B)$.  When $(A_1 \vee A_2) \wedge B$ is inconsistent,
+$A_1 \wedge B$ is necessarily inconsistent.  We use $\mathcal{I}_1$ to
+denote the set of interpolants for $(A_1, B)$.  In the same manner, we
+use $\mathcal{I}_2$ for the set of interpolants for $(A_2, B)$.  We
+may naturally obtain an interpolant set by the intersection
+$\mathcal{I} = \mathcal{I}_1 \cap \mathcal{I}_2$ as a solution for
+$(A_1 \vee A_2, B)$.  When the algorithm computes $\mathcal{I}$ by the
+previous method and it turnes to be $\emptyset$, we need to build a
+less simple solution by
 \[ \mathcal{I}' = \left\lbrace \left( I_1 \vee I_2 \right) \mid
 I_1 \in \mathcal{I}_1 \wedge I_2 \in \mathcal{I}_2 \right\rbrace \]
-Any formulas in the set $\mathcal{I}'$ are the interpolant for $(A_1
-\vee A_2, B)$ because
+Any formulas in the set $\mathcal{I}'$ are the interpolant for
+$(A_1 \vee A_2, B)$ because
 \begin{align*}
 \forall I \in \mathcal{I}. & A_1 \vdash I \wedge A_2 \vdash I \wedge I \wedge B \vdash \bot \wedge \\
-& FV(I) = FV(A_1) \cup FV(A_2) \cup FV(B)
+& FV(I) = FV(A_1) \cap FV(A_2) \cap FV(B)
 \end{align*}
 
-Second, consider the case of an interpolation problem between $A$ and
-$B_1 \wedge B_2$.
-% TODO:
+Second, consider the case of an interpolation problem
+$(A, B_1 \wedge B_2)$.  Similar to the previous case, we use
+$\mathcal{I}_1, \mathcal{I}_2$ to denote the set of interpolants for
+$(A, B_1)$ and $(A, B_2)$.  We may obtain an common interpolant set by
+$\mathcal{I} = \mathcal{I}_1 \cap \mathcal{I}_2$ as a solution for
+$(A, B_1 \wedge B_2)$.  If $\mathcal{I} = \emptyset$, we build a less
+simple solution by
+\[ \mathcal{I}' = \left\lbrace \left( I_1 \wedge I_2 \right) \mid
+I_1 \in \mathcal{I}_1 \wedge I_2 \in \mathcal{I}_2 \right\rbrace \]
+Any formulas in the set $\mathcal{I}'$ are the interpolant for
+$(A, B_1 \wedge B_2)$ because
+\begin{align*}
+\forall I \in \mathcal{I}. & A \vdash I \wedge I \wedge B_1 \vdash \bot \wedge I \wedge B_2 \vdash \bot \wedge \\
+& FV(I) = FV(A) \cap FV(B_1) \cap FV(B_2)
+\end{align*}
 
-By combining these two ways of interpolant construction, we may be
-able to build a small solution for more general cases like $(A_1,
-\ldots, A_n)$ and $(B_1, \ldots, B_n)$.  In order to obtain a small
-solution, we need to wisely choose the order of combining an
-interpolant set.
+By combining these two ways of interpolant construction wisely, we may
+be able to build a small solution for more general cases like
+$(A_1, \ldots, A_n)$ and $(B_1, \ldots, B_n)$.
+
+Recently \cite{conf/cav/AlbarghouthiM13} have independently suggested
+a non-deterministic method to combine these ways to build a small
+solution in a reasonable time.  It iteratively tests the non-emptiness
+of a common interpolant set $\mathcal{I}$ by heuristics.

### 2013-09-26 article/interpolation.tex
textsc

@@
 Although one may be able to obtain $x = y$ as the simplest
-interpolation between $A$ and $B$, CSIsat\cite{conf/cav/BeyerZM08}
-returns
+interpolation between $A$ and $B$,
+\textsc{CSIsat}\cite{conf/cav/BeyerZM08} returns
 \begin{align*}
@@
 \begin{alltt}
-    \(\star\)
-    \(\star\)
+    // Do nothing
 \(lbl\):
+    // Non deterministic branch
     if (\(\star\)) fail;

### 2013-09-26 article/intro.tex
textsc

@@
 
-TODO: Mention experiments
+To confirm the effectiveness, we incorporated our algorithm into
+existing software model checker \textsc{MoCHi}
+\cite{conf/pldi/KobayashiSU11}.  We replaced the conventional Horn
+clause solvers with our algorithms and tested some example programs to
+verify the performance.  As a result, our algorithm was able to
+compute a simple solution for most Horn clause solving problems.
+Although we observed some test cases have experienced deterioration in
+the number of verification cycles, there are some cases that are newly
+verified by \textsc{MoCHi}.  Additionally, our algorithm improved a
+performance in more than 70 percent of problems and achieved faster
+computation than conventional method.  Eventually we have verified
+certain effectiveness of our approach.
 

### 2013-09-26 article/jabstract.tex
textsc

@@
 近年のプログラム検証技術においては、自動定理証明が重要な役割を担っている。特にソフトウェア モデル検査の分野では、補間問題およびその一般化である Horn 節充足問題が注目を集めてきた。後者の問題は、循環した依存関係を持たない、未知の述語変数を複数含む Horn 節の集合の解を求める問題である。Horn 節充足問題の解、すなわち Horn 節集合を恒真とするような述語変数への代入は、プログラムの不変式の候補として用いられる。しかしながら、Horn 節充足問題の既存解法が出力する解は複雑かつ特殊すぎて不変式にならないことが多い。
-上述の問題を解決するために、われわれは線形代数上の Horn 節充足問題を解く新しいアルゴリズムを提案する。われわれのアルゴリズムは、プログラム不変式が単純な式で表されるという仮定のもとで、小さな述語の生成を試みる。この目的を達成するため、このアルゴリズムはある大きさの線形不等式の雛形を未知の述語変数それぞれに対して生成し、Horn 節充足問題を線形計画法に帰着させる。そして、解が見つかるまで雛形の拡大を繰り返す。この提案手法では、合理的な時間内に比較的小さな解を構成することができる。われわれは、このアルゴリズムを実装し関数型言語向けのソフトウェア モデル検査器 MoCHi に組み込んだ。実験により、提案手法について一定の有効性を確認した。
+上述の問題を解決するために、われわれは線形代数上の Horn 節充足問題を解く新しいアルゴリズムを提案する。われわれのアルゴリズムは、プログラム不変式が単純な式で表されるという仮定のもとで、小さな述語の生成を試みる。この目的を達成するため、このアルゴリズムはある大きさの線形不等式の雛形を未知の述語変数それぞれに対して生成し、Horn 節充足問題を線形計画法に帰着させる。そして、解が見つかるまで雛形の拡大を繰り返す。この提案手法では、合理的な時間内に比較的小さな解を構成することができる。われわれは、このアルゴリズムを実装し関数型言語向けのソフトウェア モデル検査器 \textsc{MoCHi} に組み込んだ。実験により、提案手法について一定の有効性を確認した。
 \end{jabstract}

### 2013-09-26 article/related.tex
textsc

@@
 for the model checking by CEGAR with predicate abstraction.  The
-former method is adopted when analyzing input programs by denotational
+former method is adopted when analyzing input programs by axiomatic
 semantics such as Hoare logic, and synthesizing invariants from
@@
 interpolant from counterexample's refutation proof derivation tree.
-This work implemented the proposed method into the Blast software
-model checker \cite{conf/popl/HenzingerJMS02,journals/sttt/BeyerHJM07}
-and verified the effectiveness of interpolants for predicate
-abstraction through an experiment of C program verification.
+This work implemented the proposed method into the \textsc{Blast}
+software model checker
+\cite{conf/popl/HenzingerJMS02,journals/sttt/BeyerHJM07} and verified
+the effectiveness of interpolants for predicate abstraction through an
+experiment of C program verification.
 
@@
 generation over the theory of linear arithmetic and uninterpreted
-function symbols (LI+UIF) theory \cite{journals/tcs/McMillan05}.  At the same
-time, various applications of interpolants for model checking was
-proposed \cite{conf/tacas/McMillan05}. They incorporated the algorithm
-into Z3 SMT solver \cite{conf/tacas/MouraB08} and it is possible to
-generate an interpolant from proofs generated by
-Z3 \cite{conf/fmcad/McMillan11}.
+function symbols (LI+UIF) theory \cite{journals/tcs/McMillan05}.  At
+the same time, various applications of interpolants for model checking
+was proposed \cite{conf/tacas/McMillan05}. They incorporated the
+algorithm into \textsc{Z3} SMT solver \cite{conf/tacas/MouraB08} and
+it is possible to generate an interpolant from proofs generated by
+\textsc{Z3} \cite{conf/fmcad/McMillan11}.
 
@@
 verifying higher-order functional programs.  Interpolation is used as
-a sub procedure of solving Horn clauses for dependent type
-inference \cite{conf/ppdp/UnnoK09, conf/pepm/SatoUK13} in this
-context. In MoCHi, the refinement types are given for values and
+a sub procedure of solving Horn clauses for dependent type inference
+\cite{conf/ppdp/UnnoK09, conf/pepm/SatoUK13} in this context. In
+\textsc{MoCHi}, the refinement types are given for values and
 variables for program abstraction.  Horn clause solving is necessary
 for inferring the refinment types.  \cite{conf/popl/Terauchi10}
-suggested a method for dependent type inference algorithm as
-well. Their algorithm is complete in the sense that it eventually
-returns a solution which terminates the verification process.
+suggested a method for dependent type inference algorithm as well.
 

### 2013-09-26 article/horn.tex
Small fixes
Stash
Interpolation
Horn

--- 2013-08-28 article/horn.tex
@@
 \label{chap:horn}
+
+The problems are generated by encoding constraints over execution
+paths with unknown predicates which describe functions' constraints
+between parameters and return values.
 
@@
 This construction of the algorithm enables us to compute a relatively
-simpler solution compared to existing algorithms while preserving the
+simpler solution compared to existing algorithms while restraining the
 computation cost in a reasonable time.
@@
 
+\begin{figure}\label{fig:ex1}
+\begin{center}
+\includegraphics[scale=0.8]{figures/ex1.eps}
+\end{center}
+\vspace{-18pt}
+\caption{Problem visualization for tree interpolation example}
+\end{figure}
+
 \paragraph {Tree interpolation}
 First, consider the following example.
-
 \begin{align*}
@@
 \end{align*}
-
 In this example, we would like to compute linear arithmetic formulas
 for the predicate variables $A$ and $B$.  They appear only once on
-both left-hand and right-hand globally in the problem.
+both left-hand and right-hand globally in the problem.  For ease of
+understanding, we visualized the problem by a graph in Figure~\ref{fig:ex1}.
 
@@
 parameterized.  Assume that they have following predicate templates.
-
 \begin{align*}
@@
 obtain a linear constraint
-$\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0$.
+$(\lambda_1 = c_1) \wedge (- \lambda_1 \geq c_2) \wedge (\lambda_1 \geq 0)$.
 Any assignment to $ c_1, c_2, \lambda_1 $ which satisfies this will
 make $A(x)$ valid for the first clause.  In the same manner, we obtain
-$\lambda_2 = c_3 \wedge - 2 \lambda_2 \geq c_4 \wedge \lambda_2 \geq 0$
+$(\lambda_2 = c_3) \wedge (- 2 \lambda_2 \geq c_4) \wedge (\lambda_2 \geq 0)$
 from the second clause, and 
-$c_1 = 3 \lambda_3 \wedge c_3 = 2 \lambda_3 \wedge c_2 + c_4 \geq -10 \lambda_3 \wedge \lambda_3 > 0$
+$(c_1 = 3 \lambda_3) \wedge (c_3 = 2 \lambda_3) \wedge (c_2 + c_4 \geq -10 \lambda_3) \wedge (\lambda_3 > 0)$
 from the third.
@@
 One of the models satisfying all the constraints above is
-$( c_1, c_2, c_3, c_4, \lambda_1, \lambda_2, \lambda_3 ) = (3, -3, 2, -4, 3, 2, 1)$.
+\[( c_1, c_2, c_3, c_4, \lambda_1, \lambda_2, \lambda_3 ) = (3, -3, 2, -4, 3, 2, 1) \].
 By assigning them to the template, we obtain the following solution.
-
 \begin{align*}
@@
 \paragraph {DAG problem solving}
+
+\begin{figure}\label{fig:ex21}
+\begin{center}
+\includegraphics[scale=0.8]{figures/ex2-1.eps}
+\end{center}
+\caption{Problem visualization for DAG structured example}
+\end{figure}
+
 Second, we explain a case that some predicate variables appear
 multiple times on left-hand side in different Horn clauses.
-
 \begin{align*}
@@
 \end{align*}
+Problem visualization is available in Figure~\ref{fig:ex21}. Note
+that the graph now takes a DAG structure.
 
@@
 constraint for the first clause as the previous example.
-
-\begin{align} \label{eq:constr1}
-\lambda_1 = c_1 \wedge - \lambda_1 \geq c_2 \wedge \lambda_1 \geq 0
-\end{align}
+\begin{align*} \label{eq:constr1}
+(\lambda_1 = c_1) \wedge (- \lambda_1 \geq c_2) \wedge (\lambda_1 \geq 0)
+\end{align*}
 
@@
 multiple occurrences in order to compute the simplest solution, it may
-make the whole constraints unsatisfiable if the simplest solution does
-not exist.
+generally make the whole constraints unsatisfiable if the simplest
+solution does not exist.
 
-Therefore, we generate a weaker constraint for the second clause that
-allows different expressions of $A$ for multiple occurrence.  If the
-whole constraint is satisfiable in the case, we try to solve in a
-stronger constraint to generate a simple solution.
+The above case eventually generate the constraint
+\begin{align*}
+& (\lambda_1 = c_1) \wedge (- \lambda_1 \geq c_2) \wedge (\lambda_1 \geq 0) \wedge \\
+& (c_1 = \lambda_2) \wedge (2 c_2 \geq 2 \lambda_2) \wedge (\lambda_2 > 0)
+\end{align*}
+and we may get a model $(c_1, c_2, \lambda_1) = (1, -1, 1)$ to
+obtain $Ax - 1 \leq 0$ for $A(x)$.
+
+\begin{figure}[htb]
+  \begin{minipage}[t]{.47\textwidth}
+  \centering
+  \includegraphics[scale=0.8]{figures/ex2-2.eps}
+  \label{fig:ex22}
+  \caption{Interpolation}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[t]{.47\textwidth}
+  \centering
+  \includegraphics[scale=0.8]{figures/ex2-3.eps}
+  \label{fig:ex23}
+  \caption{Interpolation}
+  \end{minipage}
+\end{figure}
+
+Now we slightly change the problem.
+\begin{align*}
+x \leq 1 & \implies A(x) \\
+A(x) \wedge A(y) & \implies x+2y \leq 3
+\end{align*}
+It is visualized in Figure~\ref{fig:ex22}.  The modifed constraint is
+\begin{align*}
+& (\lambda_1 = c_1) \wedge (- \lambda_1 \geq c_2) \wedge (\lambda_1 \geq 0) \wedge \\
+& (c_1 = \lambda_2) \wedge (c_1 = 2\lambda_2) \wedge (2 c_2 \geq 3 \lambda_2) \wedge (\lambda_2 > 0)
+\end{align*}
+No solution exists for this constraint.  Therefore we generate a
+weaker constraint for the second clause that allows different
+expressions of $A$ for multiple occurrence.  If the whole constraint
+is satisfiable in the case, we try to solve in a stronger constraint
+to generate a simple solution.
 
@@
 following problem in a weaker constraint.
-
 \begin{align*}
@@
 \end{align*}
-
-By duplicating the constraint~\ref{eq:constr1} of $A$ for newly
-created $A_1$ and $A_2$, we obtain the whole constraint as follows.
-
+Figure~\ref{fig:ex23} illustrates above.  By duplicating the
+constraint~\ref{eq:constr1} of $A$ for newly created $A_1$ and $A_2$,
+we obtain the whole constraint as follows.
 \begin{align*}
@@
 disjunctions on left-hand side.
-
+\begin{figure}\label{fig:ex31}
+  \begin{center}
+    \includegraphics[scale=0.8]{figures/ex3-1.eps}
+  \end{center}
+  \caption{Visualization of disjunctive Horn clause problem}
+\end{figure}
 \begin{align*}
@@
 \end{align*}
-
-Same as the previous examples, templates for predicate variables are
-prepared.
-
+See Figure~\ref{fig:ex31}.  Same as the previous examples, templates
+for predicate variables are prepared.
 \begin{align*}
@@
 \end{align*}
-
 The constraints generated are follows.
-
 \begin{align*}
@@
 \end{align*}
-
 The constraints here become unsatisfiable, and one possible
 unsatisfiable core is:
-
 \begin{align*}
@@
 
-Now we slightly change the problem and the template.
-
+Now we slightly change the problem and the template as in Figure~\ref{fig:ex32}.
 \begin{align*}
@@
 
+\begin{figure}\label{fig:ex32}
+  \begin{center}
+    \includegraphics[scale=0.8]{figures/ex3-2.eps}
+  \end{center}
+  \caption{Example}
+\end{figure}
 When we construct the constraints, as same as previous, we again
 obtain an unsatisfiable core:
-
 \begin{align*}
@@
 \text{Unsatisfiable core:} \\
-0 \geq c_2 \\
-& \lambda_3 = c_{3,1} \wedge \lambda_3 \geq c_{4,1} \wedge \\
-& - \lambda_4 = c_{3,2} \wedge \\
-& c_1 + c_{3,1} = 0 \wedge c_2 + c_{4,1} > 0 \wedge \\
+& 0 \geq c_2 \\
+& \lambda_3 = c_{3,1} \\
+& \lambda_3 \geq c_{4,1} \\
+& - \lambda_4 = c_{3,2} \\
+& c_1 + c_{3,1} = 0 \\ & c_2 + c_{4,1} > 0 \\
 & c_1 + c_{3,2} = 0
 \end{align*}
-
 This time, the unsatisfiable core contains the constraints that
-originate from
-
+originate from conjunction.  We need to modify the problem again.
 \begin{align*}
@@
 \end{align*}
+See Figure~\ref{fig:ex33}.
+
+\begin{figure}\label{fig:ex33}
+  \begin{center}
+    \includegraphics[scale=0.8]{figures/ex3-3.eps}
+  \end{center}
+  \caption{Example}
+\end{figure}
 
 After this, we obtain the following constraint:
-
 \begin{align*}
@@
 \end{align*}
-
-By continuing this process, the process in the end generates a
-predicate template that the solution will become.  Eventually, we
-obtain the following solution.
-
+By solving this, we eventually obtain the following solution.
 \begin{align*}
@@
 \begin{align*}
-% TODO:
 & \exists v_a, v_b \in V_T; v_P \leadsto v_a \wedge v_P \leadsto v_b \wedge \\
@@
 
-<fig...>
+\vspace{48pt}
+\begin{figure}[htb]\label{fig:conj1}
+  \begin{minipage}[t]{.47\textwidth}
+  \centering
+  \includegraphics[scale=0.8]{figures/conj1.eps}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[t]{.47\textwidth}
+  \centering
+  \includegraphics[scale=0.8]{figures/conj2.eps}
+  \end{minipage}
+  \caption{Conjunctive split}
+\end{figure}
 
@@
 
-<fig...>
+\vspace{48pt}
+\begin{figure}[htb]\label{fig:disj1}
+  \begin{minipage}[t]{.47\textwidth}
+  \centering
+  \includegraphics[scale=0.8]{figures/disj1.eps}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[t]{.47\textwidth}
+  \centering
+  \includegraphics[scale=0.8]{figures/disj2.eps}
+  \end{minipage}
+  \caption{Disjunctive split}
+\end{figure}
 
@@
 
-Consider these two ways of transformation, we show a pseudo-code for
-\linebreak \transGraph~procedure.  The procedure is given the graph
-$G$, the predicate templateset $T$, and the unsatisfiable core
-$\mathcal{U}$. It returns the transformed graph $G'$ and the new
-templateset $T'$.
+%Consider these two ways of transformation, we show a pseudo-code for
+%\linebreak \transGraph~procedure.  The procedure is given the graph
+%$G$, the predicate templateset $T$, and the unsatisfiable core
+%$\mathcal{U}$. It returns the transformed graph $G'$ and the new
+%templateset $T'$.
 
-\begin{algorithm}
-\caption{$ \transGraph (G, T, \mathcal{U}) $}\label{alg:transGraph}
-\begin{algorithmic}
+%\begin{algorithm}
+%\caption{$ \transGraph (G, T, \mathcal{U}) $}\label{alg:transGraph}
+%\begin{algorithmic}
 % TODO:
-\RETURN {$\left( G', T' \right)$}
-\end{algorithmic}
-\end{algorithm}
+%\RETURN {$\left( G', T' \right)$}
+%\end{algorithmic}
+%\end{algorithm}
 
@@
 expressions.
-% TODO: Can't we replace e_i with actual expressions?
 \begin{align*}
@@
 problems separately.
-\begin{itemize}
+\vspace*{-8pt}\begin{itemize}\itemsep-3pt
 \item $e_1 \implies P_1 \qquad e_3 \implies Q_1 \qquad P_1 \wedge Q_1 \implies \bot$
@@
 Then the solution is built by follows.
-\begin{itemize}
+\vspace*{-4pt}\begin{itemize}\itemsep-3pt
 \item $P = ( P_1 \vee P_2 ) \wedge ( P_3 \vee P_4 )$
@@
 \end{itemize}
-However, this method is not efficient.
+However, it is obvious that this method is not efficient.
 
@@
 coefficient parameter $\mathbf{a}_P$ and the constant parameter $b_P$
-are eliminated.  Due to the constraints generation process, the
-simplified constraint is also linear and can be efficiently check the
-satisfiability.
+are eliminated.  Because the constraints are conjunctive, the
+qunatifier elimination is reduced to the problem of polyhedron
+projection.
 
@@
     \IF {$\varphi (v_p)$ is a predicate variable}
-      \STATE {$C \gets C \cup \genConstr (G, v_p, visited)$}
+      \STATE {$C' \gets \genConstr (G, v_p, V_\star)$}
+      \IF {$v_p \in V_\star$}
+        \STATE {$C' \gets \textsc{Simplify}(C')$}
+      \ENDIF
+      \STATE {$C \gets C \cup C'$}
     \ELSE
-      \STATE { $id$ }
+      \STATE {$C \gets C \cup \genConstr (G, v_p, V_\star)$}
     \ENDIF
@@
 \ENDFOR
+\RETURN {$C$}
 \end{algorithmic}

### 2013-09-26 article/experiment.tex
Experiment
Fix reference

@@
 interpolation \cite{conf/ppdp/UnnoK09}.  Second, we completely
-replaced MoCHi's Horn solving algorithm with our algorithm.
+replaced \textsc{MoCHi}'s Horn solving algorithm with our algorithm.
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=1.1]{figures/int1.eps}
+\end{center}
+\caption{Complex problem 1}
+\label{fig:int1}
+\end{figure}
+
 Finally, we showed some test cases that may occur in practical interpolation.
-
+\begin{itemize}
+\item Problem in Figure~\ref{fig:int1}
 \[\begin{array}{r c}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=1]{figures/int1.eps}
-\end{center}
-\caption{Interpolation}
-\end{figure}
-
-\bigskip\centerline{}
-
+\item Problem in Figure~\ref{fig:int2}
 \[\begin{array}{r c}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=1]{figures/int2.eps}
-\quad
-\includegraphics[scale=1]{figures/int3.eps}
-\end{center}
-\caption{Interpolation}
-\end{figure}
-
-\bigskip\centerline{}
-
 \textsc{CSIsat}
@@
 
+\item Problem in Figure~\ref{fig:int3}
 \[\begin{array}{r c}
@@
 
+\end{itemize}
+
+\begin{figure}[htb]
+  \begin{minipage}[t]{.47\textwidth}
+  \centering
+  \includegraphics[scale=1.1]{figures/int2.eps}
+  \caption{Complex problem 2}
+  \label{fig:int2}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[t]{.47\textwidth}
+  \centering
+  \includegraphics[scale=1.1]{figures/int3.eps}
+  \caption{Complex problem 3}
+  \label{fig:int3}
+  \end{minipage}
+\end{figure}
+
 \section{Horn clause solving}
 
-As an Horn clause solver we incorporated our Horn clause solver to MoCHi.
+As an Horn clause solver we incorporated our Horn clause solver to
+existing software model checker \textsc{MoCHi}.  It generates a Horn
+clause solving problem while performing dependent type inference for
+input program.
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=.6]{figures/unsat2.eps}
-\quad
-\includegraphics[scale=.6]{figures/unsat3.eps}
-\end{center}
-\caption{Result obtained from \textsc{MoCHi} with Horn clause solver \textsc{CSIsat} and
-\textsc{Yhorn}}
-\end{figure}
+The previous \textsc{MoCHi} used Horn clause solving algorithm
+proposed by Unno et al. \cite{conf/ppdp/UnnoK09}.  Its implementation
+adopts \textsc{CSIsat} for solving interpolation problems that appear
+as sub problems of Horn clause solving.
 
-We obtained better result in some cases compared to the one we had in \textsc{CSIsat}.
+We configured \textsc{MoCHi} to use \textsc{TRecS} version 1.30 for
+higher order model checking of abstract model of input programs.  The
+experiments are performed for comparison between Unno's algorithm with
+convex hull optimization in predicates (\texttt{-gchi} option) and our
+algorithm (\texttt{-yhorn} option).  The test cases are available in
+\textsc{MoCHi}'s development repository or online web interface.
+
+For each test program, we compared the number of CEGAR cycles for
+verification.  Additionally, for each Horn clause solving problem, we
+inspected the quality of solutions.  The quality is evaluated by the
+number of conjunctions and disjunctions in total.  Finally, we
+compared the time consumed for solving each problem.
 

### 2013-09-26 article/experiment.tex
Time

@@
 
+\begin{center}
+  \begin{tabular}{ | l || c | c | c | l || c | c | }
+\hline
+Name & gchi & yhorn & \multirow{25}{*}{Defenders} & Name & gchi & yhorn \\ \hline
+ack & 1 & 1 & fold_fun_list & 2 & 99 \\ \hline
+a-copy-print & 6 & 99 & fold_left & 2 & 2 \\ \hline
+a-dotprod & 4 & 4 & fold_right & 2 & 2 \\ \hline
+a-init & 99 & 7 & harmonic & 1 & 1 \\ \hline
+bcopy & 3 & 3 & hors & 1 & 1 \\ \hline
+bcopy-without-size & 4 & 4 & hrec & 1 & 2 \\ \hline
+copy_intro & 2 & 99 & inc & 1 & 1 \\ \hline
+dotprod_ & 4 & 4 & isnil & 2 & 2 \\ \hline
+enc-rev_accum & 2 & 4 & iter & 1 & 1 \\ \hline
+enc-rev_append & 5 & 99 & length & 2 & 4 \\ \hline
+enc-zip & 2 & 2 & map & 1 & 1 \\ \hline
+enc-zip_map & 2 & 99 & mc91 & 2 & 2 \\ \hline
+enc-zip_map2 & 4 & 99 & mem & 3 & 3 \\ \hline
+enc-zip_unzip & 2 & 99 & mult & 3 & 4 \\ \hline
+enc-zip2 & 2 & 99 & nth & 3 & 99 \\ \hline
+enc-zipmap & 4 & 99 & nth0 & 3 & 2 \\ \hline
+exc-fact & 2 & 3 & queen & 3 & 4 \\ \hline
+fact_exn & 2 & 3 & rec_error & 2 & 99 \\ \hline
+fact_notpos & 2 & 2 & recursive & 1 & 2 \\ \hline
+fact_notpos-e & 1 & 1 & repeat & 2 & 2 \\ \hline
+fib & 2 & 99 & sigma_sum & 3 & 3 \\ \hline
+file & 6 & 99 & sum & 1 & 1 \\ \hline
+file1 & 1 & 1 & twice_rec & 99 & 3 \\ \hline
+ &  &  & zip & 3 & 2 \\ \hline
+\end{tabular}
+\end{center}
+
 %\begin{figure}
